---
title: 'OHI 2018 - Targeted harvest pressure'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

[REFERENCE RMD FILE: https://cdn.rawgit.com/OHI-Science/ohiprep/master/globalprep/np/v2017/targetharvest_dataprep.html]

#Summary
This analysis converts FAO capture production data into the OHI 2018 targeted harvest pressure data.  


#Updates from previous assessment
One more year of data

The species2group.csv file was udated: South America Sea Lion is not a cetacean therefore was assigned the pinniped order and excluded from the target list. 


***

#Data Source 
    http://www.fao.org/fishery/statistics/software/fishstatj/en#downlApp
     Release date: March 2018 
FAO Global Capture Production Quantity 1950_2016

**Downloaded**: Aug 1 2018

**Description**:  Quantity (tonnes) of fisheries capture for each county, species, year.

**Time range**: 1950-2016

***

```{r, eval=FALSE}

# load libraries, set directories
library(ohicore)  #devtools::install_github('ohi-science/ohicore@dev')
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(plotly)

## comment out when knitting
#setwd("globalprep/prs_targetedharvest/v2018")


### Load FAO-specific user-defined functions
source('../../../src/R/fao_fxn.R') # function for cleaning FAO files
source('../../../src/R/common.R') # directory locations


```


# Read in the raw data
This includes the FAO capture production data and a list of the "target" species.

```{r, eval=FALSE}

## FAO capture production data
fis_fao_csv <-  read.csv(file.path(dir_M, 'git-annex/globalprep/_raw_data/FAO_capture/d2018/Global_capture_production_Quantity_1950-2016.csv'))


# species list 
sp2grp <-  read.csv('raw/species2group.csv') %>%
  dplyr::filter(incl_excl == 'include') %>%
  dplyr::select(target, species); head(sp2grp)


```


# Clean the FAO data

```{r, eval=FALSE}

m <- fis_fao_csv %>%
  dplyr::rename(country = Country..Country.,
         species = Species..ASFIS.species.,
         area = Fishing.area..FAO.major.fishing.area.,
         Unit = Unit..Unit.) %>%
  dplyr::select(-Unit)

m <- m %>%
  tidyr::gather("year", "value", -(1:3)) %>%
  dplyr::mutate(year = gsub("X", "", year)) %>%
    fao_clean_data() 

m <- m %>%
  dplyr::mutate(species = as.character(species)) %>%
  dplyr::mutate(species = ifelse(stringr::str_detect(species, "Henslow.*s swimming crab"), "Henslow's swimming crab", species))


```

# Identify the target species
This analysis only includes target species.  The warning messages need to be checked and, if necessary, changes should be made to the raw/species2group.csv  

```{r, eval=FALSE}

# check for discrepancies in species list
## seals are no longer included (so these errors can be ignored):
spgroups <-  sort(as.character(unique(m$species)))
groups <-  c('turtle', 'seal', 'whale', 'sea lion', 'dolphin', 'porpoise')

for (group in groups) { #group='dolphin'
possibles <- spgroups[grep(group, spgroups)]
d_missing_l <-  setdiff(possibles, sp2grp$species)
  if (length(d_missing_l)>0){
    cat(sprintf("\nMISSING in the lookup the following species in target='%s'.\n    %s\n", 
                group, paste(d_missing_l, collapse='\n    ')))
  }
}

# check for species in lookup not found in data
l_missing_d <-  setdiff(sp2grp$species, spgroups)
if (length(l_missing_d)>0){
  cat(sprintf('\nMISSING: These species in the lookup are not found in the FAO data \n'))
  print(l_missing_d)
}


## filter data to include only target species ----
m2 <-  m %>%
  dplyr::filter(species %in% sp2grp$species)

unique(m2$area) # confirm these are all marine


```


# Summarize data

```{r, eval=FALSE}

# spread wide to expand years
m_w = m2 %>%
  tidyr::spread(year, value) %>%
  dplyr::left_join(sp2grp, by='species'); head(m_w)


# gather long by target
m_l = m_w %>%
  dplyr::select(-area) %>%
  tidyr::gather(year, value, -country, -species, -target, na.rm=T) %>%
  dplyr::mutate(year = as.integer(as.character(year))) %>%
  dplyr::arrange(country, target, year); head(m_l)

#Temporary note: data for Gabn goes oly to 2009


# explore Japan[210]
m_l %>% 
  dplyr::group_by(country, target, year) %>%
  dplyr::summarize(value = sum(value)) %>% 
  dplyr::filter(country == 'Japan', target == 'cetacean', year >= 2000) 

# summarize totals per region per year
m_sum = m_l %>%
  dplyr::group_by(country, year) %>%
  dplyr::summarize(value = sum(value, na.rm=TRUE)) %>%
  dplyr::filter(value != 0) %>%
  dplyr::ungroup(); head(m_sum) 

```

# Assign country names to OHI regions

```{r, eval=FALSE}


m_sum <- m_sum %>%
  dplyr::mutate(country = as.character(country)) %>%
  dplyr::mutate(country = ifelse(stringr::str_detect(country, "C.*te d'Ivoire"), "Ivory Coast", country))


### Function to convert to OHI region ID
m_sum_rgn <- name_2_rgn(df_in = m_sum, 
                       fld_name='country', 
                       flds_unique=c('year'))

# these are duplicates for the same region
dplyr::filter(m_sum_rgn, country %in% c("Guadeloupe", "Martinique"))

# They will be summed:
m_sum_rgn <- m_sum_rgn %>%
  dplyr::group_by(rgn_id, rgn_name, year) %>%
  dplyr::summarize(value = sum(value)) %>%
  dplyr::ungroup()

```

# Scale the data and save files
Data is rescaled by dividing by the 95th quantile of values across all regions from 2011 to 2014.

```{r, eval=FALSE}

target_harvest <- m_sum_rgn %>%
  dplyr::mutate(quant_95 = quantile(value[year %in% 2011:2016], 0.95, na.rm = TRUE)) %>%
  dplyr::mutate(score = value / quant_95) %>% 
  dplyr::mutate(score = ifelse(score>1, 1, score)) %>%
  dplyr::select(rgn_id, year, pressure_score = score) %>%
  dplyr::arrange(rgn_id, year); head(target_harvest); summary(target_harvest)
  
#quant_95= 160761


  # any regions that did not have a catch should have score = 0 
rgns <-  rgn_master %>%
  dplyr::filter(rgn_typ == "eez") %>%
  dplyr::select(rgn_id = rgn_id_2013) %>%
  dplyr::filter(rgn_id < 255) %>%
  base::unique() %>%
  dplyr::arrange(rgn_id)


  
rgns <- expand.grid(rgn_id = rgns$rgn_id, year = min(target_harvest$year):max(target_harvest$year))
  
target_harvest <-  rgns %>%
  dplyr::left_join(target_harvest) %>%
  dplyr::mutate(pressure_score = ifelse(is.na(pressure_score), 0, pressure_score)) %>%
  dplyr::arrange(rgn_id); head(target_harvest); summary(target_harvest)

  write.csv(target_harvest, 'output/fao_targeted.csv', row.names = FALSE)
  
target_harvest_gf <- target_harvest %>%
  dplyr::mutate(gapfill = 0) %>%
  dplyr::select(rgn_id, year, gapfill)
  
  write.csv(target_harvest_gf, 'output/fao_targeted_gf.csv', row.names = FALSE)

```

# Data check
The data from last year and this year should be the same unless there were changes to underlying FAO data or the master species list.

In this case, all of the regions looked very similar.

```{r, eval=FALSE}
new <- read.csv("output/fao_targeted.csv") %>% 
  filter(year==2015)

old <- read.csv("../v2017/output/fao_targeted.csv") %>%
  #mutate(year== year-2) %>% 
  dplyr::filter(year == 2015) %>%
  dplyr::select(rgn_id, year, pressure_score_old=pressure_score) %>%
  dplyr::left_join(new, by=c("rgn_id", "year"))
old

plot(pressure_score ~ pressure_score_old, data=old)
abline(0, 1, col="red")


compare_plot <- ggplot(data = old, aes(x=pressure_score_old, y= pressure_score, label=rgn_id))+
  geom_point()+
  geom_abline(color="red")

plot(compare_plot)
ggplotly(compare_plot)

```

