---
title: "OHI 2018 - Soft bottom pressure data prep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../src/templates/ohi_hdr.html'
pdf_document:
  toc: true
---

[REFERENCE RMD FILE: ADD!!!! after pushing the changes in this file]

# Summary
The habitat pressure data layer is created from catch by gear data of industrial and non industrial fishing profided by Watson (2018). Catches are calculated for each OHI region and then standarized by soft-bottom habitat in each region. Then rescale!!!!!!!!??????

# Updates from previous assessment
This year we are using the Watson (2018) data to calculate the soft bottom pressure. Data reports catch by gear. 
To each gear we assing a multiplier acording to how much the affect the sea bottom. Here the table with the multipliers (1 is extramly harmfull and 0 is not harmful)

Gear  | Multiplier
------|-----------
Trawl | 1
Dredge | 1
Gillnet | 0.5
Trap  | 0.5
Other | 0.25
Trawl midwater | 0
Pole and Line | 0
Longline | 0
Purse seine | 0
Seine | 0

Also this year artisanal catches were included in our analysis becasue source provided the necesary data. Finally, there where changes in the metos of the source data. 


***

# Data Source 
**Reference**: Watson, R. A. and Tidd, A. 2018. Mapping nearly a century and a half of global marine fishing: 1869–2015. Marine Policy, 93, pp. 171-177. [(Paper URL)](https://www.sciencedirect.com/science/article/pii/S0308597X18300605?via%3Dihub)

**Downloaded**: July 17, 2018 from IMAS portal (see prs_fish layer for more details)

**Description**:  Global fisheries landings data per cell separated by Industrial versus Non-Industrial catch, IUU, and discards.

**Native data resolution**: 

**Time range**: 1950 - 2015

**Format**:  csv format

***
  
# Methods 
First create one raster a year with all catches (industrial and non industrial) using gear that harm soft-bottom. 
Then esxtract catches per OHI regions and standarize by soft-bottom habitat of each region.
Finally, rescale!!!!!!!!!!!!!!!!!

***

## Setup

```{r}
library(plyr)
library(tidyverse)
library(parallel)
library(foreach)
library(doParallel)
library(raster)
library(rasterVis)
library(seaaroundus)
library(RColorBrewer)
library(cowplot)
library(stringr)
library(colorspace)
library(sp)
library(rgdal)
library(sf)

#setwd("~/github/ohiprep_v2018/globalprep/hab_prs_hd_subtidal_soft_bottom/v2018")

registerDoParallel(4) # registering cores for parallel processing


source('https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/master/src/R/spatial_common.R')


## color palette
cols = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255)) # rainbow color scheme
mytheme=rasterTheme(region=cols)

## Set template ocean raster and mollweide projection CRS
ocean <- raster::raster(file.path(dir_M, 'model/GL-NCEAS-Halpern2008/tmp/ocean.tif'))
mollCRS=crs('+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs')

options(scipen=999)


```

## Aggregate annual industrial catch by type of gear

First get the template raster with a resolution of 0.5 degree cells. The `getcells()` function comes from the [seaaroundus R package](https://github.com/ropensci/seaaroundus).

The values of these cells are the Cell ID numbers. In the fisheries catch dataset, these Cell ID numbers match up with the "Seq" numbers.

```{r}

saup_cells <- getcells("POLYGON ((-180 90, 180 90, 180 -90, -180 -90, -180 90))")

saup_rast <- raster(ncol=720, nrow=360)
saup_rast[] <- saup_cells

plot(saup_rast,col=cols,main = "SAUP Cell IDs")

```


Then create one raster per year for industrial and non industrial catches using gear that harm soft bottom. 


```{r}
years <-  c(2003:2015)

multipliers <- data.frame(FleetGearName = c("Dredge", "Gillnet", "Lines Non-tuna", "Longline Non-tuna", "LonglineTuna", "Other", "Pole and Line Tuna", "Purse seine Non-tuna", "Purse seine Tuna", "Seine", "Trap", "Trawl", "Trawl midwater"), multiplier = c(1, 0.5, 0, 0, 0, 0.25, 0, 0, 0, 0, 0.5, 1, 0))


foreach(yr = years) %dopar%{ #yr = 2015
  
  #read in raw data for the year
  raw_ind <- readRDS(paste0(file.path(dir_M,'git-annex/globalprep/prs_fish/v2018/int/annual_catch/CatchInd_'),yr,'.rds'))
  
  catch_ind <- raw_ind %>%
    dplyr::rowwise() %>%
    dplyr::mutate(catch = sum(Reported, IUU, Discards, na.rm=TRUE)) %>%
    dplyr::left_join(multipliers, by = "FleetGearName") %>% 
    dplyr::mutate(catch_discount = catch*multiplier) %>% 
    dplyr::group_by(Cell) %>%
    dplyr::summarise(cell_catch_ind = sum(catch_discount)) 
  
  raw_Nind <- readRDS(paste0(file.path(dir_M,'git-annex/globalprep/prs_fish/v2018/int/annual_catch/CatchNInd_'),yr,'.rds'))
  
  catch_Nind <- raw_Nind %>%
    rowwise() %>%
    dplyr::mutate(catch = sum(Reported, IUU, Discards, na.rm=TRUE)) %>%
    left_join(multipliers, by = "FleetGearName") %>% 
    dplyr::mutate(catch_discount = catch*multiplier) %>% 
    dplyr::group_by(Cell) %>%
    dplyr::summarise(cell_catch_Nind = sum(catch_discount))
  
  catch_total <- catch_ind %>% 
    dplyr::full_join(catch_Nind, by= "Cell") %>% 
    dplyr::mutate(total = cell_catch_ind + cell_catch_Nind) %>% 
    dplyr::mutate(total = ifelse(is.na(total), cell_catch_ind, total)) %>% 
    dplyr::select(Cell, total)
  
  
  catch_total <- catch_ind %>% 
    dplyr::full_join(catch_Nind, by= "Cell") %>%
    dplyr::rowwise() %>% 
    dplyr::mutate(total = sum(cell_catch_ind, cell_catch_Nind, na.rm = TRUE)) %>%
    dplyr::select(Cell, total)
     
  
  #rasterize catch by swapping cell ids with (calls the suap raster we made above)
  raster::subs(saup_rast, catch_total, by = 'Cell', which = 'total', subsWithNA=TRUE, filename = file.path(dir_M, paste0('git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2018/int/catch_count_',yr,'.tif')), overwrite=TRUE) #saves raster directly to Mazu
  
}

##Reading in the raster to make sure it looks as expected
test<- raster(file.path(dir_M, ('git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2018/int/catch_count_2015.tif')))

plot(test) ##Looks good!


```


## Extract values for each OHI region

```{r}

## Read tif as rasters for all catches using gear that harm soft bottom

## Reference raster
catch_crs <- raster(file.path (dir_M,'git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2018/int/catch_count_2003.tif'))

## OHI region polygons
ohi_regions <- st_read(dsn = file.path(dir_M, "git-annex/globalprep/spatial/v2017"), layer="regions_2017_update")

## ASK MEL ABOUT THIS LINE
ohi_regions_wgs <-  st_transform(ohi_regions, proj4string(catch_crs))

## List of all tifs with catches that need to be read in the loop
catch_rasts <- list.files(file.path(dir_M, "git-annex/globalprep/hab_prs_hd_subtidal_soft_bottom/v2018/int"), pattern = 'catch_count_.*.tif', full.names = TRUE)

##catch_rasts <- catch_rasts[c(1,2,3,4,5,6,7,8,9,10,11,13)] Filter for the 12 missing rasters to run in the loop

#registering cores for parallel processing
registerDoParallel(4)

foreach(sb_catch = catch_rasts) %dopar% { # sb_catch = catch_rasts[12]

cat(sb_catch)

catch_year <- str_sub(sb_catch, -8, -5) ##extract the correspending year for each file

catch <- raster(sb_catch)

data <- raster::extract(catch, ohi_regions, weights = TRUE, normalizeWeights = FALSE, progress = 'text') 

names(data) <- paste(ohi_regions$type_w_ant, ohi_regions$rgn_ant_id, sep="_") 
sb_catch_rgn   <- plyr::ldply(data, rbind)

# the following code keeps the raster information when the raster lands partially on the land polygons
sb_catch_rgn <- sb_catch_rgn %>%
  tidyr::separate(.id, c("rgn_type", "rgn_id"), sep="_") %>%
  dplyr::mutate(tonnes = value*weight) %>%
  dplyr::group_by(rgn_type, rgn_id) %>%
  dplyr::summarize(tonnes = sum(tonnes, na.rm=TRUE)) %>%
  dplyr::group_by(rgn_id) %>%
  dplyr::summarize(tonnes = sum(tonnes, na.rm=TRUE)) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(as.numeric(as.character(rgn_id))) %>%
  dplyr::mutate(year = catch_year)

write.csv(sb_catch_rgn, sprintf("int/sb_catch_rgn_%s.csv", catch_year), row.names=FALSE)

}


```

## Standarized catches per soft-bottom area

```{r}
area <-  read.csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/master/globalprep/hab_prs_hd_subtidal_soft_bottom/v2016/output/habitat_extent_softbottom.csv")


## Read in csv for each year and combine them in one data frame

files <- list.files("int", pattern = "sb_catch_rgn_.*.csv", full.names = TRUE) 


sb_catch_all <- data.frame()

for(file in files) { # file = files[1]
  tmp <- read.csv(file)
  sb_catch_all <- rbind(sb_catch_all, tmp)
}


data_density <- sb_catch_all %>%
  left_join(area, by="rgn_id") %>%
  filter(rgn_id <= 250) %>%
  mutate(density = tonnes/km2)


## find max density across all years
max <- data_density %>%
  group_by(year) %>%
  summarize(maxDensity = max(density, na.rm=TRUE)) %>%
  data.frame()

##max density for 2014: 6792.767

write.csv(max, "int/reference_point_max.csv", row.names=FALSE) ##Make sure you are working in the correct working directory


## Referenc point = max values accroass all years
ref_point <- read.csv("int/reference_point_max.csv")
ref_point_max <- max(ref_point$maxDensity)


## cheking for distribution of data. In the past density data happens to be very skewed
hist(data_density$density) ## Yes, very skewed data

## rescale the density: 
## density is rescaled twice to reduce skew

data_rescaled <- data_density %>%
  mutate(density_rescaled_max = log(density + 1)/log(ref_point_max + 1)) %>% #pressure-type measure
  mutate(inv_dens_rescaled_max = 1 - density_rescaled_max)  # health-type measure

hist(data_rescaled$density_rescaled_max)

## Find the second rescaling point
ref_median <- data_rescaled %>%
  group_by(year) %>%
  summarize(ref_median = median(inv_dens_rescaled_max, na.rm=TRUE)) %>%
  data.frame()

write.csv(ref_median, "int/reference_point_median.csv", row.names=FALSE)

ref_point_median <- min(ref_median$ref_median) # 0.8999136

## Rescale for second time:  
  data_rescaled_2 <- data_rescaled %>%
    mutate(density_rescaled_median = inv_dens_rescaled_max/ref_point_median) %>%
    mutate(density_rescaled_median_capped = ifelse(density_rescaled_median > 1, 
                                                   1, 
                                                   density_rescaled_median))
hist(data_rescaled_2$density_rescaled_median_capped)
  
 
  hist(data_rescaled_2$density_rescaled_median_capped[data_rescaled_2$year==2010])
  filter(data_rescaled_2, rgn_id==163)
  filter(data_rescaled_2, year==2010)



```

## Comparing to old data

```{r}

  ## check against old data
old <- read.csv("../v2017/output/habitat_health_softbottom.csv") %>%
  dplyr::filter(year == 2014) %>% 
  dplyr::select(rgn_id, old_health=health)
  

test <- data_rescaled_2 %>%
  ##filter(year==2010) %>%
  dplyr::select(rgn_id, health = density_rescaled_median_capped) %>%
  left_join(old)

plot(health ~old_health, data=test)
abline(0,1, col = 'red')
## Not a great correlation, but different years data and fisheries data is very different
## some new NA values, but it doesn't seem like these have soft-bottom habitat
## (at least according to our raster data)




## LAST YEAR COMPARISON
V2016 <- read.csv("../v2016/output/habitat_health_softbottom_v2010.csv") %>%
  dplyr::select(rgn_id, v16_health=health)

V2017 <- read.csv("../v2017/output/habitat_health_softbottom.csv") %>%
  filter(year==2010) %>%
  dplyr::select(rgn_id, v17_health=health)

test <- V2016 %>%
  left_join(V2017)

plot(v17_health ~v16_health, data=test)
abline(0,1, col= 'red')


```







