---
title: 'OHI 2019 - World Governance Indicator'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
  pdf_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

[REFERENCE RMD FILE](https://cdn.rawgit.com/OHI-Science/ohiprep/master/globalprep/np/v2019/WGI_dataprep.html)

# Summary
This script downloads WGI data and prepares it for a pressures (1 - WGI) and resilience data layer.  

# Updates from previous assessment

No methods updates; additional year of data added.  

Consider this improvement for future assessments: create a linear model to estimate missing data rather than just taking averages of years with data (~ line 230).

***

# Data Source 
**Reference**:  http://info.worldbank.org/governance/wgi/index.aspx#home    

**Downloaded**: March 11 2019 (data updated Sep 21 2018)

**Description**:  
The Worldwide Governance Indicators (WGI) project reports aggregate and individual governance indicators for 215 economies over the period 1996–2017, for six dimensions of governance:

* Voice and Accountability
* Political Stability and Absence of Violence
* Government Effectiveness
* Regulatory Quality
* Rule of Law
* Control of Corruption

**Time range**: 1996-2017

***

```{r, eval=FALSE}

library(ohicore) # devtools::install_github('ohi-science/ohicore@dev')
library(tools)
library(dplyr)
library(tidyr)
library(WDI) # install.packages('WDI')  # used to extract World Development Indicator (World Bank) data 
library(stringr)
library(here)
library(ggplot2)
library(plotly)


# check website to see what years are available: http://info.worldbank.org/governance/wgi/index.aspx#home
yr_start = 1996
yr_end   = 2017


```

# Obtain the WGI data
Download each of the 6 WGI indicators:
```{r, eval=FALSE}

## access data through the WDI package

## get description of variables:
indicators <-  data.frame(WDI_data[[1]])
indicators[grep("VA.EST", indicators$indicator), ]
indicators[grep("PV.EST", indicators$indicator), ]
indicators[grep("GE.EST", indicators$indicator), ]
indicators[grep("RQ.EST", indicators$indicator), ]
indicators[grep("RL.EST", indicators$indicator), ]
indicators[grep("CC.EST", indicators$indicator), ]

# identify the six indicators
# WDIsearch('violence')# general search
key_voice = WDI(
  indicator = WDIsearch('Voice and Accountability: Estimate', field='name')['indicator'],
  country = 'all', start = yr_start, end=yr_end)

key_polst = WDI(
  WDIsearch('Political Stability and Absence of Violence/Terrorism: Estimate', field='name')['indicator'],
  country='all',start = yr_start, end=yr_end)

key_gvtef = WDI(
  WDIsearch('Government Effectiveness: Estimate', field='name')['indicator'],
  country='all',start = yr_start, end=yr_end)

key_regqt = WDI(
  WDIsearch('Regulatory Quality: Estimate', field='name')['indicator'],
  country='all',start = yr_start, end=yr_end)

key_rolaw = WDI(
  WDIsearch('Rule of Law: Estimate', field='name')['indicator'],
  country='all',start = yr_start, end=yr_end)

key_corrp = WDI(
  WDIsearch('Control of Corruption: Estimate', field='name')['indicator'],
  country='all',start = yr_start, end=yr_end)

```

Combine the indicators into a single table, with a column for each indicator, and rows for each country-year pair.

```{r, eval=FALSE}
d = key_voice %>% 
  select(country, year, indicator) %>%
  left_join(key_polst %>% select(-iso2c), by=(c('country', 'year'))) %>%
  left_join(key_gvtef %>% select(-iso2c), by=(c('country', 'year'))) %>%
  left_join(key_regqt %>% select(-iso2c), by=(c('country', 'year'))) %>%
  left_join(key_rolaw %>% select(-iso2c), by=(c('country', 'year'))) %>%
  left_join(key_corrp %>% select(-iso2c), by=(c('country', 'year'))); head(d); summary(d); sapply(d, class)

colnames(d) <- c("country", "year", "VA.EST", "PV.EST", "GE.EST",	"RQ.EST",	"RL.EST", "CC.EST")
```

## Save a record of any new raw data for archival purposes

Uncomment the code chunk lines when updating WGI data, this will most likely occur when calculating for new assessment year:

```{r, eval=FALSE}

date <- Sys.Date()
write.csv(d, here(sprintf('globalprep/prs_res_wgi/v2019/raw/worldbank_wgi_from_wdi_api_%s.csv', date)), row.names=FALSE)


# This dataset currently has non-OHI regions included
```


# Gapfill, part 1: filling missing years of data for indicators, within countries
The first gapfilling occurs when we use the average of previous years data within each region/indicator.  This occurs when a region has data for an indicator, but not for all years.


**Read in WGI data - change appended date in file name to reflect the most recent version of the saved WGI data:**
```{r, eval=FALSE}
d <- read.csv(here('globalprep/prs_res_wgi/v2019/raw/worldbank_wgi_from_wdi_api_2019-04-15.csv')) # change appended date so that it only changes if you do so actively

```

```{r, eval=FALSE}

d <- gather(d, "indicator", "value", VA.EST:CC.EST)


## each country has 18 years of data
d_gap_fill  <- d %>%
  group_by(country, year) %>%
  mutate(NA_count_c_y = sum(is.na(value))) %>% # gf record: NA values within a region/year prior to gapfilling, max value is 6 (meaning that a country has no data)
  ungroup() %>%
  group_by(country, indicator) %>% # gapfill missing data with mean of values across years within the same region/indicator
  mutate(ind_mean_c_i = mean(value, na.rm=TRUE)) %>%
  ungroup() %>%
  mutate(value = ifelse(is.na(value), ind_mean_c_i, value)) %>%
  group_by(country, year) %>% 
  mutate(NA_count_post_gf1 = sum(is.na(value))) # gf record: NA values within a region/year after within region/indicator gapfilling (i.e. indicator is gapfilled by other years of data), used to cut regions <4 indicators (below)    

```

## Safeguard: cut regions with < 4 indicators (if any) to calculate score.
Once gapfilling is complete, the WGI scores are calculated as an average of the 6 indicators.  However, if a country is missing 4 or more of the indicators within a year the average would be very biased.  In these cases, a different method should be used to gapfill these data 

(NOTE: for the 2019 assessment all regions had at least 3 of the 6 indicators).

```{r purge datapoor region, eval=FALSE}

countries_no_data <- d_gap_fill %>%
  filter(NA_count_post_gf1 > 3)

countries_no_data <- unique(countries_no_data$country)
countries_no_data

## In this case, the countries with minimal data (< 4 indicators ever calculated) are deleted.  
## These will be gap-filled later on if they are deleted now.
d_gap_fill <- d_gap_fill %>%
  filter(!(country %in% countries_no_data))

```

# Calculate overall WGI score for each country

This involves:

* taking the average of the 6 indicators (assuming there are at least 4 of the 6 indicators)
* rescaling the data from 0 to 1 

```{r, eval=FALSE}

d_calcs  <-  d_gap_fill %>%
  group_by(country, year) %>%
  summarize(score_wgi_scale = mean(value, na.rm=T),
            NA_start = mean(NA_count_c_y), # initial mean number of NA across indicators, pre-gapfill 
            NA_post_gf_1 = mean(NA_count_post_gf1)) %>% # number of NA across indicators, post-gapfill across year gapfill within region/indicator
  ungroup() 

```

## Check that the values in `scores_wgi_scale` fall within the wgi range specified below:
```{r, eval=FALSE}

# summary(d_calcs) - checking to make sure NA values make sense
wgi_range = c(-2.5, 2.5) # historic values have been between -2.5 and 2.5

```

```{r, eval=FALSE}

d_calcs <- d_calcs %>%
  mutate(score =  (score_wgi_scale - wgi_range[1]) / (wgi_range[2] - wgi_range[1])) %>%
  ungroup(); head(d_calcs); summary(d_calcs)
# converts the scores between 0 and 1, using min/max to scale the data

# document gapfilling
d_calcs <- d_calcs %>%
  mutate(gap_fill = NA_start - NA_post_gf_1,   # if there are values in NA_post_gf_1, it means these weren't gapfilled
         gap_fill = ifelse(is.na(score), 0, gap_fill)) %>% # number of values that were gapfilled
  select(-NA_start, -NA_post_gf_1)

```

## Explore & check intermediate d_calcs data table

```{r, eval=FALSE}

d_calcs[d_calcs$gap_fill>0, ]     
d_calcs[d_calcs$country == "New Caledonia", ]  # no data, was deleted earlier
d_calcs[d_calcs$country == "Niue", ] # should have gap-fill values between 0-6
d_calcs[d_calcs$country == "American Samoa",]


```

## Save intermediate file 

```{r, eval=FALSE}

## save intermediate file of wgi scores pre-gapfilling (for OHI+ use)
write.csv(d_calcs %>%
            select(country, year, score_wgi_scale, score_ohi_scale = score), 
          here('globalprep/prs_res_wgi/v2019/intermediate/wgi_combined_scores_by_country.csv'),
          row.names = FALSE)

```


# Convert country names to ohi regions

```{r, eval=FALSE}

## We report these regions at a greater spatial resolution:

## Aruba is part of the Netherlands Antilles, but it is reported separately
country_split_1 <- data.frame(country = "Netherlands Antilles", region = c('Bonaire', 'Curacao', 'Saba', 'Sint Maarten', 'Sint Eustatius'))
country_split_2 <- data.frame(country = "Jersey, Channel Islands", region = c('Jersey', 'Guernsey'))
country_split <- rbind(country_split_1, country_split_2)

country_split_data <- country_split %>%
  left_join(d_calcs) %>%
  select(-country) %>%
  rename(country = region)

d_calcs <- d_calcs %>%
  filter(!(country %in% c("Netherlands Antilles", "Jersey, Channel Islands"))) %>%
  rbind(country_split_data)  %>%
  mutate(country = as.character(country))

d_calcs$country[grep("Korea, Dem.", d_calcs$country)] <- "North Korea"
# Maybe in future update package with country synonym list


## Function to convert to OHI region ID
d_calcs_rgn <- name_2_rgn(df_in = d_calcs, 
                       fld_name='country', 
                       flds_unique=c('year'))
# Eswatini is a landlocked country (aka Swaziland) in Southern Africa

## Combine the duplicate regions (we report these at lower resolution)
## In this case, we take the weighted average
population_weights <- data.frame(country = c("Virgin Islands (U.S.)", "Puerto Rico",
                                             "China", "Hong Kong SAR, China", "Macao SAR, China"),
                                 population = c(107270, 3337180,
                                         1386395000, 7391700, 622570))
# updated population values on 1 Apr 2019 (source: World Bank website, 2017 values)


d_calcs_rgn <- d_calcs_rgn %>%
  left_join(population_weights, by="country") %>%
  mutate(population = ifelse(is.na(population), 1, population)) %>% 
  group_by(rgn_id, year) %>%
  summarize(score = weighted.mean(score, population),
            gapfill_within_rgn = weighted.mean(gap_fill, population)) %>%
  ungroup() %>%
  filter(rgn_id <= 250)

summary(d_calcs_rgn)

```


# Gapfill, part 2: Filling in missing territorial region value

Assigning territorial region value to be the mean of parent country value and territorial regions with data (using same `sov_id`).

```{r, eval=FALSE}

## data that describes territories of countries
territory = rgn_master %>% 
  select(rgn_id = rgn_id_2013,
         sov_id) %>%               
  group_by(rgn_id) %>% # remove duplicated countries from this rgn_id list
  summarize(sov_id = mean(sov_id, na.rm=T)) %>% # duplicates always have the same sov_id (r2 value)
  filter(rgn_id <= 250, rgn_id != 213)

	
## expand to include all years of data
territory <- data.frame(year=yr_start:yr_end) %>% 
  merge(territory, by=NULL) 


## assign territories the values of their sovereign country
d_sovs = d_calcs_rgn %>% 
  full_join(territory, by = c('rgn_id', 'year')) %>%
  group_by(sov_id, year) %>%
  mutate(score_gf_territory = mean(score, na.rm=TRUE),
         gapfill_within_rgn = mean(gapfill_within_rgn, na.rm=TRUE))%>%
   filter(!is.na(gapfill_within_rgn)) %>%
  ungroup()

# filter(d_sovs, rgn_id %in% c(1,2,3,16) & year == 2017) check in console to make sure Australia and territories have the same score and sov_id
```


Define new data object from d_sovs which includes gapfill method and gapfilled scores:

```{r, eval=FALSE}

d_gf2 <- d_sovs %>%
  mutate(gapfill_territory = ifelse(is.na(score) & !is.na(score_gf_territory), "territory", "NA")) %>%
  mutate(score = ifelse(is.na(score), score_gf_territory, score)) %>%
  select(rgn_id, year, score, gapfill_within_rgn, gapfill_territory)

```

Add region names and clean the region data, and make sure we have all the regions:

```{r, eval=FALSE}
# get region names
regions <- rgn_master %>%
  filter(rgn_typ == "eez") %>%
  filter(rgn_id_2013 <= 250) %>% # > 250 are either FAO or a disputed region
  filter(rgn_id_2013 != 213) %>% # 213 is antarctica
  select(rgn_id = rgn_id_2013, rgn_name=rgn_nam_2013) %>%
  unique() %>%
  arrange(rgn_id) 

d_gf2 <- regions %>%
  left_join(d_gf2)
  
```


## Look at data table for the territories (gapfilled)

```{r, eval=FALSE}
head(d_sovs)
summary(d_sovs)
```

## Look at table with scores and gapfilling methods

```{r, eval=FALSE}

## check for NA values within "score" variable
## if so, need to gapfill using UN geopolitical regions
summary(d_gf2)

```

# Check data 

Comparing this year's values against last year's.  These should be the same unless there have been updates to WGI source data or a change to methods.  For this year, there was a small change that effected a few territorial regions.  In the past, we used the sovereign country value, but in the case, we averaged the sovereign country and the available territorial values.

Plot most recent shared year between last year and this years data, and look for a relationship close to a 1:1 relationship. If data are significantly off the line, look at the original (raw) data to investigate.

```{r, eval=FALSE}

new2016 <- d_gf2 %>%
  filter(year==2016) %>% # update to most recent shared year for comparison
  select(rgn_id, score)

old2016 <- read.csv(here('globalprep/prs_res_wgi/v2018/output/wgi_res.csv')) %>% # change filepath to the previous assessment year's, rename variable for clarity
  filter(year == 2016) %>%
  select(rgn_id, old_score=resilience_score) %>%
  full_join(new2016)


score_compare_plot <- ggplot(old2016, aes(x = old_score, y = score, text = rgn_id)) +
  geom_point() +
  geom_abline(slope=1, intercept = 0, col = "red")
   
ggplotly(score_compare_plot) # looks good (no outliers), no real need for comparison


```


Look at top/bottom 10 regions to make sure these seem reasonable:

```{r, eval=FALSE}

## Top/Bottom 10 scorers:

tmp <- d_gf2 %>%
  filter(year==2017) %>%
  arrange(score) %>%
  select(rgn_id, score) %>%
  left_join(regions)

tmp[1:10, ]
tmp[211:220, ]

hist(tmp$score)

```

Look at a summary to confirm scores are between 0 and 1, there are 220 regions, and there are no NAs (for this particular dataset):

```{r, eval=FALSE}
summary(d_gf2) # scores are between 0-1, gapfilled info present, no NAs
length(unique(d_gf2$rgn_id)) # 220 regions
c(min(d_gf2$score), max(d_gf2$score)) # checking for score min and max; making sure they're between 0-1

```

# Save the data
Save gapfilling and data for this assessment year.


```{r, eval=FALSE}

tmp_data_res <- d_gf2 %>%
  select(rgn_id, year, resilience_score = score)
write.csv(tmp_data_res, here("globalprep/prs_res_wgi/v2019/output/wgi_res.csv"), row.names=FALSE)

tmp_data_prs <- d_gf2 %>%
  mutate(score = 1 - score) %>%
  select(rgn_id, year, pressure_score = score)
write.csv(tmp_data_prs, here("globalprep/prs_res_wgi/v2019/output/wgi_prs.csv"), row.names=FALSE)

# gapfilled data
tmp_gf <- d_gf2 %>%
  select(rgn_id, year, gapfill_within_rgn, gapfill_territory) %>%
  mutate(gapfill_within_rgn = ifelse(gapfill_within_rgn == 0, NA, gapfill_within_rgn)) %>%
  mutate(gapfill_within_rgn = ifelse(!is.na(gapfill_within_rgn), 
                                            paste0("gapfill_within_rgn: ", gapfill_within_rgn), NA)) %>%
  mutate(gapfill_territory = ifelse(gapfill_territory == "territory", "territory, mean of admin countries", NA)) %>%
  mutate(method = paste(gapfill_within_rgn, gapfill_territory, sep="; ")) %>%
  mutate(method = gsub("NA; ", "", method)) %>%
  mutate(method = gsub("; NA", "", method)) %>%
  mutate(method = gsub("NA", NA, method)) %>%
  mutate(gapfilled = ifelse(is.na(method), 0, 1)) %>%
  select(rgn_id, year, gapfilled, method)


write.csv(tmp_gf, here("globalprep/prs_res_wgi/v2019/output/wgi_gf.csv"), row.names=FALSE)
  # change format to match data SOP, add method column


```

Checking on outliers in 2019 assessment:
```{r outliers, eval=FALSE}

worldbank_wgi_from_wdi_api <- read_csv("globalprep/prs_res_wgi/v2019/raw/worldbank_wgi_from_wdi_api_2019-04-15.csv")

worldbank_2016 <- worldbank_wgi_from_wdi_api %>% 
  filter(year == 2016) %>% 
  filter(country == "Australia" | country == "Marshall Islands")

# Compare to data from 2018:
worldbank_wgi_from_wdi_api_2018 <- read_csv("globalprep/prs_res_wgi/v2018/raw/worldbank_wgi_from_wdi_api_2018-04-13.csv")

worldbank_old <- worldbank_wgi_from_wdi_api_2018 %>% 
  filter(year == 2016) %>% 
  filter(country == "Australia" | country == "Marshall Islands")

```



