---
title: 'OHI 2018: Mariculture Population Pressure Layers'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../src/templates/ohi_hdr.html'
pdf_document:
  toc: true
---

[REFERENCE RMD FILE](https://cdn.rawgit.com/OHI-Science/ohiprep/master/globalprep/mar/v2018/mar_prs_pop_dataprep.html)


#Summary

This document describes the steps for obtaining and wrangling the data used to calculate the mariculture population pressure subdimension for the 2018 global assessment.

The general calculation is:

The following data are used:

#Updates from previous assessment
No updates were made to this methodology, from the previous assessment.

***

#Data Sources

##CIESE and CIAT Gridded World Population

**Reference**:

**Downloaded**:

**Description**:

**Native data resolution**:

**Time range**:

**Format**:

***

#Setup

Load all relevant libraries including parallel processing packages, 

```{r setup, message=FALSE, warning=FALSE, verbose=FALSE}

## set options for all chunks in code
knitr::opts_chunk$set(warning=FALSE, message=FALSE,fig.width = 6, fig.height = 4, fig.path = 'figs/')

## comment out when knitting
# setwd("globalprep/mar_prs_population/v2018")

source("../../../src/R/common.R")
goal <- "globalprep/mar_prs_population"
scenario <- "v2018" # change to reflect assessment year
data_yr <- "d2017" # change to reflect year of most recent data

if(!requireNamespace("raster", quietly = FALSE))
  install.packages("raster")
if(!requireNamespace("ohicore", quietly = FALSE))
  devtools::install_github("ohi-science/ohicore")
if(!requireNamespace("dplyr", quietly = FALSE))
  install.packages("tidyverse")
if(!requireNamespace("sf", quietly = FALSE))
  install.packages("sf")
if(!requireNamespace("fasterize", quietly = FALSE))
  install.packages("fasterize")
if(!requireNamespace("foreach", quietly = FALSE))
  install.packages("foreach")

library(ohicore)

library(raster)
library(rgdal)
library(sp)
library(sf)
library(fasterize)

library(dplyr)
library(stringr)

library(parallel)
library(foreach)
library(doParallel)

## define path names where data will be read from/written to; change to reflect most recent data (d) and current assessment year (v)
spdata_filepath <- "git-annex/globalprep/spatial/v2017/"
raw_gpw_popdata_filepath <- "git-annex/globalprep/_raw_data/CIESEandCIAT_population/d2017"
un_popdata_filepath <- "git-annex/globalprep/_raw_data/UnitedNations_population/v2017/"
marprspop_int_filepath <- "git-annex/globalprep/mar_prs_population/v2017/int/"
globalrgn_filepath <- "git-annex/Global/NCEAS-Regions_v2014/data"


```

#Methods and Calculations

##Import Raw Data

Raw datasets used in this data prep include: raw CIESE/CIAT population, ocean raster for (purpose, extracting coastal?), and UN population

```{r import raw data}

## will match projection and spatial resolution to this
ocean = raster(file.path(dir_M,'git-annex/globalprep/spatial/ocean.tif'))

## reproject and resample density data
raw <- list.files(file.path(dir_M, "git-annex/globalprep/_raw_data/CIESEandCIAT_population/d2017"), 
                  full.names = TRUE, pattern = "\\.tif$",
                  recursive = TRUE)
raw <- raw[grep("density", raw)] # all tiffs in CIESEandCIAT_population/d2017 folder, with 'density' in the name

## UN population data used to confirm derived spatial population
pop <- read.csv(file.path(dir_M,
        "git-annex/globalprep/_raw_data/UnitedNations_population/v2017/UN_pop_clean.csv"), 
        strip.white = TRUE, stringsAsFactors = FALSE)

## eez raster used in creating 25 mile inland raster
eez_raster <- raster(file.path(dir_M, "git-annex/globalprep/spatial/v2017/regions_eez_with_fao_ant.tif"))

```

##Project, Resample, Interpolate...

Gridded world population densities from CIESE and CIAT needs to be projected into the (?) coordinate reference system and resampled using nearest neighbor method, and missing years must be filled by interpolation

```{r manipulate spatial data}

## year to numeric value, project to correct crs, resample, and overwrite the data with these updates (?)
for(rast in raw){ # for each raster in raw
  yr <- as.numeric(as.character(str_sub(rast, -8, -5)))
  raster(rast) %>%
    projectRaster(crs = crs(ocean), over=TRUE) %>%
    resample(ocean, method = 'ngb',
    filename = file.path(dir_M, sprintf("git-annex/globalprep/mar_prs_population/v2017/int/human_density_%s_mol.tif", yr)),
    overwrite = TRUE) 
  }

## function to calculate yearly change in population density
files <- list.files(file.path(dir_M, 
                              "git-annex/globalprep/mar_prs_population/v2017/int/"), pattern = "_mol.tif", full = TRUE)

yearly_diff <- function(year_min, year_max, density_files = files){
  rast_diff <- stack(density_files[grep(year_max, density_files)], density_files[grep(year_min, density_files)]) %>%
  overlay(fun = function(x, y){(x - y)/5},
  filename = file.path(dir_M,
            sprintf('git-annex/globalprep/mar_prs_population/v2017/yearly_change_%s_to_%s.tif', year_min, year_max)), 
            overwrite = TRUE)
}

yearly_diff(year_min = 2000, year_max = 2005)
yearly_diff(year_min = 2005, year_max = 2010)
yearly_diff(year_min = 2010, year_max = 2015)
yearly_diff(year_min = 2015, year_max = 2020)

## function to interpolate population densities between years
yearly_interpolate <- function(year, start_raster, yearly_change){
  
stack(start_raster, yearly_change) %>%
  overlay(fun = function(x, y){(x + y)},
          filename = file.path(dir_M,
                               sprintf('git-annex/globalprep/mar_prs_population/v2017/int/human_density_%s_mol.tif', 
                                       (year + 1))), 
          overwrite = TRUE)

stack(start_raster, yearly_change) %>%
  overlay(fun = function(x, y){(x + 2*y)},
          filename = file.path(dir_M,
                               sprintf('git-annex/globalprep/mar_prs_population/v2017/int/human_density_%s_mol.tif', 
                                       (year + 2))), 
          overwrite = TRUE)

stack(start_raster, yearly_change) %>%
  overlay(fun = function(x, y){(x + 3*y)},
          filename = file.path(dir_M,
                               sprintf('git-annex/globalprep/mar_prs_population/v2017/int/human_density_%s_mol.tif', 
                                       (year + 3))), 
          overwrite = TRUE)

stack(start_raster, yearly_change) %>%
  overlay(fun = function(x, y){(x + 4*y)},
          filename = file.path(dir_M,
                               sprintf('git-annex/globalprep/mar_prs_population/v2017/int/human_density_%s_mol.tif', 
                                       (year + 4))), 
          overwrite = TRUE)
}

## interpolate missing years
files <- list.files(file.path(dir_M, 
                              "git-annex/globalprep/mar_prs_population/v2017/int/"), pattern = ".tif", full = TRUE)

yearly_interpolate(2000, start_raster = files[grep("2000_mol", files)], yearly_change = files[grep("2000_to_2005", files)])
yearly_interpolate(2005, start_raster = files[grep("2005_mol", files)], yearly_change = files[grep("2005_to_2010", files)])
yearly_interpolate(2010, start_raster = files[grep("2010_mol", files)], yearly_change = files[grep("2010_to_2015", files)])
yearly_interpolate(2015, start_raster = files[grep("2015_mol", files)], yearly_change = files[grep("2015_to_2020", files)])

```

##Convert Density to Population 

Density data are converted to population by (some function, meaning of the numeric values in for loop?). GWPv4 provides both densities and population counts. However, we derive population counts from density because (?) Density of the raw datasets is in units (people per km2?), and population will be in units of (?)

```{r convert density data to population count}

den_raster <- list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/direct_human/int'), 
                         pattern = "mol.tif", full = TRUE)

den_raster <- den_raster[!grepl("log", den_raster)]
den_raster <- den_raster[!grepl("tif.", den_raster)]
den_raster # check to make sure this looks correct

for(rast in den_raster){
  nm <- basename(rast)
  yr <- str_sub(nm,-12,-9)
  cat(nm)
  tmp <- raster(rast)
  calc(tmp, fun = function(x){x * (934.4789 * 934.4789 * 0.000001)}, # these values come from (?)
       filename = file.path(dir_M, 
                  sprintf("git-annex/globalprep/mar_prs_population/v2017/int/human_count_%s_mol.tif", yr)),
       overwrite=TRUE)
}
```

##Create EEZ plus 25-mile Inland Raster
```{r }
## need to add in Fiji (appears to get cut): include all of Fiji land falling within 50 mile boundary
fiji <- sf::read_sf(dsn = file.path(dir_M, "git-annex/globalprep/spatial/v2017"),
                              layer ="regions_2017_update") %>%
  filter(rgn_type == "land" & rgn_id==18) %>%
  select(rgn_id, geometry)

# create the 25 mile inland raster and add Fiji
inland <- sf::read_sf(dsn = file.path(dir_M, "git-annex/Global/NCEAS-Regions_v2014/data"),
                                 layer = "sp_inland25mi_gcs") %>%
  select(rgn_id, geometry)

inland <- st_transform(inland, st_crs(fiji))
inland <- rbind(inland, fiji)

## save shapefile for future reference
#if (!file.exists("raw")){ # dir_M, "git-annex/globalprep/spatial/v2017", layer = "EEZ_inland_50mi"...
#  dir.create("raw")
#} 
# st_write(inland, dsn = file.path(dir_M, "git-annex/globalprep/spatial/v2017",
#                                      layer = "EEZ_inland_50mi"), driver="ESRI Shapefile")

inland_raster <- fasterize::fasterize(inland, eez_raster, field = 'rgn_id')

tmp <- raster::freq(inland_raster) 

tmp2 <- data.frame(tmp) %>%
  dplyr::arrange(value)

raster::merge(eez_raster, inland_raster, filename=file.path(dir_M, 
                      "git-annex/globalprep/mar_prs_population/v2017/int/eez_25mi_inland.tif"))
```

#Tidy/Wrangle

##Extract by or Add OHI Regions

```{r extract population data from all population layers}

zones <- raster(file.path(dir_M, "git-annex/globalprep/mar_prs_population/v2017/int/eez_25mi_inland.tif"))
pop_rasts <- list.files(file.path(dir_M, "git-annex/globalprep/mar_prs_population/v2017/int"), 
           pattern = "mol.", full=TRUE)
pop_stack <- raster::stack(pop_rasts)
coastal_pop <- raster::zonal(pop_stack, zones, fun="sum", progress="text")

write.csv(coastal_pop, "globalprep/mar_prs_population/v2017/int", row.names=FALSE)

coastal_pop2 <- data.frame(coastal_pop) %>%
  tidyr::gather("year", "coastal_pop_25mi", -1) %>%
  dplyr::select(rgn_id=zone, year=year, popsum=coastal_pop_25mi) %>%
  dplyr::mutate(year = as.numeric(as.character(substring(year, 13, 16)))) %>%
  dplyr::filter(rgn_id <= 250)

write.csv(coastal_pop2, "globalprep/mar_prs_population/v2017/output/mar_pop_25mi.csv", row.names = FALSE)

```

#Gapfilling

No gapfilling was completed as part of the data prep methodology.

#Data Checks and/or Meta-analysis

##Check that Zonal Counts match UN Populations

```{r check full calculated population against latest UN population data}

pop_rast <- raster(file.path(dir_M, 
          "git-annex/globalprep/mar_prs_population/v2017/int/human_count_2015_mol.tif")) # update year in filename accordingly
zones <- raster(file.path(dir_M,"git-annex/globalprep/spatial/v2017/regions_land_ocean.tif"))
pop_counts <- zonal(pop_rast, zones, fun = "sum", na.rm=TRUE)

pop_UN <- read.csv("globalprep/mar_prs_population/v2017/output/UN_population.csv") %>%
  dplyr::filter(year==2015) %>%
  dplyr::select(rgn_id, pop_UN=population)

compare_pop <- data.frame(pop_counts) %>%
  dplyr::select(rgn_id = zone, pop_rast = sum) %>%
  dplyr::left_join(pop_UN, by="rgn_id")

plot(log(compare_pop$pop_rast), log(compare_pop$pop_UN))
abline(0,1, col="red")

```

```{r quick compare with previous data}

old <- read.csv("globalprep/mar_prs_population/v2016/output/mar_pop_25mi.csv") %>%
  dplyr::select(rgn_id, year, popsum_old=popsum)

tmp <- coastal_pop2 %>%
  dplyr::left_join(old, by=c("rgn_id", "year"))

plot(log(tmp$popsum), log(tmp$popsum_old))
abline(0,1, col="red")
```


***

