---
title: "OHI 2019: Old vs New Reference Point Methods"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
  pdf_document:
    toc: true
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(eval=FALSE)

### libraries useful for data wrangling
library(dplyr)
library(tidyr)
library(tidyverse)

## libraries useful for spatial data
library(raster)       
library(rgdal)        
library(sf)         
library(fasterize)

## data visualization
library(RColorBrewer)
library(ggplot2)
library(rasterVis)    
library(maps)

## path management
library(here)


## some OHI files
source('http://ohi-science.org/ohiprep_v2019/workflow/R/common.R')
dir_git_ref <- file.path(here(), "globalprep/mar/v2019/archive_reference_point")
```

# Reference Point/ Status Comparison

```{r}

# read in data
scores_2019 <- read_csv(file.path(dir_git_ref, "/check/scores.csv"))

mar_status_v2019 <- scores_2019 %>%
  filter(dimension == "status",
         goal == "MAR") 

# calculate some statistics to help compare 
sum(mar_status_v2019$score == 0, na.rm = TRUE) #732 out of 1776 are 0
sum(is.na(mar_status_v2019$score)) #168 out of 1776 are NA

  #harvest_tonnes <-
  #  AlignDataYears(layer_nm = "mar_harvest_tonnes", layers_obj = layers)
  
# read in v2019 data
harvest_tonnes_v2019 <- read_csv("/home/sgclawson/github/ohiprep_v2019/globalprep/mar/v2019/output/mar_harvest_tonnes.csv")  

  #sustainability_score <-
  #  AlignDataYears(layer_nm = "mar_sustainability_score", layers_obj = layers)

sustainability_score_v2019 <- read_csv("/home/sgclawson/github/ohiprep_v2019/globalprep/mar/v2019/output/mar_sustainability.csv") %>%
  dplyr::select(-year)
  
pot_aq_final <- read_csv(file.path(dir_git_ref, "/production_output/aq_potential_final.csv"))
  
#read in v2019 pop data
mar_pop_25mi_v2019 <- read_csv(file.path("/home/sgclawson/github/ohiprep_v2019/globalprep/mar_prs_population/v2019/output/mar_pop_25mi.csv"))
```


## Calculate scores the OLD WAY (taken (and modified) from functions.R file in OHI Global)
```{r}

  rky_old <-  harvest_tonnes_v2019 %>%
    dplyr::left_join(sustainability_score_v2019,
              by = c('rgn_id', 'taxa_code')) %>%
    dplyr::select(rgn_id, year, taxa_code, taxa_group, tonnes, sust_coeff)
  summary(rky_old)
  
  # fill in gaps with no data
  rky_old <- tidyr::spread(rky_old, year, tonnes)
  rky_old <- tidyr::gather(rky_old, "year", "tonnes",-(1:4)) %>%
    dplyr::mutate(year = as.numeric(year))
  summary(rky_old)
  
  # adjustment for seaweeds based on protein content
  rky_old <- rky_old %>%
    dplyr::mutate(tonnes = ifelse(taxa_group == "AL", tonnes*0.2, tonnes)) %>%
    dplyr::select(-taxa_group)
  
  # 4-year rolling mean of data
  m_old <- rky_old %>%
    dplyr::group_by(rgn_id, taxa_code, sust_coeff) %>%
    dplyr::arrange(rgn_id, taxa_code, year) %>%
    dplyr::mutate(sm_tonnes = zoo::rollapply(tonnes, 4, mean, na.rm = TRUE, partial =
                                        TRUE, align = "right")) %>%
    dplyr::ungroup()
  summary(m_old)
  
  # smoothed mariculture harvest * sustainability coefficient
  m_old <- m_old %>%
    dplyr::mutate(sust_tonnes = sust_coeff * sm_tonnes)
  summary(m_old)
  # aggregate all weighted timeseries per region, and divide by coastal human population
  ry_old = m_old %>%
    dplyr::group_by(rgn_id, year) %>%
    dplyr::summarize(sust_tonnes_sum = sum(sust_tonnes, na.rm = TRUE)) %>%  #na.rm = TRUE assumes that NA values are 0
    dplyr::left_join(mar_pop_25mi_v2019, by = c('rgn_id', 'year')) %>%
    dplyr::mutate(mar_pop = sust_tonnes_sum / popsum) %>%
    dplyr::ungroup()
  summary(ry_old)
  tail(ry_old)
  # get reference quantile based on argument years
  
  ref_95pct_old <- quantile(ry_old$mar_pop, 0.95, na.rm = TRUE)
 
    
ry_old = ry_old %>%
    dplyr::mutate(status = ifelse(mar_pop / ref_95pct_old > 1,
                           1,
                           mar_pop / ref_95pct_old)) 
summary(ry_old)

  ## Add all other regions/countries with no mariculture production to the data table
  ## Uninhabited or low population countries that don't have mariculture, should be given a NA since they are too small to ever be able to produce and sustain a mariculture industry.
  ## Countries that have significant population size and fishing activity (these two are proxies for having the infrastructure capacity to develop mariculture), but don't produce any mariculture, are given a '0'.

georegions <- ohicore::georegions
  all_rgns_old <- expand.grid(rgn_id = georegions$rgn_id, year = min(ry_old$year):max(ry_old$year))
  
  all_rgns_old <- all_rgns_old[!(all_rgns_old$rgn_id %in% ry_old$rgn_id),]
  table(all_rgns_old$rgn_id)
  
  uninhabited_old <- read.csv("https://raw.githubusercontent.com/OHI-Science/ohiprep/master/globalprep/spatial/v2017/output/rgn_uninhabited_islands.csv")
  
  uninhabited_old <- uninhabited_old %>% 
    dplyr::filter(rgn_nam != "British Indian Ocean Territory") # remove British Indian Ocean Territory which has fishing activity and a population size of 3000 inhabitants
  
  ## Combine all regions with mariculture data table
  ry_all_rgns_old <- all_rgns_old %>% 
    dplyr::mutate(status = 0) %>% 
    dplyr::mutate(status = ifelse(rgn_id %in% uninhabited_old$rgn_id, NA, status)) %>% 
    dplyr::bind_rows(ry_old) %>% 
    dplyr::arrange(rgn_id) 
 summary(ry_all_rgns_old) 
  
  status_old <- ry_all_rgns_old %>%
    dplyr::filter(year == 2017) %>%
    dplyr::mutate(dimension = "status") %>%
    dplyr::select(region_id = rgn_id, score = status, dimension, year) %>%
    dplyr::mutate(score = round(score * 100, 2))

  summary(status_old)

  #in this case, 2019 in the ohi global scores we already calculated is actually 2017...
  
  write_csv(status_old, file.path(dir_git_ref, "/check/status_old_ref.csv"))
```


## Calculate scores the NEW WAY (taken (and modified) from functions.R file in OHI Global)
```{r}
  rky_new <-  harvest_tonnes_v2019 %>%
    dplyr::left_join(sustainability_score_v2019,
              by = c('rgn_id', 'taxa_code')) %>%
    dplyr::select(rgn_id, year, taxa_code, taxa_group, tonnes, sust_coeff)
  
  # fill in gaps with no data
  rky_new <- tidyr::spread(rky_new, year, tonnes)
  rky_new <- tidyr::gather(rky_new, "year", "tonnes",-(1:4)) %>%
    dplyr::mutate(year = as.numeric(year))
  
  # adjustment for seaweeds based on protein content
  rky_new <- rky_new %>%
    dplyr::mutate(tonnes = ifelse(taxa_group == "AL", tonnes*0.2, tonnes)) %>%
    dplyr::select(-taxa_group)
  
  # 4-year rolling mean of data
  m_new <- rky_new %>%
    dplyr::group_by(rgn_id, taxa_code, sust_coeff) %>%
    dplyr::arrange(rgn_id, taxa_code, year) %>%
    dplyr::mutate(sm_tonnes = zoo::rollapply(tonnes, 4, mean, na.rm = TRUE, partial =
                                        TRUE, align = "right")) %>%
    dplyr::ungroup()
  
  
  # smoothed mariculture harvest * sustainability coefficient
  m_new <- m_new %>%
    dplyr::mutate(sust_tonnes = sust_coeff * sm_tonnes)
  
  
  # aggregate all weighted timeseries per region, and divide by potential aquaculture tonnes
  ry_new = m_new %>%
    dplyr::group_by(rgn_id, year) %>%
    dplyr::summarize(sust_tonnes_sum = sum(sust_tonnes, na.rm = TRUE)) %>%  #na.rm = TRUE assumes that NA values are 0
    dplyr::left_join(pot_aq_final, by = c('rgn_id')) %>%
       # dplyr::mutate(mar_pop = sust_tonnes_sum / popsum) %>%
    dplyr::ungroup()
  summary(ry_new)
  
  ry_new = ry_new %>%
    dplyr::mutate(status = ifelse(sust_tonnes_sum / potential_mar_tonnes  > 1,
                           1,
                           sust_tonnes_sum / potential_mar_tonnes)) %>%
    dplyr::mutate(status = ifelse(is.na(status),
                                  0,
                                  status)) ## correct for R making 0/0 = NA. 
  
 summary(ry_new) 
  
  
  ## Add all other regions/countries with no mariculture production to the data table
  ## Uninhabited or low population countries that don't have mariculture, should be given a NA since they are too small to ever be able to produce and sustain a mariculture industry.
  ## Countries that have significant population size and fishing activity (these two are proxies for having the infrastructure capacity to develop mariculture), but don't produce any mariculture, are given a '0'.
  
  georegions <- ohicore::georegions
  all_rgns_new <- expand.grid(rgn_id = georegions$rgn_id, year = min(ry_new$year):max(ry_new$year))
  
  all_rgns_new <- all_rgns_new[!(all_rgns_new$rgn_id %in% ry_new$rgn_id),]
  
  uninhabited <- read.csv("https://raw.githubusercontent.com/OHI-Science/ohiprep/master/globalprep/spatial/v2017/output/rgn_uninhabited_islands.csv")
  
  uninhabited <- uninhabited %>% 
    dplyr::filter(rgn_nam != "British Indian Ocean Territory") # remove British Indian Ocean Territory which has fishing activity and a population size of 3000 inhabitants
  
  ## Combine all regions with mariculture data table
  ry_all_rgns_new <- all_rgns_new %>% 
    dplyr::mutate(status = 0) %>% 
    dplyr::mutate(status = ifelse(rgn_id %in% uninhabited$rgn_id, NA, status)) %>% 
    dplyr::bind_rows(ry_new) %>% 
    dplyr::arrange(rgn_id)
  summary(ry_all_rgns_new)
  
  status_new <- ry_all_rgns_new %>%
    dplyr::filter(year == 2017) %>%
    dplyr::mutate(dimension = "status") %>%
    dplyr::select(region_id = rgn_id, score = status, dimension, year) %>%
    dplyr::mutate(score = round(score * 100, 2))

  write_csv(status_new, file.path(dir_git_ref, "/check/status_new_ref.csv"))
```

## Explore some potential rules
 - Those with low sust_tonnes_sum (i.e. rgns 10, 219, and 232) shouldn't be rewarded for "out producing" their potential aquaculture.
 - A rule could be: sust_tonnes_sum < 100 & potential_mar_tonnes < 100, then NA.
 - If potential aquaculture is low, then mariculture should not contribute to a score. 

```{r}
  rky_new_rules <-  harvest_tonnes_v2019 %>%
    dplyr::left_join(sustainability_score_v2019,
              by = c('rgn_id', 'taxa_code')) %>%
    dplyr::select(rgn_id, year, taxa_code, taxa_group, tonnes, sust_coeff)
  
  # fill in gaps with no data
  rky_new_rules <- tidyr::spread(rky_new_rules, year, tonnes)
  rky_new_rules <- tidyr::gather(rky_new_rules, "year", "tonnes",-(1:4)) %>%
    dplyr::mutate(year = as.numeric(year))
  
  # adjustment for seaweeds based on protein content
  rky_new_rules <- rky_new_rules %>%
    dplyr::mutate(tonnes = ifelse(taxa_group == "AL", tonnes*0.2, tonnes)) %>%
    dplyr::select(-taxa_group)
  
  # 4-year rolling mean of data
  m_new_rules <- rky_new_rules %>%
    dplyr::group_by(rgn_id, taxa_code, sust_coeff) %>%
    dplyr::arrange(rgn_id, taxa_code, year) %>%
    dplyr::mutate(sm_tonnes = zoo::rollapply(tonnes, 4, mean, na.rm = TRUE, partial =
                                        TRUE, align = "right")) %>%
    dplyr::ungroup()
  
  
  # smoothed mariculture harvest * sustainability coefficient
  m_new_rules <- m_new_rules %>%
    dplyr::mutate(sust_tonnes = sust_coeff * sm_tonnes)
  
  
  # aggregate all weighted timeseries per region, and divide by potential aquaculture tonnes
  ry_new_rules = m_new_rules %>%
    dplyr::group_by(rgn_id, year) %>%
    dplyr::summarize(sust_tonnes_sum = sum(sust_tonnes, na.rm = TRUE)) %>%  #na.rm = TRUE assumes that NA values are 0
    dplyr::left_join(pot_aq_final, by = c('rgn_id')) %>%
       # dplyr::mutate(mar_pop = sust_tonnes_sum / popsum) %>%
    dplyr::ungroup()
  
  ry_new_rules = ry_new_rules %>%
    dplyr::mutate(status = ifelse(sust_tonnes_sum / potential_mar_tonnes  > 1,
                           1,
                           sust_tonnes_sum / potential_mar_tonnes)) %>%
    dplyr::mutate(status = ifelse(is.na(status),
                                  0,
                                  status)) %>% ## correct for R making 0/0 = NA.
    dplyr::mutate(status = ifelse(sust_tonnes_sum < 100 & potential_mar_tonnes < 100, 
                                  NA, 
                                  status))
  
 summary(ry_new_rules) 
  
  
  ## Add all other regions/countries with no mariculture production to the data table
  ## Uninhabited or low population countries that don't have mariculture, should be given a NA since they are too small to ever be able to produce and sustain a mariculture industry.
  ## Countries that have significant population size and fishing activity (these two are proxies for having the infrastructure capacity to develop mariculture), but don't produce any mariculture, are given a '0'.
  
  georegions <- ohicore::georegions
  all_rgns_new_rules <- expand.grid(rgn_id = georegions$rgn_id, year = min(ry_new_rules$year):max(ry_new_rules$year))
  
  all_rgns_new_rules <- all_rgns_new_rules[!(all_rgns_new_rules$rgn_id %in% ry_new_rules$rgn_id),]
  
  uninhabited <- read.csv("https://raw.githubusercontent.com/OHI-Science/ohiprep/master/globalprep/spatial/v2017/output/rgn_uninhabited_islands.csv")
  
  uninhabited <- uninhabited %>% 
    dplyr::filter(rgn_nam != "British Indian Ocean Territory") # remove British Indian Ocean Territory which has fishing activity and a population size of 3000 inhabitants
  
  ## Combine all regions with mariculture data table
  ry_all_rgns_new_rules <- all_rgns_new_rules %>% 
    dplyr::mutate(status = 0) %>% 
    dplyr::mutate(status = ifelse(rgn_id %in% uninhabited$rgn_id, NA, status)) %>% 
    dplyr::bind_rows(ry_new_rules) %>% 
    dplyr::arrange(rgn_id)
  summary(ry_all_rgns_new_rules)
  
  status_new_rules <- ry_all_rgns_new_rules %>%
    dplyr::filter(year == 2017) %>%
    dplyr::mutate(dimension = "status") %>%
    dplyr::select(region_id = rgn_id, score = status, dimension, year) %>%
    dplyr::mutate(score = round(score * 100, 2))

  write_csv(status_new_rules, file.path(dir_git_ref, "/check/status_new_ref_rules.csv"))
```


## Compare New and Old Scores

```{r, eval = TRUE, warning = FALSE}
#read in status data
region_data()

status_new <- read_csv(file.path(dir_git_ref, "/check/status_new_ref.csv")) %>%
  left_join(rgns_eez, by = c("region_id" = "rgn_id")) %>%
  select(region_id:rgn_name)
status_old <- read_csv(file.path(dir_git_ref, "/check/status_old_ref.csv")) %>%
  left_join(rgns_eez, by = c("region_id" = "rgn_id")) %>%
  select(region_id:rgn_name)

#join status data
compare <- status_old %>%
  full_join(status_new, by = c("region_id", "rgn_name")) %>%
  rename(old_score = score.x, new_score = score.y)

#calculate summary statistics for the differences 
compare_stats <- compare %>%
  mutate(score_diff = old_score - new_score) %>%
  mutate(score_diff_abs = abs(score_diff)) %>%
  select(-dimension.x, -dimension.y, -year.x, -year.y)

sum(compare_stats$score_diff > 0, na.rm = TRUE) # 81 scores got worse
sum(compare_stats$score_diff < 0, na.rm = TRUE) # 25 scores got better 
sum(compare_stats$score_diff == 0 & (compare_stats$new_score != 0|compare_stats$old_score != 0), na.rm = TRUE) # 3 scores didn't change that were not already 0
sum(compare_stats$score_diff == 0 & (compare_stats$new_score == 0|compare_stats$old_score == 0), na.rm = TRUE)# 91 scores didn't change that were already 0. 
sum(is.na(compare_stats$score_diff)) # 20 NAs
mean(compare_stats$score_diff > 0, na.rm = TRUE)
mean(compare_stats$score_diff < 0, na.rm = TRUE)

#arrange to see the top scorers for old and new reference point
top_10_old <- status_old %>%
  arrange(-score) %>%
  select(region_id, rgn_name, score)
DT::datatable(top_10_old)

top_10_new <- status_new %>%
  arrange(-score) %>%
  select(region_id, rgn_name, score)
DT::datatable(top_10_new)
```

```{r, eval=TRUE}
##explore the aquaculture data

#graph to see the actual differences 
library(plotly)
library(ggplot2)
p <- ggplot(compare, aes(x=old_score, y=new_score)) +
  geom_point(aes(text = paste0("rgn = ", region_id)), shape=19) +
  geom_abline(slope=1, intercept=0) +
  theme_bw()

#ggsave(file.path(dir_git_ref, "/check/compare_reference_graph.png"), p)

plotly_fig <- plotly::ggplotly(p)
plotly_fig


# plot histograms to determine a good cutoff for potential_mar_tonnes. Some regions are scoring too well because they have 0 potential for aquaculture. For example, Nauru barely harvests any seafood, and has no potential for aquaculture, yet they score a 100. 
bw <- 2 * IQR(pot_aq_final$potential_mar_tonnes) / length(pot_aq_final$potential_mar_tonnes)^(1/3)

ggplot(pot_aq_final, aes(x = potential_mar_tonnes)) +
  geom_histogram(binwidth = bw) +
  theme_bw() +
  scale_x_continuous(expand = c(0,0), breaks = seq(0, 40000000, by  = 2500000)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(axis.text.x = element_text(angle = -45, hjust = 1))

#check for those with potential tonnes <= 100000
ggplot(filter(pot_aq_final, potential_mar_tonnes <= 100000), aes(x = potential_mar_tonnes)) +
  geom_histogram() +
  theme_bw() +
  scale_x_continuous(breaks = seq(0, 100000, by  = 5000)) +
  theme(axis.text.x = element_text(angle = -45, hjust = 1))

summary(pot_aq_final)  # median potential is 66996 tonnes....
summary(harvest_tonnes_v2019) # median is 102 tonnes....

less_100000 <- pot_aq_final %>%
  filter(potential_mar_tonnes <= 100000) %>%
  left_join(rgns_eez, by = "rgn_id") %>%
  select(rgn_id, rgn_name, potential_mar_tonnes) %>%
  arrange(-potential_mar_tonnes)



library(knitr)
  #pretty table
  print(kable(less_100000))

# View in a way so that we can see sust_tonnes_sum, potential_mar_tonnes, and score. This shows us the regions that might have an unfair advantage with the new scoring system. 
  status_new_check <- ry_all_rgns_new %>%
    dplyr::filter(year == 2017) %>%
    dplyr::mutate(dimension = "status") %>%
    dplyr::select(region_id = rgn_id, sust_tonnes_sum, potential_mar_tonnes, score = status, dimension, year) %>%
    dplyr::mutate(score = round(score * 100, 2)) %>% 
    filter(score == 100, potential_mar_tonnes == 0)
```

## Compare new rules scores and old scores
```{r, eval = TRUE, warning = FALSE}
#read in status data
region_data()

status_old <- read_csv(file.path(dir_git_ref, "/check/status_old_ref.csv")) %>%
  left_join(rgns_eez, by = c("region_id" = "rgn_id")) %>%
  select(region_id:rgn_name)
status_new_rules <- read_csv(file.path(dir_git_ref, "/check/status_new_ref_rules.csv")) %>% 
  left_join(rgns_eez, by = c("region_id" = "rgn_id")) %>%
  select(region_id:rgn_name)

#join status data
compare_rules <- status_old %>%
  full_join(status_new_rules, by = c("region_id", "rgn_name")) %>%
  rename(old_score = score.x, new_score = score.y)

compare_new_rule_change <- status_new_rules %>%
  full_join(status_new_rules, by = c("region_id", "rgn_name")) %>%
   rename(old_score = score.x, new_score = score.y)

#calculate summary statistics for the differences 
compare_stats_rules <- compare_rules %>%
  mutate(score_diff = old_score - new_score) %>%
  mutate(score_diff_abs = abs(score_diff)) %>%
  select(-dimension.x, -dimension.y, -year.x, -year.y)

compare_stats_rules_change_new <- compare_new_rule_change %>%
  mutate(score_diff = old_score - new_score) %>%
  mutate(score_diff_abs = abs(score_diff)) %>%
  select(-dimension.x, -dimension.y, -year.x, -year.y)

sum(compare_stats_rules$score_diff > 0, na.rm = TRUE) # 81 scores got worse
sum(compare_stats_rules$score_diff < 0, na.rm = TRUE) # 22 scores got better 
sum(compare_stats_rules$score_diff == 0 & (compare_stats_rules$new_score != 0|compare_stats_rules$old_score != 0), na.rm = TRUE) # 3 scores didn't change that were not already 0
sum(compare_stats_rules$score_diff == 0 & (compare_stats_rules$new_score == 0|compare_stats_rules$old_score == 0), na.rm = TRUE)# 88 scores didn't change that were already 0. 
sum(is.na(compare_stats_rules$score_diff)) # 26 NAs
mean(compare_stats_rules$score_diff > 0, na.rm = TRUE) # 0.4175258 average loss in score
mean(compare_stats_rules$score_diff < 0, na.rm = TRUE) # 0.1134021 average gain in score

#arrange to see the top scorers for old and new reference point
top_10_old <- status_old %>%
  arrange(-score) %>%
  select(region_id, rgn_name, score)
DT::datatable(top_10_old)

top_10_new_rules <- status_new_rules %>%
  arrange(-score) %>%
  select(region_id, rgn_name, score)
DT::datatable(top_10_new_rules)
```

```{r, eval=TRUE}
##explore the aquaculture data with new reference rules established above 

#graph to see the actual differences 
library(plotly)
library(ggplot2)
p <- ggplot(compare_rules, aes(x=old_score, y=new_score)) +
  geom_point(aes(text = paste0("rgn = ", region_id)), shape=19) +
  geom_abline(slope=1, intercept=0) +
  theme_bw()

p

```


