---
title: "Data Download"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
library(xlsx)
```

# New Global Fisheries Catch Data

* Industrial Catch (1950 - 2015) - reported, iuu, and discard catch data for each cell location and unique identifier
* Non-Industrial Catch (1950 - 2015) - reported, iuu, and discard catch data for each cell location and unique identifier
* Master Index File (Index.csv) - information associated with the unique identifiers. This is split into two files, "IndexInd" (industrial) and "IndexNInd" (non-industrial).
* DATA CODE DEFINITIONS (gear/taxa/country codes and cell lat/lon references)

```{r}

source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2019/gh-pages/workflow/R/common.R")
  
## Save download url suffixes 
web_years <- c("Ind1950_1954", "Ind1955_1959", "Ind1960_1964", "Ind1965_1969", 
               "Ind1970_1974", "Ind1975_1979", "Ind1980_1984", "Ind1985_1989",
               "Ind1990_1994", "Ind1995_1999", "Ind2000_2004", "Ind2005_2009", 
               "Ind2010_2014", "Ind2015_2015", "NInd1950_1954", "NInd1955_1959",
               "NInd1960_1964", "NInd1965_1969", "NInd1970_1974", "NInd1975_1979",
               "NInd1980_1984", "NInd1985_1989", "NInd1990_1994", "NInd1995_1999", 
               "NInd2000_2004", "NInd2005_2009", "NInd2010_2014", "NInd2015_2015")

reference <- c("IndexInd", "IndexNInd")

## Download reference data from web and save into mazu
for(ref in reference){  #ref <- "IndexInd"
  
  data <- read.csv(sprintf("https://data.imas.utas.edu.au/attachments/5c4590d3-a45a-4d37-bf8b-ecd145cb356d/%s.csv", ref))
  
  write.csv(data, file.path(dir_M, sprintf("git-annex/globalprep/_raw_data/test_IMAS/d2019/%s.csv", ref)), row.names=F)
}

## Download catch data from web and save into mazu
for(web_year in web_years){  
  #web_year <- "Ind1950_1954"

data <- read.csv(sprintf("https://data.imas.utas.edu.au/attachments/5c4590d3-a45a-4d37-bf8b-ecd145cb356d/Catch%s.csv", web_year))

saveRDS(data, file.path(dir_M, sprintf("git-annex/globalprep/_raw_data/test_IMAS/d2019/Catch%s.rds", web_year)))

}
```

# How to download the data code definitions. 

For this, it is saved as an .xlsx file (Codes.xlsx). This file has 4 sheets in it which contain meta data that explain columns in the Index files, and one sheet with is "Spatial Cells Reference - contains geospatial information associated wtih the Industrial Catch data". 
To download this data, go to the IMAS website: http://data.imas.utas.edu.au/portal/search?uuid=ff1274e1-c0ab-411b-a8a2-5a12eb27f2c0 and dowload it manually.

This .xlsx file contains these sheets:

* Spatial Cells Reference (Cells) - contains geospatial information associated wtih the Industrial Catch data
* Gear Reference (Gear) - contains information regarding how different fishing gear is classified in the index datasets.
* Taxa Reference (Taxa) - contains information regarding how different taxa is classified in the index datasets.
* Country Reference (Country) - contains informations regarding how different countries are labeled in the index datasets. 



