---
title: 'OHI 2018 - Iconic species subgoal'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohiprep_v2018/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---


[REFERENCE RMD FILE](https://rawgit.com/OHI-Science/ohi-global/draft/global_supplement/Supplement.html#591_iconic_species_(subgoal_of_sense_of_place))


# Summary:  Iconic Species Subgoal (Sense of Place)

This script prepares scores (status and trend) for Iconic Species in each global coastal region.  For each iconic marine species, the countries of occurrence are pulled from the IUCN API.  Extinction risk categories for each species are pulled based on current and past assessments; by tracking the assessed extinction risk over time, we can understand the trends of extinction risk for iconic species directly rather than using the "population trend" method from prior OHI assessments.

The Iconic Species sub-goal model calculates a region's status based upon an unweighted average of species health for all 'iconic' species found within each reporting region.

From Halpern et al (2012):

> Iconic species are those that are relevant to local cultural identity through a species’ relationship to one or more of the following: 1) traditional activities such as fishing, hunting or commerce; 2) local ethnic or religious practices; 3) existence value; and 4) locally-recognized aesthetic value (e.g., touristic attractions/common subjects for art such as whales). Habitat-forming species are not included in this definition of iconic species, nor are species that are harvested solely for economic or utilitarian purposes (even though they may be iconic to a sector or individual). ...

> Ultimately, almost any species can be iconic to someone, and so the intent with this goal was to focus on those species widely seen as iconic within a country, and iconic from a cultural or existence value (rather than for a livelihoods or extractive reason). ...

> The reference point is to have the risk status of all assessed species as Least Concern (i.e., a goal score = 1.0)

The Status of this sub-goal (X~ICO~) is then the % of iconic species in each threat category (as defined by the IUCN Red List), such that:

$$X_{ICO} = \frac{\displaystyle\sum_{category}S_{cat}*w_{cat}}{\displaystyle\sum_{category}S_{cat}}$$

where for each IUCN threat category:

* *S~cat~* is the number of assessed species in the category
* *w~cat~* is the status weight assigned for that category (note, these are the inverse of the risk value used in the SPP calculations):
    * 'LC' = 1.0, 'NT' = 0.8, 'VU' = 0.6, 'EN' = 0.4, 'CR' = 0.2, 'EX' = 0.0

ICO trend is calculated in a similar manner, but weightings are assigned according to IUCN population trend: 'Decreasing' = -0.5, 'Stable' = 0.0, 'Increasing' = +0.5.  

# Updates from previous assessment

No changes have been made to the data prep methods for this goal, since the 2016 assessment, though there were some slight changes to improve the code. API key is read from a file, the 'get from API' function now includes error checking (it reports a little on what the error is, whether it's a zero-length data frame or a try error, etc) and multiple attempts (if it gets an error it tries again a few times), and there is also a delay parameter to space out requests to the API. 

Changes _were_ made to the iconic species trend calculation. Trends are now calculated (in ohi-global, `functions.R`) using status values based only on species with two or more IUCN evaluations in the last 20 years. This change was made because including species which have not been re-evaluated for a long time tends to positively bias trend values, or at least dampen the trends by essentially including as zeros, values that are not truely known to be zero... To accommodate this change in the trend calculation, years 1992-present are included in the prepped data as well as in the ohi-global `scenario_data_years.csv`.

## Future improvements?

* Update list of iconic species... see issue #671 (not 2018 assessment, see previous assesment issues)


# Data Sources

**List of iconic species:**

**Species native country information:**

* __Reference__: 
    * IUCN 2016. IUCN Red List of Threatened Species. Version 2016-1 <www.iucnredlist.org>
        * __Accessed__: `r date()` by `r Sys.info()['user']` 
* __Native data resolution__: Country level (by country name)
* __Time range__: 1965-2015 (discrete past assessments by species)
* __Format__:  JSON

***
  
# Methods

## Setup

```{r setup, echo = FALSE, message = FALSE, warning = FALSE, eval=FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = FALSE, message = FALSE, warning = FALSE)

library(jsonlite)
library(readr)
library(tidyverse)
library(parallel)

## set up provenance tracking for this script:
# source(file.path('~/github/ohibc', 'src/R/prov.R'))  

goal          <- 'globalprep/ico'
scenario      <- 'v2018'
previous_scen <- '2017'
source(sprintf('~/github/ohiprep_%s/src/R/common.R', scenario))

dir_anx      <- file.path(dir_M, 'git-annex', goal)
dir_goal     <- file.path(sprintf('~/github/ohiprep_%s', scenario), goal, scenario)
dir_previous <- file.path(sprintf('~/github/ohiprep_%s/%s/v%s', scenario, goal, previous_scen))
source(file.path(dir_goal, 'ico_fxn.R'))

if(!file.exists(file.path(dir_goal, 'README.md'))) {
  warning(sprintf('No README detected in %s', dir_goal))
}
if(!file.exists(file.path(dirname(dir_goal), 'README.md'))) {
  warning(sprintf('No README detected in %s', file.path(dirname(dir_goal))))
}
```


Using the IUCN API, we accessed the full IUCN species list at http://apiv3.iucnredlist.org/api/v3/speciescount?token=<token>.  With some minor formatting, this list contains the following variables:

iucn_sid | kingdom | phylum | class | order | family | genus | sciname | population | category


## Download Species List from IUCN

Get all pages and bind into total species list, using the IUCN API.

``` {r get_spp_info, eval= F}

api_file <- file.path(dir_anx, 'api_key.csv')
api_key <- scan(api_file, what = 'character') # fb71ae836f415f04f41176f6d30c4a9e4cea620d46b9e5021bf2fb142ea51bf5

spp_list_from_api_file <- file.path(dir_goal, 'int/spp_list_from_api.csv')
reload <- TRUE

if(!file.exists(spp_list_from_api_file) | reload) {
  message('Using API to create full species list from scratch')
  
  spp_npage_url <- sprintf('http://apiv3.iucnredlist.org/api/v3/speciescount?token=%s', api_key)
  n_spp <- fromJSON(spp_npage_url) %>%
    .$count %>% as.integer()
  n_pages <- ceiling(n_spp/10000)
  
  spp_page_url <- 'http://apiv3.iucnredlist.org/api/v3/species/page/%s?token=%s'
  spp_df_all <- mc_get_from_api(spp_page_url, c(0:(n_pages - 1)), api_key) # mc_get_from_api function is defined in the ico_fxn.R script
  
  spp_df_all <- spp_df_all %>%
    dplyr::select(-infra_rank, -infra_name, -count, -page) %>%
    rename(iucn_sid = taxonid, sciname = scientific_name) %>%
    setNames(names(.) %>%
               str_replace('_name', ''))
  
  write_csv(spp_df_all, spp_list_from_api_file)
} else {
  message('reading file of API species list: \n  ', spp_list_from_api_file)
  spp_df_all <- read_csv(spp_list_from_api_file)
}

```



## Get master list of Iconic Species

The list of Iconic Species is based upon the original ICO list generated in 2011, using species identified as globally iconic (WWF Flagship species and Priority species) or regionally iconic (based upon WWF regional/local priority species and nation-specific lists).


``` {r get_ico_spp_list, eval=FALSE}
ico_list_raw <- get_ico_list(reload = TRUE) # get_ico_list is defined in ico_fxn.R

## head of ico_list_raw for v2018:

## comname                          sciname               ico_gl ico_rgn_id
##   <chr>                            <chr>                 <lgl>       <int>
## 1 Blue Shark                       Prionace glauca       T              NA
## 2 Whale Shark                      Rhincodon typus       T              NA
## 3 Shortfin Mako                    Isurus oxyrinchus     T              NA
## 4 Olive Ridley Turtle              Lepidochelys olivacea T              NA
## 5 Irrawaddy Dolphin                Orcaella brevirostris T              NA
## 6 Humphead wrasse, Napoleon Wrasse Cheilinus undulatus   T              NA

## ico_rgn_id: rgn_id in which species is iconic by regional/national lists; 
## if globally iconic, ico_rgn_id <- NA

write_csv(ico_list_raw, file.path(dir_goal, 'raw/ico_list_raw.csv'))
```

```{r, eval=F}
DT::datatable(ico_list_raw, caption = 'Iconic Species List')
```



***

## Identify countries with extant ICO species populations

Filtering the complete IUCN species list to include only the identified Iconic Species, we then use the IUCN API to access the list of countries in which each species occurs, from http://apiv3.iucnredlist.org/api/v3/species/countries/id/<species ID>?token=<token>.  The country list identifies whether the species' presence in that country is "Extant", "Extinct Post-1500", or "Possibly Extinct"; the "Extinct Post-1500" presence will be used later to identify locally extinct populations.

``` {r combine_iucn_spp_info_with_ico_list, eval= F}

spp_df_all <- read_csv(file.path(dir_goal, 'int/spp_list_from_api.csv'))
ico_list_raw <- read_csv(file.path(dir_goal, 'raw/ico_list_raw.csv')) # static list; same each year

spp_ico <- spp_df_all %>% 
  filter(sciname %in% ico_list_raw$sciname) 

spp_missing <- ico_list_raw %>% 
  filter(!sciname %in% spp_ico$sciname) # for v2018, missing only Sprattus sprattus

ico_list <- ico_list_raw %>%
  left_join(spp_ico %>% 
              select(iucn_sid, sciname, subpop = population, cat = category),
            by = 'sciname') %>%
  filter(!is.na(iucn_sid))

write_csv(ico_list, file.path(dir_goal, 'int/ico_list_prepped.csv')) # 144 in v2018
```

For each of these species, use the IUCN API to gather a list of countries in which it is present.

``` {r get_country_list, eval= F}

## for each species ID, get country list
ico_country_url <- 'http://apiv3.iucnredlist.org/api/v3/species/countries/id/%s?token=%s'

ico_spp_countries <- mc_get_from_api(ico_country_url, ico_list$iucn_sid, api_key)

rgn_iucn2ohi <- read_csv(file.path(dir_goal, 'raw/rgns_iucn2ohi.csv'))

ico_spp_rgn_raw <- ico_spp_countries %>%
  select(-code, -count, iucn_sid = name, iucn_rgn_name = country) %>% 
  mutate(iucn_sid = as.integer(iucn_sid),
         iucn_rgn_name  = str_trim(iucn_rgn_name)) %>% 
  left_join(rgn_iucn2ohi,
            by = 'iucn_rgn_name')

## error check on region name matching; landlocked --> not matched
non_match <- ico_spp_rgn_raw %>%
  filter(is.na(ohi_rgn_name))
if(nrow(non_match) > 0) {
  cat('The following IUCN countries did not match with OHI region names:\n  ')
  print(paste(non_match$iucn_rgn_name %>% unique(), collapse = ', '))
}

ico_spp_rgn_raw <- ico_spp_rgn_raw %>%
  rename(rgn_name = ohi_rgn_name) %>%
  select(-iucn_rgn_name) %>%
  filter(!is.na(rgn_id)) %>%
  distinct()

write_csv(ico_spp_rgn_raw, file.path(dir_goal, 'raw/ico_spp_rgn_raw.csv'))

```


``` {r attach_country_list_to_spp_info_list, eval=F}

ico_spp_rgn_raw <- read_csv(file.path(dir_goal, 'raw/ico_spp_rgn_raw.csv'))
ico_list <- read_csv(file.path(dir_goal, 'int/ico_list_prepped.csv'))

ico_spp_rgn_prepped <- ico_spp_rgn_raw %>%
  left_join(ico_list, by = 'iucn_sid')

## filter this for species who are global (so all instances are iconic) 
##  OR ico_rgn_id matches rgn_id (so locally iconic matches with location)
ico_spp_rgn_prepped <- ico_spp_rgn_prepped %>%
  filter(ico_gl == TRUE | ico_rgn_id == rgn_id)

write_csv(ico_spp_rgn_prepped, file.path(dir_goal, 'int/ico_spp_rgn_prepped.csv'))

ico_spp_rgn_prepped <- ico_spp_rgn_prepped %>% 
  select(iucn_sid, sciname, comname, presence, origin, distribution_code, cat, ico_gl, rgn_id, rgn_name, subpop, ico_rgn_id, api_error, param_id)

```


```{r, eval=F}
`ico_spp_rgn_prepped` saved at `~/github/ohiprep/globalprep/ico/v2016/int/ico_spp_rgn_prepped.csv`
```


---

```{r, eval=F}
DT::datatable(ico_spp_rgn_prepped, width = 400)
```

***

## Identify extinction risk from current and past assessments

We accessed the IUCN API to determine past IUCN assessments for each of the identified iconic species: http://apiv3.iucnredlist.org/api/v3/species/history/id/<species id>?token=<token>.

Each assessment includes a year and an extinction risk, along with additional information on the assessment.

``` {r get_category_list, eval=F}

ico_past_assess_url <- 'http://apiv3.iucnredlist.org/api/v3/species/history/id/%s?token=%s'
ico_list <- read_csv(file.path(dir_goal, 'int/ico_list_prepped.csv'))

## for each species ID, get past assessments
ico_assess_raw <- mc_get_from_api(ico_past_assess_url, ico_list$iucn_sid, api_key)

ico_assess_raw <- ico_assess_raw %>% 
  rename(iucn_sid = name) %>%
  mutate(iucn_sid = as.integer(iucn_sid),
         year     = as.integer(year)) %>% 
  left_join(ico_list %>% 
              select(iucn_sid, sciname) %>%
              distinct(),
            by = 'iucn_sid')

write_csv(ico_assess_raw, file.path(dir_goal, 'raw/ico_assessments_raw.csv'))

```

```{r, eval=F}
DT::datatable(ico_assess_raw, caption = 'ICO species and past IUCN assessments')
```

These raw assessments are cleaned up to standardize codes and categories, and category scores are assigned based on a scale from Least Concern = 1 to Extinct = 0, as per the [Species OHI subgoal](https://rawgit.com/OHI-Science/ohi-global/draft/global_supplement/Supplement.html#591_iconic_species_(subgoal_of_sense_of_place)).  Note that past assessments used different coding, especially prior to 1994; we reclassified older codes according to this chart:

---

\begin{eqnarray*}
\textbf{New Category} & & \longleftarrow \textbf{Original Category/Description}\\
NT & & \longleftarrow \text{"Lower risk/Near threatened ($LR/NT$)"}\\
T  & & \longleftarrow \text{"Threatened ($T$)" treat as "$EN$"}\\
VU & & \longleftarrow \text{"Vulnerable ($V$)"}\\                  
EN & & \longleftarrow \text{"Endangered ($E$)" }\\
LR/CD & & \longleftarrow \text{"Lower risk/Conservation dependent ($LR/CD$)" treat as between $VU$ and $NT$}\\
CR & & \longleftarrow \text{"Very rare and believed to be decreasing in numbers"}\\
T & & \longleftarrow \text{"Less rare but believed to be threatened-Requires watching"}\\
DD & & \longleftarrow \text{"Insufficiently known ($K$)"}\\
DD & & \longleftarrow \text{"Indeterminate ($I$)"}\\
DD & & \longleftarrow \text{"Status inadequately known-Survey required or data sought"}\\
NE & & \longleftarrow \text{"Not recognized ($NR$)"}
\end{eqnarray*}

---

## Clean up the time series

iucn_sid | year | code | category | sciname


``` {r clean_up_category_list, eval= F}

ico_assess_raw <- read_csv(file.path(dir_goal, 'raw/ico_assessments_raw.csv'))

ico_assess <- ico_assess_raw %>%
  rename(cat = code, cat_txt = category) %>%
  mutate(cat = toupper(cat),
         cat = str_replace(cat, 'LR/', ''),
         cat = ifelse(cat %in% c('K', 'I'), 'DD', cat),
         cat = ifelse(cat == 'NR', 'NE', cat),
         cat = ifelse(str_detect(toupper(cat_txt), 'VERY RARE'), 'CR', cat),
         cat = ifelse(str_detect(toupper(cat_txt), 'LESS RARE'), 'T', cat),
         cat = ifelse(str_detect(toupper(cat_txt), 'STATUS INADEQUATELY KNOWN'), 'DD', cat),
         cat = ifelse(cat == 'V', 'VU', cat), 
         cat = ifelse(cat == 'E', 'EN', cat))

pop_cat <- data.frame(cat       = c("LC", "NT", "VU", "EN", "CR", "EX", "T", "CD", "NE", "DD"), 
                      cat_score = c(   0,  0.2,  0.4,  0.6,  0.8,  1.0, 0.6,  0.3,   NA,  NA),
                      stringsAsFactors = FALSE)
  
ico_assess <- ico_assess %>% 
  left_join(pop_cat, by = 'cat') %>%
  filter(!is.na(cat_score)) %>%
  distinct() %>%
  arrange(iucn_sid, year)

write_csv(ico_assess, file.path(dir_goal, 'int/ico_assess_clean.csv'))

```

***

Using `tidyr::complete()` and `tidyr::fill()`, we create a full time series for all species from the earliest assessment to the most recent year.

``` {r fill_out_category_time_series, eval= F}

ico_assess <- read_csv(file.path(dir_goal, 'int/ico_assess_clean.csv'))
ico_list <- read_csv(file.path(dir_goal, 'int/ico_list_prepped.csv'))

ico_assess_full <- ico_assess %>%
  mutate(eval_yr = year) %>% 
  select(-sciname) %>%
  arrange(iucn_sid, year) %>%
  complete(year = full_seq(year, 1), nesting(iucn_sid)) %>%
  group_by(iucn_sid) %>%
  fill(cat, cat_txt, cat_score, eval_yr) %>% ## fills all the way to latest year
  ungroup()

ico_spp_cat <- ico_list %>% 
  rename(cat_2016 = cat) %>%
  left_join(ico_assess_full, by = c('iucn_sid'))

## if no time series available, time series years will be NA.  Assign a list to
## those NAs, then unnest it to create observations for those years.
ico_spp_cat <- ico_spp_cat %>%
  mutate(year = ifelse(is.na(year), 
                       list(c(min(year, na.rm = TRUE):max(year, na.rm = TRUE))), 
                       year)) %>% 
  unnest(year)

## NAs will be filled backward in time by starting from the most recent non-NA.
## To do this, we'll swap any current-year NAs with the cat_score (meaning no
## time series fill), and fill upwards instead of downwards.
ico_spp_cat <- ico_spp_cat %>%
  left_join(pop_cat %>% 
              rename(cat_2016 = cat, cat_2016_score = cat_score), 
            by = 'cat_2016') %>%
  mutate(cat_score = ifelse(year == max(year, na.rm = TRUE) & is.na(cat), 
                            cat_2016_score, 
                            cat_score)) %>%
  arrange(iucn_sid, year) %>%
  group_by(iucn_sid) %>%
  fill(cat, cat_score, cat_txt, .direction = 'up') %>%
  ungroup() %>%
  distinct()
  
write_csv(ico_spp_cat, file.path(dir_goal, 'int/ico_spp_cat.csv'))

```

## Combine IUCN risk category time series with country <-> species lookup table

Using `dplyr::full_join()` we combine the `ico_spp_rgn` dataframe (iconic species by OHI region) with the `ico_spp_cat` dataframe (iconic species by category and year, with species info, year, and category info).  

``` {r combine_ico_spp_cat_with_countries, eval= F}

ico_cat_ts_abbr <- read_csv(file.path(dir_goal, 'int/ico_spp_cat.csv')) %>%
  select(iucn_sid, sciname, year, cat, cat_score, eval_yr) %>%
  filter(year >= 1992)

ico_spp_rgn <- read_csv(file.path(dir_goal, 'int/ico_spp_rgn_prepped.csv')) %>%
  select(rgn_id, rgn_name, iucn_sid, comname, sciname, ico_gl, ico_rgn_id, presence)

ico_spp_rgn_cat <- ico_cat_ts_abbr %>% 
  full_join(ico_spp_rgn, by = c('iucn_sid', 'sciname'))

### No information on when species go extinct locally, so just set all years to extinct for that region
ico_spp_rgn_cat <- ico_spp_rgn_cat %>%
  mutate(cat = ifelse(str_detect(presence, '^Extinct'), 'EX', cat), # ^ indicates start of string
         cat_score = ifelse(cat == 'EX', 1, cat_score)) %>%
  filter(ico_gl | ico_rgn_id == rgn_id) %>% # keep (all globally iconic) and (regionally iconic in region only)
  distinct()

write_csv(ico_spp_rgn_cat, file.path(dir_goal, 'int/ico_spp_rgn_cat.csv'))

```

```{r, eval=FALSE}
knitr::kable(head(ico_spp_rgn_cat, 10))
```

***

## Prep dataframe for toolbox; estimate status and trend

The toolbox wants `rgn_id`, species `sciname`, and extinction risk `category` for the basic calculations.  Since some regions contain multiple subpops (or parent/subpop) we also include `iucn_sid` to differentiate.  This information is included for each `year`, filtered back to the year 2000.

While the official calculations are performed in the toolbox, we perform the same basic calcs here to get a sense of the ICO status and trend ahead of time. Report and summarize estimate of regional iconic species status:

``` {r report_status_and_trend, eval= F}

ico_spp_rgn_cat <- read_csv(file.path(dir_goal, 'int/ico_spp_rgn_cat.csv'))

## Report out for toolbox format (rgn_id | sciname | category or popn_trend for each species within a region).
## Note: in toolbox, group_by(rgn_id, sciname) and then summarize(category = mean(category)) to
##   average any parent/subpop species listings before aggregating to overall average per region.

ico_status_raw <- ico_spp_rgn_cat %>%
  select(rgn_id, rgn_name, sciname, iucn_sid, cat, cat_score, year, eval_yr) %>%
  arrange(rgn_id, desc(year), sciname) %>%
  ungroup()

ico_status_calc <- ico_status_raw %>%
  group_by(rgn_id, rgn_name, sciname, year) %>%
  filter(!is.na(cat_score)) %>% # remove any DDs
  summarize(cat_score = mean(cat_score)) %>% 
  group_by(rgn_id, rgn_name, year) %>%
  summarize(mean_cat = round(mean(cat_score), 5), 
            ico_status = (1 - mean_cat) * 100,
            n_spp = n()) %>%
  ungroup()

ico_trend <- data.frame()
for (i in 1993:max(ico_status_calc$year, na.rm = TRUE)) { # i <- 2013
  tmp_status <- ico_status_calc %>%
    filter(year <= i & year > (i - 10)) # trend based on 10-year average since assessments are sporadic
  tmp_trend <- tmp_status %>%
    group_by(rgn_id) %>%
    do(trend_lm = lm(ico_status ~ year, data = .)$coefficients[2]) %>%
    mutate(year = i,
           trend_lm  = as.numeric(trend_lm)/100, # status is 0 - 100; trend should be +1 to -1
           ico_trend = round(trend_lm * 5, 5)) %>% # trend prediction five years out
    ungroup()
  ico_trend <- ico_trend %>%
    bind_rows(tmp_trend)
}

ico_sum <- ico_status_raw %>%
  left_join(ico_status_calc, by = c('rgn_id', 'rgn_name', 'year')) %>%
  left_join(ico_trend, by = c('rgn_id', 'year'))

write_csv(ico_sum, file.path(dir_goal, 'summary/ico_summary.csv'))
### Report out for finalized status and trend values per region

ico_status_raw1 <- ico_status_raw %>%
  dplyr::select(rgn_id, sciname, iucn_sid, year, eval_yr, category = cat)

write_csv(ico_status_raw1, file.path(dir_goal, 'output/ico_spp_iucn_status.csv'))
write_csv(ico_status_calc, file.path(dir_goal, 'output/ico_status_calc.csv'))
write_csv(ico_trend,       file.path(dir_goal, 'output/ico_trend.csv'))

ico_status_raw1[duplicated(ico_status_raw1 ), ] 
### NOTE: if iucn_sid were removed, this would show duplicates due to subpops with same category.
table(ico_status_raw1$category)

```

### Iconic Species full list (year == 2015)

```{r, eval=F}
DT::datatable(ico_status_raw %>% filter(year == 2017))
```


### Iconic Species processed status and trend by region

```{r, eval=FALSE}
DT::datatable(ico_status_calc %>% filter(year == 2017), caption = 'ICO status')

DT::datatable(ico_trend %>% filter(year == 2017) %>% select(-trend_lm), caption = 'ICO trend')
```

***

## Plot scores time series

To examine results of the new methods (including API-based data scraping and trend based on time series), we plot the estimated status and trend over time.

``` {r spp_plot_scores_over_time, fig.height = 4, fig.width = 6, fig.align = 'center', eval=FALSE}

library(ggplot2)
library(plotly)

status_ts_plot <- ggplot(ico_sum %>%
                           filter(!is.na(rgn_id)),
                         aes(x = year, y = ico_status, color = rgn_id, group = rgn_id)) +
  geom_line(size = 1.2, alpha = .4) +
  labs(x = 'year',
       y = 'ICO status',
       title = 'ICO status over time',
       color = 'Region')
ggplotly(status_ts_plot)

trend_ts_plot <- ggplot(ico_sum %>%
                           filter(!is.na(rgn_id) &!is.na(ico_trend)),
                         aes(x = year, y = ico_trend, color = rgn_id, group = rgn_id)) +
  geom_line(size = 1.2, alpha = .4) +
  labs(x = 'year',
       y = 'ICO trend',
       title = 'ICO trend over time',
       color = 'Region')
ggplotly(trend_ts_plot)
```

***

## Compare OHI 2018 vs OHI 2017 assessment results

Plot the estimated status and trends for 2016 from assessments v2017 and v2018.


``` {r spp_plot_scores_vs_2015, fig.height = 4, fig.width = 6, fig.align = 'center', eval=FALSE}

status_trend <- ico_sum %>% # status and trend in this year's assessment
  filter(year == max(year)-2) %>%  # max year of ico_sum is 2018 not 2017
  select(rgn_id, ico_status, ico_trend, n_spp) %>% 
  distinct() # because removing iucn_sid, cat, cat_score, and species name, results in duplicates
  
ico_compare <- read_csv(file.path(dir_previous, 'summary/ico_summary.csv')) %>% 
  filter(year == max(year)) %>% # max year of v2017 ico_summary is 2016; make sure to update for v2019
  select(rgn_id, rgn_name, ico_status, ico_trend, n_spp) %>% 
  distinct() %>% 
  rename(previous_status = ico_status, previous_trend = ico_trend, previous_nspp = n_spp) %>% 
  left_join(status_trend, by = 'rgn_id') %>% 
  mutate(d_nspp = as.factor(previous_nspp - n_spp))

compare_status_plot <- ggplot(ico_compare) + 
  geom_point(aes(previous_status, ico_status, color = d_nspp, label = rgn_id), alpha = 0.6) + 
    geom_abline(slope = 1, intercept = 0, color = 'red') +
  labs(x = sprintf('ICO status v%s', previous_scen),
       y = sprintf('ICO status %s', scenario),
       title = 'ICO Status Comparison')
ggplotly(compare_status_plot)

trend_compare_plot <- ggplot(ico_compare) +
  geom_point(aes(previous_trend, ico_trend, color = d_nspp, label = rgn_id), alpha = .6) +
  geom_abline(slope = 1, intercept = 0, color = 'red') +
  labs(x = sprintf('ICO trend v%s', previous_scen),
       y = sprintf('ICO trend %s', scenario),
       title = 'ICO Trend Comparison')
ggplotly(trend_compare_plot)

```

***
