---
title: 'OHI 2019: Natural Products'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---


```{r setup, warning=FALSE, message=FALSE}

knitr::opts_chunk$set(eval=FALSE)

library(here)
library(tidyverse)
source(here('workflow/R/common.R'))
```

## Load fish feed/oil catch and b/bmsy data
```{r}
forage_fish_catch <- read_csv(file.path(here(),"globalprep/fis/v2019/int/catch_fish_feed.csv")) %>%
  dplyr::select(-inWatson, -catch, -catch_human) %>%
  mutate(catch_fish_feed = ifelse(catch_fish_feed == 0, NA, catch_fish_feed))

fis_bbmsy <- read_csv(file.path(here(),"globalprep/fis/v2019/output/fis_bbmsy.csv")) %>%
  mutate(bbmsy = ifelse(bbmsy > 1 , 1, bbmsy))
```

## Align forage fish catch with b/bmsy data and calculate weighted mean
```{r}
weighted_mean_bbmsy_forage <- forage_fish_catch  %>%
  left_join(fis_bbmsy, by = c("stock_id", "rgn_id", "year")) %>%
  dplyr::filter(!is.na(bbmsy)) %>%
  group_by(year, rgn_id) %>%
  summarise(catch_weighted_mean = weighted.mean(bbmsy, catch_fish_feed)) %>%
  ungroup()

head(weighted_mean_bbmsy_forage, 10)
tail(weighted_mean_bbmsy_forage, 10)
```

## Sum forage fish catch by region and compare to NP fish oil production

```{r}

sum_forage <- forage_fish_catch %>%
  group_by(rgn_id) %>%
  summarise(total_forage_catch = sum(catch_fish_feed, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(total_forage_catch_log = log(total_forage_catch + 1))

fish_oil_tonnes <- read_csv(file.path("../output/np_harvest_tonnes.csv")) %>%
  filter(product == "fish_oil") %>%
  group_by(rgn_id) %>%
  summarise(total_fish_oil = sum(tonnes)) %>%
  ungroup() %>%
  mutate(total_fish_oil_log = log(total_fish_oil + 1))

combined <- sum_forage %>%
  left_join(fish_oil_tonnes, by = "rgn_id")

plot(combined$total_forage_catch_log, combined$total_fish_oil_log)
abline(0,1, col = "red")

```

## See how much catch is is caught in high seas per each country 

*This was taken from catch_data_prep.Rmd in the fis subgoal and modified to include high seas regions in the mean catch data*
```{r}
file <- file.path(dir_M,'git-annex/globalprep/fis/v2019/int/stock_catch_by_rgn_taxa.csv')

catch <- read_csv(file) %>%
  rename(common = CommonName, fao_id = fao_rgn, species=TaxonName)

summary(catch)


## filter out non ohi eez regions 
catch <- catch %>%
  filter(!is.na(rgn_id)) %>%
  filter(!is.na(fao_id)) ## keep all regions because want to see high seas regions 


## calculate total annual catch for each stock
catch <- catch %>%
  dplyr::select(year, rgn_id, fao_id, stock_id, TaxonKey, tons) %>%
  group_by(rgn_id, fao_id, TaxonKey, stock_id, year) %>%
  summarize(catch = sum(tons)) %>%
  ungroup()
```

For years with no reported catch, add zero values (after first reported catch)

```{r, eval=FALSE}

## these data have no zero catch values, so add years with no reported catch to data table:
catch_zeros <- catch %>%
  spread(year, catch) %>%
  data.frame() %>%
  gather("year", "catch", num_range("X", min(catch$year):max(catch$year))) %>%
  mutate(year = as.numeric(gsub("X", "", year))) %>%
  mutate(catch = ifelse(is.na(catch), 0, catch))

## this part eliminates the zero catch values prior to the first reported non-zero catch   
catch_zeros <- catch_zeros %>%
  group_by(fao_id, TaxonKey, stock_id, rgn_id) %>%
  arrange(year) %>%
  mutate(cum_catch = cumsum(catch)) %>%
  filter(cum_catch > 0) %>%
  dplyr::select(-cum_catch) %>%
  ungroup()

```

## Calculate Mean Catch
Calculate mean catch for ohi regions (using data from 1980 onward). These data are used to weight the RAM b/bmsy values 
```{r, eval=FALSE}

## correcting for forage fish used as feed/fish oil
## We have traditionally included all fisheries catch in the Food Provision goal. However, a substantial portion of catch is used in animal feed. Our plan is to remove a portion of catch of these species from the fisheries goal.

## read in list of species used for feed
forage_fish_taxa_list <- read_csv(file.path(here(), "globalprep/fis/v2019/raw/forage_fish_taxa_list.csv"))


taxon_key_info <- read_csv(file.path(dir_M,'git-annex/globalprep/fis/v2019/int/stock_catch_by_rgn_taxa.csv'))

## need to get TaxonKey's for each species to join with catch_zeros
forage_fish_taxa_list <- forage_fish_taxa_list %>%
  left_join(taxon_key_info, by = c("forage_fish" = "TaxonName")) %>%
  dplyr::select(forage_fish, inWatson, TaxonKey) %>%
  unique()

 
prop_human_cons <- (1 - (19/31))
## join this with catch_zeros by species, and multiply by 1 - (19/31) =  0.3870968... this is the proportion of catch used for humans 
catch_zero_minus_fish_feed <- forage_fish_taxa_list %>%
  left_join(catch_zeros, by = "TaxonKey") %>%
  mutate(catch_human = prop_human_cons*catch,
         catch_fish_feed = catch*(1-prop_human_cons))
write_csv(catch_zero_minus_fish_feed, "int/catch_fish_feed.csv")


#join catch_zeros with catch_zero_minus_fish_feed
catch_zeros <- catch_zeros %>%
  left_join(catch_zero_minus_fish_feed) %>%
  mutate(catch_human = case_when(
    is.na(catch_human) ~ catch,
    !is.na(catch_human) ~ catch_human
  )) %>%
  dplyr::select(-forage_fish, -inWatson)

mean_catch <- catch_zeros %>%
  filter(year >= 1980) %>%
  group_by(rgn_id, fao_id, TaxonKey, stock_id) %>%
  mutate(mean_catch = mean(catch, na.rm=TRUE),
         mean_catch_human = mean(catch_human, na.rm = TRUE)) %>% # mean catch for each stock (in a specific ohi-fao region)
  filter(mean_catch != 0,
         mean_catch_human != 0)  %>%      ## some stocks have no reported catch for time period
  ungroup()

```

Check out the data
```{r, eval=FALSE}

data.frame(dplyr::filter(catch, stock_id == "Zygochlamys_patagonica-87" & rgn_id==172))
data.frame(filter(mean_catch, stock_id == "Marine_fishes_not_identified-57" & rgn_id==1)) # includes finfishes (100139) and other marine fishes (100039)
data.frame(filter(mean_catch, stock_id == "Clupeiformes-57" & rgn_id==1)) # look at one which is in forage_fish
```

## Toolbox formatting and save

```{r, eval=FALSE}
options(scipen = 999) # to prevent taxonkey from turning into scientific notation

mean_catch_toolbox <- mean_catch %>%
  mutate(stock_id_taxonkey = paste(stock_id, TaxonKey, sep="_")) %>%
  dplyr::select(rgn_id, stock_id_taxonkey, year, mean_catch) %>%
  filter(year >= 2001) %>%  # filter to include only analysis years
  data.frame()

write.csv(mean_catch_toolbox, "mean_catch_highseas.csv", row.names=FALSE)

```

