---
title: "OHI 2019 - Artisanal Opportunities: Preparing need data"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: console
---


# Summary

This script generates the "need" layer for the artisanal opportunities goal.  The "access" layer was updated in v2019 with new Fisheries Management Index data, and is located here: , which is not updated due to a lack of a data source, is located here: globalprep/res_fmi/v2019/output/fmi_res.csv.


## Updates from previous assessment
One more year of data. Uninhabited and low population regions now given an NA value (vs. a gapfilled value). Also switched to using the here() package to avoid using setwd()

***
## Data Source 

**Downloaded**: 7/10/2019

**Description**:  
GDP adjusted per capita by PPP (ppppcgdp)
http://data.worldbank.org/indicator/NY.GDP.PCAP.PP.KD
Reported at country scale.

GDP per capita based on purchasing power parity (PPP). PPP GDP is gross domestic product converted to international dollars using purchasing power parity rates. An international dollar has the same purchasing power over GDP as the U.S. dollar has in the United States. GDP at purchaser's prices is the sum of gross value added by all resident producers in the economy plus any product taxes and minus any subsidies not included in the value of the products. It is calculated without making deductions for depreciation of fabricated assets or for depletion and degradation of natural resources. Data are in constant international dollars based on the 2011 ICP round.

Data is available directly to R through the WDI package.

**Time range**: 1990-2018


***
# Methods

## Setup

``` {r setup, message = FALSE, warning = FALSE, eval=F}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/', message = FALSE, warning = FALSE)

library(ohicore) # devtools::install_github('ohi-science/ohicore@dev')
library(dplyr)
library(tidyr)
library(WDI) # install.packages('WDI')
library(stringr)
library(readr)
library(here)
library(tidyverse)
library(plotly)


# directory paths and relevant files
source('https://raw.githubusercontent.com/OHI-Science/ohiprep_v2019/gh-pages/workflow/R/common.R')


```

## Download and save data
*Skip if you have already downloaded data.*
```{r load_data, eval=F}

# check website to see what years are available
yr_start = 1990
yr_end   = 2018


# get description of variables (NOTE: these descriptions appear out of date, they aren't in sync with the definitions of the World Bank):
indicators <-  data.frame(WDI_data[[1]])

head(indicators)
head(WDI_data)
str(WDI_data)
# indicators[grep("NY.GDP.PCAP.PP.CD", indicators$indicator), ]  # current dollars (influenced by inflation, not being used)
indicators[grep("NY.GDP.PCAP.PP.KD", indicators$indicator), ]  # constant dollars. grep helps identify rows to select based on a string. (used this data)


# download the data using the WDI package - This is the data we are going to work with. Create a variable for the data frame

gdppcppp_raw <-  WDI(country = "all",
               indicator = "NY.GDP.PCAP.PP.KD", 
               start = yr_start, end=yr_end)
summary(gdppcppp_raw)

# check if 'raw', 'intermediate', and 'final' folders exist in the current assessment folder, if not, then create the folder in the working directory
if (!file.exists(here("globalprep/ao/v2019/raw"))){
  dir.create(here("globalprep/ao/v2019/raw"))
} 

if (!file.exists(here("globalprep/ao/v2019/intermediate"))){
  dir.create(here("globalprep/ao/v2019/intermediate"))
}

if (!file.exists(here("globalprep/ao/v2019/output"))){
  dir.create(here("globalprep/ao/v2019/output"))
}

date <- Sys.Date()

#Save the file into the raw folder
write_csv(gdppcppp_raw, here(sprintf('globalprep/ao/v2019/raw/raw_gdppcppp_%s.csv', date))) # Save file with date, as WDI data changes over even short periods of time. For instance, the Mauritania GDP data changed by an order of magnitude over the course of a week. We want to preserve the date it was downloaded so that data is not being overwritten everytime we run the script. 


```


```{r raw check, eval=FALSE}

last_saved_date <- "2019-07-15" # update when new data are downloaded.

new <- read_csv(here(sprintf('globalprep/ao/v2019/raw/raw_gdppcppp_%s.csv', last_saved_date)))

old <- read_csv(here('globalprep/ao/v2018/raw/raw_gdppcppp.csv')) %>%
  select(country, old_value = NY.GDP.PCAP.PP.KD, year)

compare <- left_join(new, old, by = c("country", "year"))  %>%
  filter(year==2017)
 
ggplot(compare, aes(x=NY.GDP.PCAP.PP.KD, y=old_value)) + 
  geom_point() + 
  geom_abline(slope = 1, intercept = 0, color="red")
  
```
## Gapfilling 1: Linear Regression within a country's data

When a country has some values (but not a complete series), a within-country regression model is used to predict the missing values.

```{r, eval=F}

# Reorganize to create cells for countries that have missing values for some years

last_saved_date <- "2019-07-15" # update when new data are downloaded.

gdppcppp_clean <- read_csv(here(sprintf('globalprep/ao/v2019/raw/raw_gdppcppp_%s.csv', last_saved_date))) %>% # change date here also to match filename above 
  dplyr::select(country, value=NY.GDP.PCAP.PP.KD, year) %>%
  dplyr::filter(year >= 2005) %>%
  tidyr::spread(year, value) %>%
    # spread to fill in potentially missing values with NA
  data.frame() %>% # this will add an X in front of the column names, allowing us to gather the values
  tidyr::gather(year, value, starts_with("X")) %>%
  dplyr::mutate(year = gsub("X", "", year)) %>% #substitute X for "" (nothing) in the column year
  dplyr::mutate(year = as.numeric(year)) #convert the year column into a numeric format

head(gdppcppp_clean)
summary(gdppcppp_clean) # 357 NAs

# For the first case, if there is only one value use this value for all years
# This is not ideal, but likely better than other forms of gapfilling
gdppcppp_val <- gdppcppp_clean %>%
  dplyr::group_by(country) %>%
  dplyr::mutate(value_num = sum(!is.na(value))) %>% # counts the numbers of non-missing values for each country (logical TRUEs regarded as one)
  dplyr::filter(value_num > 0) %>%    # filter out the countries with no data between 2005 and 2018 (21 regions in v2019)
  dplyr::mutate(value_num_gf = ifelse(value_num==1, mean(value, na.rm=TRUE), NA)) %>%  # mean() function is used on regions with one year of data, applies that single value to all NAs for that region (this is only gapfilling one region in v2019)
  dplyr::ungroup() %>%
  dplyr::mutate(value = ifelse(is.na(value), value_num_gf, value)) %>% # if no value is missing, leave it, otherwise gapfill; still have NAs where 13 > value_num > 1
  dplyr::select(country, year, value, value_num)  # select just these columns; to eliminate extraneous value_num_gf column


head(gdppcppp_val)
summary(gdppcppp_val) 


# Predict values using a linear regression with 'year' as an independent variable 
# Create new column with these predicted values
gdppcppp_gf <- gdppcppp_val %>%
  dplyr::group_by(country) %>%
  dplyr::do({ 
    mod <- lm(value ~ year, data =.)
    value_pred <- predict(mod, newdata =.[c('year')]) # value_pred = country-grouped mod$fitted.values?
    data.frame(., value_pred) # do loop applies the model fitting and prediction to each country group
  }) %>% 
  dplyr::ungroup()

summary(gdppcppp_gf)
head(gdppcppp_gf)


# Fill in the remaining NA values using the predicted values
gdppcppp_gf2 <- gdppcppp_gf %>%
  dplyr::ungroup() %>% # why would we need to ungroup here?
  dplyr::mutate(gapfilled = ifelse(is.na(value), 1, 0)) %>% # Create column 'gapfilled', if value is currently NA, it will be gapfilled, indicated by a 1 in the gapfill column
  dplyr::mutate(gapfilled = ifelse(value_num == 1, 1, gapfilled)) %>% # if value_num is 1 it was gapfilled previously and gets a 1 in the gapfill column
  dplyr::mutate(value = ifelse(is.na(value), value_pred, value)) %>% # if NA in value column, input the value in value_pred column
  dplyr::mutate(method = ifelse(gapfilled==1, paste("lm based on N years data:", value_num, sep=" "), NA)) %>% # Create column 'method' that indicates method of gapfilling; this puts message "lm based..." even in some rows gapfilled with one year of data
  dplyr::mutate(method = ifelse(value_num == 1, "gapfilled using one year of data", method)) # this overwrites/corrects method "lm based..." for rows actually gapfilled with "one-year of data"" method
  
summary(gdppcppp_gf2) # no more NAs because everything has been gap-filled.
```

## Calculate rescaled values
This is performed by taking the natural log of each value and then dividing by the 95th quantile of values across all years (from 2005 to current data year). 

```{r, eval=F}

# Values at the 95th Quantile or greater are given a rescaled score of '1' (the highest value)
gdppcppp_rescale <- gdppcppp_gf2 %>%
  dplyr::mutate(quantile_95 = quantile(value, probs=0.95)) %>% # gives a single value - the 95th quant (v2019=53582.14; without current year of data quant95=53208.67 - these should be different)
  dplyr::mutate(value_stand = value/quantile_95) %>% # where does value scale relative to 95th quantile
  dplyr::mutate(value_stand = ifelse(value_stand > 1, 1, value_stand)) %>% 
  dplyr::select(country, year, value, score=value_stand, gapfilled, method) # rename value_stand 'score'


summary(gdppcppp_rescale)
head(gdppcppp_rescale)


# Check to see if scores make sense - anything above reference point (quant_95) should = 1, everything below it should have a value between 0 and 1 

rescale_vis <- ggplot(gdppcppp_rescale, aes(x =value , y = score, label = country)) +
   geom_point()
ggplotly(rescale_vis)


```


## Convert country names to ohi regions

```{r, eval=F}

# Function to add OHI region ID based on country name
d_stand_rgn <- name_2_rgn(df_in = gdppcppp_rescale, 
                       fld_name='country', 
                       flds_unique=c('year'))

## v2019: Lots of error messages about missing regions from lookup table; lots of them are broad areas (e.g. "Arab World" and "fragile regions"), some are landlocked areas like N. Macedonia and Eswatini. Check to make sure there aren't any regions that need to be reported at different scales 
## China, Hong Kong and Macao are all reported separately, combine into one

data.frame(filter(d_stand_rgn, rgn_id == 209))
##This should match the duplicate regions


# Combine the duplicate regions (we report these at lower resolution)
# In this case, we take the average score weighted by population
population_weights <- read.csv('https://raw.githubusercontent.com/OHI-Science/ohiprep_v2017/master/src/LookupTables/Pop_weight_ChinaSAR_USVIslPRico.csv')


# Weight the `score`, `value`, and `gapfilled` column by population
d_stand_rgn <- d_stand_rgn %>%
  dplyr::left_join(population_weights, by="country") %>% # does it make sense to backfill population data with static data?
  dplyr::mutate(population = ifelse(is.na(population), 1, population)) %>% # If no value available, input 1 (these values will not change)
  dplyr::group_by(rgn_id, year, method, gapfilled) %>% 
  dplyr::summarize(score = weighted.mean(score, population), # weight the single score value by pop.
            value = weighted.mean(value, population)) %>%
  ungroup() 

# check again:
data.frame(filter(d_stand_rgn, rgn_id == 209))


# Removed `Azerbaijan` (255) because the adjacent body of water is a sea not the ocean - is this not done in names2region? 
d_stand_rgn <- d_stand_rgn %>%
  filter(rgn_id <= 250)

summary(d_stand_rgn) # no NAs


# save the cleaned gdppcppp for other goals
gdppcppp_data <- d_stand_rgn %>%
  select(rgn_id, year, value)

write_csv(gdppcppp_data, here("globalprep/ao/v2019/intermediate/gdppcppp_ohi.csv"))

```

## Gapfilling: part 2
In this case, we gapfill regions with no data using means based on UN-designated geopolitical levels.

```{r, eval=F}
UNgeorgn() # how is this different from Mel's georegion function in ohicore? 
head(UNgeorgn)
summary(UNgeorgn)
# Create dataframe pairing each UN geopolitical region id with a year from 2005 to current

# Assign georegion labels to each region for each level (r0, r1, r2)
d_stand_gf <- data.frame(year=min(d_stand_rgn$year):max(d_stand_rgn$year)) %>% 
  merge(UNgeorgn, by=NULL) 
summary(d_stand_gf)
head(d_stand_gf)

# Combine the two data frames by region id and year
# Calculate means across increasing geopolitical levels (e.g. r2, r1), using the highest resolution possible
d_stand_gf <- d_stand_gf %>%  
  left_join(d_stand_rgn, by = c("rgn_id", "year")) %>%
  group_by(r2_label, year) %>%
  mutate(r2_value = mean(score, na.rm=TRUE)) %>%
  ungroup() %>%
  group_by(r1_label, year) %>%
  mutate(r1_value = mean(score, na.rm=TRUE)) %>%
  ungroup() %>%
  group_by(r0_label, year) %>%
  mutate(r0_value = mean(score, na.rm=TRUE)) %>%
  ungroup()
summary(d_stand_gf) 


low_pop()
summary(low_pop)

low_pop <- low_pop %>%
  filter(est_population < 3000 | is.na(est_population)) #filter out regions that have populations > 3000 and keep NA values 
summary(low_pop)

low_pop_vector <- c(low_pop$rgn_id) #make a vector of low population areas 

# For `score` cells that still have NA values (still several hundred):
# Check to see if r2 has a value, if so use that to gapfill `score`, otherwise use r1, otherwise use r0
d_stand_gf <- d_stand_gf %>%
  mutate(gapfilled = ifelse(is.na(score) & !is.na(r2_value), "1", gapfilled)) %>%
  mutate(method = ifelse(is.na(score) & !is.na(r2_value), "UN_geopolitical region avg, r2", method)) %>%
  mutate(score = ifelse(is.na(score), r2_value, score)) %>%
  mutate(gapfilled = ifelse(is.na(score) & !is.na(r1_value), "1", gapfilled)) %>%
  mutate(method = ifelse(is.na(score) & !is.na(r1_value), "UN_geopolitical region avg, r1", method)) %>%
  mutate(score = ifelse(is.na(score), r1_value, score)) %>%
  mutate(gapfilled = ifelse(is.na(score) & !is.na(r0_value), "1", gapfilled)) %>%
  mutate(method = ifelse(is.na(score) & !is.na(r0_value), "UN_geopolitical region avg, r0", method)) %>%
  mutate(score = ifelse(is.na(score), r0_value, score))
  

d_stand_gf$score[d_stand_gf$rgn_id %in% low_pop_vector] <- NA
#should now have NA values in score column for low popuation areas

summary(d_stand_gf)

```


## Save the data

```{r, eval=F}

# Save dataframe with adjusted, gapfilled, and rescaled score information
final <- d_stand_gf %>%
  select(rgn_id, year, value=score)

write_csv(final, here("globalprep/ao/v2019/output/wb_gdppcppp_rescaled.csv"))


# Save dataframe with gapfilled method and status information
final_gf <- d_stand_gf %>%
  select(rgn_id, year, gapfilled, method)

write_csv(final_gf, here("globalprep/ao/v2019/output/wb_gdppcppp_rescaled_gf.csv"))


```


## Compare data to previous year
Use most recent data year shared by current and previous assessment.
```{r, eval=F}
old_gdppcppp <- read.csv(here("globalprep/ao/v2018/output/wb_gdppcppp_rescaled.csv")) %>% 
  dplyr::rename(old_value=value) %>% 
  dplyr::filter(year == 2017)
summary(old_gdppcppp) # No NAs  

compare <- read.csv(here("globalprep/ao/v2019/output/wb_gdppcppp_rescaled.csv")) %>%
  dplyr::filter(year == 2017) %>% 
  dplyr::left_join(old_gdppcppp, by="rgn_id") %>% 
  select(rgn_id, value, old_value)
summary(compare) # still 20 NAs - this is because of converting unpopulated/low population regions to NAs


ggplotly(ggplot(compare, aes(x = value, y = old_value, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red")) # if anything was off of abline then something went wrong; however in v2019 data were positively skewed and we attributed this to source data changes. 



####### Investigating outlier (v2019)
old_220 <- old_gdppcppp %>% 
  filter(rgn_id == 220) 

new_220 <- d_stand_gf %>%
  filter(rgn_id == 220) %>%
  filter(year < 2018) %>% 
  rename(new_value = score) %>% 
  select(rgn_id, year, gapfilled, method, new_value)

old_gf_220 <- read_csv(here("globalprep/ao/v2018/output/wb_gdppcppp_rescaled_gf.csv")) %>% 
  filter(rgn_id == 220) %>%
  left_join(old_220, by = c("rgn_id", "year"))


#### Result: looks like v2018 gapfilled all of the data for rgn 220, and v2019 only gapfilled for a few years of data, thus causing the score to change.
  
```

