---
title: "OHI 2017 - Artisanal Opportunities: Preparing need data"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../src/templates/ohi_hdr.html'
pdf_document:
  toc: true
---


# Summary

This script generates the "need" layer for the artisanal opportunities goal.  The "access" layer, which is not updated due to a lack of a data source, is located here: globalprep/res_mora_ao/v2013/data/r_mora_s4_2013a.csv.


## Updates from previous assessment
None

***
#Data Source 

**Downloaded**: 3/9/2018

**Description**:  
GDP adjusted per capita by PPP (ppppcgdp)
http://data.worldbank.org/indicator/NY.GDP.PCAP.PP.KD
Reported at country scale.

GDP per capita based on purchasing power parity (PPP). PPP GDP is gross domestic product converted to international dollars using purchasing power parity rates. An international dollar has the same purchasing power over GDP as the U.S. dollar has in the United States. GDP at purchaser's prices is the sum of gross value added by all resident producers in the economy plus any product taxes and minus any subsidies not included in the value of the products. It is calculated without making deductions for depreciation of fabricated assets or for depletion and degradation of natural resources. Data are in constant international dollars based on the 2011 ICP round.

**Time range**: 1990-2016


***
# Methods

## Setup

``` {r setup, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/', message = FALSE, warning = FALSE)

# setting up provenance
# devtools::install_github('oharac/provRmd')
# library(provRmd)
# prov_setup()

library(ohicore) # devtools::install_github('ohi-science/ohicore@dev')
library(dplyr)
library(tidyr)
library(WDI) # install.packages('WDI')
library(stringr)
library(readr)

# comment out when knitting:
# setwd("globalprep/ao/v2018")

# directory paths
source('../../../src/R/common.R')

```

## Download and save data

```{r load_data,eval=F}

# check website to see what years are available
yr_start = 1990
yr_end   = 2016


# get description of variables (NOTE: these descriptions appear out of date, they aren't in sync with the definitions of the World Bank):
indicators <-  data.frame(WDI_data[[1]])

head(indicators)


# indicators[grep("NY.GDP.PCAP.PP.CD", indicators$indicator), ]  # current dollars (influenced by inflation, not being used)
indicators[grep("NY.GDP.PCAP.PP.KD", indicators$indicator), ]  # constant dollars. grep helps identify rows to select based on a string. (used this data)


# download the data
gdppcppp = WDI(country = "all",
               indicator = "NY.GDP.PCAP.PP.KD", 
               start = yr_start, end=yr_end)


# check if 'raw', 'intermediate', and 'final' exists, if not, then create the folder in the working directory
if (!file.exists("raw")){
  dir.create("raw")
} 

if (!file.exists("intermediate")){
  dir.create("intermediate")
} 

if (!file.exists("output")){
  dir.create("output")
} 

write.csv(gdppcppp, 'raw/raw_gdppcppp.csv', row.names=FALSE) # once you've made sure the 'raw' folder exists, you can save your file

```


## Within country gapfilling

When a country has some values (but not a complete series), a within country regression model is used to predict the missing values.

```{r}

# Reorganize to create cells for countries that have missing values for some years
d <- read.csv('raw/raw_gdppcppp.csv') %>%
  select(country, value=NY.GDP.PCAP.PP.KD, year) %>%
  filter(year >= 2005) %>%
  spread(year, value) %>%  # fill in potentially missing values with NA
  data.frame() %>% # this will add an X in front of the column names, allowing us to gather the values
  gather(year, value, starts_with("X")) %>%
  mutate(year = gsub("X", "", year)) %>% #substitute X for "" (nothing) in the column year
  mutate(year = as.numeric(year)) #convert the year column into a numeric format


# For the first case, if there is only one value use this value for all years
# This is not ideal, but likely better than other forms of gapfilling
d <- d %>%
  group_by(country) %>%
  mutate(value_num = sum(!is.na(value))) %>% # counts the numbers of non-missing values for each country (logical TRUEs regarded as one)
  filter(value_num > 0) %>%    # filter out the countries with no data between 2005 and 2016
  mutate(value_num_gf = ifelse(value_num==1, mean(value, na.rm=TRUE), NA)) %>%  # mean() function is just used to get the single value and remove all NAs in that group
  ungroup() %>%
  mutate(value = ifelse(is.na(value), value_num_gf, value)) %>% # if no value is missing, leave it, otherwise gapfill
  select(country, year, value, value_num)


# Predict values using a linear regression with 'year' as an independent variable 
# Create new column with these predicted values
d_gf <- d %>%
  group_by(country) %>%
  do({
    mod <- lm(value ~ year, data=.) # why is country not included in the lm?
    value_pred <- predict(mod, newdata=.['year'])
    data.frame(., value_pred)
  })


# Fill out the remaining NA values using the predicted values
# These are those countries where number of non-NA values > 1
d_gf <- d_gf %>%
  ungroup() %>%
  mutate(gapfilled = ifelse(is.na(value), 1, 0)) %>% # Create column 'gapfilled' that indicates whether gapfilling occurred
  mutate(gapfilled = ifelse(value_num == 1, 1, gapfilled)) %>%
  mutate(value = ifelse(is.na(value), value_pred, value)) %>% # if missing value still, input the value in value_pred column
  mutate(method = ifelse(gapfilled==1, paste("lm based on N years data:", value_num, sep=" "), NA)) %>% # Create column 'method' that indicates method of gapfilling
  mutate(method = ifelse(value_num == 1, "gapfilled using one year of data", method))
  
```

## Calculate rescaled values
This is performed by taking the natural log of each value and then dividing by the 95th quantile of values across all years (from 2005 to 2016). 

```{r}

# Values at the 95th Quantile or greater are given a rescaled score of '1' (the highest value)
d_rescale <- d_gf %>%
  mutate(quantile_95 = quantile(value, probs=0.95)) %>% # gives a single value - the 95th quant
  mutate(value_stand = value/quantile_95) %>% # where does value scale relative to 95th quantile
  mutate(value_stand = ifelse(value_stand > 1, 1, value_stand))

d_rescale <- d_rescale %>%
  select(country, year, value, score=value_stand, gapfilled, method) # rename value_stand 'score'


# check the data with some plots
# plotData <- d_rescale %>%
#   mutate(need = 1-score) %>%
#   dplyr::select(country, year, need)

# library(googleVis)
#
# Motion = gvisMotionChart(plotData,
# idvar="country", timevar="year")
# 
# plot(Motion)
# 
# print(Motion, file = file.path('ao_need_95q.html'))

```


# Convert country names to ohi regions

```{r}

# Function to add OHI region ID based on country name
d_stand_rgn <- name_2_rgn(df_in = d_rescale, 
                       fld_name='country', 
                       flds_unique=c('year')) # cannot load look up tables


# Combine the duplicate regions (we report these at lower resolution)
# In this case, we take the average score weighted by population
population_weights <- read.csv('../../../../ohiprep_v2018/src/LookupTables/Pop_weight_ChinaSAR_USVIslPRico.csv')


# Weight the `score`, `value`, and `gapfilled` column by population
d_stand_rgn <- d_stand_rgn %>%
  left_join(population_weights, by="country") %>%
  mutate(population = ifelse(is.na(population), 1, population)) %>% # If no value available, input 1
  group_by(rgn_id, year) %>%
  summarize(score = weighted.mean(score, population), # weight the single score value by pop.
            value = weighted.mean(value, population),
            gapfilled = weighted.mean(gapfilled, population),
            method = method[1]) %>% 
  ungroup() 


# Removed `Azerbaijan` (255) and `Aruba` (250)
d_stand_rgn <- d_stand_rgn %>%
  filter(rgn_id <= 250)

summary(d_stand_rgn)


# save the cleaned gdppcppp for other goals
gdppcppp_data <- d_stand_rgn %>%
  select(rgn_id, year, value)

write_csv(gdppcppp_data, "intermediate/gdppcppp_ohi.csv")

```

# Gapfilling: part 2
In this case, we gapfill regions with no data using UN geopolitical means.

```{r}

# Create dataframe pairing each UN geopolitical region id with a year from 2005 to current
d_stand_gf <- data.frame(year=min(d_stand_rgn$year):max(d_stand_rgn$year)) %>% 
  merge(georegions, by=NULL) 


# Combine the two data frames by region id and year
# Calculate means across different geopolitical levels (e.g. r2, r1)
d_stand_gf <- d_stand_gf %>%  
  left_join(d_stand_rgn, by = c("rgn_id", "year")) %>%
  group_by(r2, year) %>%
  mutate(r2_value = mean(score, na.rm=TRUE)) %>%
  ungroup() %>%
  group_by(r1, year) %>%
  mutate(r1_value = mean(score, na.rm=TRUE)) %>%
  ungroup() %>%
  group_by(r0, year) %>%
  mutate(r0_value = mean(score, na.rm=TRUE)) %>%
  ungroup()


# For `score` cells that still have NA values (still several hundred):
# Check to see if r2 has a value, if so use that to gapfill `score`, otherwise use r1, otherwise use r0
d_stand_gf <- d_stand_gf %>%
  mutate(gapfilled = ifelse(is.na(score) & !is.na(r2_value), "1", gapfilled)) %>%
  mutate(method = ifelse(is.na(score) & !is.na(r2_value), "UN_geopolitical region avg, r2", method)) %>%
  mutate(score = ifelse(is.na(score), r2_value, score)) %>%
  mutate(gapfilled = ifelse(is.na(score) & !is.na(r1_value), "1", gapfilled)) %>%
  mutate(method = ifelse(is.na(score) & !is.na(r1_value), "UN_geopolitical region avg, r1", method)) %>%
  mutate(score = ifelse(is.na(score), r1_value, score)) %>%
  mutate(gapfilled = ifelse(is.na(score) & !is.na(r0_value), "1", gapfilled)) %>%
  mutate(method = ifelse(is.na(score) & !is.na(r0_value), "UN_geopolitical region avg, r0", method)) %>%
  mutate(score = ifelse(is.na(score), r0_value, score))

# should no longer have NAs in the score column!

```


# Save the data

```{r}

# Save dataframe with adjusted, gapfilled, and rescaled score information
final <- d_stand_gf %>%
  select(rgn_id, year, value=score)

write.csv(final, "output/wb_gdppcppp_rescaled.csv", row.names=FALSE)


# Save dataframe with gapfilled method and status information
final_gf <- d_stand_gf %>%
  select(rgn_id, year, gapfilled, method)

write.csv(final_gf, "output/wb_gdppcppp_rescaled_gf.csv", row.names=FALSE)


```
