---
title: "OHI 2019: Aquaculture Country Production Estimate"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../../workflow/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

[REFERENCE RMD FILE: https://raw.githubusercontent.com/OHI-Science/ohiprep_v2019/gh-pages/globalprep/mar/v2019/reference_point/CountryProductionEstimate.Rmd]

# Summary
This analysis produces potential tonnes of aquaculture from growth potential estimates of finfish and bivalves. These aquaculture numbers will be used for the reference point for the mariculture subgoal. 

***

# Data Source 

## Growth Potential (Phi) data 

**Reference**:      
https://knb.ecoinformatics.org/view/doi:10.5063/F1CF9N69
Rebecca Gentry, Halley Froehlich, Dietmar Grimm, Peter Kareiva, Michael Parke, et al. SNAPP - Mapping the Global Potential for Marine Aquaculture. Knowledge Network for Biocomplexity. doi:10.5063/F1CF9N69. 

**Downloaded**: 7/3/2019

**Description**: Growth Potential estimate raster for global cells 

***
  
# Methods
```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(eval=FALSE)
#######regression inputs from VBGF_Fish.r

# This is a heavily modified script from KNB:
# https://knb.ecoinformatics.org/view/doi:10.5063/F1CF9N69

# From this paper:
# https://www.nature.org/content/dam/tnc/nature/en/documents/Mapping_the_global_potential_for_marine_aquaculture.pdf


### libraries useful for data wrangling
library(dplyr)
library(tidyr)
library(tidyverse)

## libraries useful for spatial data
library(raster)       
library(rgdal)        
library(sf)         
library(fasterize)

## data visualization
library(RColorBrewer)
library(ggplot2)
library(rasterVis)    
library(maps)

## path management
library(here)


## some OHI files
source('http://ohi-science.org/ohiprep_v2019/workflow/R/common.R')
dir_git_ref <- file.path(here(), "globalprep/mar/v2019/archive_reference_point")
```

## Establish variables and coefficients
```{r}
## Variables needed to get from PHI to tonnes of production
## MRF: assumes 1 farm is 1km2  ##
# (seems easier to just units of km2, rather than farm)

F_estCoef <- c(7.6792, (-5.8198)) #from regression estimated in VBGF_Fish_Final.r
B_estCoef <- c(2.9959,(-1.6659)) #from regression estimate in VBGF_Bivalves.r
density <- 20 #juveniles per m3
cagesize <- 9000 #m3
cagesperfarm <- 24 #located atleast 1 km apart.. cagesperkm2
bivperfarm <- 130000000 #(100 longlines/km2) * 4 km * (100 bivalves seeded/0.0003 km) = 133333333 bivalve/km2
weight35cm <- 554.8  ## in grams see VBGF_Fish_Final. Paper reports 548 grams 
```

## Finfish Production
```{r}
## Global tiff file of PHI (Growth Potential) estimates
FishPhiALLConstraints <- raster(file.path(dir_M, "git-annex/globalprep/mar/v2019/Spatial_Data/NewLayersWOHypoxia/FishPhiALLConstraints95LT2.tiff"))
plot(FishPhiALLConstraints)
FishPhiVector=getValues(FishPhiALLConstraints)


## Convert Phi raster to number of years it takes to grow a 35 cm fish

LogFishYears <- calc(FishPhiALLConstraints, fun=function(x){F_estCoef[1]+F_estCoef[2]*log(x)})
LogFishYears
plot(LogFishYears)

FishYears <- calc(LogFishYears, fun=function(x){exp(x)})

FishYears
plot(FishYears)
writeRaster(FishYears,file.path(dir_M, "git-annex/globalprep/mar/v2019/Spatial_Data/NewLayersWOHypoxia/FishYearsbyCell.tif"), overwrite=TRUE)

FishYears <- raster(file.path(dir_M, "git-annex/globalprep/mar/v2019/Spatial_Data/NewLayersWOHypoxia/FishYearsbyCell.tif"))
FishYearsVector=getValues(FishYears)



#OHI 2018 regions (original analysis used older regions file)

## call spatial file from sourced file
regions_shape()
OHIcountries <- regions %>%
  filter(type_w_ant == "eez")
OHIcountries <- st_transform(OHIcountries, crs(FishPhiALLConstraints))
OHIcountries_raster <- fasterize(OHIcountries, FishPhiALLConstraints, field="rgn_id")  
CountryVector=getValues(OHIcountries_raster)




### area of each cell (each cell is different given lat/long coordinate reference system)
areaPerCell <- area(FishPhiALLConstraints, weights=FALSE, na.rm=TRUE)
areaPerCellVector <- getValues(areaPerCell)

### Make a dataframe with raster values that includes cells: Country, area, Phi, and Years to Harvest
productionDF <- data.frame(CellID = 1:933120000,
                           Country = CountryVector,
                           AreaKM2 = areaPerCellVector,
                           PhiPrime = FishPhiVector, 
                           YearsToHarvest = FishYearsVector)

head(productionDF)

summary(FishYearsVector)
summary(areaPerCellVector) ##they seem to match



## calculate production for each cell
productionDFFishCells <- productionDF %>%
  filter(!is.na(YearsToHarvest)) %>%
  mutate(F_yieldperfarmMT = (weight35cm * density * cagesize * cagesperfarm)/1000000) %>%  # MRF: units yieldperkm2? 554.8 grams * 20 juv/m3 * 9000 m3 * 24 cages/km2 = grams/km2
  mutate(F_yieldpercellperyear = (F_yieldperfarmMT/YearsToHarvest) * AreaKM2) %>%
  arrange(YearsToHarvest) %>%
  mutate(YieldCumSum = cumsum(F_yieldpercellperyear)) %>%
  mutate(AreaCumSum = cumsum(AreaKM2))

write.csv(productionDFFishCells, file.path(dir_M, "git-annex/globalprep/mar/v2019/int/fish/productionDFFishCells.csv")) #save to mazu because so large. This is functionally a raster file. 

productionDFFishCells <- read.csv(file.path(dir_M, "git-annex/globalprep/mar/v2019/int/fish/productionDFFishCells.csv"))

head(productionDFFishCells)
summary(productionDFFishCells)
str(productionDFFishCells)
##cumsum area is 11,402,629 km2 -  # matches the paper 
##cumprod is 15,950,000,000MT -   # matches the paper 

### how many of these cells are not in a country?
sum(is.na(productionDFFishCells$Country))
dim(productionDFFishCells)

## MRF: with new OHI regions: 544,569 are not in a country..probably a lot are in conflicted areas


## Calculate production if 1% of top production area is used:
productionByCountryFishDF <- productionDFFishCells %>%
  filter(!is.na(Country)) %>%
  dplyr::select(-YieldCumSum, -AreaCumSum, -X) %>%
  arrange(YearsToHarvest) %>%
  mutate(ID = Country) %>%
  dplyr::arrange(ID) %>%
  group_by(ID) %>%
  mutate(CountryYieldCumSum = cumsum(F_yieldpercellperyear)) %>%
  mutate(CountryAreaCumSum = cumsum(AreaKM2)) %>%
  mutate(MaxProdPerCountry = max(CountryYieldCumSum)) %>% 
  mutate(OnePercentDevPerCountry = .01*max(CountryAreaCumSum)) #calculating 1 percent of area

write.csv(productionByCountryFishDF,file.path(dir_M, "git-annex/globalprep/mar/v2019/int/fish/FishProdByCountryByCell.csv"), row.names = FALSE) #save to mazu because so large. This is functionally a raster file. 

productionByCountryFishDF <- read.csv(file.path(dir_M, "git-annex/globalprep/mar/v2019/int/fish/FishProdByCountryByCell.csv"))

## MRF: For each area identify amount of area that corresponds to 1% of production area, 
## MRF: assume maximum production within the country for the 1% of area
CountryProdSummary <- productionByCountryFishDF %>%
  dplyr::arrange(YearsToHarvest) %>%
  dplyr::arrange(ID) %>%
  group_by(ID) %>%
  filter(CountryAreaCumSum <= OnePercentDevPerCountry) %>%
  mutate(ProdPerCountryOnePercent = max(CountryYieldCumSum)) %>%
  slice(1)


write.csv(CountryProdSummary, file.path(dir_git_ref, "/production_int/FishProdByCountrySummary.csv"), row.names = FALSE) #save to github

CountryProdSummary <- read.csv(file.path(dir_git_ref, "production_int/FishProdByCountrySummary.csv"))

# MRF: get fasted YearsToHarvest for each country
CountryProdSummaryNop <- productionByCountryFishDF %>%
  dplyr::arrange(YearsToHarvest) %>%
  dplyr::arrange(ID) %>%
  group_by(ID) %>%
  slice(1)



# Add country names

region_data()
CountryLabel <- rgns_eez %>%
  dplyr::select(ID = rgn_id, rgn_name)


## Final data
## I think the relevant value we want for the reference point is in this table:  ProdPerCountryOnePercent. Indonesia number for fish matches up with paper. > 24 million tonnes of fish if 1% of aquaculture potential developed. 
CountryProdSummaryFAO <- CountryProdSummary %>%
  ungroup %>%
  dplyr::select(ID:ProdPerCountryOnePercent) %>%
  full_join(CountryLabel, by= "ID")

write.csv(CountryProdSummaryFAO, file.path(dir_git_ref, "/production_int/FishProdByCountrySummaryFAO.csv"), row.names = FALSE) #save to github

sum(CountryProdSummaryFAO$MaxProdPerCountry, na.rm = TRUE)
#15451023277 number matches paper. > 15 billion tonnes 
```


## Bivalve Production
```{r}
##Now for Bivalves 
BivalvePhiALLConstraints=raster(file.path(dir_M, "git-annex/globalprep/mar/v2019/Spatial_Data/NewLayersWOHypoxia/BivalvePhiALLConstraints95LT1.tif"))

plot(BivalvePhiALLConstraints)
BivalvePhiVector <- getValues(BivalvePhiALLConstraints)


#OHI 2018 regions (original analysis used older regions file)

## call spatial file from sourced file
regions_shape()

OHIcountries <- regions %>%
  filter(type_w_ant == "eez")

OHIcountries <- st_transform(OHIcountries, crs(FishPhiALLConstraints))

OHIcountries_raster <- fasterize(OHIcountries, FishPhiALLConstraints, field="rgn_id")  

CountryVector <- getValues(OHIcountries_raster)

#make the value of each cell the years it takes to grow a 4 cm bivlave
LogBivalveYears <- calc(BivalvePhiALLConstraints, fun = function(x){B_estCoef[1] + B_estCoef[2]*(x)})
LogBivalveYears
plot(LogBivalveYears)

BivalveYears=calc(LogBivalveYears,fun=function(x){exp(x)})

BivalveYears
plot(BivalveYears)

writeRaster(BivalveYears,file.path(dir_M, "git-annex/globalprep/mar/v2019/Spatial_Data/NewLayersWOHypoxia/BivalveYearsByCell.tif"), overwrite=TRUE)

BivalveYears <- raster(file.path(dir_M, "git-annex/globalprep/mar/v2019/Spatial_Data/NewLayersWOHypoxia/BivalveYearsByCell.tif"))

BivalveYearsVector <- getValues(BivalveYears)


###now load in area values for each cell
#areaPerCell=raster("Spatial_Data/MiddleFiles/AreaBivalveLT1.grd")
areaPerCell <- area(BivalvePhiALLConstraints, weights=FALSE, na.rm=TRUE)
areaPerCell

areaPerCellVector=getValues(areaPerCell)


productionDFBiv <- data.frame(CellID=1:933120000,
                           Country=CountryVector,
                           AreaKM2=areaPerCellVector,
                           PhiPrime=BivalvePhiVector, 
                           YearsToHarvest=BivalveYearsVector)

head(productionDFBiv)

productionDFBivCells <- productionDFBiv %>%
  filter(!is.na(YearsToHarvest)) %>%
  mutate(B_yieldperfarmInd = bivperfarm) %>%
  mutate(B_yieldpercellperyear = (B_yieldperfarmInd / YearsToHarvest) * AreaKM2) %>%
  arrange(YearsToHarvest) %>%
  mutate(YieldCumSum = cumsum(B_yieldpercellperyear)) %>%
  mutate(AreaCumSum = cumsum(AreaKM2))

head(productionDFBivCells)
summary(productionDFBivCells)
str(productionDFBivCells)

write.csv(productionDFBivCells, file.path(dir_M, "/git-annex/globalprep/mar/v2019/int/productionDFBivCells.csv")) #save to mazu because so large. This is functionally a raster file. 

productionDFBivCells <- read.csv(file.path(dir_M, "/git-annex/globalprep/mar/v2019/int/bivalve/productionDFBivCells.csv"))

productionByCountryBivDF <- productionDFBivCells %>%
  filter(!is.na(Country)) %>%
  dplyr::select(-YieldCumSum, -X) %>%
  dplyr::select(-AreaCumSum) %>%
  arrange(YearsToHarvest) %>%
  mutate(ID = Country) %>%
  dplyr::arrange(ID) %>%
  group_by(ID) %>%
  mutate(CountryYieldCumSum = cumsum(B_yieldpercellperyear)) %>%
  mutate(CountryAreaCumSum = cumsum(AreaKM2)) %>%
  mutate(MaxPhi = max(PhiPrime)) %>%
  mutate(averagePhi = mean(PhiPrime)) %>%
  mutate(averageWeightedPhi = sum(PhiPrime*AreaKM2)/(max(CountryAreaCumSum))) %>%
  mutate(MaxDevPerCountry = max(CountryAreaCumSum)) %>%
  mutate(MaxProdPerCountry = max(CountryYieldCumSum)) %>%
  mutate(OnePercentDevPerCountry = .01*max(CountryAreaCumSum))

head(productionByCountryBivDF)

write.csv(productionByCountryBivDF,file.path(dir_M, "git-annex/globalprep/mar/v2019/int/bivalve/BivProdByCountryByCell.csv"), row.names = FALSE) #save to mazu because so large. This is functionally a raster file. 

productionByCountryBivDF <- read.csv(file.path(dir_M, "git-annex/globalprep/mar/v2019/int/bivalve/BivProdByCountryByCell.csv")) 


CountryProdSummary <- productionByCountryBivDF %>%
  filter(CountryAreaCumSum <= OnePercentDevPerCountry) %>%
  mutate(ProdPerCountryOnePercent = max(CountryYieldCumSum)) %>%
  slice(1)

write.csv(CountryProdSummary,file.path(dir_git_ref, "production_int/BivalveProdByCountrySummary.csv"), row.names = FALSE) #save to github


region_data()
CountryLabel <- rgns_eez %>%
  dplyr::select(ID = rgn_id, rgn_name)

head(CountryLabel)

## I think the relevant value we want for the reference point is in this table:  ProdPerCountryOnePercent. Indonesia number for bivalves is higher than the paper, however the paper does use the phrase "over 3.9*10^11 tonnes". Data says 4.7 * 10^11 million tonnes of bivalves if 1% of aquaculture potential developed, which is greater than 3.9 * 10^11 tonnes.  
CountryProdSummaryFAO <- CountryProdSummary %>%
  ungroup %>%
  dplyr::select(ID:ProdPerCountryOnePercent) %>%
  full_join(CountryLabel, by = "ID")

write.csv(CountryProdSummaryFAO, file.path(dir_git_ref, "production_int/BivalveProdByCountrySummaryFAO.csv"), row.names = FALSE) #save to github 

sum(CountryProdSummaryFAO$MaxDevPerCountry, na.rm = TRUE)
#1491404 km2. Matches paper ~1,500,000 km2
```


## Unit Conversion and Gap Filling
To compare potential vs harvest, we need to convert bivalve units to metric tonnes, they are currently in units of individual bivalves.  Some figures:

* Scallops: [40-50 per kg (~22g ea)](http://www.thefoodcoach.com.au/food/?FoodID=241)
* Mussels: [15-20 per 1 lb blue (~25g ea)](https://www.pacseafood.com/products/details/seafood,shellfish/mussels/)
* Oysters: [12-15 per 1 lb (~33g ea) aquaculture halfshell](http://www.mobjackbayseafood.com/products.html)
* Clams: [20-30 per 1 lb Manila (~18g ea), 8-14 per 1 lb littlenecks (~40g ea)](https://www.pacseafood.com/products/details/seafood,shellfish/clams/)

Averaging these gives about 27.5g per piece.

``` {r}
aq_mass_per_pc <- 0.0275 * 1e-3 ### mass of bivalve piece in tonnes

pot_b <- read_csv(file.path(dir_git_ref, 'production_int/BivalveProdByCountrySummaryFAO.csv')) %>%
  mutate(potential_prod_one_percent_b = ProdPerCountryOnePercent * aq_mass_per_pc,
         potential_prod_max_b = MaxProdPerCountry * aq_mass_per_pc,
         aq_type = 'shellfish',
         units   = 'tonnes') %>%
  dplyr::select("rgn_id" = "ID", rgn_name, potential_prod_one_percent_b, potential_prod_max_b)

pot_f <- read_csv(file.path(dir_git_ref, 'production_int/FishProdByCountrySummaryFAO.csv')) %>%
  mutate(aq_type = 'finfish',
         units   = 'tonnes') %>%
  dplyr::select("rgn_id" = "ID", rgn_name, "potential_prod_one_percent_f" = "ProdPerCountryOnePercent", "potential_prod_max_f" = "MaxProdPerCountry")

pot_aq_int <- full_join(pot_b, pot_f) ## makes sense. 1% of potential AREA... not 1% of potential PRODUCTION

write_csv(pot_aq, file.path(dir_git_ref, 'production_int/aq_potential_int.csv')) ##intermediate file that might come in handy in the future

DT::datatable(pot_aq_int)


pot_aq_final <- pot_aq_int %>%
  mutate(potential_prod_one_percent_b = replace_na(potential_prod_one_percent_b, 0),
         potential_prod_one_percent_f = replace_na(potential_prod_one_percent_f, 0)) %>%
  mutate(potential_mar_tonnes = potential_prod_one_percent_b + potential_prod_one_percent_f) %>%
  arrange(rgn_id) %>%
  dplyr::select(rgn_id, potential_mar_tonnes)

write_csv(pot_aq_final, file.path(dir_git_ref, 'production_output/aq_potential_final.csv'))

DT::datatable(pot_aq_final)

## make gapfilling dataset 
pot_aq_final_gf <- pot_aq_final %>%
  mutate(gapfilled = case_when(
    potential_mar_tonnes == 0 ~ 1,
    potential_mar_tonnes > 0 ~ 0
  ), 
  method = case_when(
    potential_mar_tonnes == 0 ~ "missing regions given 0 value",
    potential_mar_tonnes > 0 ~ ""
  )) %>%
  dplyr::select(rgn_id, gapfilled, method)

write_csv(pot_aq_final_gf, file.path(dir_git_ref, "/production_output/aq_potential_gf.csv"))
```