---
title: "OHI 2018 - Food Provision: Preparing catch data for BBmsy calculations and mean catch (weights in FIS model) and FP weights"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohiprep/src/templates/ohi_hdr.html'
pdf_document:
  toc: true
---

# Summary

This script takes the Watson 2018 catch data, provided at a resolution of half-degree cells globally, and aggregates catch to **stock levels**. For the Ocean Health Index, we assume a stock is represented by the FAO region in which the species is caught.

In order to aggregate to FAO regions, we associate each cell to the FAO region and the OHI region in which it is located.

An example of our aggregation proces: New Zealand is located entirely in FAO region 81. All catch reported by New Zealand will be aggregated by species to the FAO region. If a species was reported as caught in both New Zealand waters and in the High Seas of area 81, these two records will be combined into one by summing the catch.

## Updates from previous assessment

Using Watson 2018 data this year. Catch data now goes from 1950 - 2015. Previously catch data was only available through 2014 from the Sea Around Us (SAUP). New Watson data doesn't have Taxon Key information. Add it in from Watson 2017 data after aggregating catch to work with a relativey smaller dataframe. No longer using SAUP to create cells.csv - took a raster template of Watson (watson_cell_id_rast.tif) created by Casey for a different project.

***

# Data Source

**Reference**: Watson, R. A. and Tidd, A. 2018. Mapping nearly a century and a half of global marine fishing: 1869–2015. Marine Policy, 93, pp. 171-177. [(Paper URL)](https://doi.org/10.1016/j.marpol.2018.04.023)

**Downloaded**: July 17, 2018 from [IMAS portal](http://data.imas.utas.edu.au/portal/search?uuid=ff1274e1-c0ab-411b-a8a2-5a12eb27f2c0) - clikc on download tab, step 3

**Description**:  Global fisheries landings data per cell separated by Industrial versus Non-Industrial catch, IUU, and discards.

**Native data resolution**:   

**Time range**: 1950 - 2015

**Format**:  CSV format

**Additional Information**: [Metadata](http://metadata.imas.utas.edu.au/geonetwork/srv/eng/metadata.show?uuid=ff1274e1-c0ab-411b-a8a2-5a12eb27f2c0), [Supplementary Material](https://ars.els-cdn.com/content/image/1-s2.0-S0308597X18300605-mmc1.docx)

***

## Setup

Note: the same data was used to prepare fishing pressures (prs_fish). We will be using annual catch .rds files prepared in the mazu prs_fish folder
``` {r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',message = FALSE, warning = FALSE)
```

```{r, eval=FALSE}
## Libraries
library(readr)
library(dplyr)
library(raster)
library(parallel)
library(purrr)
library(stringr)
library(tidyr)
library(foreach)
library(here)
library(sf)

setwd(here::here("globalprep/fis/v2018"))
source('../../../src/R/common.R')
source('../../../src/R/spatial_common.R')

## Paths for data
path_data = file.path(dir_M,"git-annex/globalprep/prs_fish/v2018/int")
fis_path = file.path(dir_M,"git-annex/globalprep/fis/v2018/int")

```

***

# Load Data 

## Combine Industrial and Non-Industrial Catch

* The raw Watson data is separated into industrial and non-industrial fishing. Combine both types for each year from 1950-2015.

```{r, eval=FALSE}

years <- c(1950:2015)
data_files <- list.files(file.path(path_data, "annual_catch"), full.names = T)

doParallel::registerDoParallel(3)
getDoParWorkers()

foreach(yr = years) %dopar% { # yr = 2015
  
  ## find file path of the respective year of data
  yr <- as.character(yr)
  ## Select the catch data for the respective year
  datanames <- data_files[which(str_detect(data_files, yr))]
  ## read in the two data tables
  list_data <- map(datanames, readRDS)
  
  ## combine the two data tables in your list
  combined <- bind_rows(list_data)
  
  ## save to fis folder in mazu
  saveRDS(combined, paste0(fis_path, "annual_catch/", sprintf("Catch_%s.rds",yr)))
  
  }

```

Look at catch data
```{r, eval=FALSE}

## read in one of the catch data
catch <- readRDS(file.path(fis_path, "annual_catch","Catch_2015.rds"))

head(catch)
dim(catch)
summary(catch)

```

## Read in Cells Datatable

Since we are using a new data source, we recreate the cells.csv file in **clean_cells.Rmd**, which will include cell ids and corresponding OHI and FAO region ids, which is later used to align catch data with appropriate regions.

These files are large so using the data.table package is recommended due to R memory limitations. Check that the cell values match up with the cell values in the catch data.

```{r load_data, eval=F}

cells <- read_csv('cells.csv') %>%
  dplyr::select(-X1)

# cells_old <- read_csv('../v2017/cells.csv') %>%
#   dplyr:: select(-X1)

```

# Aggregate catch

Aggregate catch per OHI region and FAO area. This catch will be used twice. 

(1) The catch is used to weight scores per region. For this we need to use catch records, including those not reported at the species level. See note below.

(2) The catch data at species level is used to calculate stock status (BBmsy) per stock (remember that our definition of a stock is a species caught within a single FAO area).


Note: Saved two versions of catch aggregation. One using IUU and Reported only (`Landings`) as the catch sum and one using IUU, Reported, and Discards (`CatchTotal`). 

(IWEN 8/27 added `na.rm=TRUE` to `summarise( )` is that correct?)

**Total Catch** - First aggregation
```{r, eval=FALSE}

## list all data files
data_files <- list.files(file.path(fis_path, "annual_catch"), full.names = T)

## function to wrangle data into what we need (total catch per OHI region per stock)
stock_rgn_total <- function(file) {

  catch <- readRDS(file)
  
output_df <- catch %>% 
    dplyr::mutate(CatchTotal = IUU + Reported + Discards) %>% 
    dplyr::select(Year, TaxonName, CommonName, Cell, CatchTotal) %>%
    dplyr::rename(CellID = Cell) %>% # match what is in cells.csv
    dplyr::left_join(cells, by = "CellID") %>%
    dplyr::mutate(catch_prop = CatchTotal * area) %>% # NAs produced here due to lack of assigned area, ok
    dplyr::group_by(Year, rgn_id, fao_id, TaxonName, CommonName) %>%
    dplyr::summarise(catch = sum(catch_prop, na.rm=TRUE)) %>% # do we want to keep NAs?
    dplyr::ungroup() %>%
    dplyr::mutate(stock_id = gsub(" ", "_", paste(TaxonName, fao_id, sep='-'), fixed=TRUE))%>%
    dplyr::rename(fao_rgn  = fao_id,
                  tons     = catch)

return(output_df)

}

total_catch <- map_df(data_files, stock_rgn_total)

write.csv(total_catch, file = file.path(dir_M,'git-annex/globalprep/fis/v2018/int/stock_catch_by_rgn.csv'), row.names=FALSE)

```


**Landings** - Second aggregation
```{r, eval=F}

## function to wrangle data into what we need (total landings per OHI region per stock)
stock_rgn_landings <- function(file) {

output_df <- readRDS(file) %>%
        dplyr::select(Year, TaxonName, CommonName, Cell, Landings) %>%
        dplyr::rename(CellID = cell_id) %>% # change cellid name to match what is in cells.csv
        dplyr::left_join(cells, by = "CellID") %>%
        dplyr::mutate(catch_prop = Landings * area) %>% # NAs produced here due to lack of assigned area, ok
        dplyr::group_by(Year, rgn_id, fao_id, TaxonName, CommonName) %>%
        dplyr::summarise(catch = sum(catch_prop, na.rm=TRUE)) %>% # do we want to keep NAs?
        dplyr::ungroup() %>%
        dplyr::mutate(stock_id = gsub(" ", "_", paste(taxon_scientific_name, fao_id, sep='-'), fixed=TRUE)) %>%
        dplyr::rename(fao_rgn  = fao_id,
               tons     = catch)

return(output_df)

}

## use purrr::map to apply the function to all yearly datasets
landings <- map_df(data_files, stock_rgn_landings)

## this is a large file (150 MB) so it is saved on the server
write.csv(landings, file =  file.path(dir_M,'git-annex/globalprep/fis/v2018/int/stock_catch_by_rgn_landings.csv'), row.names=FALSE)

```


## Add Taxon Key Information

Need taxon key to easily remove higher level (e.g. genus) taxonomic catch data. Unique taxon key was extracted from Watson 2017 catch data in `catch_taxon_key.Rmd`. Add them to the stock catch by region.

Combine taxa with all catch regions data
```{r, eval=FALSE}

## read in taxon key and stock catch (tons) by region data
taxonkey <- read.csv("int/watson_taxon_key.csv")
stock_rgn <- read_csv(file.path(dir_M,'git-annex/globalprep/fis/v2018/int/stock_catch_by_rgn.csv'))

stock_rgn_taxa <- stock_rgn %>% 
  left_join(taxonkey, by = c("TaxonName","CommonName")) # some stocks won't have matches and will get an NA in the column TaxonKey, will assess manually later

write.csv(stock_rgn_taxa, file.path(dir_M,'git-annex/globalprep/fis/v2018/int/stock_catch_by_rgn_taxa.csv'), row.names=FALSE)

```

Combine taxa with landings version as well
```{r, eval=FALSE}
stock_rgn <- read_csv(file.path(dir_M,'git-annex/globalprep/fis/v2018/int/stock_catch_by_rgn_landings.csv'))

stock_rgn_taxa <- stock_rgn %>% 
  left_join(taxonkey, by = c("TaxonName","CommonName"))

write.csv(stock_rgn_taxa, file.path(dir_M,'git-annex/globalprep/fis/v2018/int/stock_catch_by_rgn_landings_taxa.csv'), row.names=FALSE)
```

Look at which entries that don't have a Taxon key match
```{r, eval=FALSE}
df_na <- stock_rgn_taxa %>% 
  filter(is.na(Taxonkey))
noTaxonKey <- data.frame(taxa = unique(df_na$TaxonName))

## create list of catch data without taxon key info that is not at species-level
noTaxonKey <- filter(noTaxonKey, taxa != "Triglopsis quadricornis" & taxa != "Bolinus brandaris")
highTaxa <- paste(noTaxonKey[[1]], collapse="|") # used for easy filtering in B/Bmsy prep

```


## Data Check

Take a look at catch data with missing ohi and fao regions in **stock_catch_by_rgn_taxa**. These have taxon key matches, but no ohi or fao regions assigned to them. 
```{r, eval=FALSE}
df <- read_csv(file.path(dir_M,'git-annex/globalprep/fis/v2018/int/stock_catch_by_rgn_taxa.csv'))

## more NAs in this table where stock is combined with taxa info than in the table from output below
df_na <- df %>% 
  filter(is.na(rgn_id))
```

Check NA values before taxa was added
```{r, eval=FALSE}
## before adding in taxa info
stock_rgn <- read_csv(file.path(dir_M,'git-annex/globalprep/fis/v2018/int/stock_catch_by_rgn.csv'))
stock_na <- stock_rgn %>%
  filter(is.na(rgn_id))
View(stock_na)
```

Look at summary info for original catch file and output after joining to cells.csv
```{r, eval=FALSE}
catch <- readRDS(file.path(fis_path, "annual_catch","Catch_2014.rds"))
summary(catch) # no NAs

output_df <- catch %>% 
    dplyr::mutate(CatchTotal = IUU + Reported + Discards) %>% 
    dplyr::select(Year, TaxonName, CommonName, Cell, CatchTotal) %>%
    dplyr::rename(CellID = Cell) %>% # match what is in cells.csv
    dplyr::left_join(cells)
summary(output_df) # 31,367 NAs

output_na <- output_df %>% 
  filter(is.na(rgn_id)) # extract just the rows with NAs
```

Look at which cells we are missing ohi and fao regions for in the 2014 catch. Looks like a lot of the cells in Watson catch with missing OHI and FAO regions are on land along the coastline except an area above Russia and the Netherlands. Should be alright? (Iwen 8/27)

```{r, eval=FALSE}
watson_rast <- raster(extent(c(-180, 180, -90, 90)), res = 0.5, crs = '+init=epsg:4326')
values(watson_rast) <- 1:ncell(watson_rast) # should have 260 rows and 720 columns

cell_na <- unique(data.frame(cell_id = output_na$CellID, value = 1)) # set random value for viewing
cell_na_plot <- raster::subs(watson_rast, cell_na, by = "cell_id", which = "value", subsWithNA=TRUE)

map('legacy_world') 
plot(cell_na_plot, add=T)

```

***

# Prep data for B/Bmsy calculations

Catch-MSY is the model we use to estimate stock status for all global stocks. This model requires information about the resilience of each species in addition to the catch data for each year.

Load taxonomic resilience information

```{r resilience,eval=F}

## add the taxon_resilence data to catch for b/bmsy calculations
taxon_res = read_csv('../v2017/data/taxon_resilience_lookup.csv') %>%
              #mutate(common = ifelse(common %in% "Silver croaker", paste(common, sciname, sep=" "), common)) %>%
              dplyr::select(CommonName=common, Resilience)

```

## Wrangle

Filter out all stocks that don't meet our conditions:

(1) Add TaxonKey information from 2017 Watson data using `catch_taxon_key.Rmd`
(2) Keep all stocks that have at least an average annual harvest of 1000 tons
(3) Keep all stocks with time series of 20 years or more

There are some missing OHI and FAO regions, don't worry? (Iwen 8/27) 
```{r prep_data_bbmsy, eval=F}

## set variables to filter by
min_yrs = 20
min_tons = 1000

## read in catch data created above
df <- read_csv(file.path(dir_M,'git-annex/globalprep/fis/v2018/int/stock_catch_by_rgn_taxa.csv'))
# old_df <- read_csv(file.path(dir_M,'git-annex/globalprep/fis/v2017/int/stock_catch_by_rgn.csv')) # for comparison



## create dataset ready to run through catch-only models
    stks <- df %>%
              filter(Taxonkey >= 600000,               #remove all records of catch reported at higher taxonomic levels than species
                     tons     > 0) %>%                  #remove records of 0 catch
              filter(!TaxonName %in% highTaxa) %>%    # remove leftover records of catch reported at higher taxonomomic levels 
              dplyr::select(-rgn_id) %>%                       #remove rgn_id since we aggregate stocks to the FAO level   
              dplyr::group_by(stock_id, Year, fao_rgn, TaxonName, CommonName, Taxonkey) %>%
              dplyr::summarise(tons = sum(tons)) %>%           #calculate total tons per stock and year
              ungroup() %>%
              dplyr::group_by(stock_id) %>%
              dplyr::mutate(nyrs = n(),                       #get the total number of years the stock has records for   
                     avg_ann_catch = mean(tons)) %>%    #calculate the mean catch over all catch years for each stock
              dplyr::ungroup() %>%
              dplyr::filter(avg_ann_catch >= min_tons,        #keep only those stocks that meet our conditions
                              nyrs >= min_yrs) %>%
              dplyr::left_join(taxon_res, by = "CommonName") %>%                  #add resilience information
              dplyr::select(Year, TaxonName, CommonName, fao_rgn, stock_id, Taxonkey, Resilience, tons)

## check on stocks that don't have a resilience
no_res <- filter(stks, is.na(Resilience)) %>%
          dplyr::select(TaxonName, CommonName) %>%
          distinct()
    
nrow(no_res) # 270 species do not have a Resilience. These will get assigned a Medium Resilience by default by the CMSY model.

write.csv(stks, file = 'data/stock_catch.csv', row.names = FALSE)

```

Take a look at the stock data datatable
```{r, eval=FALSE}

stks = read.csv('data/stock_catch.csv')

DT::datatable(head(stks,n=100))

```


# Prep data for mean catch 

## Wrangle

Mean catch data is used to weight the B/Bmsy values in the fishery subgoal.  

```{r mean catch, eval=F}

catch <- read_csv(file.path(dir_M,'git-annex/globalprep/fis/v2018/int/stock_catch_by_rgn_taxa.csv')) %>%
  rename(common = CommonName, fao_id = fao_rgn, species=TaxonName)

summary(catch)

## filter out non ohi eez regions 
catch <- catch %>%
  filter(!is.na(rgn_id)) %>%
  filter(!is.na(fao_id)) %>%
  filter(rgn_id <= 250) %>%
  filter(rgn_id != 213)

## calculate total annual catch for each stock
catch <- catch %>%
  dplyr::select(Year, rgn_id, fao_id, stock_id, Taxonkey, tons) %>%
  group_by(rgn_id, fao_id, Taxonkey, stock_id, Year) %>%
  summarize(catch = sum(tons)) %>%
  ungroup()

```

Take a look at a few stocks. Elasmobranchii is missing after I made the adjustments to cells and taxa matching - problem? (Iwen 8/27)
```{r, eval=FALSE}
data.frame(dplyr::filter(catch, stock_id == "Elasmobranchii-57" & rgn_id==1))
data.frame(dplyr::filter(catch, stock_id == "Carcharhinidae-57" & rgn_id==1))
data.frame(dplyr::filter(catch, stock_id == "Marine_fishes_not_identified-57" & rgn_id==1))
```

## Fill in Zeros
For years with no reported catch, add zero values (after first reported catch)
```{r, eval=FALSE}

## these data have no zero catch values, so add years with no reported catch to data table:
catch_zeros <- catch %>%
  spread(Year, catch) %>%
  data.frame() %>%
  gather("Year", "catch", num_range("X", min(catch$Year):max(catch$Year))) %>%
  mutate(Year = as.numeric(gsub("X", "", Year))) %>%
  mutate(catch = ifelse(is.na(catch), 0, catch))

## this part eliminates the zero catch values prior to the first reported non-zero catch   
catch_zeros <- catch_zeros %>%
  group_by(fao_id, Taxonkey, stock_id, rgn_id) %>%
  arrange(Year) %>%
  mutate(cum_catch = cumsum(catch)) %>%
  filter(cum_catch > 0) %>%
  dplyr::select(-cum_catch) %>%
  ungroup()

```

## Calculate Mean Catch
Calculate mean catch for ohi regions (using data from 1980 onward). These data are used to weight the RAM b/bmys values 
```{r, eval=FALSE}
mean_catch <- catch_zeros %>%
  filter(Year >= 1980) %>%
  group_by(rgn_id, fao_id, Taxonkey, stock_id) %>%
  mutate(mean_catch = mean(catch, na.rm=TRUE)) %>% # mean catch for each stock (in a specific ohi-fao region)
  filter(mean_catch != 0)  %>%      ## some stocks have no reported catch for time period
  ungroup()
```

Check out the data
```{r, eval=FALSE}
data.frame(filter(mean_catch, stock_id == "Carcharhinidae-57" & rgn_id==1))
data.frame(filter(mean_catch, stock_id == "Marine_fishes_not_identified-57" & rgn_id==1))
```

## Toolbox formatting and save
```{r, eval=FALSE}

mean_catch_toolbox <- mean_catch %>%
  mutate(stock_id_taxonkey = paste(stock_id, Taxonkey, sep="_")) %>%
  dplyr::select(rgn_id, stock_id_taxonkey, Year, mean_catch) %>%
  filter(Year >= 2001) %>%  # filter to include only analysis years
  data.frame()

write.csv(mean_catch_toolbox, "data/mean_catch.csv", row.names=FALSE)

```

Compare with last year
```{r, eval=FALSE}

new <- read.csv("data/mean_catch.csv")
new_filt <- new %>% 
  filter(stock_id_taxonkey == "Carangidae-31_400314") %>% 
  rename(new_mean_catch = mean_catch)

old <- read.csv("../v2017/data/mean_catch.csv")
old_filt <- old %>% 
  filter(stock_id_taxonkey == "Carangidae-31_400314") %>% 
  rename(Year = year)

check <- old_filt %>% 
  left_join(new_filt, by = c("rgn_id","stock_id_taxonkey","Year"))

plot(check$mean_catch,check$new_mean_catch)
abline(col="red", 1,1)

```

# Prep data for food provision weights 

These data determine the tonnes of food provided by fisheries.  Ultimately, the proportion of food from fisheries relative to mariculture will be calculated to weight the contributions of fishery and mariculture scores to final food provision scores.  

```{r tonnes fp, eval=F}

total_catch_FP <- mean_catch %>%
  group_by(rgn_id, Year) %>%
  summarize(fis_catch = sum(catch)) %>%
  dplyr::select(rgn_id, Year, fis_catch) %>%
  filter(Year >= 2005) # filter to include only the relevant analysis years

write.csv(total_catch_FP, "data/FP_fis_catch.csv", row.names=FALSE)

```

***

# Citation information

Pauly D. and Zeller D. (Editors), 2015. Sea Around Us Concepts, Design and Data (seaaroundus.org)