---
title: "Investigate Differences"
author: "Iwen Su"
date: "9/5/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Investigate Differences RAM v4.40 and RAM v3.80

## Investigate differences between `ram_spatial` and `ram_meta`

Save New and Old RAM metadata and timeseries data to unqiue variables for comparison. Old data checks out but new data has more unqiue `assessid` and `stockid` entries in the `timeseries` data than in the `metadata`. 
```{r}
## Check unique stockid in new RAM data
load(file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2018/RAM v4.40 (6-4-18)/DB Files With Assessment Data/DBdata.RData"))

## save newest metadata to unique variable 
RAM4.40_meta = as.data.frame(metadata)
RAM4.40_data = as.data.frame(timeseries)

## Check length
length(unique(RAM4.40_meta$assessid)) #1252 unique entries
length(unique(RAM4.40_meta$stockid)) # same number of stockids as assessids
length(unique(RAM4.40_meta$stocklong)) # same number of stocklong as the above two

length(unique(RAM4.40_data$assessid)) # 1814 unique entries, more than in metadata
length(unique(RAM4.40_data$stockid)) # 1290 unique, ~500 fewer..
length(unique(RAM4.40_data$stocklong)) #1291 unique

setdiff(RAM4.40_data$assessid, RAM4.40_meta$assessid) #563 diff in data not in meta
setdiff(RAM4.40_data$stockid, RAM4.40_meta$stockid) #39 stockids in data not in meta


## old stocks 
load(file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2017/RAM v3.80/DB Files With Assessment Data/DBdata.RData"))

RAM3.80_meta = as.data.frame(meta.data)
RAM3.80_data = as.data.frame(timeseries)

## Check length
length(unique(RAM3.80_meta$assessid)) # 1058
length(unique(RAM3.80_meta$stockid)) # 1058
length(unique(RAM3.80_meta$assessid)) # 1058

length(unique(RAM3.80_data$assessid)) # 1057 
length(unique(RAM3.80_data$stockid)) # 1057
length(unique(RAM3.80_data$stocklong)) # 1057

setdiff(RAM3.80_data$assessid, RAM3.80_meta$assessid)
setdiff(RAM3.80_data$stockid, RAM3.80_meta$stockid)

```

## Investigate New RAM Database filterd for B/Bmsy data
```{r}

ram_bmsy <- RAM4.40_data %>%
  dplyr::filter(tsid == "BdivBmsytouse-dimensionless") %>%
  dplyr::filter(!is.na(tsvalue)) %>%
  dplyr::mutate(tsyear = as.numeric(as.character(tsyear))) %>%
  dplyr::filter(tsyear > 1979) %>%
  dplyr::select(assessid, year=tsyear, ram_bmsy = tsvalue, stockid, stocklong)

length(unique(ram_bmsy$assessid)) # 473 unique entries
length(unique(ram_bmsy$stockid)) # 353
length(unique(ram_bmsy$stocklong)) # 353

setdiff(ram_bmsy$assessid, RAM4.40_meta$assessid) # 155 diffs
setdiff(ram_bmsy$stockid, RAM4.40_meta$stockid) # 8 diffs


```

## Check a stock that is in `ram_spatial` not `raw_meta`

Try looking at stockid BIGEYEWPO, in ram_spatial but not raw_meta. ram_spatial combines last year's stock and this year's stock. Check to see if BIGEYEWPO is in old RAM database
```{r}

## See if BIGEYEWPO is in this vector - it is not
new_stockid <- unique(ram_bmsy$stockid)

## Check to see if BIGEYEWPO is in old stock data
ram_bmsy_old <- data.frame(RAM3.80_data) %>%
  dplyr::filter(tsid == "BdivBmsytouse-dimensionless") %>%
  dplyr::filter(!is.na(tsvalue)) %>%
  dplyr::mutate(tsyear = as.numeric(as.character(tsyear))) %>%
  dplyr::filter(tsyear > 1979) %>%
  dplyr::select(assessid, year=tsyear, ram_bmsy = tsvalue, stockid, stocklong)

## YES, BIGEYEWPO is in the old stock data. Maybe we should be be using the new RAM stock data instead of joining the added stocks to the old stocks?
old_stockid <-  unique(ram_bmsy_old$stockid)


## Compare stockid in raw_meta for v4.40 RAM data versus stockid in timeseries.. should match
setdiff(ram_bmsy$stockid,ram_meta$stockid)

```

***

# Check Stock Name Changes

Do comparison with documented stock name changes. 
```{r}

## from fao_ohi_rgns.Rmd
newStocks <- data.frame(stockid = setdiff(new_stockid, old_stockid)) # there are 101 diff stock names

## import stock name changes
## gather to make it easier to search for updated stocks
changes <- read_csv(file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2018/RAM v4.40 (6-4-18)/DB Documents/Stock_Change_History_060418.csv"))

changes <- changes %>% 
  gather("temp","stock_changes", contains("stock")) %>% 
  select(outdated, stock_changes) %>% 
  na.omit()

## List of species that have documented stock changes - reconcile with the Stock Change History file
stock_change_match <- newStocks %>% 
  filter(stockid %in% changes$stock_changes)

## from RAM_data_prep.Rmd
ram_spatial <- read.csv("int/RAM_fao_ohi_rgns.csv", stringsAsFactors = FALSE)
ram_meta <- data.frame(metadata) %>%
  dplyr::select(assessid, stockid, scientificname)

diff <- setdiff(ram_spatial$stockid, ram_meta$stockid)

stock_change_match <- stock_change_match %>% 
  filter(stockid %in% diff)
```


***

Look through stock metadata with Mel

# Investigate `assessid` and `stockid` in RAM v4.40 Database

## Identify the number of assessments per stock
```{r}
## see unique pairs of assess and stockid
test <- RAM4.40_data %>% 
  select(assessid, stockid) %>% 
  unique() %>% 
  group_by(stockid) %>% 
  summarize(num = length(assessid))
```

## Check stocks with multiple assessments

Check `ALPLAICBSAI` which has 2 assessments
```{r}
test <- ram_bmsy %>% 
  filter(stockid == "ALPLAICBSAI") %>% 
  spread(assessid, ram_bmsy)

plot(test$`AFSC-ALPLAICBSAI-1972-2010-STACHURA`,test$`AFSC-ALPLAICBSAI-1975-2015-SISIMP2016`)
abline(0,1, col="red")
```

ANGLVIIIc-IXa has 6 assessments, take a look
```{r}
test2 <- ram_bmsy %>% 
  filter(stockid == "ANGLVIIIc-IXa")

unique(test2$assessid)

test3 <- test2 %>% 
  spread(assessid, tsvalue)
```
