---
title: "OHI 2018: Food Provision, Identify FAO and OHI regions for RAM stocks"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohiprep/src/templates/ohi_hdr.html'
pdf_document:
  toc: true
---

```{r}
knitr::opts_chunk$set(eval=FALSE)
```

## Summary

Assign FAO and OHI region ids for each additional stock (`stockid`) listed. Some stocks have multiple assessments, assign them the same region ids and retain that column information (`assessid`). 


## Setup

```{r}

## Libraries
library(dplyr)
library(tidyr)
library(readr)
library(sf)
library(ggplot2)
library(stringr)
library(here) 
setwd(here::here("globalprep","fis","v2018"))

source('../../../src/R/common.R')

```

## Identify Newly Added Stocks

Compare old to new stocks to see what additional stocks have been added.

Old Stocks: Used `timeseries` table, because in RAM v3.80, this table only contained single assessment for each unique stock
```{r}

## old stocks
load(file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2017/RAM v3.80/DB Files With Assessment Data/DBdata.RData"))

ram_bmsy_old <- data.frame(timeseries) %>%
  dplyr::filter(tsid == "BdivBmsytouse-dimensionless") %>%
  dplyr::filter(!is.na(tsvalue)) %>%
  dplyr::mutate(tsyear = as.numeric(as.character(tsyear))) %>%
  dplyr::filter(tsyear > 1979) %>%
  dplyr::select(assessid, year=tsyear, ram_bmsy = tsvalue, stockid, stocklong) 

## check unique values
old_stockid <- unique(ram_bmsy_old$stockid) # 257 unique entries

```

New Stocks: Used `timeseries_values_views` table, because in RAM v4.40, this is the table that contains the most recent assessment for each unique stock. The `timeseries` table has all assessments conducted so there are multiple `assessid` per unique `stockid`.
```{r}
## new stocks
load(file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2018/RAM v4.40 (6-4-18)/DB Files With Assessment Data/DBdata.RData"))

# What we used to use
ram_bmsy_ts <- data.frame(timeseries) %>%
  dplyr::filter(tsid == "BdivBmsytouse-dimensionless") %>%
  dplyr::filter(!is.na(tsvalue)) %>%
  dplyr::mutate(tsyear = as.numeric(as.character(tsyear))) %>%
  dplyr::filter(tsyear > 1979) %>%
  dplyr::select(assessid, year=tsyear, ram_bmsy = tsvalue, stockid, stocklong)

new_stockid_ts <- unique(ram_bmsy_ts$stockid) #353 unique entries

# What we're trying to use now
ram_bmsy <- timeseries_values_views %>%
  select(stockid, stocklong, year, TBdivTBmsy, SSBdivSSBmsy) %>% 
  mutate(ram_bmsy = ifelse(!is.na(TBdivTBmsy), TBdivTBmsy, SSBdivSSBmsy)) %>% 
  filter(!is.na(ram_bmsy)) %>% 
  dplyr::filter(year > 1979) %>%
  dplyr::select(stockid, stocklong, year, ram_bmsy)

## check number of unique entries
new_stockid <- unique(ram_bmsy$stockid) # 345 unique entries

```


Check differences

Make sure to check for stock name changes from old to new. Check documentation in mazu (e.g. Stock Change History). A few outdated stock names. Can remove unused `stockid`s in RAM_data_prep.Rmd if needed. The following 4 stocks were matched using unique `timeseries` stocks that are not needed in unique `timeseries_values_views` stocks: "OROUGHYCR", "STMARLINNPAC", "HOGFISHEGM", "HOGFISHSEFL".
```{r}

## find newly added stocks, when using timeseries_values_views
newStocks <- setdiff(new_stockid, old_stockid) # there are 97 diff stock names

## find newly added stocks when using timeseries
newStocks2 <- setdiff(new_stockid_ts, old_stockid) # there are 101 differences

## Compare diff when using `timeseries` table with diff when using `timeseries_values_views` for RAMv4.40 
setdiff(newStocks, newStocks2) # nothing in new that's not in old..
setdiff(newStocks2, newStocks) # 4 in old not in new

```

Subset for just new stocks in RAM v4.40 data
```{r}

ram_bmsy <- ram_bmsy %>% 
  filter(stockid %in% newStocks)

id_new <- ram_bmsy %>% 
  select(stockid, stocklong) %>% 
  distinct()

```


## Add Stock Info to Old RAM_fao_ohi_rgns

Combine the data table Mel created `RAM_fao_ohi_rgns` that contains fao id, ohi id, and assessid from RAM with the old RAM data to match each row with the appropriate stockid and stocklong information
```{r}
## Grab last year's fao-ohi-assessid data table
RAM_fao_ohi_rgns <- read.csv("../v2017/int/RAM_fao_ohi_rgns.csv") 

## create unique values of assessid, stock id, and stock long
id_old <- ram_bmsy_old %>% 
  select(assessid, stockid, stocklong) %>% 
  distinct()

## join tables
RAM_rgns_old <- RAM_fao_ohi_rgns %>% 
  full_join(id_old, by="assessid") %>% 
  select(-assessid) # no longer using this id

```

## Combine New and Old Stock Data to Compare

Essentially adds new stocks to the bottom of the old stocks data table.
```{r}

RAM_rgns_new <- RAM_rgns_old %>% 
  full_join(id_new, by = c("stockid","stocklong"))

```

## Fill Region IDs for New Stocks

Investigate what FAO/OHI region each stock belongs to: 
* Full intermediate stock data table [here](https://github.com/OHI-Science/ohiprep_v2018/blob/master/globalprep/fis/v2018/int/RAM_fao_ohi_rgns.csv)
* Primarily used www.fishsource.org to find stock distribution information. ICES sub-region codes found here: https://www.researchgate.net/figure/Stock-units-defined-by-ICES-for-white-anglerfish-southern-stock-in-Divisions-VIIIc-and_fig1_31198841 and http://gis.ices.dk/sf/
* Referenced F_CODE/FAO_AREAS (filtered for MAJOR in F_LEVEL) in ohi_fao_rgns shapefile (fis/v2017/int) from mazu (can use other map visuals to identify stock extent). 
* Searched and tested key words in data table viewer (`DT::datatable`) to use in `str_detect`. See which stocks have the same FAO id assignment(s).

Some regions have more than 1 `fao_id` or `rgn_id` (e.g. Alaska skate Bering Sea and Aleutian Islands,	Kamchatka flounder Bering Sea and Aleutian Islands have `fao_id` 61 and 67).

Note: In `str_detect`, it doesn't always work when you type out "Bering Sea | Gulf of Alaska$". Using `paste` with the `collapse = "|"` argument seems to work best.

```{r}

## Subset for rows with missing FAO and OHI ids
RAM_filt <- RAM_rgns_new %>% 
  filter(is.na(fao_id))

RAM <- RAM_filt %>% 
  ## ADDING FAO REGION ID
  mutate(fao_id = 
           case_when(
             str_detect(stocklong, paste(c("Bering Sea","Gulf of Alaska$"), collapse = "|")) ~ 61,
             str_detect(stocklong, "Sicily$") ~ 37,
             str_detect(stocklong, paste(c("^Jackass morwong Western","^Jackass morwong Eastern","Redfish Eastern Australia","Blue grenadier New South Wales"), collapse = "|")) ~ 57,
             str_detect(stocklong, paste(c("Scotian Shelf","Haddock","NAFO", "Georges Bank", "Atlantic coast", "Atlantic Coast"), collapse = "|")) ~ 21,
             str_detect(stocklong, paste(c("Japan", "Mikawa Bay", "East China Sea", "Tsushima Strait","Kichiji", "^Pacific cod North", "Northwest Pacific", "Snow crab North Pacific", "Willowy flounder North Pacific"), collapse = "|")) ~ 61,
             str_detect(stocklong, ".*Chile$") ~ 87,
             str_detect(stocklong, ".*Argentina$") ~ 41,
             str_detect(stocklong, paste(c("China rockfish Northern", "China rockfish Central", "Blue shark North Pacific", "^Black rockfish", "Aurora rockfish", "Brown rockfish Pacific Coast", "Copper rockfish Pacific Coast", "Striped marlin"), collapse = "|")) ~ 67,
             str_detect(stocklong, paste(c("China rockfish Southern", "Black rockfish California"), collapse = "|")) ~ 77,
             str_detect(stocklong, paste(c("^Black bellied angler", "^Megrim", "Roundnose grenadier", "Greenland halibut", "Northern shrimp"), collapse = "|")) ~ 27,
             str_detect(stocklong, "Abalone South Africa") ~ 47,
             str_detect(stocklong, paste(c("Gulf of Mexico", "Hogfish Southeast Florida"), collapse = "|")) ~ 31,
             str_detect(stocklong, "Orange roughy Chatham Rise") ~ 81
           )) %>%
  
  mutate(fao_id_2 =
           case_when(
             str_detect(stocklong, "Bering Sea") ~ 67,
             str_detect(stocklong,  paste(c("^Jackass morwong Eastern","Redfish Eastern Australia","Blue grenadier New South Wales"), collapse = "|")) ~ 81,
             str_detect(stocklong, paste(c("Southern blue whiting Chile", "Southern hake Chile", "Patagonian Toothfish Chile"), collapse = "|")) ~ 41,
             str_detect(stocklong, paste(c("China rockfish Central","Blue shark North Pacific", "Black rockfish California", "Aurora rockfish", "Brown rockfish Pacific Coast", "Copper rockfish Pacific Coast", "Striped marlin"), collapse = "|")) ~ 77,
             str_detect(stocklong, paste(c("Blueline tilefish", "Cobia Southern"), collapse = "|")) ~ 31
             )) %>% 
  
  mutate(fao_id_3 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific", "Striped marlin"), collapse = "|")) ~ 61
           )) %>% 
  
  mutate(fao_id_4 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific", "Striped marlin"), collapse = "|")) ~ 71
           )) %>%
  
    mutate(fao_id_5 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific", "Striped marlin"), collapse = "|")) ~ 87
           )) %>%
  
  ## ADDING OHI REGION ID
  mutate(rgn_id = 
           case_when(
             str_detect(stocklong, paste(c("Bering Sea","Gulf of Alaska$", "Gulf of Mexico", "^Haddock Georges", "^China rockfish.*Pacific Coast$", "Blue shark North Pacific", "^Monkfish", "Yellowtail flounder","Atlantic Coast","Atlantic coast", "Black rockfish", "Aurora rockfish", "Brown rockfish Pacific Coast", "Copper rockfish Pacific Coast", "Hogfish Southeast Florida", "Striped marlin"), collapse = "|")) ~ 163,
             str_detect(stocklong, "Sicily$") ~ 184,
             str_detect(stocklong, "Australia$") ~ 16,
             str_detect(stocklong, paste(c("Scotian Shelf","NAFO"), collapse = "|")) ~ 218,
             str_detect(stocklong, paste(c("Japan","Tsushima Strait", "East China Sea$", "^Kichiji", "^Pacific cod North", "Northwest Pacific", "Snow crab North Pacific", "Willowy flounder North Pacific"), collapse = "|")) ~ 210,
             str_detect(stocklong, ".*Chile$") ~ 224,
             str_detect(stocklong, ".*Argentina$") ~ 172,
             str_detect(stocklong, "^Black bellied angler") ~ 182,
             str_detect(stocklong, paste(c("^Megrim", "Roundnose grenadier", "Greenland halibut"), collapse = "|")) ~ 181,
             str_detect(stocklong, "Northern shrimp") ~ 180,
             str_detect(stocklong, "Abalone South Africa") ~ 102,
             str_detect(stocklong, "Orange roughy Chatham Rise") ~ 162
           )) %>% 
  
  mutate(rgn_id_2 =
           case_when(
             str_detect(stocklong, paste(c("Bering Sea", "Japanese anchovy Pacific Coast", "Japanese sardine Pacific Coast", "Northwest Pacific", "Snow crab North Pacific"), collapse = "|")) ~ 73,
             str_detect(stocklong, "Sicily$") ~ 68,
             str_detect(stocklong, paste(c("Japanese flying squid", "Tsushima Strait", "East China Sea$", "Willowy flounder North Pacific"), collapse="|")) ~ 20,
             str_detect(stocklong, paste(c("Southern blue whiting Chile", "Southern hake Chile", "Patagonian Toothfish Chile"), collapse = "|")) ~ 172,
             str_detect(stocklong, ".*Northern Argentina$") ~ 173,
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 218,
             str_detect(stocklong, "^Black bellied angler") ~ 183,
             str_detect(stocklong, paste(c("^Megrim", "Roundnose grenadier", "Greenland halibut"), collapse = "|")) ~ 180,
             str_detect(stocklong, "Northern shrimp") ~ 223,
             str_detect(stocklong, "Striped marlin") ~ 137
           )) %>% 
  
  mutate(rgn_id_3 =
           case_when(
              str_detect(stocklong, "Sicily$") ~ 61,
              str_detect(stocklong, paste(c("Japanese flying squid", "Tsushima Strait","East China Sea$", "Willowy flounder North Pacific"), collapse="|")) ~ 209,
              str_detect(stocklong, "Snow crab North Pacific") ~ 20,
              str_detect(stocklong, ".*Northern Argentina$") ~ 171,
              str_detect(stocklong, "Patagonian Toothfish Chile") ~ 173,
              str_detect(stocklong, paste(c("Blue shark North Pacific", "Striped marlin"), collapse = "|")) ~ 135,
              str_detect(stocklong, paste(c("Roundnose grenadier"), collapse = "|")) ~ 179,
              str_detect(stocklong, "Greenland halibut") ~ 143,
              str_detect(stocklong, "Northern shrimp") ~ 175
           )) %>% 
  
  mutate(rgn_id_4 = 
           case_when(
             str_detect(stocklong, "Snow crab North Pacific") ~ 21,
             str_detect(stocklong, "Patagonian Toothfish Chile") ~ 95,
             str_detect(stocklong, paste(c("Blue shark North Pacific", "Striped marlin"), collapse = "|")) ~ 136,
             str_detect(stocklong, paste(c("Roundnose grenadier", "Greenland halibut"), collapse = "|")) ~ 141,
             str_detect(stocklong, "Northern shrimp") ~ 222
           )) %>% 
  mutate(rgn_id_5 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific", "Striped marlin"), collapse = "|")) ~ 134,
             str_detect(stocklong, "Greenland halibut") ~ 145
           )) %>% 
  mutate(rgn_id_6 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific", "Striped marlin"), collapse = "|")) ~ 131
           )) %>% 
  mutate(rgn_id_7 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific", "Striped marlin"), collapse = "|")) ~ 130
           )) %>% 
  mutate(rgn_id_8 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific", "Striped marlin"), collapse = "|")) ~ 129
           )) %>% 
  mutate(rgn_id_9 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific", "Striped marlin"), collapse = "|")) ~ 159
           )) %>% 
  mutate(rgn_id_10 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 150,
             str_detect(stocklong, "Striped marlin") ~ 132
           )) %>% 
  mutate(rgn_id_11 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 207,
             str_detect(stocklong, "Striped marlin") ~ 148
           )) %>% 
  mutate(rgn_id_12 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 206
           )) %>% 
  mutate(rgn_id_13 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 25
           )) %>% 
  mutate(rgn_id_14 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 14
           )) %>% 
  mutate(rgn_id_15 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 15
           )) %>% 
  mutate(rgn_id_16 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 8
           )) %>% 
  mutate(rgn_id_17 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 9
           )) %>% 
  mutate(rgn_id_18 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 13
           )) %>% 
  mutate(rgn_id_19 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 210
           )) %>% 
  mutate(rgn_id_20 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 12
           )) %>% 
  mutate(rgn_id_21 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 11
           )) %>% 
  mutate(rgn_id_22 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 73
           ))
             
DT::datatable(RAM,rownames = F)

```

## Tidy Final Datatable

### Gather FAO and OHI ids into long format

```{r}

RAM_temp <- RAM %>% 
  gather(fao, fao_values, contains("fao_id")) %>% 
  rename(fao_id = fao_values) %>% 
  select(-fao) %>% 
  gather(ohi, ohi_values, contains("rgn_id")) %>% 
  rename(rgn_id = ohi_values) %>% 
  select(-ohi) %>% 
  filter(!is.na(fao_id), !is.na(rgn_id)) 

```

### Fix Incorrect FAO-OHI Matches

Each row must have both an FAO id and a rgn id. Fix duplicate/incorrect fao-ohi matches by using ohi_fao_rgns spatial file
```{r}

## Spatial file with fao and ohi regions, F_CODE is the FAO id
fao_ohi <- st_read(dsn = file.path(dir_M, "git-annex/globalprep/fis/v2017/int"),
                   layer = "ohi_fao_rgns")
st_geometry(fao_ohi) <- NULL # removes geometry

fao_ohi_id <- fao_ohi %>%
  select(rgn_id, fao_id = F_CODE) %>% 
  arrange(rgn_id) %>% 
  mutate(fao_id = as.numeric(as.character(fao_id))) %>% 
  mutate(rgn_id = as.numeric(as.character(rgn_id))) %>% 
  distinct()

## Filter for correct fao-ohi pairs in the RAM regions table 
RAM_rgns_new_final <- fao_ohi_id %>% 
  left_join(RAM_temp, by = c("rgn_id", "fao_id")) %>% 
  filter(!is.na(stockid)) %>% # keep matches only
  select(-RAM_area_m2) 

```

Compare with unfiltered RAM regions table to check on fao-ohi pairs that were dropped
```{r}
## Number of unique stocks after joining
nrow(RAM_rgns_new_final) #203
## Number of unique stocks before joining
RAM_temp <- RAM_temp %>% select(rgn_id, fao_id, stockid, stocklong) %>% distinct()
nrow(RAM_temp) #282 

## add identifier for the two RAM table versions
RAM_temp$idtemp <- 1:nrow(RAM_temp)
RAM_rgns_new_final$idtemp2 <- 1:nrow(RAM_rgns_new_final)

## view the ohi-fao papirs that would be removed
combine <- RAM_temp %>% 
  full_join(RAM_rgns_new_final, by = c("rgn_id", "fao_id", "stockid", "stocklong")) %>% 
  filter(is.na(idtemp2))

```

Tidy up RAM data table
```{r}

RAM_final <- RAM_rgns_new_final %>% 
  select(-idtemp2)

## check that there are still 97 unique stock ids
length(unique(RAM_final$stockid))

```

### Combine with full dataset

Combine newly added stocks with ohi and fao region information to the old stock data table. Make sure there are no NAs!
```{r}

RAM_fao_ohi_rgn <- RAM_rgns_old %>% 
  full_join(RAM_final, by = c("rgn_id", "fao_id", "stockid", "stocklong")) 

write.csv(RAM_fao_ohi_rgn, "int/RAM_fao_ohi_rgns.csv", row.names=FALSE)

```

Summary: combined dataset contains old stock-spatial information from RAM v3.80 database and Christopher M. Free's spatial work at Rutgers as well as newly added stocks from RAM v4.40 with manually assigned ohi and fao region ids. Dataset only contains stocks with B/Bmsy data.
