---
title: "Untitled"
author: "Iwen Su"
output: html_document
---


```{r}

## Libraries
library(dplyr)
library(tidyr)
library(readr)
library(sf)
library(ggplot2)
library(stringr)
library(here) 
setwd(here::here("globalprep","fis","v2018"))

source('../../../src/R/common.R')

```

## Identify Newly Added Stocks
```{r}

## old stocks
load(file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2017/RAM v3.80/DB Files With Assessment Data/DBdata.RData"))

ram_bmsy_old <- data.frame(timeseries) %>%
  dplyr::filter(tsid == "BdivBmsytouse-dimensionless") %>%
  dplyr::filter(!is.na(tsvalue)) %>%
  dplyr::mutate(tsyear = as.numeric(as.character(tsyear))) %>%
  dplyr::filter(tsyear > 1979) %>%
  dplyr::select(assessid, year=tsyear, ram_bmsy = tsvalue, stockid, stocklong)

old_stocklong <- unique(ram_bmsy_old$stocklong)
old_stockid <- unique(ram_bmsy_old$stockid)
old_assessid <- unique(ram_bmsy_old$assessid)

## new stocks
load(file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2018/RAM v4.40 (6-4-18)/DB Files With Assessment Data/DBdata.RData"))

ram_bmsy <- data.frame(timeseries) %>%
  dplyr::filter(tsid == "BdivBmsytouse-dimensionless") %>%
  dplyr::filter(!is.na(tsvalue)) %>%
  dplyr::mutate(tsyear = as.numeric(as.character(tsyear))) %>%
  dplyr::filter(tsyear > 1979) %>%
  dplyr::select(assessid, year=tsyear, ram_bmsy = tsvalue, stockid, stocklong)

new_stocklong <- unique(ram_bmsy$stocklong)
new_stockid <- unique(ram_bmsy$stockid)
new_assessid <- unique(ram_bmsy$assessid)

## find diff
setdiff(new_stocklong, old_stocklong)
newStocks <- setdiff(new_stockid, old_stockid); length(newStocks)
setdiff(new_assessid, old_assessid)



```


Subset for just new stocks in 2018 RAM data
```{r}

ram_bmsy <- ram_bmsy %>% 
  filter(stockid %in% newStocks)

## some stockid values are repeated in more than one observation due to multiple assessid years
id_new <- ram_bmsy %>% 
  select(assessid, stockid, stocklong) %>% 
  distinct()

```


## Add Stock Info to Old RAM_fao_ohi_rgns

Combine the data table Mel created `RAM_fao_ohi_rgns` that contains fao id, ohi id, and assessid from RAM with the old RAM data to match each row with the appropriate stockid and stocklong information
```{r}
## Grab last year's fao-ohi-assessid data table
RAM_fao_ohi_rgns <- read.csv("../v2017/int/RAM_fao_ohi_rgns.csv")

## create unique values of assessid, stock id, and stock long
id_old <- ram_bmsy_old %>% 
  select(assessid, stockid, stocklong) %>% 
  distinct()

## join tables
all_old <- RAM_fao_ohi_rgns %>% 
  full_join(id_old, by="assessid")

```

## Combine New and Old Stock Data to Compare

```{r}

RAM_rgns_new <- all_old %>% 
  full_join(id_new, by = c("assessid","stockid","stocklong"))

```

## Fill Region IDs for New Stocks

Investigate what FAO/OHI region each stock belongs to: 
* Full intermediate stock data table [here](https://github.com/OHI-Science/ohiprep_v2018/blob/master/globalprep/fis/v2018/int/RAM_fao_ohi_rgns.csv)
* Primarily used www.fishsource.org to find stock distribution information. ICES sub-region codes found here: https://www.researchgate.net/figure/Stock-units-defined-by-ICES-for-white-anglerfish-southern-stock-in-Divisions-VIIIc-and_fig1_31198841 and http://gis.ices.dk/sf/
* Referenced FAO_AREAS (filtered for MAJOR) and ohi_fao_rgns shapefile from mazu on ArcGIS (can use other map visuals to identify stock extent). 
* Searched and tested key words in data table viewer (`DT::datatable`) to use in `str_detect`. See which stocks have the same FAO id assignment(s).

Some regions have more than 1 `fao_id` or `rgn_id` (e.g. Alaska skate Bering Sea and Aleutian Islands,	Kamchatka flounder Bering Sea and Aleutian Islands have `fao_id` 61 and 67).

Note: In `str_detect`, it doesn't always work when you type out "Bering Sea | Gulf of Alaska$". Using `paste` with the `collapse = "|"` argument seems to work best.

```{r}

## Subset for rows with missing FAO and OHI ids
RAM_filt <- RAM_rgns_new %>% 
  filter(is.na(fao_id)) %>% 
  select(-RAM_area_m2)

RAM <- RAM_filt %>% 
  ## ADDING FAO REGION ID
  mutate(fao_id = 
           case_when(
             str_detect(stocklong, paste(c("Bering Sea","Gulf of Alaska$"), collapse = "|")) ~ 61,
             str_detect(stocklong, "Sicily$") ~ 37,
             str_detect(stocklong, paste(c("^Jackass morwong Western","^Jackass morwong Eastern","Redfish Eastern Australia","Blue grenadier New South Wales"), collapse = "|")) ~ 57,
             str_detect(stocklong, paste(c("Scotian Shelf","Haddock","NAFO", "Georges Bank", "Atlantic Coast", "Atlantic coast"), collapse = "|")) ~ 21,
             str_detect(stocklong, paste(c("Japan", "Mikawa Bay", "East China Sea", "Tsushima Strait","Kichiji", "^Pacific cod North", "Northwest Pacific", "Snow crab North Pacific", "Willowy flounder North Pacific"), collapse = "|")) ~ 61,
             str_detect(stocklong, ".*Chile$") ~ 87,
             str_detect(stocklong, ".*Argentina$") ~ 41,
             str_detect(stocklong, paste(c("China rockfish Northern", "China rockfish Central", "Blue shark North Pacific", "^Black rockfish", "Aurora rockfish", "Brown rockfish Pacific Coast", "Copper rockfish Pacific Coast", "Striped marlin"), collapse = "|")) ~ 87,
             str_detect(stocklong, paste(c("China rockfish Southern", "Black rockfish California"), collapse = "|")) ~ 77,
             str_detect(stocklong, paste(c("^Black bellied angler", "^Megrim", "Roundnose grenadier", "Greenland halibut", "Northern shrimp"), collapse = "|")) ~ 27,
             str_detect(stocklong, "Abalone South Africa") ~ 47,
             str_detect(stocklong, paste(c("Gulf of Mexico", "Hogfish Southeast Florida"), collapse = "|")) ~ 31,
             str_detect(stocklong, "Orange roughy Chatham Rise") ~ 81
           )) %>%
  
  mutate(fao_id_2 =
           case_when(
             str_detect(stocklong, "Bering Sea") ~ 67,
             str_detect(stocklong,  paste(c("^Jackass morwong Eastern","Redfish Eastern Australia","Blue grenadier New South Wales"), collapse = "|")) ~ 81,
             str_detect(stocklong, paste(c("Southern blue whiting Chile", "Southern hake Chile", "Patagonian Toothfish Chile"), collapse = "|")) ~ 41,
             str_detect(stocklong, paste(c("China rockfish Central","Blue shark North Pacific", "Black rockfish California", "Aurora rockfish", "Brown rockfish Pacific Coast", "Copper rockfish Pacific Coast", "Striped marlin"), collapse = "|")) ~ 77,
             str_detect(stocklong, paste(c("Blueline tilefish", "Cobia Southern"), collapse = "|")) ~ 31
             )) %>% 
  
  mutate(fao_id_3 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific", "Striped marlin"), collapse = "|")) ~ 61
           )) %>% 
  
  mutate(fao_id_4 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific", "Striped marlin"), collapse = "|")) ~ 71
           )) %>%
  
  ## ADDING OHI REGION ID
  mutate(rgn_id = 
           case_when(
             str_detect(stocklong, paste(c("Bering Sea","Gulf of Alaska$", "Gulf of Mexico", "^Haddock Georges", "^China rockfish.*Pacific Coast$", "Blue shark North Pacific", "^Monkfish", "Yellowtail flounder","Atlantic Coast","Atlantic coast", "Black rockfish", "Aurora rockfish", "Brown rockfish Pacific Coast", "Copper rockfish Pacific Coast", "Hogfish Southeast Florida", "Striped marlin"), collapse = "|")) ~ 163,
             str_detect(stocklong, "Sicily$") ~ 184,
             str_detect(stocklong, "Australia$") ~ 16,
             str_detect(stocklong, paste(c("Scotian Shelf","NAFO"), collapse = "|")) ~ 218,
             str_detect(stocklong, paste(c("Japan","Tsushima Strait", "East China Sea$", "^Kichiji", "^Pacific cod North", "Northwest Pacific", "Snow crab North Pacific", "Willowy flounder North Pacific"), collapse = "|")) ~ 210,
             str_detect(stocklong, ".*Chile$") ~ 224,
             str_detect(stocklong, ".*Argentina$") ~ 172,
             str_detect(stocklong, "^Black bellied angler") ~ 182,
             str_detect(stocklong, paste(c("^Megrim", "Roundnose grenadier", "Greenland halibut"), collapse = "|")) ~ 181,
             str_detect(stocklong, "Northern shrimp") ~ 180,
             str_detect(stocklong, "Abalone South Africa") ~ 102,
             str_detect(stocklong, "Orange roughy Chatham Rise") ~ 162
           )) %>% 
  
  mutate(rgn_id_2 =
           case_when(
             str_detect(stocklong, paste(c("Bering Sea", "Japanese anchovy Pacific Coast", "Japanese sardine Pacific Coast", "Northwest Pacific", "Snow crab North Pacific"), collapse = "|")) ~ 73,
             str_detect(stocklong, "Sicily$") ~ 68,
             str_detect(stocklong, paste(c("Japanese flying squid", "Tsushima Strait", "East China Sea$", "Willowy flounder North Pacific"), collapse="|")) ~ 20,
             str_detect(stocklong, paste(c("Southern blue whiting Chile", "Southern hake Chile", "Patagonian Toothfish Chile"), collapse = "|")) ~ 172,
             str_detect(stocklong, ".*Northern Argentina$") ~ 173,
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 218,
             str_detect(stocklong, "^Black bellied angler") ~ 183,
             str_detect(stocklong, paste(c("^Megrim", "Roundnose grenadier", "Greenland halibut"), collapse = "|")) ~ 180,
             str_detect(stocklong, "Northern shrimp") ~ 223,
             str_detect(stocklong, "Striped marlin") ~ 137
           )) %>% 
  
  mutate(rgn_id_3 =
           case_when(
              str_detect(stocklong, "Sicily$") ~ 61,
              str_detect(stocklong, paste(c("Japanese flying squid", "Tsushima Strait","East China Sea$", "Willowy flounder North Pacific"), collapse="|")) ~ 209,
              str_detect(stocklong, "Snow crab North Pacific") ~ 20,
              str_detect(stocklong, ".*Northern Argentina$") ~ 171,
              str_detect(stocklong, "Patagonian Toothfish Chile") ~ 173,
              str_detect(stocklong, paste(c("Blue shark North Pacific", "Striped marlin"), collapse = "|")) ~ 135,
              str_detect(stocklong, paste(c("Roundnose grenadier"), collapse = "|")) ~ 179,
              str_detect(stocklong, "Greenland halibut") ~ 143,
              str_detect(stocklong, "Northern shrimp") ~ 175
           )) %>% 
  
  mutate(rgn_id_4 = 
           case_when(
             str_detect(stocklong, "Snow crab North Pacific") ~ 21,
             str_detect(stocklong, "Patagonian Toothfish Chile") ~ 95,
             str_detect(stocklong, paste(c("Blue shark North Pacific", "Striped marlin"), collapse = "|")) ~ 136,
             str_detect(stocklong, paste(c("Roundnose grenadier", "Greenland halibut"), collapse = "|")) ~ 141,
             str_detect(stocklong, "Northern shrimp") ~ 222
           )) %>% 
  mutate(rgn_id_5 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific", "Striped marlin"), collapse = "|")) ~ 134,
             str_detect(stocklong, "Greenland halibut") ~ 145
           )) %>% 
  mutate(rgn_id_6 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific", "Striped marlin"), collapse = "|")) ~ 131
           )) %>% 
  mutate(rgn_id_7 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific", "Striped marlin"), collapse = "|")) ~ 130
           )) %>% 
  mutate(rgn_id_8 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific", "Striped marlin"), collapse = "|")) ~ 129
           )) %>% 
  mutate(rgn_id_9 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific", "Striped marlin"), collapse = "|")) ~ 159
           )) %>% 
  mutate(rgn_id_10 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 150,
             str_detect(stocklong, "Striped marlin") ~ 132
           )) %>% 
  mutate(rgn_id_11 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 207,
             str_detect(stocklong, "Striped marlin") ~ 148
           )) %>% 
  mutate(rgn_id_12 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 206
           )) %>% 
  mutate(rgn_id_13 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 25
           )) %>% 
  mutate(rgn_id_14 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 14
           )) %>% 
  mutate(rgn_id_15 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 15
           )) %>% 
  mutate(rgn_id_16 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 8
           )) %>% 
  mutate(rgn_id_17 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 9
           )) %>% 
  mutate(rgn_id_18 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 13
           )) %>% 
  mutate(rgn_id_19 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 210
           )) %>% 
  mutate(rgn_id_20 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 12
           )) %>% 
  mutate(rgn_id_21 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 11
           )) %>% 
  mutate(rgn_id_22 =
           case_when(
             str_detect(stocklong, paste(c("Blue shark North Pacific"), collapse = "|")) ~ 73
           ))
             
DT::datatable(RAM,rownames = F)
```

## Tidy Final Datatable

### Gather FAO and OHI ids

```{r}

RAM_final <- RAM %>% 
  gather(fao, fao_values, contains("fao_id")) %>% 
  rename(fao_id = fao_values) %>% 
  select(-fao) %>% 
  gather(ohi, ohi_values, contains("rgn_id")) %>% 
  rename(rgn_id = ohi_values) %>% 
  select(-ohi) %>% 
  na.omit()


```

### Combine with full dataset

```{r}

RAM_rgns_new <- all_old %>% 
  full_join(RAM_final, by = c("rgn_id", "fao_id", "assessid", "stockid", "stocklong")) %>% 
  select(-RAM_area_m2)

write.csv(RAM_rgns_new, "int/RAM_fao_ohi_rgns.csv", row.names=FALSE)

```

