---
title: "Untitled"
author: "Iwen Su"
output: html_document
---


```{r}

## Libraries
library(dplyr)
library(tidyr)
library(readr)
library(sf)
library(ggplot2)
library(stringr)
library(here) 
setwd(here::here("globalprep","fis","v2018"))

source('../../../src/R/common.R')

```

## Check Newly Added Stocks
```{r}

## old stocks
load(file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2017/RAM v3.80/DB Files With Assessment Data/DBdata.RData"))

ram_bmsy_old <- data.frame(timeseries) %>%
  dplyr::filter(tsid == "BdivBmsytouse-dimensionless") %>%
  dplyr::filter(!is.na(tsvalue)) %>%
  dplyr::mutate(tsyear = as.numeric(as.character(tsyear))) %>%
  dplyr::filter(tsyear > 1979) %>%
  dplyr::select(assessid, year=tsyear, ram_bmsy = tsvalue, stockid, stocklong)

old_stocklong <- unique(ram_bmsy_old$stocklong)
old_stockid <- unique(ram_bmsy_old$stockid)
old_assessid <- unique(ram_bmsy_old$assessid)

## new stocks
load(file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2018/RAM v4.40 (6-4-18)/DB Files With Assessment Data/DBdata.RData"))

ram_bmsy <- data.frame(timeseries) %>%
  dplyr::filter(tsid == "BdivBmsytouse-dimensionless") %>%
  dplyr::filter(!is.na(tsvalue)) %>%
  dplyr::mutate(tsyear = as.numeric(as.character(tsyear))) %>%
  dplyr::filter(tsyear > 1979) %>%
  dplyr::select(assessid, year=tsyear, ram_bmsy = tsvalue, stockid, stocklong)

new_stocklong <- unique(ram_bmsy$stocklong)
new_stockid <- unique(ram_bmsy$stockid)
new_assessid <- unique(ram_bmsy$assessid)

## find diff
setdiff(new_stocklong, old_stocklong)
newStocks <- setdiff(new_stockid, old_stockid); length(newStocks)
setdiff(new_assessid, old_assessid)



```


Subset for just new stocks in 2018 RAM data - for some reason, some of the id prefixes specifying RAM regions are not found in the RLSADB v4.40 excel file
```{r}

ram_bmsy <- ram_bmsy %>% 
  filter(stockid %in% newStocks)

## some stockid values are repeated in more than one observation due to multiple assessid years
id_new <- ram_bmsy %>% 
  select(assessid, stockid, stocklong) %>% 
  distinct()

```


## Combine old RAM_fao_ohi_rgns with other stock information

Combine the data table Mel created `RAM_fao_ohi_rgns` that contains fao id, ohi id, and assessid from RAM with the old RAM data to match each row with the appropriate stockid and stocklong information
```{r}
## Grab last year's fao-ohi-assessid data table
RAM_fao_ohi_rgns <- read.csv("../v2017/int/RAM_fao_ohi_rgns.csv")

## create unique values of assessid, stock id, and stock long
id_old <- ram_bmsy_old %>% 
  select(assessid, stockid, stocklong) %>% 
  distinct()

## join tables
all_old <- RAM_fao_ohi_rgns %>% 
  full_join(id_old, by="assessid")

```

## Combine new stocks with old stock metadata

```{r}

RAM_rgns_new <- all_old %>% 
  full_join(id_new, by = c("assessid","stockid","stocklong"))

write.csv(RAM_rgns_new, "int/RAM_fao_ohi_rgns.csv", row.names=FALSE)
```

## Fill in OHI and FAO regions

Investigate what FAO/OHI region each stock belongs to: 
* Full intermediate stock data table [here](https://github.com/OHI-Science/ohiprep_v2018/blob/master/globalprep/fis/v2018/int/RAM_fao_ohi_rgns.csv)
* Primarily used www.fishsource.org to find stock distribution information. 
* Referenced FAO_AREAS (filtered for MAJOR) and ohi_fao_rgns shapefile from mazu on ArcGIS (can use other map visuals to identify stock extent). 
* Searched and tested key words in data table viewer (`DT::datatable`) to use in `str_detect`. See which stocks have the same FAO id assignment(s).

Some regions have more than 1 `fao_id` or `rgn_id` (e.g. Alaska skate Bering Sea and Aleutian Islands,	Kamchatka flounder Bering Sea and Aleutian Islands have `fao_id` 61 and 67)

```{r}

RAM_filt <- RAM_rgns_new %>% 
  filter(is.na(fao_id)) %>% 
  select(-RAM_area_m2)


RAM <- RAM_filt %>%
  ## Adding FAO region id
  mutate(fao_id  = 
           ifelse(str_detect(stocklong, paste(c("Bering Sea","Gulf of Alaska$"), collapse = "|")), 61, 
           ifelse(str_detect(stocklong, "Sicily$"), 37,
           ifelse(str_detect(stocklong, paste(c("^Jackass morwong Western","^Jackass morwong Eastern","Redfish Eastern Australia","Blue grenadier New South Wales"), collapse = "|")), 57,
           ifelse(str_detect(stocklong, paste(c("Scotian Shelf","Haddock","NAFO"), collapse = "|")), 21,
           ifelse(str_detect(stocklong, paste(c("Japan","Tsushima Strait"), collapse = "|")), 61,
           
           NA)))))) %>% 
  mutate(fao_id_2  = 
           ifelse(str_detect(stocklong, "Bering Sea"), 67, 
           ifelse(str_detect(stocklong,  paste(c("^Jackass morwong Eastern","Redfish Eastern Australia","Blue grenadier New South Wales"), collapse = "|")), 81,

           
           NA))) %>% 
  ## Adding OHI region id
  mutate(rgn_id  = 
           ifelse(str_detect(stocklong, paste(c("Bering Sea","Gulf of Alaska$"), collapse = "|")), 163,
           ifelse(str_detect(stocklong, "Sicily$"), 184,
           ifelse(str_detect(stocklong, "Australia$"), 16,
           ifelse(str_detect(stocklong, paste(c("Scotian Shelf","NAFO"), collapse = "|")), 218,
           ifelse(str_detect(stocklong, "^Haddock Georges"), 163,
           ifelse(str_detect(stocklong, "Japan"), 210,


           
           NA))))))) %>% 
  mutate(rgn_id2  = 
           ifelse(str_detect(stocklong, "Bering Sea"), 73, #russia
           ifelse(str_detect(stocklong, "Sicily$"), 68,
           ifelse(str_detect(stocklong, "Japanese flying squid"), 20,
           ifelse(str_detect(stocklong, paste(c("Japanese anchovy Pacific Coast", "Japanese sardine Pacific Coast"), collapse = "|")), 73, # extend to russia
                  

           
           NA))))) %>% 
  mutate(rgn_id3 = 
           ifelse(str_detect(stocklong, "Sicily$"), 61,
           ifelse(str_detect(stocklong, "Japanese flying squid"), 209, 
                  
           
           NA)))


DT::datatable(RAM,rownames = F)




```

