---
title: "OHI 2017 - Food Provision: Preparing RAM B/Bmsy data"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../src/templates/ohi_hdr.html'
pdf_document:
  toc: true
---

# Summary

This script prepares the RAM B/Bmsy data:
1. Relevant data are collected from the RAM database
2. Missing years are gapfilled when appropriate
3. RAM and Watson species names are harmonized in a few cases
4. RAM stocks are associated with the corresponding OHI and FAO regions


## Updates from previous assessment
This year we have additional stocks without spatial information from Christopher Free. We combined ohi and fao region id information to the additional stocks in `fao_ohi_rgns.Rmd` and saved the file in `int/RAM_fao_ohi_rgns.csv`. Watson species names should be similar if not the same as the SAUP species names used prior this the 2018 assessment (Iwen 8/30). Data describing the spatial boundaries of the RAM stocks were not previously available. We meant to gapfill RAM B/Bmsy values last year, but due to an oversight we did not use gapfilled data.

The updated RAM database has multiple assessments for the same stock. Looking into using `stockid` when preparing the data instead of `assessid`. 

***

# Data
## B/Bmsy values from stock assessments 

Reference: RAM Legacy Stock Assessment Database v4.41. 2018 Shared with us by folks at UW.  http://ramlegacy.org.

* Description: B/Bmsy value by stock and year (other data, which we do not use, are also available in the database)
* Native data resolution: stock (fish stock, species and region specific)
* Time range: 1950 - 2016
* Format: R data files (.rds)


## Stock range data 
Reference: Christopher M. Free. 2017. Mapping fish stock boundaries for the original Ram Myers stock-recruit database. https://marine.rutgers.edu/~cfree/mapping-fish-stock-boundaries-for-the-original-ram-myers-stock-recruit-database/. downloaded 9/25/2017.

* Downloaded: 8/20/2018
* Description: Shapefiles for each stock describing their distribution
* Native data resolution: Spatial shapefiles
* Format: Shapefiles

***

## Setup

``` {r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',message = FALSE, warning = FALSE)
```

```{r}
## highlight out when knitting
# setwd("globalprep/fis/v2017")

## Libraries
library(dplyr)
library(tidyr)
library(readr)
library(sf)
library(ggplot2)
library(here) 
setwd(here::here("globalprep","fis","v2018"))

source('../../../src/R/common.R')

## Paths for data
# path_raw_data = file.path(dir_M, "git-annex/globalprep/_raw_data/SAUP/d2017")
path_raw_data = file.path(dir_M, "git-annex/globalprep/fis/v2018/int/annual_catch")

```

## Obtain RAM B/Bmsy data

The data is stored as a relational database in an R object. Check that the names of each element have not changed from last year! Update as appropriate in the below list. 

The following tables are included (for full list, see loadDBdata.r in mazu):  

1. timeseries  
The time series data is a matrix object with the following headers/columns:  
(1) assessid (2) stockid (3) stocklong (4) tsid (5) tsyear (6) tsvalue

2. bioparams  
The time series data is a matrix object with the following headers/columns:  
(1) assessid (2) stockid (3) stocklong (4) bioid (5) biovalue (6) bioyear (7) bionotes  

3.  timeseries_values_views  
This stores the timeseries values, using the most recent assessment available, with timeseries type along the columns (stockid, stocklong, year, TBbest, TCbest, ERbest, BdivBmsypref, UdivUmsypref, BdivBmgtpref, UdivUmgtpref, TB, SSB, TN, R,
TC, TL, F, ER, TBdivTBmsy, SSBdivSSBmsy, FdivFmsy, ERdivERmsy, Btouse, Ctouse, Utouse, B/Bmsytouse, U/Umsytouse,
TB/TBmgt, SSB/SSBmgt, F/Fmgt, ER/ERmgt, B/Bmgttouse, U/Umgttouse) and stocks along the rows

4. timeseries_units_views  
This stores the timeseries units (or time series source for touse time series), with timeseries type
along the columns (TB, SSB, TN, R, TC, TL, F, ER) and stocks along the rows

5. timeseries_id_views  
This stores the timeseries ids with timeseries id along the columns (TB, SSB, TN, R,
TC, TL, F, ER, TB/TBmsy, SSB/SSBmsy, F/Fmsy, ER/ERmsy, Btouse, Ctouse, Utouse, B/Bmsytouse, U/Umsytouse,
TB/TBmgt, SSB/SSBmgt, F/Fmgt, ER/ERmgt, B/Bmgttouse, U/Umgttouse) and stocks along the rows

6. bioparams_values_views  
This stores the bioparams values, with bioparam type along the columns
(TBmsy, SSBmsy, Nmsy, MSY, Fmsy, ERmsy, TB0, SSB0, M, Bmsytouse, Umsytouse, TBmgt, SSBmgt, Fmgt, ERmgt,
Bmgttouse, Umgttouse) and stocks along the rows

7. bioparams_units_views  
This stores the bioparams units (or parameter source for touse parameters), with bioparam type
along the columns (TBmsy, SSBmsy, Nmsy, MSY, Fmsy, ERmsy, TB0, SSB0, M, TBmgt, SSBmgt, Fmgt, ERmgt) and
stocks along the rows

8. bioparams_ids_views  
This stores the bioparams ids, with bioparam id along the columns
(TBmsy, SSBmsy, Nmsy, MSY, Fmsy, ERmsy, TB0, SSB0, M, Bmsytouse, Umsytouse, TBmgt, SSBmgt, Fmgt, ERmgt,
Bmgttouse, Umgttouse) and stocks along the rows

9. metadata  
This stores assorted metadata associated with the stock, with datatypes along the columns
(assessid, stockid, stocklong, scientificname, FisheryType, region, areaid, areaname,
assessorid, mgmt, management authority) and stock by row

```{r load_data}

load(file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2018/RAM v4.40 (6-4-18)/DB Files With Assessment Data/DBdata.RData"))


ram_bmsy <- timeseries_values_views %>%
  dplyr::filter(!is.na(BdivBmsypref)) %>%
  dplyr::filter(year > 1979) %>%
  dplyr::select(stockid, year, ram_bmsy = BdivBmsypref)

```

## Adjust for Stocks with Multiple Assessments

For each stock with multiple assessmnets, use `assess_yr` to select the assessment with the largest number of data years. For those that don't have the largest range of data years, assign it an NA to remove.

1. Create a list of assessments that we want to keep (`assess_keep`)
2. Use that list to filter `ram_bmsy`
```{r}

# assess_keep <- ram_bmsy %>%
#   filter(assess_num > 1) %>% # ignore stocks with only one stock assessment 
#   select(assessid, stockid, assess_yr) %>% 
#   distinct() %>% 
#   group_by(stockid) %>% 
#   mutate(best_assess = ifelse(assess_yr == max(assess_yr), assessid, NA)) %>% 
#   ungroup()
# 
# ## Manually check that you correctly selected for the assessments with the largest range of data years (e.g. filter for "DSOLEGA" and "ALPLAICBSAI")
# DT::datatable(assess_keep)
# 
# library(validate)
# check_NA <- check_that(assess_keep, is.na(best_assess)) 
# summary(check_NA) # 118 of 231 items are NAs, so after na.omit should only have 113 values
# 
# assess_keep <- assess_keep %>% 
#   na.omit() # removes any row with NA values
# 
# dim(assess_keep) # great, only 113 values

```

Filter for just assessments we want
```{r}

# ram_bmsy <- ram_bmsy %>% 
#   filter(assessid %in% assess_keep$assessid) 
# 
# length(unique(ram_bmsy$assessid)) # check that after filter the number of unique assessid matches what you had above (113); great!
# length(unique(ram_bmsy$stockid)) # check length; 111
# 
# ## Looks like two stock ids have two assessments with the same number of data years, identify which ones
# multiple_assess <- ram_bmsy %>% 
#   group_by(stockid) %>% 
#   mutate(assess_num = length(unique(assessid))) %>% 
#   filter(assess_num > 1) %>% 
#   select(-year, -ram_bmsy) %>% 
#   distinct()

```

## Gapfill RAM data when there are missing years

For each stock: 
- Missing years are gapfilled using a linear regression model that includes data from 2001 to 2015 (2015 is the final year of Watson data). To be included in the gapfilling, there have to be 5 or more years of B/Bmsy data occuring over the last 11 years of data, from 2005 to 2015.
- We convert any predicted RAM B/Bmsy value less than the minimum observed B/Bmsy value to that the minimum observed value, as there are some negative predicted values.

Summary: 
- There are 389 RAM stocks with at least 5 years of B/Bmsy data from 2005 to 2015. 
- 205 of these stocks have at least 1 year of gapfilled data.  
- A few of the predicted B/Bmsy values go below zero. We convert anything with a RAM B/Bmsy value < 0.0026 to 0.0026, which is the minimum observed B/Bmsy value in the data.

```{r gapfill}

## gapfill ram_bmsy
ram_gf_check <- ram_bmsy %>%
  filter(year >= 2001) %>%
  spread(year, ram_bmsy) 
## based on this it seams reasonable to gap-fill missing values

ram_bmsy_gf <- ram_bmsy %>%
  filter(year >= 2001 & year <= 2015) %>%   # 2015 corresponds to the final year of Watson catch data
  group_by(stockid) %>%
  mutate(years_data = length(ram_bmsy[year >= 2005])) %>%
  ungroup() %>%
  filter(years_data >= 5) %>%
  spread(year, ram_bmsy) %>%
  gather("year", "ram_bmsy", -stockid, -years_data) %>%
  mutate(year = as.numeric(year)) %>%   
  mutate(gapfilled = NA) %>%
  mutate(gapfilled = ifelse(years_data == 11, gapfilled, paste(11 - years_data, "years gf", sep = " ")))

## see unique values of stocks
tmp <- ram_bmsy_gf %>%
  dplyr::select(stockid, gapfilled) %>%
  unique()

## check out gapfilling stats
length(tmp$gapfilled)   # 385 stocks with at least 5 years of data in past 11 years
table(tmp$gapfilled)    
sum(table(tmp$gapfilled))  # 234 stocks have at least one year of B/Bmsy values gapfilled

## regression model for prediction for each stock
ram_bmsy_gf <- ram_bmsy_gf %>%
  group_by(stockid) %>%
  do({
    mod <- lm(ram_bmsy ~ year, data=.)  
    ram_bmsy_predict <- predict(mod, newdata=.[c('year')])
    data.frame(., ram_bmsy_predict)
  }) %>%
  ungroup()

summary(ram_bmsy_gf) # a few of the predicted B/Bmsy values go below zero.  

## We convert anything with a RAM BBmsy value < 0.0026 to 0.0026, which is the minimum observed B/Bmsy value in the data; add method documentation
ram_bmsy_gf <- ram_bmsy_gf %>%
  mutate(ram_bmsy_predict = ifelse(ram_bmsy_predict < 0.0026, 0.0026, ram_bmsy_predict)) %>%
  mutate(method = ifelse(is.na(ram_bmsy), paste0("lm, ", gapfilled), NA)) %>%
  mutate(gapfilled = ifelse(is.na(ram_bmsy), "1", "0")) %>%
  mutate(ram_bmsy = ifelse(is.na(ram_bmsy), ram_bmsy_predict, ram_bmsy)) %>%
  dplyr::select(stockid, year, ram_bmsy, gapfilled, method) 

write.csv(ram_bmsy_gf, "int/ram_stock_bmsy_gf.csv", row.names=FALSE)

```

The following code can be used to estimate the amount of error for each year due to gapfilling. 
- In the first example, the gapfilling model is estimated using all the data prior to 2015. The model is then used to estimate the 2015 data and examine how accurate the estimates are. In this case, we assume all years of data prior to 2015 is available (i.e., only one missing year of data), which often isn't the case becasue stocks are often missing multiple years of data.
- The second example examines the error in the 2015 data if the B/Bmsy data only go to 2010 (i.e., 5 missing years of data - 2011 to 2015).

```{r gf error est}

# Example 1
ram_bmsy_gf_check1 <- ram_bmsy_gf %>%
  group_by(stockid) %>%
  do({
    mod <- lm(ram_bmsy ~ year, data= filter(., year < 2015))  ### change year to test different accuracy levels based on # years of data used in lm
    ram_bmsy_predict_error <- predict(mod, newdata=.[c('year')])
    data.frame(., ram_bmsy_predict_error)
  })

ggplot(data = filter(ram_bmsy_gf_check1, year==2015), aes(y=ram_bmsy_predict_error, x=ram_bmsy)) +
  geom_point() +
  theme_bw() +
  geom_abline(slope=1, intercept=0, color="red")

summary(lm(ram_bmsy_predict_error ~ ram_bmsy, data=filter(ram_bmsy_gf_check1, year==2015)))

# Example 2
ram_bmsy_gf_check2 <- ram_bmsy_gf %>%
  group_by(stockid) %>%
  do({
    mod <- lm(ram_bmsy ~ year, data= filter(., year <= 2010))  ### change year to test different accuracy levels based on # years of data used in lm
    ram_bmsy_predict_error <- predict(mod, newdata=.[c('year')])
    data.frame(., ram_bmsy_predict_error)
  })

plot <- ggplot(data = filter(ram_bmsy_gf_check2, year==2015), aes(y=log(ram_bmsy_predict_error+1), x=log(ram_bmsy+1), col = stockid)) +
  geom_point() +
  theme_bw() +
  geom_abline(slope=1, intercept=0, color="red")

library(plotly)
ggplotly(plot)

summary(lm(ram_bmsy_predict_error ~ ram_bmsy, data=filter(ram_bmsy_gf_check2, year==2015)))

```

## Standardize species names

In most cases, the RAM and Watson data use the same species names, but there are a few exceptions.  The following code identifies species in the RAM data that are not in the Watson data. In these cases, different species names may be used (although not necessarily because some of the species may be present in RAM, but not Watson, for other reasons). For these species, I used fishbase to explore synonyms and create a table to harmonize the RAM species names with the Watson species names (saved as: int/RAM_species_to_Watson.csv).

```{r name harmonize}

# get list of RAM species, scientific name
ram_sp <- ram_bmsy_gf %>%
  left_join(data.frame(metadata), by = "stockid") %>%
  dplyr::select(scientificname) %>%
  unique() %>%
  arrange(scientificname)

# Watson species, sci name (read in the datatable that includes TaxonKey)
wat_sp <- read_csv(file.path(dir_M,'git-annex/globalprep/fis/v2018/int/stock_catch_by_rgn_taxa.csv')) %>% 
  dplyr::rename(wat_scientificname = TaxonName) %>%
  dplyr::select(wat_scientificname) %>%
  unique() %>%
  arrange(wat_scientificname)


# compare names - what's in RAM that's not in Watson (only 6 species differing)
tmp <- sort(setdiff(ram_sp$scientificname, wat_sp$wat_scientificname))
write.csv(tmp, "int/unmatched_RAM_species.csv", row.names=FALSE)

# Then I hand-looked up each of the missing ones to generate this list - most still unmatched
ram_name_corr <- read.csv("int/RAM_species_to_Watson.csv", stringsAsFactors = FALSE) %>%
   filter(!is.na(Watson_species))  # Watson to RAM name conversion

ram_name_corr # matched species, only 6

```


## Identify FAO and OHI regions for RAM stocks

Identify the FAO/OHI regions where each RAM stock is located. This step is completed in [fao_ohi_rgns.Rmd](https://mazu.nceas.ucsb.edu/rstudio/files/github/ohiprep_v2018/globalprep/fis/v2018/fao_ohi_rgns.html).



## Combine all the data

### Check Differences

Final formatting of the data is performed. If there are many differences between RAM spatial file and RAM metadata, check the fao_ohi_rgns.Rmd prep again.

```{r final}

## Read in RAM spatial stocks file
ram_spatial <- read.csv("int/RAM_fao_ohi_rgns.csv", stringsAsFactors = FALSE)

ram_meta <- metadata %>% 
  dplyr::select(assessid, stockid, scientificname)

setdiff(ram_spatial$stockid, ram_meta$stockid) # make sure all the spatial data has corresponding metadata (should be 0); 9 diff

```

### Investigate Differences

Complete in-depth investigation in data_check.Rmd

* BIGEYEWPO
* MORWONGSE
* PCODBSAI
* BLACKROCKSPCOAST - from OLD  RAM 3.80, outdated stock
* BLACKROCKNPCOAST - from OLD RAM 3.80, outdated stock
* OROUGHYCR
* STMARLINNPAC
* HOGFISHEGM
* HOGFISHSEFL

```{r}

## Check that outdated stocks are only in OLD RAM 3.80
RAM_old <- read.csv("../v2017/int/RAM_fao_ohi_rgns.csv")
DT::datatable(RAM_old) # serach stocks

## Checked newly added stocks 


```


### Combine
```{r}
# join with metadata
ram_spatial <- ram_spatial %>%
  left_join(ram_meta, by = "assessid") %>%
  rename(RAM_species = scientificname)

# correct names in a few cases to match with Watson names
ram_name_corr <- read.csv("int/RAM_species_to_Watson.csv", stringsAsFactors = FALSE) %>%
  filter(!is.na(Watson_species))  # Watson to RAM name conversion


ram_spatial <- ram_spatial %>%
  left_join(ram_name_corr, by="RAM_species") %>%
  dplyr::mutate(species = ifelse(!is.na(SAUP_species), SAUP_species, RAM_species)) %>%
  dplyr::select(assessid, stockid_ram = stockid, rgn_id, fao_id, species, RAM_area_m2)

length(unique(ram_spatial$stockid_ram)) # 257 RAM stocks with B/Bmsy data

## filter out the regions that are not in an eez
ram_spatial <- filter(ram_spatial, rgn_id<250)
length(unique(ram_spatial$stockid_ram))    # 256 RAM stocks with B/Bmsy data and are found in EEZ regions


## Combine RAM spatial data with B/Bmsy data

ram_bmsy <- read.csv("int/ram_stock_bmsy_gf.csv")

ram_data <- ram_spatial %>% 
  left_join(ram_bmsy_gf, by="assessid") %>%
  dplyr::mutate(stock_id = paste(species, fao_id, sep="-")) %>%
  dplyr::mutate(stock_id = gsub(" ", "_", stock_id)) %>%
  dplyr::filter(!is.na(year)) %>%   ## these are the ones that didn't have enough ram data to make an accurate guess
  ungroup() %>%
  dplyr::select(rgn_id, stock_id, assessid, year, RAM_area_m2, ram_bmsy, gapfilled, method) %>%
  unique()

write.csv(ram_data, "int/ram_bmsy.csv", row.names=FALSE)


```
