---
title: "OHI 2018: Food Provision/Fisheries, Preparing RAM B/Bmsy data"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../src/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: console
---

# Summary

This script prepares the RAM B/Bmsy data:
1. Relevant data are collected from the RAM database
2. Missing years are gapfilled when appropriate
3. RAM and Watson species names are harmonized in a few cases
4. RAM stocks are associated with the corresponding OHI and FAO regions


## Updates from previous assessment

After Jamie consulted with Daniel Hively, the research scientist currently working on the RAM Legacy Database, we decided it was best to use the `timeseries_values_views` table instead of the `timeseries` table. The former contains data using the most recent year of assessments for each stock. The original `timeseries` table now contains multiple assessments for the same stock. Since this table didn't have an `assessid` we shifted to using `stockid` to join the relational databases. Furthermore, to extract the previous `BdivBmsytouse-dimensionless` values from `timeseries`, we now combined `TBdivTBmsy` and `SSBdivSSBmsy` values based on the definition of the `BdivBmsytouse-dimensionless` in the `tsmetrics` table: **"General biomass time series relative to msy reference point (TB/TBmsy, or SSB/SSBmsy otherwise)"**.

This year we have additional stocks without spatial information from Christopher Free (2017). We manually assigned ohi and fao region id information to the additional stocks in `fao_ohi_rgns.Rmd` using best available information on stock distribution and saved the file in `int/RAM_fao_ohi_rgns.csv`. Watson species names should be similar if not the same as the SAUP species names used prior this the 2018 assessment. 


***

# Data
## B/Bmsy values from stock assessments 

Reference: [RAM Legacy Stock Assessment Database](http://ramlegacy.org) v4.40 (6-4-2018) Shared with us by folks at UW.

* Description: B/Bmsy value by stock and year (other data, which we do not use, are also available in the database)
* Native data resolution: stock (fish stock, species and region specific)
* Time range: 1950 - 2016
* Format: R data files (.rds)


## Stock range data 
Reference: Christopher M. Free. 2017. Mapping fish stock boundaries for the original Ram Myers stock-recruit database. https://marine.rutgers.edu/~cfree/mapping-fish-stock-boundaries-for-the-original-ram-myers-stock-recruit-database/. downloaded 9/25/2017.

* Downloaded: 8/20/2018
* Description: Shapefiles for each stock describing their distribution
* Native data resolution: Spatial shapefiles
* Format: Shapefiles

***

## Setup

``` {r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',message = FALSE, warning = FALSE)
```

```{r, eval = FALSE}

## Libraries
library(dplyr)
library(tidyr)
library(readr)
library(sf)
library(ggplot2)
library(here) 

## highlight out when knitting
#setwd(here::here("globalprep","fis","v2018"))

source('../../../src/R/common.R')

## Paths for data
path_raw_data = file.path(dir_M, "git-annex/globalprep/fis/v2018/int/annual_catch")

```

## Obtain RAM B/Bmsy data

The data is stored as a relational database in an R object. Check that the names of each element have not changed from last year! Update as appropriate in the below list. 

The following tables are included (for full list, see loadDBdata.r in mazu):  

1. timeseries  
The time series data is a data frame containing all assessments conducted per stock with the following headers/columns:  
(1) assessid (2) stockid (3) stocklong (4) tsid (5) tsyear (6) tsvalue

2. bioparams  
The time series data is a data frame with the following headers/columns:  
(1) assessid (2) stockid (3) stocklong (4) bioid (5) biovalue (6) bioyear (7) bionotes  

3.  timeseries_values_views  
This stores the timeseries values, using the most recent assessment available, with timeseries type along the columns (stockid, stocklong, year, TBbest, TCbest, ERbest, BdivBmsypref, UdivUmsypref, BdivBmgtpref, UdivUmgtpref, TB, SSB, TN, R,
TC, TL, F, ER, TBdivTBmsy, SSBdivSSBmsy, FdivFmsy, ERdivERmsy, Btouse, Ctouse, Utouse, B/Bmsytouse, U/Umsytouse,
TB/TBmgt, SSB/SSBmgt, F/Fmgt, ER/ERmgt, B/Bmgttouse, U/Umgttouse) and stocks along the rows

4. timeseries_units_views  
This stores the timeseries units (or time series source for touse time series), with timeseries type
along the columns (TB, SSB, TN, R, TC, TL, F, ER) and stocks along the rows

5. timeseries_id_views  
This stores the timeseries ids with timeseries id along the columns (TB, SSB, TN, R,
TC, TL, F, ER, TB/TBmsy, SSB/SSBmsy, F/Fmsy, ER/ERmsy, Btouse, Ctouse, Utouse, B/Bmsytouse, U/Umsytouse,
TB/TBmgt, SSB/SSBmgt, F/Fmgt, ER/ERmgt, B/Bmgttouse, U/Umgttouse) and stocks along the rows

6. bioparams_values_views  
This stores the bioparams values, with bioparam type along the columns
(TBmsy, SSBmsy, Nmsy, MSY, Fmsy, ERmsy, TB0, SSB0, M, Bmsytouse, Umsytouse, TBmgt, SSBmgt, Fmgt, ERmgt,
Bmgttouse, Umgttouse) and stocks along the rows

7. bioparams_units_views  
This stores the bioparams units (or parameter source for touse parameters), with bioparam type
along the columns (TBmsy, SSBmsy, Nmsy, MSY, Fmsy, ERmsy, TB0, SSB0, M, TBmgt, SSBmgt, Fmgt, ERmgt) and
stocks along the rows

8. bioparams_ids_views  
This stores the bioparams ids, with bioparam id along the columns
(TBmsy, SSBmsy, Nmsy, MSY, Fmsy, ERmsy, TB0, SSB0, M, Bmsytouse, Umsytouse, TBmgt, SSBmgt, Fmgt, ERmgt,
Bmgttouse, Umgttouse) and stocks along the rows

9. metadata  
This stores assorted metadata associated with the stock, with datatypes along the columns
(assessid, stockid, stocklong, scientificname, FisheryType, region, areaid, areaname,
assessorid, mgmt, management authority) and stock by row

10. tsmetrics
Contains metadata, such as definitions and units, of `tsid` values in `timeseries` table.


For this data prep we primarily use and consult `timeseries_values_views`, `tsmetrics`, and `metadata`
```{r load_data, eval = FALSE}

load(file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2018/RAM v4.40 (6-4-18)/DB Files With Assessment Data/DBdata.RData"))

ram_bmsy <- timeseries_values_views %>%
  select(stockid, year, TBdivTBmsy, SSBdivSSBmsy) %>% 
  mutate(ram_bmsy = ifelse(!is.na(TBdivTBmsy), TBdivTBmsy, SSBdivSSBmsy)) %>% 
  dplyr::filter(year > 1979) %>%
  filter(!is.na(ram_bmsy)) %>% 
  dplyr::select(stockid, year, ram_bmsy)

```

For future reference: Consider using management B/Bmsy values in next year's assessment.

Note: In the past, we have used the B/Bmsy to use, dimensionless (BdivBmsytouse-dimensionless) values in the `timeseries_values_views` data table. According to `tsmetrics` table, this consists of TB/TBmsy (TBdivTBmsy) otherwise if NA, SSB/SSBmsy (SSBdivSSBmsy). B/Bmsy preferred values (BdivBmsypref) uses B/Bmsy (BdivBmsytouse) preferably, but if NA, then uses B/Bmgt (BdivBmgttouse) values as a substitute. The below table filters for the stocks that use BdivBmgttouse values, not currently incorporated in OHI Global from RAM. Any missing fisheries stocks are currently gapfilled instead. Investigate use of these B/Bmgt values next year. Many of them are around Iceland or European stocks.
```{r, eval = FALSE}

## Filter for rows where there is no B/Bmsy to use (TB/Tbmsy or SSB/SSBmsy) values.
## For B/Bmsypref values, when B/Bmsy is NA, then it takes B/Bmgt values instead
test <- timeseries_values_views %>% 
  filter(!is.na(BdivBmsypref) & is.na(SSBdivSSBmsy) & is.na(TBdivTBmsy))

## All these stocks are those that have B/Bmgt values
unique(test$stockid)

## Save into a data frame
species_list <- test %>% 
  select(stockid,stocklong) %>% 
  unique() #96 unique stocks that don't currently get used in the OHI GLobal. Consider using next year

write.csv(species_list, "int/RAM_addl_mgt_stocks.csv", row.names = FALSE)

```

## Gapfill RAM data when there are missing years

For each stock: 
- Missing years are gapfilled using a linear regression model that includes data from 2001 to 2015 (2015 is the final year of Watson data). To be included in the gapfilling, there have to be 5 or more years of B/Bmsy data occuring over the last 11 years of data, from 2005 to 2015.
- We convert any predicted RAM B/Bmsy value less than the minimum observed B/Bmsy value to that the minimum observed value, as there are some negative predicted values.

Summary: 
- There are 389 RAM stocks with at least 5 years of B/Bmsy data from 2005 to 2015. 
- 205 of these stocks have at least 1 year of gapfilled data.  
- A few of the predicted B/Bmsy values go below zero. We convert anything with a RAM B/Bmsy value < 0.0026 to 0.0026, which is the minimum observed B/Bmsy value in the data.

```{r gapfill, eval = FALSE}

## gapfill ram_bmsy
ram_gf_check <- ram_bmsy %>%
  filter(year >= 2001) %>%
  spread(year, ram_bmsy) 
## based on this it seams reasonable to gap-fill missing values

# identify stocks for gapfilling (those with 5 or more years of data since 2005).
# NOTE: we potentially gapfill to 2001, but we want stocks with adequate *recent* data 
ram_bmsy_gf <- ram_bmsy %>%
  filter(year >= 2001 & year <= 2015) %>%   # 2015 corresponds to the final year of Watson catch data
  group_by(stockid) %>%
  mutate(years_data_2005_now = length(ram_bmsy[year >= 2005])) %>%
  mutate(years_data_2001_now = length(ram_bmsy[year >= 2001])) %>%
  ungroup() %>%
  filter(years_data_2005_now >= 5)

## Get rows for stocks/years with no B/Bmsy (identified as NA B/Bmsy value for now)
ram_bmsy_gf <- ram_bmsy_gf %>%
  spread(year, ram_bmsy) %>% 
  gather("year", "ram_bmsy", -stockid, -years_data_2005_now, -years_data_2001_now) %>%
  mutate(year = as.numeric(year)) 

## gapfilling record keeping
ram_bmsy_gf <- ram_bmsy_gf %>%   
  mutate(gapfilled = NA) %>%
  mutate(gapfilled = ifelse(years_data_2001_now == 15, gapfilled, 
                            paste(15 - years_data_2001_now, "years gf", sep = " ")))

## see unique values of stocks
tmp <- ram_bmsy_gf %>%
  dplyr::select(stockid, gapfilled) %>%
  unique()

## check out gapfilling stats
length(tmp$gapfilled)   # 296 stocks with at least 5 years of data in past 11 years
table(tmp$gapfilled)    
sum(table(tmp$gapfilled))  # 216 stocks have at least one year of B/Bmsy values gapfilled

## regression model for prediction for each stock
ram_bmsy_gf <- ram_bmsy_gf %>%
  group_by(stockid) %>%
  do({
    mod <- lm(ram_bmsy ~ year, data=.)  
    ram_bmsy_predict <- predict(mod, newdata=.[c('year')])
    data.frame(., ram_bmsy_predict)
  }) %>%
  ungroup()

summary(ram_bmsy_gf) # a few of the predicted B/Bmsy values go below zero.  

## We convert anything with a RAM BBmsy value < 0.0026 to 0.0026, which is the minimum observed B/Bmsy value in the data; add method documentation
ram_bmsy_gf <- ram_bmsy_gf %>%
  mutate(ram_bmsy_predict = ifelse(ram_bmsy_predict < 0.0026, 0.0026, ram_bmsy_predict)) 

## gapfilling record keeping
ram_bmsy_gf_final <- ram_bmsy_gf %>%
  mutate(method = ifelse(is.na(ram_bmsy), paste0("lm, ", gapfilled), NA)) %>%
  mutate(gapfilled = ifelse(is.na(ram_bmsy), "1", "0")) %>%
  mutate(ram_bmsy = ifelse(is.na(ram_bmsy), ram_bmsy_predict, ram_bmsy)) %>%
  dplyr::select(stockid, year, ram_bmsy, gapfilled, method) 

write.csv(ram_bmsy_gf_final, "int/ram_stock_bmsy_gf.csv", row.names=FALSE)

```

Get a general idea of how well the model predicts missing data based on observed and model predicted values.  This model appears to do fairly well. 
```{r gf error est, eval=FALSE}

plot(ram_bmsy_gf$ram_bmsy, ram_bmsy_gf$ram_bmsy_predict)
abline(0,1, col="red")

plot(log(ram_bmsy_gf$ram_bmsy), log(ram_bmsy_gf$ram_bmsy_predict))
abline(0,1, col="red")

mod <- lm(ram_bmsy ~ ram_bmsy_predict, data=ram_bmsy_gf)
summary(mod)
```

## Standardize species names

In most cases, the RAM and Watson data use the same species names, but there are a few exceptions.  The following code identifies species in the RAM data that are not in the Watson data. In these cases, different species names may be used (although not necessarily because some of the species may be present in RAM, but not Watson, for other reasons). For these species, I used fishbase to explore synonyms and create a table to harmonize the RAM species names with the Watson species names (saved as: int/RAM_species_to_Watson.csv).

```{r name harmonize, eval = FALSE}

# get list of RAM species, scientific name
ram_sp <- ram_bmsy_gf_final %>%
  left_join(data.frame(metadata), by = "stockid") %>%
  dplyr::select(scientificname) %>%
  unique() %>%
  arrange(scientificname)

# Watson species, sci name (read in the datatable that includes TaxonKey)
wat_sp <- read_csv(file.path(dir_M,'git-annex/globalprep/fis/v2018/int/stock_catch_by_rgn_taxa.csv')) %>% 
  dplyr::rename(wat_scientificname = TaxonName) %>%
  dplyr::select(wat_scientificname) %>%
  unique() %>%
  arrange(wat_scientificname)

# compare names - what's in RAM that's not in Watson
tmp <- sort(setdiff(ram_sp$scientificname, wat_sp$wat_scientificname))
tmp2 <- sort(setdiff(wat_sp$wat_scientificname, ram_sp$scientificname))

write.csv(tmp, "int/unmatched_RAM_species.csv", row.names=FALSE)
write.csv(tmp2, "int/Watson_species_no_RAM.csv", row.names=FALSE)

# Then I hand-looked up each of the missing ones to generate this list - most still unmatched
ram_name_corr <- read.csv("int/RAM_species_to_Watson.csv", stringsAsFactors = FALSE) %>%
   filter(!is.na(Watson_species))  # Watson to RAM name conversion

ram_name_corr # matched species, only 6

```


## Identify FAO and OHI regions for RAM stocks

Identify the FAO/OHI regions where each RAM stock is located (fao and ohi regions are assigned to RAM Data in [fao_ohi_rgns.Rmd](https://mazu.nceas.ucsb.edu/rstudio/files/github/ohiprep_v2018/globalprep/fis/v2018/fao_ohi_rgns.html).

If there are many differences between RAM spatial file and RAM metadata, check the fao_ohi_rgns.Rmd prep again. 


```{r final, eval = FALSE}

## Read in RAM spatial stocks file
ram_spatial <- read.csv("int/RAM_fao_ohi_rgns.csv", stringsAsFactors = FALSE)

ram_meta <- metadata %>% 
  dplyr::select(stockid, stocklong, scientificname)

setdiff(ram_spatial$stockid, ram_meta$stockid) # make sure all the spatial data has corresponding metadata (should be 0)

# join with metadata to get scientific name
ram_spatial <- ram_spatial %>%
  select(-stocklong) %>%
  left_join(ram_meta, by = c("stockid")) %>%
  rename(RAM_species = scientificname)

```



### Final formatting

Harmonize names between RAM and Watson data.
```{r, eval=FALSE}

# correct names in a few cases to match with Watson names
ram_name_corr <- read.csv("int/RAM_species_to_Watson.csv", stringsAsFactors = FALSE) %>%
  filter(!is.na(Watson_species))  # Watson to RAM name conversion

ram_spatial <- ram_spatial %>%
  left_join(ram_name_corr, by="RAM_species") %>%
  dplyr::mutate(species = ifelse(!is.na(Watson_species), Watson_species, RAM_species)) %>%
  dplyr::select(rgn_id, fao_id, stockid, stocklong, species, RAM_area_m2)

length(unique(ram_spatial$stockid)) # 345 RAM stocks with B/Bmsy data

## filter out the regions that are not in an eez
# ram_spatial <- filter(ram_spatial, rgn_id < 250)
# length(unique(ram_spatial$stockid))
```


Re-name `stockid` column to `stockid_ram` and create new column `stockid` that matches with the `stockid` column in the CMSY data table prepared in [calculate_bbmsy.Rmd](https://cdn.rawgit.com/OHI-Science/ohiprep_v2018/master/globalprep/fis/v2018/calculate_bbmsy.html).

```{r, eval = FALSE}
## Combine RAM spatial data with B/Bmsy data
ram_bmsy_gf <- read.csv("int/ram_stock_bmsy_gf.csv")

# check every stock has a location:
setdiff(ram_bmsy_gf$stockid, ram_spatial$stockid) # should be 0: every ram stock should have ohi/fao rgn
setdiff(ram_spatial$stockid, ram_bmsy_gf$stockid) # these are stocks that were dropped due to insufficient years of data

ram_data <- ram_bmsy_gf %>% 
  left_join(ram_spatial, by="stockid") %>%
  rename(stockid_ram = stockid) %>% 
  dplyr::mutate(stockid = paste(species, fao_id, sep="-")) %>%
  dplyr::mutate(stockid = gsub(" ", "_", stockid)) %>%
  dplyr::select(rgn_id, stockid, stockid_ram, stocklong, year, RAM_area_m2, ram_bmsy, gapfilled, method) %>%
  unique()

write.csv(ram_data, "int/ram_bmsy.csv", row.names=FALSE)

```
