---
title: "OHI 2019: Food Provision, Fix Cell Issues"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: console
---

# Summary

From Jamie's clean_cells.R script in v2017. This script fixes the issue of some half degree cells that slighlty overlap the OHI regions shapefile, leaving them with a proportional area less than 1. This would cause us to lose catch when assigning catch to cells. To fix this, we define a vector of cellids that have a proportionArea <1 and are NOT duplicated (i.e. the other portion of the area missing is not accounted for) and assign a proportionArea of 1 to these cells.

Also takes code from [saup_rasters_to_ohi_rgns.R](https://github.com/OHI-Science/ohiprep_v2019/globalprep/fis/v2016/saup_rasters_to_ohi_rgns.R) to re-write fao and ohi regions with cell information.

The output is saved as int/cells.csv

# Setup 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```

```{r}
#library(reshape2)
library(tidyverse)
#library(seaaroundus)
library(rgdal)
library(raster)
library(here)
library(readxl)

setwd(here::here("globalprep/fis/v2019"))

source('../../../workflow/R/common.R')


IMAS_d2019 <- file.path(dir_M, "git-annex/globalprep/_raw_data/IMAS_GlobalFisheriesLandings/d2019")
```

# Methods

## Set up a watson cell ID raster

Metadata for Watson 2019 Cells data [here](http://metadata.imas.utas.edu.au/geonetwork/srv/eng/metadata.show). The Cells data is within Codes.xlsx in a sheet named "Cell".
```{r}
#create a raster of Cell numbers
## This Codes.xlsx was downloaded from the same place as the raw Watson data.
cells <- read_excel(file.path(IMAS_d2019, "Codes.xlsx"), sheet = "Cell") %>%
  dplyr::rename(x = LonCentre,  y = LatCentre, z = Cell) %>% #I renamed these xyz to use in the rasterFromXYZ() function below
  dplyr::select(x,y,z)

#turn the lat/long points into a raster
cells_raster <- rasterFromXYZ(cells)

crs(cells_raster) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
crs(cells_raster)
#check out the cells raster to make sure it looks good
plot(cells_raster)

```

## Match Cell IDs with OHI regions

Script taken from [v2016 saup_rasters_to_ohi_rgns.R script](https://github.com/OHI-Science/ohiprep_v2019/globalprep/fis/v2016/saup_rasters_to_ohi_rgns.R). 

Result of `ohi_to_wats_raster` will be a list of length 526 (the same number of rows in the shapefile **regions**). Each element in the list will be associated with at least one cell number in a range from 1 to `ncell(watson_rast)` or 259,200. 

Note: When using spatial polygons (e.g. **regions**) in `extract`, the extract method returns the values of the cells of a Raster object (e.g. **cells_raster**) that are covered by a polygon. A cell is covered if its center is inside the polygon, but setting `weights = TRUE` and `normalizeWeights = FALSE` accounts from the fraction of each cell that is covered by each polygon.

Extract values to create raster (this may take a while)
```{r}

## associate each cell value with ohi and fao regions
ohi_to_wats_raster <- raster::extract(cells_raster, regions, weights = TRUE, normalizeWeights = FALSE, progress = 'text')

## assign rgn type and id (e.g. eez_100) to each element in the list
names(ohi_to_wats_raster) <- paste(regions$type_w_ant, regions$rgn_ant_id, sep="_")

## combine all 526 elements into a data frame with col names '.id', 'value', and 'weight'
region_prop_df <- plyr::ldply(ohi_to_wats_raster, rbind)

## splits up rgn type and id, rename col
cells_raw <- region_prop_df %>%
  separate(.id, c('rgn_typ', 'rgn_id'), sep = '_') %>%
  rename(wat_cell_id = value, 
         proportionArea = weight)

## save all the cell associations
write.csv(cells_raw, file.path(dir_M, "git-annex/globalprep/fis/v2019/raw/watson_rasters_to_ohi_rgns.csv"), row.names=FALSE)

```

### Compare OHI-CellID Matching 

Check this year's ohi-cellID match with last year's to see what changed.
```{r}
## Check raw file
cells_raw <- read.csv(file.path(dir_M, "git-annex/globalprep/fis/v2019/raw/watson_rasters_to_ohi_rgns.csv")) %>%
  arrange(rgn_id)

old_cells_raw <- read.csv(file.path(dir_M, "git-annex/globalprep/fis/v2018/raw/watson_rasters_to_ohi_rgns.csv")) %>% 
  arrange(rgn_id)

# new all_df has 446 more rows
dim(cells_raw) # 282519 rows
dim(old_cells_raw) # 291209 rows
length(setdiff(cells_raw$wat_cell_id, old_cells_raw$saup_cell_id)) #149929 differences

```

## A few modifications to the data

### Merge land and eez areas
Originally, we included only water areas, but some rasters were excluded because the cells raster doesn't exactly align with our region map.

```{r}

ohi_cells <- read.csv(file.path(dir_M, "git-annex/globalprep/fis/v2019/raw/watson_rasters_to_ohi_rgns.csv")) %>%
  rename(CellID = wat_cell_id) %>%
  group_by(CellID, rgn_id) %>%   # groups land and eez data (that way the cell catch is fully applied to the region...rather than cutting the portion that overlaps land) 
  dplyr::summarise(area = sum(proportionArea)) %>%
  mutate(area = ifelse(area > 1, 1, area))%>%  # this corrects when there is land/eez overlap within the same region resulting in cell area >1
  ungroup()  

```


### Check Cell Proportions

Calculate total proportion coverage of each cell, which should add up to 1.  When the total area doesn't equal 1, it means there is overlap of polygons and a portion of the cell gets assigned to multiple regions. 

There is a bit of weirdness, but nothing that seems overly concerning to me.  The number of cells with this issue is fairly small, and I believe many of the biggest offenders are land discrepancies.
```{r}

bad_ones <- ohi_cells %>%
  group_by(CellID) %>%
  mutate(total_area = sum(area)) %>%
  filter(total_area > 1) %>%
  arrange(CellID) %>%
  arrange(-total_area)

data.frame(bad_ones)
table(bad_ones$rgn_id)
hist(bad_ones$total_area)

```


### Correct for cells with proportion <1
The opposite issue is when a cell has only partial coverage (i.e., proportion is <1).  In these cases, we will adjust to equal 1.  

Get a list of cells that are not duplicated and have areas <1, and adjust these to equal 1. These are the cells that were cut off prematurely due to edge effects, etc. If area is <1 and the cell is not duplicated, the area will be converted to 1.

```{r}
## ID singleton cells:
dups <- ohi_cells$CellID[duplicated(ohi_cells$CellID)]
nodups <- ohi_cells$CellID[-which(ohi_cells$CellID %in% dups)]
length(unique(ohi_cells$CellID))
length(nodups)
length(dups)

tmp <- ohi_cells %>% 
  filter(CellID %in% nodups & area < 1)
head(tmp)
sum(duplicated(tmp$CellID))

ohi_cells <- ohi_cells %>%
  mutate(area = ifelse(CellID %in% nodups & area < 1, 1, area))

write.csv(ohi_cells, file.path(dir_M, "git-annex/globalprep/fis/v2019/raw/watson_rasters_to_ohi_rgns_v2.csv"), 
          row.names=FALSE)

```


## Match Cell IDs with FAO Regions

Similar steps were used for extracting cell values for FAO regions as the OHI shapefile. This is already done in [saup_rasters_to_ohi_rgns.R](https://github.com/OHI-Science/ohiprep_v2019/globalprep/fis/v2016/saup_rasters_to_ohi_rgns.R)

Since data source is the same, just read in the FAO raster - check with the new dataframe I created above.
```{r}

fao_regions <- readOGR(dsn = file.path(dir_M, "git-annex/Global/NCEAS-Regions_v2014/orig/FAO_AREAS"), layer="FAO_AREAS")  
fao_regions <- fao_regions[fao_regions$F_LEVEL == "MAJOR", ]
plot(fao_regions)

```

*** 

## Final Wrangling of Cells

Read in the dataset matching each cell to an FAO region (need both OHI and FAO region for analysis)

Note: Currently CellID in `ohi_cells` range from 3210 to 252062. CellID in `fao_cells` ranges from 1 to 252,083. There are many OHI regions with cells that are not assigned FAO regions (likely due to being along the land-ocean boundary). It's okay to ignore, very small percentage.
```{r}

fao_cells <- read.csv(file.path(dir_M, "git-annex/globalprep/fis/v2015/raw/saup_rasters_to_fao_rgns.csv")) %>%
  rename(CellID = saup_cell_id)
sum(duplicated(fao_cells$CellID)) # good: no duplicates

ohi_cells <- read.csv(file.path(dir_M, "git-annex/globalprep/fis/v2019/raw/watson_rasters_to_ohi_rgns_v2.csv")) 

ohi_fao_cells <- ohi_cells %>%
  full_join(fao_cells, by="CellID") %>%
  arrange(CellID) %>%
  mutate(area = ifelse(is.na(area), 1, area))  # these are fao regions with no corresponding ohi


summary(ohi_fao_cells) # 84,218 NAs in fao_id - v2018
                       # 758 NAs in fao_id - v2019; probably due changes in watson data

fao_id_na <- ohi_fao_cells %>%
  filter(is.na(fao_id))

```

From `summary(ohi_fao_cells)` we see that there are 758 NAs in the `fao_id` field. The corresponding `ocean` field is also an NA.

One issue of concern: Some regions aren't assigned an FAO region value (in places where land > 50% of cell cover)...and we need both for the analysis.  Previous analyses indicate that most of these are land issues.  So not worrying about for now, but might want to do gapfilling in the future.
```{r}

## see what percentage have no FAO region assigned
sum(is.na(ohi_fao_cells$fao_id))/max(ohi_fao_cells$CellID, na.rm=TRUE) # 758/252,083 or 0.3% is NA

```

***

## Save File
```{r} 
write.csv(ohi_fao_cells, file = "int/cells.csv", row.names=FALSE)
```
***

## Check Output

Compare this year's cells.csv with last year's cells.csv output
```{r}

## new cells output
cells <- read.csv('int/cells.csv')  %>%
  arrange(CellID) %>%
  mutate(cell_fao_ohi = paste(CellID, fao_id, rgn_id, sep="_"))

## checking old cells.csv files
old_cells <- read.csv('../v2018/int/cells.csv') %>%
  arrange(CellID) %>% 
  mutate(cell_fao_ohi = paste(CellID, fao_id, rgn_id, sep="_"))

old_2017_cells <- read.csv('../v2017/cells.csv') %>%
  arrange(CellID) %>%
  mutate(cell_fao_ohi = paste(CellID, fao_id, rgn_id, sep = "_"))

dim(cells)[1]-dim(old_cells)[1] # there are 83743 less rows due to deleting land

## Diff in IDs in new cells csv versus old
new_old <- setdiff(cells$cell_fao_ohi, old_cells$cell_fao_ohi)
## Diff in IDs in old cells csv versus new
old_new <- setdiff(cells$cell_fao_ohi,old_cells$cell_fao_ohi)

## Take a look at locations of new cell IDs, using cells_raster as the template
library(maps)

cell_diff <- data.frame(cell_id = substr(new_old, 1,6), value = 1)
cell_diff$cell_id <- as.numeric(gsub("_","",cell_diff$cell_id))
unique_cell_diff <- unique(cell_diff)
diff_plot <- raster::subs(cells_raster, unique_cell_diff, by = "cell_id", which = "value", subsWithNA=TRUE)
map('legacy_world'); plot(diff_plot, add=T) # not sure crs aligns but just for visualizing

```
