---
title: "OHI 2018: Food Provision, Investigate Differences"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
---

# Summary

Data check during data prep.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# Setup
```{r}

## Libraries
library(dplyr)
library(tidyr)
library(readr)
library(sf)
library(ggplot2)
library(stringr)
library(here) 
setwd(here::here("globalprep","fis","v2019"))

source('../../../workflow/R/common.R')

```

# Initial Exploration of RAM v4.44 and RAM v4.40 Differences

Note: When loading old and new, make sure to clear environment if there is ever a situation where you are unsure whether the table you are wrangling with is from the new or old data (they have diff table names, so not all get replaced when loading a new version of RAM).

## Compare `metadata` to `timeseries` for both new and old versions

Save New and Old RAM metadata and timeseries data to unqiue variables for comparison. Old data checks out but new data has more unqiue `assessid` and `stockid` entries in the `timeseries` data than in the `metadata`. 
```{r}
## Check unique stockid in new RAM data
load(file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2019/RLSADB v4.44/DB Files With Assessment Data/DBdata.RData"))

## save newest metadata to unique variable 
RAM4.44_meta = as.data.frame(metadata)
RAM4.44_data = as.data.frame(timeseries)

## Check length in metadata
length(unique(RAM4.44_meta$assessid)) #1330 unique entries
length(unique(RAM4.44_meta$stockid)) # same number of stockids as assessids
length(unique(RAM4.44_meta$stocklong)) # same number of stocklong as the above two


## Check length in timeseries data
length(unique(RAM4.44_data$assessid)) # 2197 unique entries, more than in metadata
length(unique(RAM4.44_data$stockid)) # 1371 unique, more than in metadata..
length(unique(RAM4.44_data$stocklong)) #1373 unique

setdiff(RAM4.44_data$assessid, RAM4.44_meta$assessid) #868 diff in data not in meta
setdiff(RAM4.44_data$stockid, RAM4.44_meta$stockid) #42 stockids in data not in meta


## old stocks 
load(file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2018/RAM v4.40 (6-4-18)/DB Files With Assessment Data/DBdata.RData"))

RAM4.40_meta = as.data.frame(metadata)
RAM4.40_data = as.data.frame(timeseries)

## Check length
length(unique(RAM4.40_meta$assessid)) # 1252
length(unique(RAM4.40_meta$stockid)) # 1252
length(unique(RAM4.40_meta$assessid)) # 1252

length(unique(RAM4.40_data$assessid)) # 1814 
length(unique(RAM4.40_data$stockid)) # 1290
length(unique(RAM4.40_data$stocklong)) # 1292

setdiff(RAM4.40_data$assessid, RAM4.40_meta$assessid) #563
setdiff(RAM4.40_data$stockid, RAM4.40_meta$stockid) #39

```

## Investigate New RAM Database filtered for B/Bmsy data
```{r}

ram_bmsy <- RAM4.44_data %>%
  dplyr::filter(tsid == "BdivBmsytouse-dimensionless") %>%
  dplyr::filter(!is.na(tsvalue)) %>%
  dplyr::mutate(tsyear = as.numeric(as.character(tsyear))) %>%
  dplyr::filter(tsyear > 1979) %>%
  dplyr::select(assessid, year=tsyear, ram_bmsy = tsvalue, stockid, stocklong)

length(unique(ram_bmsy$assessid)) # 487 unique entries
length(unique(ram_bmsy$stockid)) # 355
length(unique(ram_bmsy$stocklong)) # 355

setdiff(ram_bmsy$assessid, RAM4.44_meta$assessid) # 169 diffs
setdiff(ram_bmsy$stockid, RAM4.44_meta$stockid) # 9 diffs


```

## Check a stock that is in `ram_spatial` not `raw_meta`

Try looking at stockid BLACKROCKNPCOAST, in ram_spatial but not raw_meta. ram_spatial combines last year's stock and this year's stock. Check to see if BLACKROCKNPCOAST is in old RAM database
```{r}
#BLACKROCKNPCOAST
## See if BLACKROCKNPCOAST is in this vector - it is 
new_stockid <- unique(ram_bmsy$stockid)

## Check to see if BLACKROCKNPCOAST is in old stock data
ram_bmsy_old <- data.frame(RAM4.40_data) %>%
  dplyr::filter(tsid == "BdivBmsytouse-dimensionless") %>%
  dplyr::filter(!is.na(tsvalue)) %>%
  dplyr::mutate(tsyear = as.numeric(as.character(tsyear))) %>%
  dplyr::filter(tsyear > 1979) %>%
  dplyr::select(assessid, year=tsyear, ram_bmsy = tsvalue, stockid, stocklong)

## YES, BLACKROCKNPCOAST is in the old stock data. 
old_stockid <-  unique(ram_bmsy_old$stockid)


## Compare stockid in raw_meta for v4.44 RAM data versus stockid in timeseries.. should match
setdiff(ram_bmsy$stockid,RAM4.44_meta$stockid)

```


***

# Investigate Multiple `assessid` in RAM v4.44 Database

## Identify the number of assessments per stock
```{r}
## see unique pairs of assess and stockid
test <- RAM4.40_data %>% 
  select(assessid, stockid) %>% 
  unique() %>% 
  group_by(stockid) %>% 
  summarize(num = length(assessid))
```

## Check stocks with multiple assessments

Check `ACADREDGOMGB` which has 2 assessments
```{r}
test <- ram_bmsy %>% 
  filter(stockid == "ACADREDGOMGB") %>% 
  spread(assessid, ram_bmsy)

plot(test$`NEFSC-ACADREDGOMGB-1913-2007-MILLER`,test$`NEFSC-ACADREDGOMGB-1913-2014-SISIMP2016`)
abline(0,1, col="red")
```

ANGLVIIIc-IXa has 6 assessments, take a look
```{r}
test2 <- ram_bmsy %>% 
  filter(stockid == "ANGLVIIIc-IXa")

unique(test2$assessid)

```

***

# Explore `timeseries_` tables

Found out that `metadata` data table contains only the most recent assessments, and `timeseries` contain all assessments ever conducted and collected by RAM. Check the following `timeseries_XXX` tables:

- timeseries_assessment
- timeseries_ids_views
- timeseries_notes_views
- timeseries_sources_views
- timeseries_units_views
- timeseries_values_views
- timeseries_years_views

Can use `ALPLAICBSAI` stock id, which has two assessments in `timeseries` to check.

Summary: `metadata` table has 1:1 assessid and stockid values.
```{r}

load(file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2018/RAM v4.40 (6-4-18)/DB Files With Assessment Data/DBdata.RData"))

# timeseries
ram_bmsy <- timeseries %>%
  dplyr::filter(tsid == "BdivBmsytouse-dimensionless") %>% 
  # filter(assessid == "AFSC-ALPLAICBSAI-1975-2015-SISIMP2016") %>% 
  select(assessid, stockid, stocklong, year = tsyear, BBmsy = tsvalue)

# timeseries views: If I only retain BdivBmsypref
ram_bmsy_new <- timeseries_values_views %>% 
  #filter(stockid == "ALPLAICBSAI") %>% 
  select(stockid, stocklong, year, BdivBmsypref)

# length of stockids in the two tables
length(unique(ram_bmsy$stockid)); length(unique(ram_bmsy_new$stockid)) # more in ts_values_views

# diff between unique stockids in the two tables
setdiff(unique(ram_bmsy$stockid), unique(ram_bmsy_new$stockid)) # 8 diffs 
setdiff(unique(ram_bmsy_new$stockid), unique(ram_bmsy$stockid)) # 86 diffs

# diff between metadata list of stockids and those in the two tables
setdiff(unique(ram_bmsy$stockid), unique(metadata$stockid)) # 8 diffs..
setdiff(unique(ram_bmsy_new$stockid), unique(metadata$stockid)) # no diffs!!



# timeseries views: If I only retain SSBdivSSBmsy
ram_bmsy_new <- timeseries_values_views %>% 
  #filter(stockid == "ALPLAICBSAI") %>% 
  select(stockid, stocklong, year, SSBdivSSBmsy) %>% 
  na.omit() 

setdiff(unique(ram_bmsy$stockid), unique(ram_bmsy_new$stockid)) # 72 diffs.. 
setdiff(unique(ram_bmsy_new$stockid), unique(ram_bmsy$stockid)) # 0 diffs


# timeseries views: If I only retain TBdivTBmsy and SSBdivSSBmsy
ram_bmsy_new <- timeseries_values_views %>% 
  #filter(stockid == "ALPLAICBSAI") %>% 
  select(stockid, stocklong, year, TBdivTBmsy, SSBdivSSBmsy, BdivBmsypref) 

setdiff(unique(ram_bmsy$stockid), unique(ram_bmsy_new$stockid)) # 72 diffs.. 
setdiff(unique(ram_bmsy_new$stockid), unique(ram_bmsy$stockid)) # 0 diffs





# compare B/Bmsy values for a single stock- change above
check <- ram_bmsy %>% 
  left_join(ram_bmsy_new, by = c("stockid", "stocklong", "year"))

plot(check$BBmsy, check$BBmsy_ts)
abline(0,1,col="red")
plot(check$BBmsy_ts, check$BBmsy_ts2)
abline(0,1,col="red")


```

***

# Investigate Using `timeseries_values_views`

Talked to Mel about grabbing most recent/best prioritization from `timeseries_values_views`. Note: After joining `timeseries` to `timeseries_values_views` by `stockid`, create a new column with the relationship we think is occurring between B/Bmsytouse, TB/TBmsy, and SSB/SSBmsy. And then compare with B/Bmsytouse column. There should be at least one assessid per stockid that matches.

1. from `timeseries_values_views`: select TB/TBmsy and SSB/SSBmsy
2. In `timeseries`, filter out the BdivBmsytouse.
3. Then join with the filtered `timeseries_values_views`. The `timeseries_values_views` should be repeated in the cases where more than one assessment was performed on a stock.
4. Create new column selecting TB/TBmsy when available, and if not available select SSB/SSBmsy
5. Subtract new column from B/Bmsytouse values. If TB/TBmsy is used preferentially for BdivBmsytouse, then every stock will have at least one assessment that is equal to zero.
5. If this is the case, then I think we can feel comfortable filtering the TB/TBmsy values from the `timeseries_values_views`. 
6. Then removing the stocks in this list from `timeseries_values_views` 
7. filtering the remaining stocks for SSB/SSBmsy.  
8. Then combining the TB/TBmsy and SSB/SSBmsy data into a BdivBmsytouse variable.


STEP 1-3
from `timeseries_values_views`: select TB/TBmsy and SSB/SSBmsy, filter out BdivBmsytouse in `timeseries`, and join.
```{r}
load(file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2018/RAM v4.40 (6-4-18)/DB Files With Assessment Data/DBdata.RData"))

## filter for TB/TBmsy and SSB/SSBmsy
tb_ssb <- timeseries_values_views %>% 
  select(stockid, stocklong, year, TBdivTBmsy, SSBdivSSBmsy)

## filter for B/Bmsy to use (BdivBmsytouse) from the timeseries table we used to use
bbmsy_touse <- timeseries %>% 
  filter(tsid == "BdivBmsytouse-dimensionless") %>% 
  select(-tsid) %>% 
  rename(year = tsyear, BdivBmsytouse = tsvalue)

# join old timeseries with new timeseries table
compare <- bbmsy_touse %>% 
  full_join(tb_ssb, by = c("stockid","stocklong","year"))
```

Check joined table info
```{r}

# number of unique asssessments and stocks
dim(compare) # 58,715 rows, 7 col
summary(compare) # 35,075 NAs in BdivBmsytouse, 47,810 NAs in TBdivTBmsy, 38,739 NAs in SSBdivSSBmsy
length(unique(compare$assessid)) # 474
length(unique(compare$stockid)) # 1259

## find diffs between the tables and against metadata table
setdiff(unique(tb_tbmsy$stockid), unique(bbmsy_touse$stockid)) # 906 diff
setdiff(unique(bbmsy_touse$stockid), unique(tb_tbmsy$stockid)) # 8 diff
setdiff(unique(compare$stockid), unique(metadata$stockid)) # 9 diff all due to stock name changes documented in "Stock Change History (6-1-18) except HOGFISHEGM and HOGFISHSEFL

## compare table
## some stockids from timseries_values_views have no assess id, aka no matching value in timeseries

## check rows where assessid = NA
check_assess_NA <- compare %>% 
  filter(is.na(assessid))

summary(check_assess_NA)

sum(is.na(check_assess_NA$assessid))

## number of missing assessments matches the number of missing BdivBmsytouse, TBdivTBmsy, and SSBdivSSBmsy. remove rows with missing assessments from the joined table

compare <- compare %>% 
  filter(!is.na(assessid))

sum(is.na(compare$assessid)) # should be 0

```

number of missing assessments matches the number of missing BdivBmsytouse, TBdivTBmsy, and SSBdivSSBmsy. remove rows with missing assessments from the joined table
```{r}

compare <- compare %>% 
  filter(!is.na(assessid))

sum(is.na(compare$assessid)) # should be 0

```

STEP 4. 
Create new column selecting TB/TBmsy when available, and if not available select SSB/SSBmsy

```{r}

check_bbmsy <- compare %>% 
  mutate(new_bbmsy = ifelse(!is.na(TBdivTBmsy), TBdivTBmsy, SSBdivSSBmsy))

```

STEP 5. 
Subtract new column from B/Bmsytouse values. If TB/TBmsy is used preferentially for BdivBmsytouse, then every stock will have at least one assessment that is equal to zero.

```{r}

check_bbmsy <- check_bbmsy %>% 
  mutate(check = new_bbmsy - BdivBmsytouse)

```

Check Results

```{r}

## remove rows where there is no BdivBmsytouse value OR a new BBMSY value (which means NA for both TB/TBmsy AND SSB/SSBmsy)
results <- check_bbmsy %>% 
  filter(!is.na(new_bbmsy) & !is.na(BdivBmsytouse))

length(unique(results$stockid)) #345
length(unique(results$assessid)) #465

## filter for rows where new_bbmsy matches BdivBmsytouse
results <- results %>% 
  filter(check == 0)

length(unique(results$stockid)) #345 !!!! SAME. GREAT. 
length(unique(results$assessid)) #387

lastcheck <- results %>% 
  group_by(year, stockid) %>% 
  mutate(assessnum = length(unique(assessid))) %>% 
  ungroup() %>% 
  filter(assessnum > 1)

unique(lastheck$stockid) ## only 5 stock ids have multiple assessments still, because the diff assess have the same values


## plot each of the five stock ids, and set color to assessid to make sure values are overlapping - should be a single dot visible per year!!
ggplot(lastcheck %>% filter(stockid == "ALBASATL"), aes(x=year , y=BdivBmsytouse, col = assessid)) +
  geom_point()

ggplot(lastcheck %>% filter(stockid == "CHAKESA"), aes(x=year , y=BdivBmsytouse, col = assessid)) +
  geom_point()

ggplot(lastcheck %>% filter(stockid == "GHALV-VI-XII-XIV"), aes(x=year , y=BdivBmsytouse, col = assessid)) +
  geom_point()

ggplot(lastcheck %>% filter(stockid == "ANGLVIIIc-IXa"), aes(x=year , y=BdivBmsytouse, col = assessid)) +
  geom_point()

ggplot(lastcheck %>% filter(stockid == "MEGSPPIVa-VIa"), aes(x=year , y=BdivBmsytouse, col = assessid)) +
  geom_point()
  

```

***

## Number of Stocks With Missing RAM Area

get old ram spatial table, join stockid info
```{r}
old_ram_spatial <- read_csv("~/github/ohiprep_v2018/globalprep/fis/v2017/int/RAM_fao_ohi_rgns.csv")

load(file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2017/RAM v3.80/DB Files With Assessment Data/DBdata.RData"))

old_timeseries <- as.data.frame(timeseries) %>% 
  select(stockid, assessid) %>% 
  distinct()

old_ram_spatial <- old_ram_spatial %>% 
  left_join(old_timeseries, by= c("assessid")) %>% 
  select(-assessid) %>% 
  rename(stockid_ram =stockid)

```

combined watson and ram bbmsy data
```{r}
ram <- read.csv("int/ram_bmsy.csv") %>% # final output from RAM_data_prep
  rename(stock_id = stockid) # to match other two data tabls

mean_catch <- read.csv("data/mean_catch.csv") %>% # final output from Watson catch
  mutate(stock_id_taxonkey = as.character(stock_id_taxonkey)) %>%
  mutate(taxon_key = str_extract(stock_id_taxonkey, "(\\d)+$")) %>% # extract ending consecutive digits
  mutate(stock_id = str_extract(stock_id_taxonkey, "^(\\w+).(\\d){1,2}")) %>% 
  rename(year = Year)
data <- mean_catch %>%
  left_join(ram, by=c('rgn_id', 'stock_id', "year"))
```

Combine RAM-Watson data table with old ram_spatial that contains area info
```{r}
c_data <- data %>%
  mutate(fao_id = as.integer(str_extract(stock_id, "(\\d+)"))) %>% 
  left_join(old_ram_spatial, by =c("rgn_id", "fao_id", "stockid_ram"))
```

Look at values with non-missing RAM bmsy data and see how many unique stock info have missing area
```{r}

summary(c_data)

## Filter for only catch with ram bbmsy data
has_ram <- c_data %>% 
  filter(!is.na(ram_bmsy)) 

# has_ram <- c_data %>% 
#   filter(!is.na(ram_bmsy)) %>% 
#   select(rgn_id, stock_id, stockid_ram, RAM_area_m2) %>% 
#   distinct()

## Filter for missing RAM Area within RAM bbmsy data
nas <- has_ram %>% 
  filter(is.na(RAM_area_m2))

dim(nas)[1]/dim(has_ram)[1] # proportion of missing RAM area


```
