---
title: "OHI 2019: Explore Biomass Variables"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: console
---

# Summary

This script explores the use of management target biomass variables (TBdivTBmgt, SSBdivSSBmgt), in conjunction with the regular maximum sustainable yield variables (TBdivTBmsy, SSBdivSSBmsy). 

## Updates from previous assessment

This is new to the 2019 assessment year. 

# Data
## B/Bmsy values from stock assessments 

Reference: [RAM Legacy Stock Assessment Database](http://ramlegacy.org) v4.44

* Downloaded: 2/28/2019
* Description: B/Bmsy value by stock and year (other data, which we do not use, are also available in the database)
* Native data resolution: stock (fish stock, species and region specific)
* Time range: 1950 - 2016
* Format: R data files (.rds)
* DOI: 10.5281/zenodo.2542919

## Setup

``` {r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',message = FALSE, warning = FALSE)
```

```{r, eval = FALSE}

## Libraries
library(dplyr)
library(tidyr)
library(readr)
library(sf)
library(ggplot2)
library(here) 
library(corrplot)

## highlight out when knitting
setwd(here::here("globalprep/fis/v2019"))
source('../../../workflow/R/common.R')

```

## Obtain RAM B/Bmsy data

The data is stored as a relational database in an R object. Check that the names of each element have not changed from last year! Update as appropriate in the below list. 

The following tables are included (for full list, see loadDBdata.r in mazu):  

1. timeseries  
The time series data is a data frame containing all assessments conducted per stock with the following headers/columns:  
(1) assessid (2) stockid (3) stocklong (4) tsid (5) tsyear (6) tsvalue

2. bioparams  
The time series data is a data frame with the following headers/columns:  
(1) assessid (2) stockid (3) stocklong (4) bioid (5) biovalue (6) bioyear (7) bionotes  

3.  timeseries_values_views  
This stores the timeseries values, using the most recent assessment available, with timeseries type. The dataframe contains the following headers/columns: stockid, stocklong, year, TBbest, TCbest, ERbest, BdivBmsypref, UdivUmsypref, BdivBmgtpref, UdivUmgtpref, TB, SSB, TN, R, TC, TL, RecC, F, ER, TBdivTBmsy, SSBdivSSBmsy, FdivFmsy, ERdivERmsy, CdivMSY, TBdivTBmsy, SSBdivSSBmsy, FdivFmgt, ERdivERmgt, Cpair, TAC, Cadvised, survB, CPUE, EFFORT, and stocks along the rows.

4. timeseries_units_views  
This stores the timeseries units (or time series source for touse time series), with timeseries type. The dataframe contains the following headers/columns: stockid, stocklong, TBbest, TCbest, ERbest, BdivBmgtpref, UdivUmsypref, BdivBmgtpret, UdivUmgtpref, TB, SSB, TN, R, TC, TL, RecC, F, ER, TBdivTBmsy, SSBdivSSBmsy, FdivFmsy, ERdivERmsy, CdivMSY, TBdivTBmgt, SSBdivSSBmgt, FdivFmgt, ERdivERmgt, Cpair, TAC, Cadvised, survB, CPUE, EFFORT, and stocks along the rows

5. timeseries_id_views  
This stores the timeseries ids with timeseries id along the columns. The dataframe contains the following headers/columns: stockid, stocklong, TBbest, TCbest, ERbest, BdivBmsypref, UdivUmsypref, BdivBmgtpref, UdivUmgtpref, TB, SSB, TN, R, TC, TL, RecC, F, ER, TBdivTBmsy, SSBdivSSBmsy, FdivFmsy, ERdivERmsy, CdivMSY, TBdivTBmgt, SSBdivSSBmgt, FdivFmgt, ERdivERmgt, Cpair, TAC, Cadvised, survB, CPUE, EFFORT, and stocks along the rows.

6. bioparams_values_views  
This stores the bioparams values, with bioparam type along the columns
(TBmsybest, ERmsybest, MSYbest, TBmgtbest, ERmgtbest, TBmsy, SSBmsy, Nmsy, MSY, Fmsy, ERmsy, TBmgt, SSBmgt, Fmgt, ERmgt, TB0, SSB0, M, TBlim, SSBlim, Flim, ERlim) and stocks along the rows.

7. bioparams_units_views  
This stores the bioparams units, with bioparam type along the columns (TBmsybest, ERmsybest, MSYbest, TBmgtbest, ERmgtbest, TBmsy, SSBmsy, Nmsy, MSY, Fmsy, ERmsy, TBmgt, SSBmgt, Fmgt, ERmgt, TB0, SSB0, M, TBlim, SSBlim, Flim, ERlim) and stocks along the rows.

8. bioparams_ids_views  
This stores the bioparams ids, with bioparam id along the columns
(TBmsybest, ERmsybest, MSYbest, TBmgtbest, ERmgtbest, TBmsy, SSBmsy, Nmsy, MSY, Fmsy, ERmsy, TBmgt, SSBmgt, Fmgt, ERmgt, TB0, SSB0, M, TBlim, SSBlim, Flim, ERlim) and stocks along the rows.

9. metadata  
This stores assorted metadata associated with the stock, with datatypes along the columns
(assessid, stockid, stocklong, assessyear, scientificname, commonname, areaname, managementauthority, assessorfull, region, FisheryType) and stock by row.

10. tsmetrics
Contains metadata, with columns tscategory, tsshort, tslong, tsunitsshort, tsunitslong, tsunique.


For this data prep we primarily use and consult `timeseries_values_views`, `tsmetrics`, and `metadata`
```{r load_data, eval = FALSE}

load(file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2019/RLSADB v4.44/DB Files With Assessment Data/DBdata.RData"))

ram_bmsy <- timeseries_values_views %>%
  dplyr::select(stockid, year, TBdivTBmsy, SSBdivSSBmsy, TBdivTBmgt, SSBdivSSBmgt)

```

Make a correlation plot of these data
```{r, eval = FALSE}
## what we currently use. TBmsy and SSBmsy
tbmsy_ssbmsy <- ram_bmsy %>%
  dplyr::select(stockid, year, TBdivTBmsy, SSBdivSSBmsy) %>%
  dplyr::filter(!is.na(TBdivTBmsy), !is.na(SSBdivSSBmsy)) #have to filter out NAs or we can't make a corrplot...

tbmsy_ssbmsy_cor <- cor(tbmsy_ssbmsy[3:4])
corrplot.mixed(tbmsy_ssbmsy_cor,
               number.cex = 0.7, 
               upper = "ellipse", 
               tl.col = "black",
               tl.pos = "d",
               tl.cex = 0.75) ## 0.83 positive correlation

tbmsy_ssbmsy_filt <- tbmsy_ssbmsy %>%
  filter(TBdivTBmsy <= 2, SSBdivSSBmsy <= 2)

plot(tbmsy_ssbmsy_filt$TBdivTBmsy, tbmsy_ssbmsy_filt$SSBdivSSBmsy)
abline(0,1, col = "red")

## compare the Total Biomass msy and management targets. 
TB_msy_mgt <- ram_bmsy %>%
  dplyr::select(stockid, year, TBdivTBmsy, TBdivTBmgt) %>%
  dplyr::filter(!is.na(TBdivTBmsy), !is.na(TBdivTBmgt))

tb_msy_mgt_cor <- cor(TB_msy_mgt[3:4])
corrplot.mixed(tb_msy_mgt_cor,
               number.cex = 0.7, 
               upper = "ellipse", 
               tl.col = "black",
               tl.pos = "d",
               tl.cex = 0.75) #0.64 positive correlation

tb_msy_mgt_filt <- TB_msy_mgt %>%
  filter(TBdivTBmsy <= 2, TBdivTBmgt <= 2)

plot(tb_msy_mgt_filt$TBdivTBmsy, tb_msy_mgt_filt$TBdivTBmgt)
abline(0,1, col = "red")

## compare the ssb msy and management targets. 
ssb_msy_mgt <- ram_bmsy %>%
  dplyr::select(stockid, year, SSBdivSSBmsy, SSBdivSSBmgt) %>%
  dplyr::filter(!is.na(SSBdivSSBmsy), !is.na(SSBdivSSBmgt))

ssb_msy_mgt_cor <- cor(ssb_msy_mgt[3:4])
corrplot.mixed(ssb_msy_mgt_cor,
               number.cex = 0.7, 
               upper = "ellipse", 
               tl.col = "black",
               tl.pos = "d",
               tl.cex = 0.75) #0.84 positive correlation

ssb_msy_mgt_filt <- ssb_msy_mgt %>%
  filter(SSBdivSSBmsy <= 2, SSBdivSSBmgt <= 2)

plot(ssb_msy_mgt_filt$SSBdivSSBmsy, ssb_msy_mgt_filt$SSBdivSSBmgt)
abline(0,1, col = "red")


## Compare tbmgt and ssbmgt 
tbmgt_ssbmgt <- ram_bmsy %>%
  dplyr::select(stockid, year, TBdivTBmgt, SSBdivSSBmgt) %>%
  dplyr::filter(!is.na(TBdivTBmgt), !is.na(SSBdivSSBmgt))

tbmgt_ssbmgt_cor <- cor(tbmgt_ssbmgt[3:4])
corrplot.mixed(tbmgt_ssbmgt_cor,
               number.cex = 0.7, 
               upper = "ellipse", 
               tl.col = "black",
               tl.pos = "d",
               tl.cex = 0.75) ## 0.79 positive correlation

tbmgt_ssbmgt_filt <- tbmgt_ssbmgt %>%
    filter(TBdivTBmgt <= 2, SSBdivSSBmgt <= 2)


plot(tbmgt_ssbmgt_filt$TBdivTBmgt, tbmgt_ssbmgt_filt$SSBdivSSBmgt)
abline(0,1, col = "red")

## compare TBmsy and SSBmgt
tbmsy_ssbmgt <- ram_bmsy %>%
  dplyr::select(stockid, year, TBdivTBmsy, SSBdivSSBmgt) %>%
  dplyr::filter(!is.na(TBdivTBmsy), !is.na(SSBdivSSBmgt))

tbmsy_ssbmgt_cor <- cor(tbmsy_ssbmgt[3:4])
corrplot.mixed(tbmsy_ssbmgt_cor,
               number.cex = 0.7, 
               upper = "ellipse", 
               tl.col = "black",
               tl.pos = "d",
               tl.cex = 0.75) ## 0.82 positive correlation

tbmsy_ssbmgt_filt <- tbmsy_ssbmgt %>%
    filter(TBdivTBmsy <= 2, SSBdivSSBmgt <= 2)

plot(tbmsy_ssbmgt_filt$TBdivTBmsy, tbmsy_ssbmgt_filt$SSBdivSSBmgt)
abline(0,1, col = "red")

## compare TBmsy, TBmgt, and SSBmsy
tbmsy_tbmgt_ssbmsy <- ram_bmsy %>%
  dplyr::select(stockid, year, TBdivTBmsy, TBdivTBmgt, SSBdivSSBmsy) %>%
  filter(!is.na(TBdivTBmsy), !is.na(TBdivTBmgt), !is.na(SSBdivSSBmsy))

tbmsy_tbmgt_ssbmsy_cor <- cor(tbmsy_tbmgt_ssbmsy[3:5])
corrplot.mixed(tbmsy_tbmgt_ssbmsy_cor,
               number.cex = 0.7, 
               upper = "ellipse", 
               tl.col = "black",
               tl.pos = "d",
               tl.cex = 0.75) ## when all three of these are present, very correllated. Nearly 1 for all. 

## compare TBmsy, TBmgt, and SSBmgt 
tbmsy_tbmgt_ssbmgt <- ram_bmsy %>%
  dplyr::select(stockid, year, TBdivTBmsy, TBdivTBmgt, SSBdivSSBmgt) %>%
  filter(!is.na(TBdivTBmsy), !is.na(TBdivTBmgt), !is.na(SSBdivSSBmgt)) ## data frame is empty. Cannot make a correlation plot from this.

## compare TBmsy, SSBmsy, and SSBmgt 
tbmsy_SSBmsy_ssbmgt <- ram_bmsy %>%
  dplyr::select(stockid, year, TBdivTBmsy, SSBdivSSBmsy, SSBdivSSBmgt) %>%
  filter(!is.na(TBdivTBmsy), !is.na(SSBdivSSBmsy), !is.na(SSBdivSSBmgt)) ## 571 observations

tbmsy_ssbmsy_ssbmgt_cor <- cor(tbmsy_SSBmsy_ssbmgt[3:5])
corrplot.mixed(tbmsy_ssbmsy_ssbmgt_cor,
               number.cex = 0.7, 
               upper = "ellipse", 
               tl.col = "black",
               tl.pos = "d",
               tl.cex = 0.75)

## compare SSBdivSSBmsy TBmgt, and SSBmgt 
ssbmsy_tbmgt_ssbmgt <- ram_bmsy %>%
  dplyr::select(stockid, year, SSBdivSSBmsy, TBdivTBmgt, SSBdivSSBmgt) %>%
  filter(!is.na(SSBdivSSBmsy), !is.na(TBdivTBmgt), !is.na(SSBdivSSBmgt)) ## data frame is empty. Cannot make a correlation plot from this.

## compare TBmsy, SSBmsy, and SSBmgt 
tbmsy_SSBmsy_ssbmgt <- ram_bmsy %>%
  dplyr::select(stockid, year, TBdivTBmsy, SSBdivSSBmsy, SSBdivSSBmgt) %>%
  filter(!is.na(TBdivTBmsy), !is.na(SSBdivSSBmsy), !is.na(SSBdivSSBmgt)) ## data frame is empty. Cannot make a correlation plot from this.


```

