---
title: "Exploring Watson v4 fishing catch data"
output: html_document
---

```{r}

library(here)
library(tidyverse)
library(plotly)


# directory paths and relevant files
source('https://raw.githubusercontent.com/OHI-Science/ohiprep_v2019/gh-pages/workflow/R/common.R')
```


```{r}

new_rawFolder <- file.path(dir_M, "git-annex/globalprep/_raw_data/IMAS_GlobalFisheriesLandings/d2019")
old_rawFolder <- file.path(dir_M, "git-annex/globalprep/_raw_data/IMAS_GlobalFisheriesLandings/d2018")


data1 <- readRDS(file.path(new_rawFolder, "CatchInd2015_2015.rds"))
data2 <- readRDS(file.path(new_rawFolder, "CatchInd2010_2014.rds"))
data3 <- readRDS(file.path(new_rawFolder, "CatchInd2005_2009.rds"))

new_data <- rbind(data1, data2, data3)


## confirming that IndexInd.csv summarizes this information...it does!
tmp <- filter(new_data, ID==667150)  
# > dim(tmp)
# [1] 730   5
# > sum(tmp$Reported)
# [1] 142.6872
# > sum(tmp$IUU)
# [1] 36.7611
# > sum(tmp$Discards)
# [1] 0.11696


```

Only need the index file if not assigning to raster cells.

```{r}

Ind_new <- read.csv(file.path(new_rawFolder, "IndexInd.csv"))
Ind_new <- gather(Ind_new, "catch_type", "tonnes", c("Reported", "IUUTotal", "Discards")) %>%
  mutate(data = "Industrial_v4")

NInd_new <- read.csv(file.path(new_rawFolder, "IndexNInd.csv"))
NInd_new <- gather(NInd_new, "catch_type", "tonnes", c("Reported", "IUUTotal", "Discards")) %>%
  mutate(data = "NonIndustrial_v4")

new_catch <- rbind(Ind_new, NInd_new) %>%
  select(ID, Year=IYear, catch_type, tonnes, data)

Ind_old <- read.csv(file.path(old_rawFolder, "Index.csv")) %>%
  select(-NIndReported, -NIndIUU, -NIndDiscards) %>%
  rename(Reported=IndReported, IUUTotal=IndIUU, Discards=IndDiscards) %>%
  gather("catch_type", "tonnes", c("Reported", "IUUTotal", "Discards")) %>%
  mutate(data = "Industrial_v3")

NInd_old <- read.csv(file.path(old_rawFolder, "Index.csv")) %>%
  select(-IndReported, -IndIUU, -IndDiscards) %>%
  rename(Reported=NIndReported, IUUTotal=NIndIUU, Discards=NIndDiscards) %>%
  gather("catch_type", "tonnes", c("Reported", "IUUTotal", "Discards")) %>%
  mutate(data = "NonIndustrial_v3")

old_catch <- rbind(Ind_old, NInd_old) %>%
  select(ID, Year, catch_type, tonnes, data) 

catch_data <- rbind(new_catch, old_catch) %>%
  group_by(Year, data, catch_type) %>%
  summarize(tonnes = sum(tonnes)) %>%
  filter(Year >= 1990) %>%
  separate(data, c("Ind_vs_NInd", "data_version"), by="_")

ggplot(filter(catch_data, Ind_vs_NInd=="Industrial"), aes(x=Year, y=tonnes, 
                       group=interaction(data_version, catch_type),
                       color=catch_type,
                       linetype=data_version,
                       shape=data_version)) +
  geom_point() +
  geom_line() +
  labs(title="Commercial Fisheries")

ggplot(filter(catch_data, Ind_vs_NInd=="NonIndustrial"), aes(x=Year, y=tonnes, 
                       group=interaction(data_version, catch_type),
                       color=catch_type,
                       linetype=data_version,
                       shape=data_version)) +
  geom_point() +
  geom_line() +
  labs(title="NonCommercial Fisheries")

summary_catch <- filter(catch_data, Year==2014) %>%
  spread(data_version, tonnes) %>%
  arrange(catch_type) 

71037203+10374083
51404298+30008229

summary_catch
  colSums(summary_catch[, c(4,5)])

## SAUP Reported + Unreported
78914000+29773000

catch_data2 <- rbind(new_catch, old_catch) %>%
  separate(data, c("Ind_vs_NInd", "data_version"), by="_") %>%
  group_by(Year, data_version, Ind_vs_NInd) %>%
  summarize(tonnes = sum(tonnes)) %>%
  filter(Year >= 1990) 



ggplot(catch_data2, aes(x=Year, y=tonnes, 
                       group= interaction(data_version, Ind_vs_NInd),
                       color=Ind_vs_NInd,
                       linetype = data_version,
                       style = data_version)) +
  geom_point() +
  geom_line() +
  labs(title="All Fisheries") +
  ylim(0,max(catch_data2$tonnes))



catch_data3 <- rbind(new_catch, old_catch) %>%
  separate(data, c("Ind_vs_NInd", "data_version"), by="_") %>%
  group_by(Year, data_version) %>%
  summarize(tonnes = sum(tonnes)) %>%
  filter(Year >= 1990) 

ggplot(catch_data3, aes(x=Year, y=tonnes, 
                       group= data_version,
                       color=data_version)) +
  geom_point() +
  geom_line() +
  labs(title="All Fisheries") +
  ylim(0,max(catch_data3$tonnes))


prop_diff <- catch_data3 %>%
  group_by(Year) %>%
  summarize(prop_diff = tonnes[data_version=='v3']/tonnes[data_version=='v4'])

filter(catch_data3, Year==2014)

