---
title: "OHI 2019: Food Provision/Fisheries, Preparing RAM B/Bmsy data"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: console
---

# Summary

This script prepares the RAM B/Bmsy data:
1. Relevant data are collected from the RAM database
2. Missing years are gapfilled when appropriate
3. RAM and Watson species names are harmonized in a few cases
4. RAM stocks are associated with the corresponding OHI and FAO regions


## Updates from previous assessment

This year we have additional stocks without spatial information from Christopher Free (2017). We manually assigned ohi and fao region id information to the additional stocks in `fao_ohi_rgns.Rmd` using best available information on stock distribution and saved the file in `int/RAM_fao_ohi_rgns.csv`. Watson species names should be similar if not the same as the SAUP species names used prior the 2018 assessment. We are also using management target calues for biomass values, which add more RAM species to our analysis. 


***

# Data
## B/Bmsy values from stock assessments 

Reference: [RAM Legacy Stock Assessment Database](http://ramlegacy.org) v4.44

* Downloaded: 2/28/2019
* Description: B/Bmsy value by stock and year (other data, which we do not use, are also available in the database)
* Native data resolution: stock (fish stock, species and region specific)
* Time range: 1950 - 2016
* Format: R data files (.rds)
* DOI: 10.5281/zenodo.2542919


## Stock range data 
Reference: Christopher M. Free. 2017. Mapping fish stock boundaries for the original Ram Myers stock-recruit database. https://marine.rutgers.edu/~cfree/mapping-fish-stock-boundaries-for-the-original-ram-myers-stock-recruit-database/. downloaded 9/25/2017.

* Downloaded: 8/20/2018
* Description: Shapefiles for each stock describing their distribution
* Native data resolution: Spatial shapefiles
* Format: Shapefiles

***

## Setup

``` {r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',message = FALSE, warning = FALSE, echo = TRUE, eval=FALSE)
```

```{r, eval = FALSE}

## Libraries
library(dplyr)
library(tidyr)
library(readr)
library(sf)
library(ggplot2)
library(here) 

## highlight out when knitting
setwd(here::here("globalprep/fis/v2019"))
source('../../../workflow/R/common.R')

## Paths for data
path_raw_data = file.path(dir_M, "git-annex/globalprep/fis/v2019/int/annual_catch")

```

## Obtain RAM B/Bmsy data

The data is stored as a relational database in an R object. Check that the names of each element have not changed from last year! Update as appropriate in the below list. 

The following tables are included (for full list, see loadDBdata.r in mazu):  

1. timeseries  
The time series data is a data frame containing all assessments conducted per stock with the following headers/columns:  
(1) assessid (2) stockid (3) stocklong (4) tsid (5) tsyear (6) tsvalue

2. bioparams  
The time series data is a data frame with the following headers/columns:  
(1) assessid (2) stockid (3) stocklong (4) bioid (5) biovalue (6) bioyear (7) bionotes  

3.  timeseries_values_views  
This stores the timeseries values, using the most recent assessment available, with timeseries type. The dataframe contains the following headers/columns: stockid, stocklong, year, TBbest, TCbest, ERbest, BdivBmsypref, UdivUmsypref, BdivBmgtpref, UdivUmgtpref, TB, SSB, TN, R, TC, TL, RecC, F, ER, TBdivTBmsy, SSBdivSSBmsy, FdivFmsy, ERdivERmsy, CdivMSY, TBdivTBmsy, SSBdivSSBmsy, FdivFmgt, ERdivERmgt, Cpair, TAC, Cadvised, survB, CPUE, EFFORT, and stocks along the rows.

4. timeseries_units_views  
This stores the timeseries units (or time series source for touse time series), with timeseries type. The dataframe contains the following headers/columns: stockid, stocklong, TBbest, TCbest, ERbest, BdivBmgtpref, UdivUmsypref, BdivBmgtpret, UdivUmgtpref, TB, SSB, TN, R, TC, TL, RecC, F, ER, TBdivTBmsy, SSBdivSSBmsy, FdivFmsy, ERdivERmsy, CdivMSY, TBdivTBmgt, SSBdivSSBmgt, FdivFmgt, ERdivERmgt, Cpair, TAC, Cadvised, survB, CPUE, EFFORT, and stocks along the rows

5. timeseries_id_views  
This stores the timeseries ids with timeseries id along the columns. The dataframe contains the following headers/columns: stockid, stocklong, TBbest, TCbest, ERbest, BdivBmsypref, UdivUmsypref, BdivBmgtpref, UdivUmgtpref, TB, SSB, TN, R, TC, TL, RecC, F, ER, TBdivTBmsy, SSBdivSSBmsy, FdivFmsy, ERdivERmsy, CdivMSY, TBdivTBmgt, SSBdivSSBmgt, FdivFmgt, ERdivERmgt, Cpair, TAC, Cadvised, survB, CPUE, EFFORT, and stocks along the rows.

6. bioparams_values_views  
This stores the bioparams values, with bioparam type along the columns
(TBmsybest, ERmsybest, MSYbest, TBmgtbest, ERmgtbest, TBmsy, SSBmsy, Nmsy, MSY, Fmsy, ERmsy, TBmgt, SSBmgt, Fmgt, ERmgt, TB0, SSB0, M, TBlim, SSBlim, Flim, ERlim) and stocks along the rows.

7. bioparams_units_views  
This stores the bioparams units, with bioparam type along the columns (TBmsybest, ERmsybest, MSYbest, TBmgtbest, ERmgtbest, TBmsy, SSBmsy, Nmsy, MSY, Fmsy, ERmsy, TBmgt, SSBmgt, Fmgt, ERmgt, TB0, SSB0, M, TBlim, SSBlim, Flim, ERlim) and stocks along the rows.

8. bioparams_ids_views  
This stores the bioparams ids, with bioparam id along the columns
(TBmsybest, ERmsybest, MSYbest, TBmgtbest, ERmgtbest, TBmsy, SSBmsy, Nmsy, MSY, Fmsy, ERmsy, TBmgt, SSBmgt, Fmgt, ERmgt, TB0, SSB0, M, TBlim, SSBlim, Flim, ERlim) and stocks along the rows.

9. metadata  
This stores assorted metadata associated with the stock, with datatypes along the columns
(assessid, stockid, stocklong, assessyear, scientificname, commonname, areaname, managementauthority, assessorfull, region, FisheryType) and stock by row.

10. tsmetrics
Contains metadata, with columns tscategory, tsshort, tslong, tsunitsshort, tsunitslong, tsunique.


For this data prep we primarily use and consult `timeseries_values_views`, `tsmetrics`, and `metadata`
```{r load_data, eval = FALSE}

load(file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2019/RLSADB v4.44/DB Files With Assessment Data/DBdata.RData"))

# ram_bmsy <- timeseries_values_views %>%
#   select(stockid,stocklong, year, TBdivTBmsy, SSBdivSSBmsy) %>% 
#   mutate(ram_bmsy = ifelse(!is.na(TBdivTBmsy), TBdivTBmsy, SSBdivSSBmsy)) %>% 
#   dplyr::filter(year > 1979) %>%
#   filter(!is.na(ram_bmsy)) %>% 
#   dplyr::select(stockid, stocklong, year, ram_bmsy)

ram_bmsy_new <- timeseries_values_views %>%
  dplyr::select(stockid, stocklong, year, TBdivTBmsy, SSBdivSSBmsy, TBdivTBmgt, SSBdivSSBmgt) %>%
  mutate(ram_bmsy = 
           ifelse(!is.na(TBdivTBmsy), TBdivTBmsy, SSBdivSSBmsy)) %>%
  mutate(ram_bmsy =
           ifelse(is.na(TBdivTBmsy) & is.na(SSBdivSSBmsy), TBdivTBmgt, ram_bmsy)) %>%
  mutate(ram_bmsy = 
           ifelse(is.na(TBdivTBmsy) & is.na(SSBdivSSBmsy) & is.na(TBdivTBmgt), SSBdivSSBmgt, ram_bmsy)) %>%
  dplyr::filter(year > 1979) %>%
  filter(!is.na(ram_bmsy)) %>%
  dplyr::select(stockid, stocklong, year, ram_bmsy)

```


## Identify FAO and OHI regions for RAM stocks added by using management values

*Note: This is only necessary for the 2019 assessment, since this is the first year we are implementing management target biomass values*
Determine if we will add management values
```{r, eval = FALSE}
ram_bmsy_check <- timeseries_values_views %>%
  select(stockid, year, TBdivTBmsy, SSBdivSSBmsy) %>% 
  mutate(ram_bmsy = ifelse(!is.na(TBdivTBmsy), TBdivTBmsy, SSBdivSSBmsy)) %>% 
  dplyr::filter(year > 1979)

sum(is.na(ram_bmsy_check$ram_bmsy)) #24480 NAs with only msy values in ram_bmsy column... 24480/35177 = ~69% of ram data is NA with these methods. 


ram_bmgt_new <- timeseries_values_views %>%
  select(stockid, stocklong, year, TBdivTBmsy, SSBdivSSBmsy, TBdivTBmgt, SSBdivSSBmgt) %>%
  mutate(ram_bmsy = 
           ifelse(!is.na(TBdivTBmsy), TBdivTBmsy, SSBdivSSBmsy)) %>%
  mutate(ram_bmsy =
           ifelse(is.na(TBdivTBmsy) & is.na(SSBdivSSBmsy), TBdivTBmgt, ram_bmsy)) %>%
  mutate(ram_bmsy = 
           ifelse(is.na(TBdivTBmsy) & is.na(SSBdivSSBmsy) & is.na(TBdivTBmgt), SSBdivSSBmgt, ram_bmsy)) %>%
  dplyr::filter(year > 1979)

sum(is.na(ram_bmgt_new$ram_bmsy)) #21195 NAs with new methods in ram_bmsy column... 21195/35177 = ~60% of ram data is NA with new methods.. better than before. Gained 3285 rows of data.

ram_bmgt_new_final <- ram_bmgt_new %>%
  filter(!is.na(ram_bmsy)) %>%
  dplyr::select(stockid,stocklong,year,ram_bmsy) 

new_species_list <- data.frame(stockid = sort(setdiff(ram_bmgt_new_final$stockid, ram_bmsy$stockid))) %>%
  unique() %>% ## there are 96 stocks added.
  left_join(ram_bmgt_new_final, by = "stockid") %>%
  dplyr::select(stockid, stocklong) %>%
  unique()

write.csv(new_species_list, "int/RAM_addl_mgt_stocks.csv", row.names = FALSE)

## based off of this, we will change our methods to include the management targets for bmsy... 
```


Here we identify the FAO/OHI regions where each RAM stock is located.  This involved the following steps:

1. Create a intersection map that identifies each FAO/OHI region.
2. Overlay each of the RAM stocks on the region map to determine where they fall.  There were 2 general challenges to doing this.  A few of the stocks did not have a complete dataframe in the shapefile and could not be joined to the other data.  A few stocks had intersecting polygons and consequently would not run.  In the first case, I added these data by hand later in the process.  For the second case, I fixed the intersections and added later.  

```{r RAM to OHI_FAO rgn, eval = FALSE}

RAM_spatial_dir <- file.path(dir_M, "git-annex/globalprep/_raw_data/RAM/d2017/ramldb_boundaries/ramldb_boundaries") 

ram_sf <- list.files(RAM_spatial_dir, pattern = "shp")

ram_sf <- gsub(".shp", "", ram_sf)

ram_sf <- intersect(ram_bmsy_gf$assessid, ram_sf) # get these spatial data

## get a sample file to convert other spatial data
tmp <- read_sf(dsn = RAM_spatial_dir, layer = ram_sf[1])

## Overlay stock data on fao-ohi region data

fao_ohi <- st_read(file.path(dir_M, "git-annex/globalprep/fis/v2017/int"), 
        layer = "ohi_fao_rgns")

ram_sf_df <- data.frame(ram_sf, stringsAsFactors=FALSE) # make ram_sf a dataframe so that we can filter for new stocks added from management values...

new_species_list <- read_csv("int/RAM_addl_mgt_stocks.csv")

ram_sf_df_mgt <- ram_sf_df %>%
  filter(grepl(paste(new_species_list$stockid, collapse = "|"), ram_sf))
ram_sf_df_mgt_vec <- c(ram_sf_df_mgt$ram_sf)


stock_fao_ohi <- NA
for(stock in ram_sf_df_mgt_vec) {    #stock = ram_sf_df_mgt$ram_sf[50]
    cat(paste0(stock, "\n"))
    tmp_poly <- read_sf(dsn = RAM_spatial_dir, layer = stock)
    tmp_fao_ohi <- st_intersection(fao_ohi, tmp_poly)
    if(sum(is.na(stock_fao_ohi))==1){
      stock_fao_ohi <- tmp_fao_ohi
    }else
    {stock_fao_ohi <- rbind(stock_fao_ohi, tmp_fao_ohi)}
}

#got nervous about losing this... writing to mazu just incase
# st_write(stock_fao_ohi, dsn = file.path(dir_M, "git-annex/globalprep/fis/v2019/int/management_additions"), 
#         layer = "stock_fao_ohi_mgt_initial", driver = "ESRI Shapefile")

stock_fao_ohi_shps <- stock_fao_ohi[st_dimension(stock_fao_ohi) == 2,]

stock_fao_ohi_shps <- stock_fao_ohi_shps[!is.na(st_dimension(stock_fao_ohi_shps)), ] #5904

## Fix GEOMETRYCOLLECTION features, http://r-spatial.org/r/2017/03/19/invalid.html

type <- st_is(stock_fao_ohi_shps, "GEOMETRYCOLLECTION")

stock_fao_ohi_shps[type, ] <- st_buffer(stock_fao_ohi_shps[type, ], 0.0) 

#check:
type <- st_is(stock_fao_ohi_shps, "GEOMETRYCOLLECTION")


## Get areas in case we want to later weight the data 
stock_fao_ohi_shps$RAM_area_m2 <- st_area(stock_fao_ohi_shps)

st_geometry(stock_fao_ohi_shps) <- NULL


stock_fao_ohi_mgt <- stock_fao_ohi_shps %>%
  dplyr::select(type_w_ant, rgn_ant_id, F_CODE, assessid, RAM_area_m2) %>%
  group_by(rgn_ant_id, F_CODE, assessid) %>%
  summarize(RAM_area_m2 = sum(RAM_area_m2)) %>%
  rename(rgn_id = rgn_ant_id, fao_id = F_CODE)

## Need to wrangle this so that we have stockid and stocklong (from 2019 ram data v4.44) instead of assessid (from 2017 ram data v3.80)

new_species_list <- read_csv("int/RAM_addl_mgt_stocks.csv")

stock_fao_ohi_mgt_final <- stock_fao_ohi_mgt %>%
  mutate(stockid = gsub("-1.*", "", assessid)) %>% #remove everything after "-1"
  mutate(stockid = gsub("-2.*", "", stockid)) %>% #remove everything after "-2"
  mutate(stockid = gsub("^[^-]*-","",stockid)) %>% #remove everything before the first "-"
  mutate(stockid = gsub("NFLD-", "", stockid)) %>% ## our above efforts did not catch all of the instances of we needed. There is one species with an extra "-" in there, so delete that... 
  left_join(new_species_list, by = "stockid") %>% # join on stockid to get the stocklong names we need...
 dplyr::select(-assessid) %>% #delete assessid
ungroup()

write.csv(stock_fao_ohi_mgt, "int/RAM_fao_ohi_rgns_mgt.csv", row.names=FALSE)

## Now add these to RAM_fao_ohi_rgns.csv file that you created in fao_ohi_rgns.Rmd

RAM_fao_ohi_rgns <- read_csv("int/RAM_fao_ohi_rgns.csv")

initial_mgt_add_fao_ohi_rgns <- rbind(RAM_fao_ohi_rgns, stock_fao_ohi_mgt_final)

### Now we need to manually add the rest of the new species from the management value change by looking each of them up individually and creating a spreadsheet with columns rgn_id, fao_id, RAM_area_m2 (these will all be NA), stockid, and stocklong. Make this csv file and upload to mazu. The code below will help as a check to make sure the rgn ids are in the correct fao rgn.

## Spatial file with fao and ohi regions, F_CODE is the FAO id
fao_ohi <- st_read(dsn = file.path(dir_M, "git-annex/globalprep/fis/v2017/int"),
                   layer = "ohi_fao_rgns")
st_geometry(fao_ohi) <- NULL # removes geometry
fao_ohi_id <- fao_ohi %>%
  select(rgn_id, fao_id = F_CODE) %>% 
  arrange(rgn_id) %>% 
  mutate(fao_id = as.numeric(as.character(fao_id))) %>% 
  mutate(rgn_id = as.numeric(as.character(rgn_id))) %>% 
  distinct()

manual_additions <- data.frame(stockid = sort(setdiff(new_species_list$stockid, initial_mgt_add_fao_ohi_rgns$stockid))) %>% 
  left_join(new_species_list, by = "stockid")


write_csv(manual_additions, file.path("int/manual_management_additions.csv"))

## after manually adding the rest, reupload into "int" folder as manual_management_additions_corr.csv

manual_additions_corr <- read_csv("int/manual_management_additions_corr.csv")

## add the manual additions to our fao_ohi_rgns dataset

RAM_fao_ohi_rgns_final <- rbind(initial_mgt_add_fao_ohi_rgns, manual_additions_corr)

## write this to int folder
write_csv(RAM_fao_ohi_rgns_final, file.path("int/RAM_fao_ohi_rgns_final.csv"))
```

## Gapfill RAM data when there are missing years

For each stock: 
- Missing years are gapfilled using a linear regression model that includes data from 2001 to 2015 (2015 is the final year of Watson data). To be included in the gapfilling, there have to be 5 or more years of B/Bmsy data occuring over the last 11 years of data, from 2005 to 2015.
- We convert any predicted RAM B/Bmsy value less than the minimum observed B/Bmsy value to that the minimum observed value, as there are some negative predicted values.

Summary: 
- There are 397 RAM stocks with at least 5 years of B/Bmsy data from 2005 to 2015. 
- 222 of these stocks have at least 1 year of gapfilled data.  
- A few of the predicted B/Bmsy values go below zero. We convert anything with a RAM B/Bmsy value < 0.0026 to 0.0026, which is the minimum observed B/Bmsy value in the data.

```{r gapfill, eval = FALSE}

## gapfill ram_bmsy
## based on this it seams reasonable to gap-fill missing values

ram_gf_check <- ram_bmsy_new %>%
   filter(year >= 2001) %>%
  spread(year, ram_bmsy) 

# identify stocks for gapfilling (those with 5 or more years of data since 2005).
# NOTE: we potentially gapfill to 2001, but we want stocks with adequate *recent* data 
ram_bmsy_gf <- ram_bmsy_new %>%
  filter(year >= 2001 & year <= 2015) %>%   # 2015 corresponds to the final year of Watson catch data
  group_by(stockid) %>%
  mutate(years_data_2005_now = length(ram_bmsy[year >= 2005])) %>%
  mutate(years_data_2001_now = length(ram_bmsy[year >= 2001])) %>%
  ungroup() %>%
  filter(years_data_2005_now >= 5)


## Get rows for stocks/years with no B/Bmsy (identified as NA B/Bmsy value for now)
ram_bmsy_gf <- ram_bmsy_gf %>%
  spread(year, ram_bmsy) %>% 
  gather("year", "ram_bmsy", -stockid, -years_data_2005_now, -years_data_2001_now, - stocklong) %>%
  mutate(year = as.numeric(year)) 


## gapfilling record keeping
ram_bmsy_gf <- ram_bmsy_gf %>%   
  mutate(gapfilled = NA) %>%
  mutate(gapfilled = ifelse(years_data_2001_now == 15, gapfilled, 
                            paste(15 - years_data_2001_now, "years gf", sep = " ")))

## see unique values of stocks
tmp <- ram_bmsy_gf %>%
  dplyr::select(stockid, gapfilled) %>%
  unique()


## check out gapfilling stats
length(tmp$gapfilled)  # 397 stocks with at least 5 years of data in past 11 years - v2019

sum(table(tmp$gapfilled))  # 222 stocks have at least one year of B/Bmsy values gapfilled - v2019


## regression model for prediction for each stock
ram_bmsy_gf <- ram_bmsy_gf %>%
  group_by(stockid) %>%
  do({
    mod <- lm(ram_bmsy ~ year, data=.)  
    ram_bmsy_predict <- predict(mod, newdata=.[c('year')])
    data.frame(., ram_bmsy_predict)
  }) %>%
  ungroup()

summary(ram_bmsy_gf) #611 NAs for ram_bmsy
sum(ram_bmsy_gf$ram_bmsy_predict < 0 )  # 19 of the predicted B/Bmsy values go below zero.  
min(ram_bmsy_gf$ram_bmsy, na.rm = TRUE) #0.00263

## We convert anything with a RAM BBmsy value < 0.00263 to 0.00263, which is the minimum observed B/Bmsy value in the data; add method documentation
ram_bmsy_gf <- ram_bmsy_gf %>%
  mutate(ram_bmsy_predict = ifelse(ram_bmsy_predict < 0.00263, 0.00263, ram_bmsy_predict)) 

## gapfilling record keeping
ram_bmsy_gf_final <- ram_bmsy_gf %>%
  mutate(method = ifelse(is.na(ram_bmsy), paste0("lm, ", gapfilled), NA)) %>%
  mutate(gapfilled = ifelse(is.na(ram_bmsy), "1", "0")) %>%
  mutate(ram_bmsy = ifelse(is.na(ram_bmsy), ram_bmsy_predict, ram_bmsy)) %>%
  dplyr::select(stockid, year, ram_bmsy, gapfilled, method) 

write.csv(ram_bmsy_gf_final, "int/ram_stock_bmsy_gf.csv", row.names=FALSE)

```

Get a general idea of how well the model predicts missing data based on observed and model predicted values.  This model appears to do fairly well. 
```{r gf error est, eval=FALSE}

plot(ram_bmsy_gf$ram_bmsy, ram_bmsy_gf$ram_bmsy_predict)
abline(0,1, col="red")

plot(log(ram_bmsy_gf$ram_bmsy), log(ram_bmsy_gf$ram_bmsy_predict))
abline(0,1, col="red")

mod <- lm(ram_bmsy ~ ram_bmsy_predict, data=ram_bmsy_gf)
summary(mod)

```


## Identify FAO and OHI regions for RAM stocks

Identify the FAO/OHI regions where each RAM stock is located (fao and ohi regions are assigned to RAM Data in [fao_ohi_rgns.Rmd](https://mazu.nceas.ucsb.edu/rstudio/files/github/ohiprep_v2019/globalprep/fis/v2019/fao_ohi_rgns.html).

If there are many differences between RAM spatial file and RAM metadata, check the fao_ohi_rgns.Rmd prep again. 


```{r final, eval = FALSE}

## Read in RAM spatial stocks file
ram_spatial <- read.csv("int/RAM_fao_ohi_rgns_final.csv", stringsAsFactors = FALSE)

ram_meta <- metadata %>% 
  dplyr::select(stockid, stocklong, scientificname)

setdiff(ram_spatial$stockid, ram_meta$stockid) # make sure all the spatial data has corresponding metadata (should be 0). It is not 0, probably because these are ones that have been removed from the RAM database since the 2017 assessment... delete these from the dataframe. 

# join with metadata to get scientific name
ram_spatial <- ram_spatial %>%
  dplyr::select(-stocklong) %>%
  left_join(ram_meta, by = c("stockid")) %>%
  rename(RAM_species = scientificname)

```

## Standardize species names

In most cases, the RAM and Watson data use the same species names, but there are a few exceptions.  The following code identifies species in the RAM data that are not in the Watson data. In these cases, different species names may be used (although not necessarily because some of the species may be present in RAM, but not Watson, for other reasons). For these species, I used fishbase to explore synonyms and create a table to harmonize the RAM species names with the Watson species names (saved as: int/RAM_species_to_Watson.csv).

```{r name harmonize, eval = FALSE}
ram_bmsy_gf_final <- read_csv(file.path("int/ram_stock_bmsy_gf.csv"))

# get list of RAM species, scientific name
ram_sp <- ram_bmsy_gf_final %>%
  left_join(data.frame(metadata), by = "stockid") %>%
  dplyr::select(scientificname) %>%
  unique() %>%
  arrange(scientificname)


# Watson species, sci name (read in the datatable that includes TaxonKey)
wat_sp <- read_csv(file.path(dir_M,'git-annex/globalprep/fis/v2019/int/stock_catch_by_rgn_taxa.csv')) %>% 
  dplyr::rename(wat_scientificname = TaxonName) %>%
  dplyr::select(wat_scientificname) %>%
  unique() %>%
  arrange(wat_scientificname)

# compare names - what's in RAM that's not in Watson
tmp <- data.frame(scientificname = sort(setdiff(ram_sp$scientificname, wat_sp$wat_scientificname))) # 50 species names

# compare names - what's in watson that's not in RAM
tmp2 <- sort(setdiff(wat_sp$wat_scientificname, ram_sp$scientificname)) # 1197 species names

write.csv(tmp, "int/unmatched_RAM_species.csv", row.names=FALSE)
write.csv(tmp2, "int/Watson_species_no_RAM.csv", row.names=FALSE)

setdiff(tmp, ram_name_corr$RAM_species)


## join ram spatial with RAM species on scientific name. We can use this to help check whether questionable species names across the ram and watson data match by region and fao id...
ram_sp_fao_ohi <- tmp %>%
  left_join(ram_spatial, by = c("scientificname" = "RAM_species")) %>%
  unique()

write_csv(ram_sp_fao_ohi, "int/new_ram_sp.csv")  

## get watson fao_ohi regions
wat_sp_fao_ohi <- read_csv(file.path(dir_M,'git-annex/globalprep/fis/v2019/int/stock_catch_by_rgn_taxa.csv')) %>% 
  dplyr::rename(wat_scientificname = TaxonName) %>%
   dplyr::filter(year > 1979)


# Then I hand-looked up each of the missing ones to generate this list - most still unmatched. See "RAM_species_to_watson_notes.csv" for our reasoning behind these changes. 
ram_name_corr <- read.csv("int/RAM_species_to_Watson.csv", stringsAsFactors = FALSE) %>%
   filter(!is.na(Watson_species))  # Watson to RAM name conversion

ram_name_corr # matched species, only 15
```


### Final formatting

Harmonize names between RAM and Watson data.
```{r, eval=FALSE}

# correct names in a few cases to match with Watson names
ram_name_corr <- read.csv("int/RAM_species_to_Watson.csv", stringsAsFactors = FALSE) %>%
  filter(!is.na(Watson_species))  # Watson to RAM name conversion

ram_spatial <- ram_spatial %>%
  left_join(ram_name_corr, by="RAM_species") %>%
  dplyr::mutate(species = ifelse(!is.na(Watson_species), Watson_species, RAM_species)) %>%
  dplyr::select(rgn_id, fao_id, stockid, stocklong, species, RAM_area_m2)

length(unique(ram_spatial$stockid)) # 452 RAM stocks with B/Bmsy data - v2019
length(unique(ram_spatial$species)) #218
```


Re-name `stockid` column to `stockid_ram` and create new column `stockid` that matches with the `stockid` column in the CMSY data table prepared in [calculate_bbmsy.Rmd](https://cdn.rawgit.com/OHI-Science/ohiprep_v2019/master/globalprep/fis/v2019/calculate_bbmsy.html).

```{r, eval = FALSE}
## Combine RAM spatial data with B/Bmsy data
ram_bmsy_gf <- read.csv("int/ram_stock_bmsy_gf.csv")


# check every stock has a location:
setdiff(ram_bmsy_gf$stockid, ram_spatial$stockid) # should be 0: every ram stock should have ohi/fao rgn
setdiff(ram_spatial$stockid, ram_bmsy_gf$stockid) # these are stocks that were dropped due to insufficient years of data

ram_data <- ram_bmsy_gf %>% 
  left_join(ram_spatial, by="stockid") %>%
  rename(stockid_ram = stockid) %>% 
  dplyr::mutate(stockid = paste(species, fao_id, sep="-")) %>%
  dplyr::mutate(stockid = gsub(" ", "_", stockid)) %>%
  dplyr::select(rgn_id, stockid, stockid_ram, stocklong, year, RAM_area_m2, ram_bmsy, gapfilled, method) %>%
  unique()

write.csv(ram_data, "int/ram_bmsy.csv", row.names=FALSE)
```
