---
title: "Learning about OHI scores"
output: html_document
---

```{r setup, include=FALSE}
## loading packages
library(dplyr)
library(tidyr)
library(tidyverse)

```


## OH NO!
While you are working at your computer at NCEAS, you receive a frantic call from someone working on an OHI project:

THEY NEED HELP!

"No problem" you say, "I'll get Julie or Melanie"...but they are no where to be found!

"Hold on, I'll get Jamie or Casey or Courtney"...but they are also gone.

It is clearly going to be up to you to save the day!!

## You've got this!

Please help them answer the following questions.  You will work from this document, using the code chunks as a work space to work explore the data. Do NOT bother keeping the code chunks neat and organized, we want to see your work process in all its messy glory.  However, I DO recommend including plenty of comments, mostly because I have found this helps me organize my thoughts.   

You will work together to check answers, discuss code, troubleshoot.  But write your own code.  

You can also ask us any questions along the way (slack or issues)!

Good luck! 

## Questions

## {.tabset .tabset-fade .tabset-pills}
### Getting the data

*Goal* Read the latest OHI global data into the R working space (call the data object "scores").

*Hint* Here is a link to the data: 
https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/scores.csv

```{r, include=FALSE}
## Working space

scores <- read.csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/scores.csv")

```

### Region explore

*Question* Why are there N=222 regions (see code below)...aren't there supposed to be 220?

*Hint 1* Here is a link to the official region list: https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv

*Hint 2* Which region has all NA values for scores?
 
*Answer* region_id 213 has all NA values for its score. This is because region_id 213 is Antarctica which is not included in the global assessment. region_id 0 is not in the official region list. It has scores for every observation. Is this the global averages? 

```{r, include=FALSE}
## Working space

length(unique(scores$region_id))


na_df <- scores %>% 
  group_by(region_id) %>%
  filter(is.na(score)) %>%
  summarize(n = length(score))
##^^ Wrong bc it eliminated all region ids without any NAs 

na_df1 <- scores %>% 
  group_by(region_id) %>%
  summarize(n = length(score[is.na(score)]))
## ^^ this gives all region ids with NAs counted. i.e. region 0 

df_213 <- scores %>%
  filter(region_id == 213)  # this one is all NAs

df_0 <- scores %>%
  filter(region_id == 0) #this one has a score for everything
```

*Question* Which region id corresponds to the United States?

### Goal explore

*Question * When I look at the number of goals, N=19 show up....I thought there were supposed to be 10 goals!!! Can you explain what all these are? Also, why do some of the goal abbreviations have 2 letters and others 3 letters?  

*Hint 1* Review the methods doc here: https://raw.githack.com/OHI-Science/ohi-global/published/global2018/Results/Supplement_Results.html

*Answer* The United States region_id is 163. It includes sub goals. 2 Letters are full goals, while 3 letters are sub goals. Not sure what goal == "Index" is. 

```{r, include=FALSE}
## Working space

unique(scores$goal)

#what is index?

index_df <- scores %>%
  filter(goal == "Index")
#what is the index 


```


### Dimension explore

*Question* There are 6 dimensions reported in this dataset (future, pressures, resilience, score, status, and trend).  When I look at the number of records in each of these categories, I notice that there are more "score" values reported than "status" values (see working space below). Given that scores are calculated using status, it seems like these should have the same number of records. Can you figure out what is going on here. Is something going wrong?  
 
*Answer* It looks like the "Index" goal is influencing the number of "score" values. There are values in any of the other dimensions for goal "Index". 

```{r, include=FALSE}
## Working space

unique(scores$dimension)
table(scores$dimension)

table(scores$goal)
table(scores$goal, scores$dimension)  #index is influencing. There are no status values for Index. 

# ok...but why are the different values for pressures/resilience/trend and future?
# addition of region 0!
ao <- filter(scores, goal=="AO")
table(scores$region_id, scores$dimension)

# all of these are the same except for region 0. 
```


### Missing data

*Question* Figure out which goals/subgoals have the most NA values for scores in 2018.  Which ones have the least?  Can you discern any reason why some goals have lots of missing values and others do not?

*Hint* Include only dimension = score, and year = 2018, and cut region_id = 0.

*Answer* Natural Products (NP) has the most NA values for 2018 with 83. Carbon Storage (CS) has the second most NA values, with 73. AO, BD, CW, FIS, ICO, Index, LSP, SP, and SPP all have 1 NA values for 2018. The reason for NP having so many NA values could be tha t some regions do not harvest any of the 6 natural product categories included in the score calculation. The goals with the most NA values could be harder to measure than the goals with the least NA values. 

```{r, include=FALSE}
## Working space

na_data <- scores %>% 
  filter(dimension == "score") %>%
  filter(region_id != 0) %>%
  filter(year == 2018) %>%
  group_by(goal) %>%
  summarise(n = length(score[is.na(score)])) #count all of the NA values 



```

*Question* Artisanal opportunities has no missing data (except for region 213, which is Antarctica).  Look over the description of the artisanal opportunities goal (http://ohi-science.org/ohi-global/goals#artisanal_fishing_opportunity).  Based on this description, in your opinion, should uninhabited islands be given an NA value? 

*Answer* Yes, uninhabited islands should be given an NA value. This goal is calculated using variables like per capita gross domestic product and average annual income. Since these islands are uninhabited, there can be no data to input for them for Artisanal Opportunities. 

### Scores

*Question* If we have a goal with a future status of 80 and status of 90...what is the score?  

*Hint* Isolate the future, status, and score values for one region and one goal and see if you can identify a pattern.

*Answer* Future and status are averaged to get a score, so the score would be 85. 

```{r, include=FALSE}
## Working space

df_future_status <- scores %>%
  filter(dimension %in% c("future", "status", "score"), region_id == 131) #filter for dimension in future, status, score and region_id = mexico. Looks like the future and status scores are just averaged.  

```


### Metadata

*Project* Based on your data exploration and other resources, provide some metadata that describes each variable.  Write it so it would be useful to you in the future as you are looking through these data.

Write it in the following table.  NOTE: Knit the document to see if the table is formatting correctly (but don't bother if you do not know how to knit a document or if you are running into errors!).

Variable   | Description                  | Values
---------- | ---------------------------- | ------------------------------
goal       | The category of ocean-derived benefit. Abbreviations with 2 letters are goals, while abbreviations with 3 letters are subgoals, which are used to calculate the goals.                              | AO, BD, CP, CS, CW, ECO, FIS, FP, HAB, ICO, Index, LE, LIV, LSP, MAR, NP, SP, SPP, TR
dimension  | Variables influencing goal calculation.                             | future, pressures, resilience, score, status, trend
region_id  | Numerical value which identifies the region or country in the assessment. | 0 to 250 
score      | The OHI score associated with an observation.                             |    0 to 100      
year       |  Scenario year               | 2012 to 2019 


```{r, include=FALSE}
## Working space

summary(scores)
unique(scores$goal) # get all goal values
unique(scores$dimension)
```


### Plot

*Project* Create a scatterplot that compares 2012 and 2018 *scores* for each region for the artisanal opportunities goal. Based on this, do scores for the artisanal opportunities goal appear to have increased or decreased over time? 

*Answer* 

```{r, include=FALSE}
## Working space

scores_12_18 <- scores %>%
  filter(year %in% c(2012, 2018), goal == "AO", dimension == "score")

ggplot(scores_12_18, aes(x = region_id, y = score)) +
  geom_point(aes(color = factor(year))) #<< this one seems to be the most inciteful graph among the three. 

ggplot(scores_12_18, aes(x = year, y = score)) +
  geom_point(aes(color = factor(region_id))) + 
  theme(legend.position = "none") #doesnt tell us anything at all 

scores_2012 <- scores %>%
  filter(year == 2012, goal == "AO", dimension == "score")

scores_2018 <- scores %>%
  filter(year == 2018, goal == "AO", dimension == "score")

plot(scores_2012$score, scores_2018$score) #doesnt tell us much 

# make two new columns for the scores corresponding to 2012 and 2018
new_scores_df <- scores_12_18 %>%
  spread(year, score) %>%
  mutate(difference = `2018` - `2012`)  

ggplot(new_scores_df, aes(x = `2012`, y = `2018`, size = difference)) +
  geom_point() +
  theme()
```

```{r, echo = FALSE, warning = FALSE}
ggplot(new_scores_df, aes(x = region_id, y = difference)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0, color = "red")+
  theme_bw() +
  coord_flip() + 
  labs(y = "Score Difference between 2018 and 2012", x = "Region")

ggplot(new_scores_df, aes(x = region_id, y = difference, color = case_when(
  difference > 0 ~ "green",
  difference < 0 ~ "red",
  difference == 0 ~ "blue"
))) +
  geom_point() +
  geom_hline(yintercept = 0, color = "black")+
  theme_bw() +
  coord_flip() + 
  labs(y = "Score Difference between 2018 and 2012", x = "Region") +
  theme(legend.position = "none")
```

**Figure 1.** The majority of the scores in 2018 are greater than the scores in 2012 for each region. This is seen in the plot, which shows the difference between the scores for each score. The two charts above display the same information. 
