---
title: "Learning about OHI scores"
output: html_document
---

```{r setup, include=FALSE}
## loading packages
library(tidyverse)
library(dplyr)
library(tidyr)

filter <- dplyr::filter
```


## OH NO!
While you are working at your computer at NCEAS, you receive a frantic call from someone working on an OHI project:

THEY NEED HELP!

"No problem" you say, "I'll get Julie or Melanie"...but they are no where to be found!

"Hold on, I'll get Jamie or Casey or Courtney"...but they are also gone.

It is clearly going to be up to you to save the day!!

## You've got this!

Please help them answer the following questions.  You will work from this document, using the code chunks as a work space to work explore the data. Do NOT bother keeping the code chunks neat and organized, we want to see your work process in all its messy glory.  However, I DO recommend including plenty of comments, mostly because I have found this helps me organize my thoughts.   

You will work together to check answers, discuss code, troubleshoot.  But write your own code.  

You can also ask us any questions along the way (slack or issues)!

Good luck!

## Questions

### Getting the data

*Goal* Read the latest OHI global data into the R working space (call the data object "scores").

*Hint* Here is a link to the data: 
https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/scores.csv

```{r, include=FALSE}
## Working space

# Before even starting, we needed to make sure we were invited to the repository and that we were logged into Github through the RStudio server. My username is "mollywillams", not mollywilliams >.<

scores <- read.csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/scores.csv")

```

### Region explore

*Question* Why are there N=222 regions (see code below)...aren't there supposed to be 220?

*Hint 1* Here is a link to the official region list: https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv

*Hint 2* Which region has all NA values for scores?

*Answer* Antarctica (Region 213) does not have score data associated with it, and there are data for region_id=0, which is not included in the list of regions linked above (maybe this is the average?)

```{r, include=FALSE}
## Working space

length(unique(scores$region_id))

# Create a data frame that just filters the regions that have NA in the score category
scores_na <- scores %>%
  filter(is.na(score)) %>% 
  group_by(region_id) %>% 
  tally()

# When organized by "n" (the number of total NA observations for that region), the data frame shows that Region 213 had 880 NA values, almost twice as high as any other region. This led us to believe that this was the region with all NA values for scores. Filtering just this region confirmed this guess:

all_na <- scores %>% 
  filter(region_id == "213")

# Region 213 is Antarctica, and is not included in the assessment. 

```

*Question* Which region id corresponds to the United States?

### Goal explore

*Question * When I look at the number of goals, N=19 show up....I thought there were supposed to be 10 goals!!! Can you explain what all these are? Also, why do some of the goal abbreviations have 2 letters and others 3 letters?  

*Hint 1* Review the methods doc here: https://raw.githack.com/OHI-Science/ohi-global/published/global2018/Results/Supplement_Results.html

*Answer* The US is Region ID 163. The 10 main goals have two-letter abbreviations, and the 8 subgoals (not every goal has a subgoal) have three-letter abbreviations. Each regigon also has an associated "Index" score.

```{r, include=FALSE}
## Working space

unique(scores$goal) # N = 19


```


### Dimension explore

*Question* There are 6 dimensions reported in this dataset (future, pressures, resilience, score, status, and trend).  When I look at the number of records in each of these categories, I notice that there are more "score" values reported than "status" values (see working space below). Given that scores are calculated using status, it seems like these should have the same number of records. Can you figure out what is going on here. Is something going wrong?  

*Answer* There are more score values than status values, because of the "Index" scores. The Index score for each region does not have a corresponding status value, only a score. This makes there appear to be more score values. 

```{r, include=FALSE}
## Working space

unique(scores$dimension)
table(scores$dimension)

table(scores$goal)
table(scores$goal, scores$dimension)

# ok...but why are the different values for pressures/resilience/trend and future?
# addition of region 0!
ao <- filter(scores, goal=="AO")
table(scores$region_id, scores$dimension)

# All of these are the same except for region zero - which is the average for future and score dimensions across regions (?)
```


### Missing data

*Question* Figure out which goals/subgoals have the most NA values for scores in 2018.  Which ones have the least?  Can you discern any reason why some goals have lots of missing values and others do not?

*Hint* Include only dimension = score, and year = 2018, and cut region_id = 0.

*Answer* With 83, the natural products goal had the most NA values in the 2018 scores. The following goals/subgoals all had just one NA value: artisanal fishing, biodiversity, clean waters, fishing, iconic species (sense of place subgoal), lasting special places (sense of place subgoals), and species condition (biodiversity subgoal). Along with natural products, carbon storage and coastal protection had high numbers of NA values. These metrics may be difficult to measure, and analogs may not readily exist for gap-filling the data. 

```{r, include=FALSE}
## Working space

na_data <- scores %>% 
  filter(dimension == "score") %>%
  filter(region_id != 0) %>%
  filter(year == 2018) %>%  # added the following commands to see NA counts by goal:
  filter(is.na(score)) %>% 
  group_by(goal) %>% 
  tally()
  

```

*Question* Artisanal opportunities has no missing data (except for region 213, which is Antarctica).  Look over the description of the artisanal opportunities goal (http://ohi-science.org/ohi-global/goals#artisanal_fishing_opportunity).  Based on this description, in your opinion, should uninhabited islands be given an NA value? 

*Answer* Yes - the score directly assesses the needs of the inhabitants, so if there are no inhabitants, the score should be NA. Further, the values that are incorporated into this score are not relevant for an uninhabited region (economic need, social progress, governance, etc)

### Scores

*Question* If we have a goal with a future status of 80 and status of 90...what is the score?  

*Hint* Isolate the future, status, and score values for one region and one goal and see if you can identify a pattern. *see comments in code chunk below!*

*Answer* Current and future status are averaged to get the score, so a goal with a future status of 80 and a status of 90 would have a score of 85.
```{r, include=FALSE}
## Working space


iran_future <- scores %>% 
  filter(region_id == 191) %>% 
  filter(dimension == "future")

ggplot(iran_future, aes(x = year, y = score)) +
  geom_line(aes(color = goal)) +
  theme_minimal()

# For Iran (region 191) the scores under the future dimension have remained consistent over time for most goals. Natural products declined sharply between 2016 and 2017 (from 100 to <25), while scores for mariculture have risen gradually (from ~10 to . Tourism and recreation also declined beween 2017 and 2018. These trends are also visible when looking at scores over time: 


iran_status <- scores %>% 
  filter(region_id == 191) %>% 
  filter(dimension == "status")

ggplot(iran_status, aes(x = year, y = score)) +
  geom_line(aes(color = goal)) +
  theme_minimal()

# Same trends are visible when looking at the score dimension over time (makes sense): 

iran_score <- scores %>% 
  filter(region_id == 191) %>% 
  filter(dimension == "score")

ggplot(iran_score, aes(x = year, y = score)) +
  geom_line(aes(color = goal)) +
  theme_minimal()



```


### Metadata

*Project* Based on your data exploration and other resources, provide some metadata that describes each variable.  Write it so it would be useful to you in the future as you are looking through these data.

Write it in the following table.  NOTE: Knit the document to see if the table is formatting correctly (but don't bother if you do not know how to knit a document or if you are running into errors!).

Variable   | Description                  | Values
---------- | ---------------------------- | ------------------------------
goal       | Category of ocean-derived benefit  | AO, BD, CP, CS, CW, ECO, FIS, FP, HAB, ICO, LE, LIV, LSP, MAR, NP, SP, SPP, TR (see table below for abbreviation classification)
dimension  |  Variables influencing goal calculation  | future, pressures, resilience, score, status, trend
region_id  |  Numerical value assigned to each country included in the assessment          | 1-250, with 30 IDs eliminated throughout for disputed and landlocked regions
score      |  Final value for each region, averaged between current and future status | 0-100         
year       |  Scenario year               | 2012 to 2019 


Goal/Subgoal   | Abbreviation                  | Classification
---------- | ---------------------------- | ------------------------------
Artisanal Fishing Opportunities | AO  | Goal
Biodiversity    | BD   | Goal 
Carbon storage  | CS | Goal
Coastal livelihoods and economies | LE | Goa
Coastal protection    | CP   | Goal 
Clean waters  | CW | Goal
Economies    | ECO   | Subgoal, LE 
Fisheries | FIS | Subgoal, FP
Food provision | FP | Goal
Habitat | HAB | Subgoal, BD
Iconic species | ICO | Subgoal, SP
Lasting special places | LSP | Subgoal, SP
Livelihoods | LIV | Subgoal, LE
Mariculture | MAR | Subgoal, FP
Natural products | NP | Goal
Sense of place | SP | Goal
Species condition | SPP | Subgoal, BD
Tourism and recreation | TR | Goal


```{r, include=FALSE}
## Working space

summary(scores)

```


### Plot

*Project* Create a scatterplot that compares 2012 and 2018 *scores* for each region for the artisanal opportunities goal. Based on this, do scores for the artisanal opportunities goal appear to have increased or decreased over time? 

*Answer* See plot below:

```{r, include=FALSE}
## Working space

# Create a new dataframe that just has AO scores for 2012 and 2018, reorganized into columnns by year:
ao_12_18 <- scores %>% 
  filter(year == "2012" | year == "2018") %>% 
  filter(goal == "AO") %>% 
  filter(dimension == "score") %>% 
  spread(year, score)

# Eliminate unnecessary columns, add new column for score change between years:

ao_12_18_reorg <- ao_12_18 %>%
  select(region_id, "2012", "2018") %>% 
  mutate(difference = `2018`-`2012`)
  


```


```{r, echo=FALSE}

ggplot(ao_12_18_reorg, aes(x = difference, y = region_id, color = case_when(difference < 0 ~"green", difference > 0 ~"red"))) +
  geom_point() +
  geom_vline(xintercept = 0) + 
  theme_minimal() +
  theme(legend.position="none")
```

Scatter plot showing change in overall OHI score between 2012 and 2018 for all regions. The majority of the values are positive (in red), indicating that most regions experienced a positive change in scores over this time period.

