---
title: 'OHI 2019 - Fisheries Management Index (Resilience)'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
editor_options: 
  chunk_output_type: console
---


# Summary
This document outlines the process for developing linear models to gapfill fisheries management index scores to use as a resilience layer for the 2019 global assessment.

# Data Sources

The following data are used:

## Fisheries Management Index Data 

[Fisheries Management Index](https://oursharedseas.com/2019-update/fisheries/#highchart-fisheries-fmindex) data were first introduced in a 2017 paper by Melnychuk et al., *[Fisheries Management Impacts on Target Species Status](https://doi.org/10.1073/pnas.1609915114)*. Scores range from 0-1 and rate management effectiveness of fisheries in distinct regions on a stock-by-stock basis. The scores are determined by expert surveys that characterize attributes of research, management, enforcement, and socioeconomic factors. The first survey was conducted in 2016 in 28 major fishing countries that collectively account for >80% of global catch. Another survey was performed in 2018 on 40 countries, the scores from which we are using for this OHI resilience layer. 

**Date retreived:** 12 July 2019

**Method:** Data are not accessible in csv format from website, so points were manually entered into excel and saved as a csv (found in v2019/raw). Because FMI scores only exist for 40 out of the 220 OHI regions, we trained linear models using scores from the AO need layer (rescaled GDP per capita per person purchasing power), GDP per capita, World Governance Index (WGI), Social Progress Indicator (SPI), and UN georegion labels. After comparing the models, we determined SPI and UN georegions to be the best predictors of FMI, and used these models to gapfill FMI scores for the remaining regions.


## GDP per capita 

Fisheries management index scores are highly correlated with GDP, so GDP per capita per person purchasing power (gdppcpppp) was used to create linear regression models to gapfill FMI data. We used World Bank data (reported at country-scale) and gapfilled any missing GDP values using CIA data. 

[World bank](http://data.worldbank.org/indicator/NY.GDP.PCAP.PP.KD) 
* Downloaded 7/24/19
* Time range: 1990-2018

[CIA World Factbook](https://www.cia.gov/library/publications/the-world-factbook/rankorder/2004rank.html)
* Downloaded 7/24/19
* Time range: 


## SPI

**Citation**: http://www.socialprogress.org/

Stern, S., A. Wares and T. Epner. 2018. Social Progress Index: 2018 Methodology Report.

**Source information**: 
http://www.socialprogress.org/ --> Download Data

**Date Downloaded**: 9/21/2018

**Time range**: 2014-2018

**Native data resolution**: country scores

**Format**:  Excel file

**Description**: Social Progress Index scores and components for countries.

## WGI

**Reference**:  http://info.worldbank.org/governance/wgi/index.aspx#home    

**Downloaded**: March 11 2019 (data updated Sep 21 2018)

**Description**:  
The Worldwide Governance Indicators (WGI) project reports aggregate and individual governance indicators for 215 economies over the period 1996–2017, for six dimensions of governance:

* Voice and Accountability
* Political Stability and Absence of Violence
* Government Effectiveness
* Regulatory Quality
* Rule of Law
* Control of Corruption

**Time range**: 1996-2017


## UN georegions

UNgeorgn() loads a dataframe from common.R with UN geopolitical designations, and is commonly used in OHI to gapfill missing data. The distinct regions are derived from the [United Nations Statistics Division](https://unstats.un.org/unsd/methodology/m49/)Each region is assigned four labels with increasing granularity/specificity: r0_label = World (1 level), r1_label = continental regions (7 levels: Oceania, Asia, Africa, Europe, Southern Islands, Latin America and the Caribbean, and Americas), r2_label = georegions (22 levels for subregions and intermediary regions).


# Updates from previous assessment
These data have not been updated since 2013, so this is an entirely new method for establishing resilience values. 



# Initial set-up code

```{r setup, message=FALSE, warning=FALSE}

#library(devtools)
#devtools::install_github("ohi-science/ohicore@dev")
library(ohicore)
library(tidyverse)
library(stringr)
library(WDI) # for accessing World Bank data 
library(here) 
library(plotly)
library(psych) # for correlation testing

source('https://raw.githubusercontent.com/OHI-Science/ohiprep_v2019/gh-pages/workflow/R/common.R')



```

# Load and wrangle FMI data 

```{r}

fmi_raw <- read_csv(here("globalprep/res_fmi/v2019/raw/FMI_data_raw.csv")) %>%
  rename("2016" = fmi_2016) %>% 
  rename("2018" = fmi_2018) %>% 
  gather(key = "year", value = "fmi", -country)

# Add region ID
fmi_rgn <- name_2_rgn(df_in = fmi_raw, 
                       fld_name='country', 
                       flds_unique=c('fmi', 'year')) %>% 
  select(rgn_id, rgn_name, year, fmi) %>% 
  filter(year == 2018) # remove 2016 points so that they don't skew the model

fmi_rgn$year <- as.numeric(fmi_rgn$year)

```

# Create linear regression models using gapfilled GDP data

## Correlation and model based on gapfilled, rescaled "score" data (calculated from gdppcppp, AO output product)
```{r}

ao_score <- read_csv(here("globalprep/ao/v2019/output/wb_gdppcppp_rescaled.csv"))

fmi_ao <- fmi_rgn %>% 
  left_join(ao_score, by=c("rgn_id","year")) %>% 
  select(-rgn_name) %>% 
  rename(ao_need_score = value)

# Look at correlation 

ggplotly(ggplot(fmi_ao, aes(x = ao_need_score, y = fmi, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red"))

fmi_ao_plot <- fmi_ao %>% 
  select(fmi, ao_need_score)
  
pairs.panels(fmi_ao_plot, density = TRUE, cor=TRUE, lm=TRUE)

# Take logs and create scatter
# Log transformations adjust for skew/exponential relationships
log_fmi_ao <- log(fmi_ao_plot+1) %>% 
  rename(log_fmi=fmi) %>% 
  rename(log_ao_need_score=ao_need_score)

pairs.panels(log_fmi_ao, density = TRUE, cor=FALSE, lm=TRUE)

# Correlation test
cor.test(log_fmi_ao$log_fmi, log_fmi_ao$log_ao_need_score, method="pearson", alternative="two.sided")
# p <0.001, reject null that true correlation=0
# cor = ~0.597 

model_ao <- lm(fmi ~ ao_need_score, data=fmi_ao)
summary(model_ao)
anova(model_ao)
# p<0.001 for ao_need_score, reject null ; AO need is significant predictor of FMI 
# R2 = 0.3971
# plot(model)

fmi_gf_ao <- fmi_ao %>%
  dplyr::group_by(rgn_id) %>%
  dplyr::do({ 
    fmi_pred <- predict(model, newdata =.[c('ao_need_score')]) 
    data.frame(., fmi_pred) # do loop applies the model fitting and prediction to each country group
  }) %>% 
  dplyr::ungroup()


```

## Model based on Social Progress Index (SPI)
```{r}
spi <- read_csv(here("globalprep/prs_res_spi/v2018/output/spi_res.csv"))
# can update this for 2019 once we complete this layer 

fmi_spi <- fmi_rgn %>% 
  left_join(spi, by=c("rgn_id","year")) %>% 
  select(-rgn_name)

# Look at correlation
ggplotly(ggplot(fmi_spi, aes(x = resilience_score, y = fmi, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red"))

fmi_spi_plot <- fmi_spi %>% 
  select(fmi, resilience_score)
  
pairs.panels(fmi_spi_plot, density = TRUE, cor=TRUE, lm=TRUE)

# Take logs and create scatter
# Log transformations adjust for skew/exponential relationships
log_fmi_spi <- log(fmi_spi_plot+1) %>% 
  rename(log_fmi=fmi) %>% 
  rename(log_resilience_score=resilience_score)

pairs.panels(log_fmi_spi, density = TRUE, cor=FALSE, lm=TRUE)

# Correlation test
cor.test(log_fmi_spi$log_fmi, log_fmi_spi$log_resilience_score, method="pearson", alternative="two.sided")
# p <0.001, reject null that true correlation=0
# cor = ~0.568 (decent strong positve)

model_spi <- lm(fmi ~ resilience_score, data=fmi_spi)
summary(model_spi)
anova(model_spi)
# p<0.001 for resilience_score, reject null ; SPI resilience score is significant predictor of FMI 
# R2 = 0.4204
# plot(model)

fmi_gf_spi <- fmi_spi %>%
  dplyr::group_by(rgn_id) %>%
  dplyr::do({ 
    fmi_pred <- predict(model, newdata =.[c('resilience_score')]) 
    data.frame(., fmi_pred) # do loop applies the model fitting and prediction to each country group
  }) %>% 
  dplyr::ungroup()

```

## Model based on world governance indicator score
**Note: v2019 data go through 2017**
```{r}
wgi <- read_csv(here("globalprep/prs_res_wgi/v2019/output/wgi_res.csv")) %>% 
dplyr::mutate(year = ifelse(stringr::str_detect(year,"2017"), "2018", year)) # making data from 2017 2018 instead so that it will merge with FMI data 

wgi$year <- as.numeric(wgi$year)

fmi_wgi <- fmi_rgn %>% 
  left_join(wgi, by=c("rgn_id","year")) %>% 
  select(-rgn_name)

# Look at correlation
ggplotly(ggplot(fmi_wgi, aes(x = resilience_score, y = fmi, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red"))

fmi_wgi_plot <- fmi_wgi %>% 
  select(fmi, resilience_score)

pairs.panels(fmi_wgi_plot, density = TRUE, cor=TRUE, lm=TRUE)

# Take logs and create scatter
# Log transformations adjust for skew/exponential relationships
log_fmi_wgi <- log(fmi_wgi_plot+1) %>% 
  rename(log_fmi=fmi) %>% 
  rename(log_resilience_score=resilience_score)

pairs.panels(log_fmi_wgi, density = TRUE, cor=FALSE, lm=TRUE)

# Correlation test
cor.test(log_fmi_wgi$log_fmi, log_fmi_wgi$log_resilience_score, method="pearson", alternative="two.sided")
# p <0.001, reject null that true correlation=0
# cor = ~0.585

model_wgi <- lm(fmi ~ resilience_score, data=fmi_wgi)
summary(model_wgi)
anova(model_wgi)
# p<0.001 for resilience_score, reject null ; SPI resilience score is significant predictor of FMI 
# R2 = 0.4298
# plot(model)

fmi_gf_wgi <- fmi_wgi %>%
  dplyr::group_by(rgn_id) %>%
  dplyr::do({ 
    fmi_pred <- predict(model, newdata =.[c('resilience_score')]) 
    data.frame(., fmi_pred) # do loop applies the model fitting and prediction to each country group
  }) %>% 
  dplyr::ungroup()
```


**How can we test to determine an acceptable level of difference between predicted an actual for the data we do have?**


## Model based on GDP per capita (constant 2010 US$; alternative to use current US$)
Source: https://data.worldbank.org/indicator/NY.GDP.PCAP.KD
**Can skip if you have already run this chunk and saved GDP data*
```{r}

# Extract data from WDI()

# check world bank website to see what years are available
yr_start = 1960
yr_end   = 2018


# get description of variables (NOTE: these descriptions appear out of date, they aren't in sync with the definitions of the World Bank):
indicators <-  data.frame(WDI_data[[1]])

# Get information about dataset 
indicators[grep("NY.GDP.PCAP.KD", indicators$indicator), ]  # constant dollars. grep helps identify rows to select based on a string. (used this data)


# download the data using the WDI package 
gdppc_raw <-  WDI(country = "all",
               indicator = "NY.GDP.PCAP.KD", 
               start = yr_start, end=yr_end)
summary(gdppc_raw)

# Can add data saving/re-importing steps in between here if we decide to use this data in the future

# Clean and wrangle GDP data
gdppc_clean <- gdppc_raw %>% 
  dplyr::select(country, value=NY.GDP.PCAP.KD, year) %>%
  dplyr::filter(year >= 2005) %>%
  tidyr::spread(year, value) %>%
    # spread to fill in potentially missing values with NA
  data.frame() %>% # this will add an X in front of the column names, allowing us to gather the values
  tidyr::gather(year, value, starts_with("X")) %>%
  dplyr::mutate(year = gsub("X", "", year)) %>% #substitute X for "" (nothing) in the column year
  dplyr::mutate(year = as.numeric(year)) #convert the year column into a numeric format

summary(gdppc_clean) # 273 NAs

# For the first gapfilling stage, if a region has only one value use this value for all years
gdppc_val <- gdppc_clean %>%
  dplyr::group_by(country) %>%
  dplyr::mutate(value_num = sum(!is.na(value))) %>% # counts the numbers of non-missing values for each country (logical TRUEs regarded as one)
  dplyr::filter(value_num > 0) %>%    # filter out the countries with no data between 2005 and 2018 
  dplyr::mutate(value_num_gf = ifelse(value_num==1, mean(value, na.rm=TRUE), NA)) %>%  # mean() function is used on regions with one year of data, applies that single value to all NAs for that region 
  dplyr::ungroup() %>%
  dplyr::mutate(value = ifelse(is.na(value), value_num_gf, value)) %>% # if no value is missing, leave it, otherwise gapfill
  dplyr::select(country, year, value, value_num)  # select just these columns; to eliminate extraneous value_num_gf column

summary(gdppc_val) # now only 39 NAs ! 


# Predict values using a linear regression with 'year' as an independent variable 
# Create new column with these predicted values
gdppc_gf <- gdppc_val %>%
  dplyr::group_by(country) %>%
  dplyr::do({ 
    mod <- lm(value ~ year, data =.)
    value_pred <- predict(mod, newdata =.[c('year')]) # value_pred = country-grouped mod$fitted.values?
    data.frame(., value_pred) # do loop applies the model fitting and prediction to each country group
  }) %>% 
  dplyr::ungroup()

summary(gdppc_gf)

# Fill in the remaining NA values using the predicted values
gdppc_gf2 <- gdppc_gf %>%
  dplyr::ungroup() %>% # why would we need to ungroup here?
  dplyr::mutate(gapfilled = ifelse(is.na(value), 1, 0)) %>% # Create column 'gapfilled', if value is currently NA, it will be gapfilled, indicated by a 1 in the gapfill column
  dplyr::mutate(gapfilled = ifelse(value_num == 1, 1, gapfilled)) %>% # if value_num is 1 it was gapfilled previously and gets a 1 in the gapfill column
  dplyr::mutate(value = ifelse(is.na(value), value_pred, value)) %>% # if NA in value column, input the value in value_pred column
  dplyr::mutate(method = ifelse(gapfilled==1, paste("lm based on N years data:", value_num, sep=" "), NA)) %>% # Create column 'method' that indicates method of gapfilling; this puts message "lm based..." even in some rows gapfilled with one year of data
  dplyr::mutate(method = ifelse(value_num == 1, "gapfilled using one year of data", method)) # this overwrites/corrects method "lm based..." for rows actually gapfilled with "one-year of data"" method
  
summary(gdppc_gf2) # no more NAs because everything has been gap-filled.

## Rescale values
# Values at the 95th Quantile or greater are given a rescaled score of '1' (the highest value)
gdppc_rescale <- gdppc_gf2 %>%
  dplyr::mutate(quantile_95 = quantile(value, probs=0.95)) %>% # gives a single value - the 95th quant (v2019=53582.14; without current year of data quant95=53208.67 - these should be different)
  dplyr::mutate(value_stand = value/quantile_95) %>% # where does value scale relative to 95th quantile
  dplyr::mutate(value_stand = ifelse(value_stand > 1, 1, value_stand)) %>% 
  dplyr::select(country, year, value, score=value_stand, gapfilled, method) # rename value_stand 'score'

summary(gdppc_rescale)

## Convert to OHI regions
d_stand_rgn <- name_2_rgn(df_in = gdppc_rescale, 
                       fld_name='country', 
                       flds_unique=c('year'))


# Combine the duplicate regions (we report these at lower resolution)
# In this case, we take the average score weighted by population
population_weights <- read.csv('https://raw.githubusercontent.com/OHI-Science/ohiprep_v2017/master/src/LookupTables/Pop_weight_ChinaSAR_USVIslPRico.csv')


# Weight the `score`, `value`, and `gapfilled` column by population
d_stand_rgn <- d_stand_rgn %>%
  dplyr::left_join(population_weights, by="country") %>% # does it make sense to backfill population data with static data?
  dplyr::mutate(population = ifelse(is.na(population), 1, population)) %>% # If no value available, input 1 (these values will not change)
  dplyr::group_by(rgn_id, year, method, gapfilled) %>% 
  dplyr::summarize(score = weighted.mean(score, population), # weight the single score value by pop.
            value = weighted.mean(value, population)) %>%
  ungroup() 

# check again:
data.frame(filter(d_stand_rgn, rgn_id == 209))


# Removed `Azerbaijan` (255) because the adjacent body of water is a sea not the ocean
d_stand_rgn <- d_stand_rgn %>%
  filter(rgn_id <= 250)

summary(d_stand_rgn) # no NAs


# save simplified cleaned gdppc data for existing regions
gdppc_data <- d_stand_rgn %>%
  select(rgn_id, year, value)

write_csv(gdppc_data, here("globalprep/res_fmi/v2019/intermediate/gdppc_ohi.csv"))



## Gapfilling regions with no GDP per capita data using UN georegions 

# Re-import cleaned/gapfilled gdppc data if you've already completed the above steps 
gdppc_data <- read_csv(here("globalprep/res_fmi/v2019/intermediate/gdppc_ohi.csv"))


# Create dataframe pairing each UN geopolitical region id with a year from 2005 to current
UNgeorgn()  

# Assign georegion labels to each region for each level (r0, r1, r2)
d_stand_gf <- data.frame(year=min(d_stand_rgn$year):max(d_stand_rgn$year)) %>% 
  merge(UNgeorgn, by=NULL) 

# Combine the two data frames by region id and year
# Calculate means across increasing geopolitical levels (e.g. r2, r1), using the highest resolution possible
d_stand_gf <- d_stand_gf %>%  
  left_join(d_stand_rgn, by = c("rgn_id", "year")) %>%
  group_by(r2_label, year) %>%
  mutate(r2_value = mean(score, na.rm=TRUE)) %>%
  ungroup() %>%
  group_by(r1_label, year) %>%
  mutate(r1_value = mean(score, na.rm=TRUE)) %>%
  ungroup() %>%
  group_by(r0_label, year) %>%
  mutate(r0_value = mean(score, na.rm=TRUE)) %>%
  ungroup()
summary(d_stand_gf) 

# Create vector to filter out low population regions so that they are NA 
low_pop()
low_pop <- low_pop %>%
  filter(est_population < 3000 | is.na(est_population)) #filter out regions that have populations > 3000 and keep NA values 
low_pop_vector <- c(low_pop$rgn_id) #make a vector of low population areas 

# For `score` cells that still have NA values (still several hundred):
# Check to see if r2 has a value, if so use that to gapfill `score`, otherwise use r1, otherwise use r0
d_stand_gf <- d_stand_gf %>%
  mutate(gapfilled = ifelse(is.na(score) & !is.na(r2_value), "1", gapfilled)) %>%
  mutate(method = ifelse(is.na(score) & !is.na(r2_value), "UN_geopolitical region avg, r2", method)) %>%
  mutate(score = ifelse(is.na(score), r2_value, score)) %>%
  mutate(gapfilled = ifelse(is.na(score) & !is.na(r1_value), "1", gapfilled)) %>%
  mutate(method = ifelse(is.na(score) & !is.na(r1_value), "UN_geopolitical region avg, r1", method)) %>%
  mutate(score = ifelse(is.na(score), r1_value, score)) %>%
  mutate(gapfilled = ifelse(is.na(score) & !is.na(r0_value), "1", gapfilled)) %>%
  mutate(method = ifelse(is.na(score) & !is.na(r0_value), "UN_geopolitical region avg, r0", method)) %>%
  mutate(score = ifelse(is.na(score), r0_value, score))
  

d_stand_gf$score[d_stand_gf$rgn_id %in% low_pop_vector] <- NA
#should now have NA values in score column for low popuation areas


# Save complete GDP per capita data 

final_gdppc <- d_stand_gf %>%
  select(rgn_id, year, value=score)

write_csv(final_gdppc, here("globalprep/res_fmi/v2019/output/wb_gdppc_rescaled.csv"))


# Save dataframe with gapfilled method and status information
final_gdppc_gf <- d_stand_gf %>%
  select(rgn_id, year, gapfilled, method)

write_csv(final_gdppc_gf, here("globalprep/res_fmi/v2019/output/wb_gdppc_rescaled_gf.csv"))
```


# Create linear model based on GDP per capita data 

```{r}
gdppc <- read_csv(here("globalprep/res_fmi/v2019/output/wb_gdppc_rescaled.csv"))

fmi_gdppc <- fmi_rgn %>% 
  left_join(gdppc, by=c("rgn_id","year")) %>% 
  select(-rgn_name) %>% 
  rename(gdppc = value)


# Look at correlation
ggplotly(ggplot(fmi_gdppc, aes(x = gdppc, y = fmi, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red"))

fmi_gdppc_plot <- fmi_gdppc %>% 
  select(fmi, gdppc)
  
pairs.panels(fmi_gdppc_plot, density = TRUE, cor=TRUE, lm=TRUE)

# Take logs and create scatter
# Log transformations adjust for skew/exponential relationships
log_fmi_gdppc <- log(fmi_gdppc_plot+1) %>% 
  rename(log_fmi=fmi) %>% 
  rename(log_gdppc=gdppc)

pairs.panels(log_fmi_gdppc, density = TRUE, cor=FALSE, lm=TRUE)

# Correlation test
cor.test(log_fmi_gdppc$log_fmi, log_fmi_gdppc$log_gdppc, method="pearson", alternative="two.sided")
# p <0.0001, reject null that true correlation=0
# cor = ~0.604

model_gdppc <- lm(fmi ~ gdppc, data=fmi_gdppc)
summary(model_gdppc)
# p<0.001 for gdppc, reject null ; GDP per capita is significant predictor of FMI 
# R2 = 0.3935 (but not a very strong predictor)
# plot(model)

# Create array of predicted FMI values 
fmi_gf_gdppc <- fmi_gdppc %>%
  dplyr::group_by(rgn_id) %>%
  dplyr::do({ 
    fmi_pred <- predict(model_gdppc, newdata =.[c('gdppc')]) 
    data.frame(., fmi_pred) # do loop applies the model fitting and prediction to each country group
  }) %>% 
  dplyr::ungroup()

# Plot predicted vs actual values 
ggplotly(ggplot(fmi_gf_gdppc, aes(x = fmi, y = fmi_pred, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red"))

```


# Create linear model based on UN georegions 

```{r}
source('https://raw.githubusercontent.com/OHI-Science/ohiprep_v2019/gh-pages/workflow/R/common.R')

# Create dataframe pairing each UN geopolitical region id with a year from 2005 to current
UNgeorgn()
georegions <- UNgeorgn %>% 
  select(rgn_id, rgn_name = rgn_label, r0_label, r1_label, r2_label)


fmi_georegions <- fmi_rgn %>% 
  left_join(georegions, by=c("rgn_id","rgn_name"))


model_gr <- lm(fmi ~ r2_label, data=fmi_georegions)
summary(model_gr)
# p<0.001 for r2_label, reject null ; smaller georegions are significant predictor of FMI 
# R2 = 0.5951 (pretty good!)
# plot(model_gr)


# Create array of predicted FMI values 
fmi_gf_gr <- fmi_georegions %>%
  dplyr::group_by(rgn_id) %>%
  dplyr::do({ 
    fmi_pred <- predict(model_gr, newdata =.[c('r2_label')]) 
    data.frame(., fmi_pred) # do loop applies the model fitting and prediction to each country group
  }) %>% 
  dplyr::ungroup()


# Plot predicted vs actual values 
ggplotly(ggplot(fmi_gf_gr, aes(x = fmi, y = fmi_pred, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red"))


model_gl <- lm(fmi ~ r1_label, data=fmi_georegions)
summary(model_gl)
# p<0.001 for r1_label, reject null ; larger georegions are significant predictor of FMI 
# R2 = 0.467 
# plot(model)

# Create array of predicted FMI values:
fmi_gf_gl <- fmi_georegions %>%
  dplyr::group_by(rgn_id) %>%
  dplyr::do({ 
    fmi_pred <- predict(model_gl, newdata =.[c('r1_label')]) 
    data.frame(., fmi_pred) # do loop applies the model fitting and prediction to each country group
  }) %>% 
  dplyr::ungroup()

# Plot predicted vs actual values 
ggplotly(ggplot(fmi_gf_gl, aes(x = fmi, y = fmi_pred, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red"))
```


# Combined linear models and testing
Variables to combine to predict FMI: SPI, WGI, 
```{r}

all_layers <- fmi_ao %>% 
  left_join(fmi_wgi, by = c("rgn_id", "year", "fmi")) %>% 
  rename(wgi = resilience_score) %>% 
  left_join(fmi_spi, by = c("rgn_id", "year", "fmi")) %>% 
  rename(spi = resilience_score) %>% 
  left_join(fmi_georegions, by = c("rgn_id", "year", "fmi")) %>% 
  select(rgn_id, rgn_name, fmi, ao_need = ao_need_score, wgi, spi, r0_label, r2_label, r1_label)
  
mod1 <- lm(fmi ~ ao_need + wgi, data=all_layers)
summary(mod1)  # r^2 = 0.3962

mod2 <- lm(fmi ~ ao_need + spi, data=all_layers)
summary(mod2)  # r^2 = 0.3812

mod3 <- lm(fmi ~ ao_need + r2_label, data=all_layers)
summary(mod3)  # r^2 = 0.6246

mod4 <- lm(fmi ~ wgi + spi, data=all_layers)
summary(mod4)  # r^2 = 0.3517

mod5 <- lm(fmi ~ wgi + r2_label, data=all_layers)
summary(mod5)  # r^2 = 0.6152

mod6 <- lm(fmi ~ spi + r2_label, data=all_layers)
summary(mod6)  # r^2 = 0.6583

mod7 <- lm(fmi ~ ao_need + wgi + spi, data=all_layers)
summary(mod7)  # r^2 = 0.3832

mod8 <- lm(fmi ~ ao_need + wgi + r2_label, data=all_layers)
summary(mod8)  # r^2 = 0.6151

mod9 <- lm(fmi ~ ao_need + spi + r2_label, data=all_layers)
summary(mod9)  # r^2 = 0.6416

mod10 <- lm(fmi ~ wgi + spi + r2_label, data=all_layers)
summary(mod10)  # r^2 = 0.6442

mod11 <- lm(fmi ~ ao_need + wgi + spi + r2_label, data=all_layers)
summary(mod11)  # r^2 = 0.6257

AIC(mod1, mod2, mod3, mod4, mod5, mod6, mod7, mod8, mod9, mod10, mod11)

mod6.5 <- lm(fmi ~ spi + r1_label, data=all_layers)
summary(mod6.5)  # r^2 = 0.5776


# Create array of predicted values from best models and visualize:


# Create array of predicted FMI values using fmi ~ spi + r2_label
fmi_gf_spi_r2 <- all_layers %>%
  dplyr::group_by(rgn_id) %>%
  dplyr::do({ 
    fmi_pred <- predict(mod6, newdata =.[c('r2_label', 'spi')]) 
    data.frame(., fmi_pred) # do loop applies the model fitting and prediction to each country group
  }) %>% 
  dplyr::ungroup()

# Plot predicted vs actual values 
ggplotly(ggplot(fmi_gf_spi_r2, aes(x = fmi, y = fmi_pred, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red"))

# Create array of predicted FMI values using fmi ~ spi + r1_label
fmi_gf_spi_r1 <- all_layers %>%
  dplyr::group_by(rgn_id) %>%
  dplyr::do({ 
    fmi_pred <- predict(mod6.5, newdata =.[c('r1_label', 'spi')]) 
    data.frame(., fmi_pred) # do loop applies the model fitting and prediction to each country group
  }) %>% 
  dplyr::ungroup()

# Plot predicted vs actual values 
ggplotly(ggplot(fmi_gf_spi_r1, aes(x = fmi, y = fmi_pred, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red"))

```


# Gapfilling regions without FMI scores using best fit model
```{r}

# Create vector to filter out low population regions so that they are NA 
low_pop()
low_pop <- low_pop %>%
  filter(est_population < 3000 | is.na(est_population)) #filter out regions that have populations > 3000 and keep NA values 
low_pop_vector <- c(low_pop$rgn_id) #make a vector of rgn_ids of low population areas 

UNgeorgn() 

fmi_gf <- UNgeorgn %>% 
  merge(spi) %>%
  filter(year==2018) %>% 
  left_join(fmi_rgn, by=c("rgn_id", "year")) %>%
  mutate(rgn_label = as.character(rgn_label)) %>% 
  mutate(rgn_label = ifelse(str_detect(rgn_label, "R_union"), "Reunion", rgn_label)) %>% 
  select(-rgn_name, -r0_label, spi=resilience_score)

# Separate out low population/uninhabited regions from the data frame to be gapfilled
low_pop_regions <- fmi_gf %>% 
  filter(rgn_id %in% low_pop_vector)

# Remove low pop regions from data to be gapfilled
fmi_gf <- fmi_gf %>% 
  filter(!rgn_id %in% low_pop_vector)


# Create array of predicted FMI values using fmi ~ spi + r1_label
fmi_gf1 <- fmi_gf %>%
  dplyr::group_by(rgn_id) %>%
  dplyr::do({ 
    fmi_pred_r1 <- predict(mod6.5, newdata =.[c('r1_label', 'spi')]) 
    data.frame(., fmi_pred_r1) # do loop applies the model fitting and prediction to each country group
  }) %>% 
  dplyr::ungroup()


# Create array of predicted FMI values using fmi ~ spi + r2_label
fmi_gf2 <- fmi_gf %>%
  filter(r2_label != "Melanesia") %>% # Melanesia does not have any FMI data points in df used to train mod6
  filter(r2_label != "Polynesia") %>% # same issue ^
  filter(r2_label != "Western Asia") %>% 
  dplyr::group_by(rgn_id) %>%
  dplyr::do({ 
    fmi_pred_r2 <- predict(mod6, newdata =.[c('r2_label', 'spi')]) 
    data.frame(., fmi_pred_r2) # do loop applies the model fitting and prediction to each country group
  }) %>% 
  dplyr::ungroup()


# Rejoin regions gapfilled with r1, assign best predicted FMI scores to populated regions without FMI data
# Check to see if r2 has a value, if so use that to gapfill `score`, otherwise use r1, otherwise use r0

fmi_gf_all <- fmi_gf1 %>%
  left_join(fmi_gf2, by=c("rgn_id", "r1_label", "r2_label", "rgn_label", "year", "spi", "fmi")) %>% 
  mutate(gapfilled = ifelse(is.na(fmi) & !is.na(fmi_pred_r2), "1", 0)) %>%
  mutate(method = ifelse(is.na(fmi) & !is.na(fmi_pred_r2), "SPI + UN_geopolitical region r2", NA)) %>%
  mutate(fmi = ifelse(is.na(fmi), fmi_pred_r2, fmi)) %>% 
  mutate(gapfilled = ifelse(is.na(fmi) & !is.na(fmi_pred_r1), "1", gapfilled)) %>%
  mutate(method = ifelse(is.na(fmi) & !is.na(fmi_pred_r1), "SPI + UN_geopolitical region r1", method)) %>%
  mutate(fmi = ifelse(is.na(fmi), fmi_pred_r1, fmi))
  
# Save the data
final_fmi <- fmi_gf_all %>%
  rbind(low_pop_regions) %>%  # these 20 regions will have NAs
  select(rgn_id, year, value=fmi)
   
write_csv(final_fmi, here("globalprep/res_fmi/v2019/output/fmi_res.csv"))

# Add gapfilling info to low pop regions
low_pop_regions_gf <- low_pop_regions %>% 
  mutate(gapfilled = 0) %>% 
  mutate(method = "Not gapfilled due to low pop") %>% 
  select(rgn_id, year, gapfilled, method)

# Save dataframe with gapfilled method and status information
final_fmi_gf <- fmi_gf_all %>%
  select(rgn_id, year, gapfilled, method) %>% 
  rbind(low_pop_regions_gf)

write_csv(final_fmi_gf, here("globalprep/res_fmi/v2019/output/fmi_res_gf.csv"))


```

## Compare to Mora data from 2013

```{r}

# Load old data (not sure which file to use for comparions?)

mora <- read_csv(here("globalprep/res_mora_ao/v2013/data/r_mora_2013a.csv"))

mora_s4 <- read_csv(here("globalprep/res_mora_ao/v2013/data/r_mora_s4_2013a.csv"))

mora_s4_updated <- read_csv(here("globalprep/res_mora_ao/v2013/data/r_mora_s4_2013a_updated.csv"))

mora_v_fmi <- final_fmi %>% 
  rename(fmi_2019=value) %>% 
  left_join(mora, by="rgn_id") %>% 
  rename(mora=value)

# Compare gapfilled FMI scores with Mora data
ggplotly(ggplot(mora_v_fmi, aes(x = mora, y = fmi_2019, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red"))


# No correlation here - compare non-gapfilled data only with Mora data:

mora_v_fmi2 <- fmi_rgn %>% 
  rename(fmi_2019=fmi) %>% 
  left_join(mora, by="rgn_id") %>% 
  rename(mora=value)

ggplotly(ggplot(mora_v_fmi2, aes(x = mora, y = fmi_2019, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red"))

# Also no correlation; lack of relationship can be attributed to source data rather than gapfilling method. 
```

