---
title: 'OHI 2019 - Fisheries Management Index (Resilience)'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
editor_options: 
  chunk_output_type: console
---


# Summary
This document describes the steps for obtaining and wrangling the data used to calculate the fisheries management index resilience layer for the 2019 global assessment.

# Data Sources

The following data are used:

## Fisheries Management Index Data 
Add some detail about this data.

* [Fisheries Management Index](https://oursharedseas.com/2019-update/fisheries/#highchart-fisheries-fmindex)

**Date retreived:** 12 July 2019

**Method:** Data are not accessible in csv format from website, so points were manually entered into excel and saved as a csv (found in v2019/raw)

## GDP per capita 

Fisheries management index scores are highly correlated with GDP, so GDP per capita per person purchasing power (gdppcpppp) was used to create linear regression models to gapfill FMI data. We used World Bank data (reported at country-scale) and gapfilled any missing GDP values using CIA data. 

[World bank](http://data.worldbank.org/indicator/NY.GDP.PCAP.PP.KD) 
* Downloaded 7/24/19
* Time range: 1990-2018

[CIA World Factbook](https://www.cia.gov/library/publications/the-world-factbook/rankorder/2004rank.html)
* Downloaded 7/24/19
* Time range: 


# Updates from previous assessment
These data have not been updated since 2013, so this is an entirely new method for establishing resilience values. 



# Initial set-up code

```{r setup, message=FALSE, warning=FALSE}

#library(devtools)
#devtools::install_github("ohi-science/ohicore@dev")
library(ohicore)
library(tidyverse)
library(stringr)
library(WDI) # for accessing World Bank data 
library(here) 
library(janitor) # for tidying column names
library(plotly)
library(psych) # for correlation testing

source('https://raw.githubusercontent.com/OHI-Science/ohiprep_v2019/gh-pages/workflow/R/common.R')



```

# Load and wrangle FMI data 

```{r}

fmi_raw <- read_csv(here("globalprep/res_fmi/v2019/raw/FMI_data_raw.csv")) %>%
  rename("2016" = fmi_2016) %>% 
  rename("2018" = fmi_2018) %>% 
  gather(key = "year", value = "fmi", -country)

# Add region ID
fmi_rgn <- name_2_rgn(df_in = fmi_raw, 
                       fld_name='country', 
                       flds_unique=c('fmi', 'year')) %>% 
  select(rgn_id, rgn_name, year, fmi) %>% 
  filter(year == 2018) # remove 2016 points so that they don't skew the model

fmi_rgn$year <- as.numeric(fmi_rgn$year)

```

# Load World Bank data
**Adapted from AO need data prep script**
**If AO is done, we can just import this data and use it ??**

```{r}
# check website to see what years are available
yr_start = 1990
yr_end   = 2018

# download the data using the WDI package - This is the data we are going to work with. Create a variable for the data frame

gdppcppp_raw <-  WDI(country = "all",
               indicator = "NY.GDP.PCAP.PP.KD", 
               start = yr_start, end=yr_end)
summary(gdppcppp_raw)

date <- Sys.Date()

#Save the file into the raw folder
write_csv(gdppcppp_raw, here(sprintf('globalprep/res_fmi/v2019/raw/raw_gdppcppp_%s.csv', date))) # Save file with date, as WDI data changes over even short periods of time. For instance, the Mauritania GDP data changed by an order of magnitude over the course of a week. We want to preserve the date it was downloaded so that data is not being overwritten everytime we run the script. 

# Do we want to just use the gdp data used for AO for consistency? 

```

# Check raw WB GDP data
**Comparing to old GDP data used for AO since this layer has not been updated since 2013.**
```{r raw check, eval=FALSE}

last_saved_date <- "2019-07-24" # update when new data are downloaded.

new <- read_csv(here(sprintf('globalprep/res_fmi/v2019/raw/raw_gdppcppp_%s.csv', last_saved_date)))

old <- read_csv(here('globalprep/ao/v2018/raw/raw_gdppcppp.csv')) %>%
  select(country, old_value = NY.GDP.PCAP.PP.KD, year)

compare <- left_join(new, old, by = c("country", "year"))  %>%
  filter(year==2017)
 
ggplot(compare, aes(x=NY.GDP.PCAP.PP.KD, y=old_value)) + 
  geom_point() + 
  geom_abline(slope = 1, intercept = 0, color="red")
  
# Looks good! 
```

## Gapfilling 1: Linear Regression within a country's data
As we only have FMI data for 40 countries, we extrapolate values for the other 180 regions in OHI using GDP data to create a predictive within-country regression model.

```{r, eval=F}

# Reorganize to create cells for countries that have missing values for some years

gdppcppp_clean <- read_csv(here(sprintf('globalprep/res_fmi/v2019/raw/raw_gdppcppp_%s.csv', last_saved_date))) %>% # change date here also to match filename above 
  dplyr::select(country, value=NY.GDP.PCAP.PP.KD, year) %>%
  dplyr::filter(year >= 2005) %>%
  tidyr::spread(year, value) %>%
    # spread to fill in potentially missing values with NA
  data.frame() %>% # this will add an X in front of the column names, allowing us to gather the values
  tidyr::gather(year, value, starts_with("X")) %>%
  dplyr::mutate(year = gsub("X", "", year)) %>% #substitute X for "" (nothing) in the column year
  dplyr::mutate(year = as.numeric(year)) #convert the year column into a numeric format

head(gdppcppp_clean)
summary(gdppcppp_clean) # 357 NAs - 44 regions with incomplete GDP data


# For the first case, if there is only one value use this value for all years
# This is not ideal, but likely better than other forms of gapfilling - what about the CIA data? 
gdppcppp_val <- gdppcppp_clean %>%
  dplyr::group_by(country) %>%
  dplyr::mutate(value_num = sum(!is.na(value))) %>% # counts the numbers of non-missing values for each country (logical TRUEs regarded as one)
  dplyr::filter(value_num > 0) %>%    # filter out the countries with no data between 2005 and 2018 (21 regions in v2019)
  dplyr::mutate(value_num_gf = ifelse(value_num==1, mean(value, na.rm=TRUE), NA)) %>%  # mean() function is used on regions with one year of data, applies that single value to all NAs for that region (this is only gapfilling one region in v2019)
  dplyr::ungroup() %>%
  dplyr::mutate(value = ifelse(is.na(value), value_num_gf, value)) %>% # if no value is missing, leave it, otherwise gapfill; still have NAs where 13 > value_num > 1
  dplyr::select(country, year, value, value_num)  # select just these columns; to eliminate extraneous value_num_gf column


head(gdppcppp_val)
summary(gdppcppp_val) 


# Predict values using a linear regression with 'year' as an independent variable 
# Create new column with these predicted values
gdppcppp_gf <- gdppcppp_val %>%
  dplyr::group_by(country) %>%
  dplyr::do({ 
    mod <- lm(value ~ year, data =.)
    value_pred <- predict(mod, newdata =.[c('year')]) # value_pred = country-grouped mod$fitted.values?
    data.frame(., value_pred) # do loop applies the model fitting and prediction to each country group
  }) %>% 
  dplyr::ungroup()

summary(gdppcppp_gf)
head(gdppcppp_gf)


# Fill in the remaining NA values using the predicted values
gdppcppp_gf2 <- gdppcppp_gf %>%
  dplyr::ungroup() %>% # why would we need to ungroup here?
  dplyr::mutate(gapfilled = ifelse(is.na(value), 1, 0)) %>% # Create column 'gapfilled', if value is currently NA, it will be gapfilled, indicated by a 1 in the gapfill column
  dplyr::mutate(gapfilled = ifelse(value_num == 1, 1, gapfilled)) %>% # if value_num is 1 it was gapfilled previously and gets a 1 in the gapfill column
  dplyr::mutate(value = ifelse(is.na(value), value_pred, value)) %>% # if NA in value column, input the value in value_pred column
  dplyr::mutate(method = ifelse(gapfilled==1, paste("lm based on N years data:", value_num, sep=" "), NA)) %>% # Create column 'method' that indicates method of gapfilling; this puts message "lm based..." even in some rows gapfilled with one year of data
  dplyr::mutate(method = ifelse(value_num == 1, "gapfilled using one year of data", method)) # this overwrites/corrects method "lm based..." for rows actually gapfilled with "one-year of data"" method
  
summary(gdppcppp_gf2) # no more NAs because everything has been gap-filled.
```


## Calculate rescaled values - do we need to do this for the FMI data? 
This is performed by taking the natural log of each value and then dividing by the 95th quantile of values across all years (from 2005 to current data year). 

```{r, eval=F}

# Values at the 95th Quantile or greater are given a rescaled score of '1' (the highest value)
gdppcppp_rescale <- gdppcppp_gf2 %>%
  dplyr::mutate(quantile_95 = quantile(value, probs=0.95)) %>% # gives a single value - the 95th quant (v2019=53582.14; without current year of data quant95=53208.67 - these should be different)
  dplyr::mutate(value_stand = value/quantile_95) %>% # where does value scale relative to 95th quantile
  dplyr::mutate(value_stand = ifelse(value_stand > 1, 1, value_stand)) %>% 
  dplyr::select(country, year, value, score=value_stand, gapfilled, method) # rename value_stand 'score'


summary(gdppcppp_rescale)
head(gdppcppp_rescale)


# Check to see if scores make sense - anything above reference point (quant_95) should = 1, everything below it should have a value between 0 and 1 

rescale_vis <- ggplot(gdppcppp_rescale, aes(x =value , y = score, label = country)) +
   geom_point()
ggplotly(rescale_vis)


```


## Convert country names to ohi regions

```{r, eval=F}

# Function to add OHI region ID based on country name
d_stand_rgn <- name_2_rgn(df_in = gdppcppp_rescale, 
                       fld_name='country', 
                       flds_unique=c('year'))

data.frame(filter(d_stand_rgn, rgn_id == 209))
##This should match the duplicate regions


# Combine the duplicate regions (we report these at lower resolution)
# In this case, we take the average score weighted by population
population_weights <- read.csv('https://raw.githubusercontent.com/OHI-Science/ohiprep_v2017/master/src/LookupTables/Pop_weight_ChinaSAR_USVIslPRico.csv')


# Weight the `score`, `value`, and `gapfilled` column by population
d_stand_rgn <- d_stand_rgn %>%
  dplyr::left_join(population_weights, by="country") %>% # does it make sense to backfill population data with static data?
  dplyr::mutate(population = ifelse(is.na(population), 1, population)) %>% # If no value available, input 1 (these values will not change)
  dplyr::group_by(rgn_id, year, method, gapfilled) %>% 
  dplyr::summarize(score = weighted.mean(score, population), # weight the single score value by pop.
            value = weighted.mean(value, population)) %>%
  ungroup() 

# check again:
data.frame(filter(d_stand_rgn, rgn_id == 209))


# Removed `Azerbaijan` (255) because the adjacent body of water is a sea not the ocean
d_stand_rgn <- d_stand_rgn %>%
  filter(rgn_id <= 250)

summary(d_stand_rgn) # no NAs


# save the cleaned gdppcppp for other goals
gdppcppp_data <- d_stand_rgn %>%
  select(rgn_id, year, value)

write_csv(gdppcppp_data, here("globalprep/res_fmi/v2019/intermediate/gdppcppp_ohi.csv"))

```


## Gapfilling: part 2
In this case, we gapfill regions with no data using means based on UN-designated geopolitical levels.
**For FMI, do this for the GDP value, not the score? 
```{r, eval=F}
# Create dataframe pairing each UN geopolitical region id with a year from 2005 to current
UNgeorgn() 

# Assign georegion labels to each region for each level (r0, r1, r2)
d_stand_gf <- data.frame(year=min(d_stand_rgn$year):max(d_stand_rgn$year)) %>% 
  merge(UNgeorgn, by=NULL) 

# Combine the two data frames by region id and year
# Calculate means across increasing geopolitical levels (e.g. r2, r1), using the highest resolution possible
d_stand_gf <- d_stand_gf %>%  
  left_join(d_stand_rgn, by = c("rgn_id", "year")) %>%
  group_by(r2_label, year) %>%
  mutate(r2_value = mean(value, na.rm=TRUE)) %>%
  ungroup() %>%
  group_by(r1_label, year) %>%
  mutate(r1_value = mean(value, na.rm=TRUE)) %>%
  ungroup() %>%
  group_by(r0_label, year) %>%
  mutate(r0_value = mean(value, na.rm=TRUE)) %>%
  ungroup()
summary(d_stand_gf) 

######################
# Do we want to remove low population areas for FMI? 
low_pop()

low_pop <- low_pop %>%
  filter(est_population < 3000 | is.na(est_population)) #filter out regions that have populations > 3000 and keep NA values 

low_pop_vector <- c(low_pop$rgn_id) #make a vector of low population areas 
########################

# For GDP `value` cells that still have NA values (still several hundred):
# Check to see if r2 has a value, if so use that to gapfill `score`, otherwise use r1, otherwise use r0
d_stand_gf <- d_stand_gf %>%
  mutate(gapfilled = ifelse(is.na(value) & !is.na(r2_value), "1", gapfilled)) %>%
  mutate(method = ifelse(is.na(value) & !is.na(r2_value), "UN_geopolitical region avg, r2", method)) %>%
  mutate(value = ifelse(is.na(value), r2_value, value)) %>%
  mutate(gapfilled = ifelse(is.na(value) & !is.na(r1_value), "1", gapfilled)) %>%
  mutate(method = ifelse(is.na(value) & !is.na(r1_value), "UN_geopolitical region avg, r1", method)) %>%
  mutate(value = ifelse(is.na(value), r1_value, value)) %>%
  mutate(gapfilled = ifelse(is.na(value) & !is.na(r0_value), "1", gapfilled)) %>%
  mutate(method = ifelse(is.na(value) & !is.na(r0_value), "UN_geopolitical region avg, r0", method)) %>%
  mutate(value = ifelse(is.na(value), r0_value, value))
  
############# Keep this for FMI? 
d_stand_gf$value[d_stand_gf$rgn_id %in% low_pop_vector] <- NA
#should now have NA values in score column for low popuation areas

summary(d_stand_gf)

```


## Save the data

```{r, eval=F}

# Save dataframe with adjusted, gapfilled, and rescaled score information
final <- d_stand_gf %>%
  select(rgn_id, rgn_label, year, value)

write_csv(final, here("globalprep/res_fmi/v2019/output/wb_gdppcppp_rescaled.csv"))


# Save dataframe with gapfilled method and status information
final_gf <- d_stand_gf %>%
  select(rgn_id, year, gapfilled, method)

write_csv(final_gf, here("globalprep/res_fmi/v2019/output/wb_gdppcppp_rescaled_gf.csv"))
# Some regions say that they are gapfilled even if they were saved as NA due to low population 

```


# Create linear regression models using gapfilled GDP data

## Load and organize the GDP data 
```{r}

gdp <- read_csv(here("globalprep/res_fmi/v2019/output/wb_gdppcppp_rescaled.csv")) %>% 
  rename(rgn_name=rgn_label) %>% 
  rename(gdppcppp=value)
  

fmi_gdp <- fmi_rgn %>% 
  left_join(gdp, by=c("rgn_id","rgn_name","year")) %>% 
  select(-rgn_name)

```

## Correlation testing
```{r}

# Look at data and check for correlation
ggplotly(ggplot(fmi_gdp, aes(x = gdppcppp, y = fmi, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red"))

fmi_gdp_plot <- fmi_gdp %>% 
  select(fmi, gdppcppp)
  
pairs.panels(fmi_gdp_plot, density = TRUE, cor=TRUE, lm=TRUE)


# Take logs and create scatter
# Log transformations adjust for skew/exponential relationships
log_fmi_gdp <- log(fmi_gdp_plot+1) %>% 
  rename(log_fmi=fmi) %>% 
  rename(log_gdppcppp=gdppcppp)

pairs.panels(log_fmi_gdp, density = TRUE, cor=FALSE, lm=TRUE)

# Correlation test
cor.test(log_fmi_gdp$log_fmi, log_fmi_gdp$log_gdppcppp, method="pearson", alternative="two.sided")
# p <0.001, reject null that true correlation=0
# cor = ~0.511 

```


## Create linear regression models for gapfilling
```{r}

### Using gapfilled GDP data (process above)
model <- lm(fmi ~ gdppcppp, data=fmi_gdp)
summary(model)
anova(model)
# p<0.001 for gdp, reject null ; GDP is significant predictor of FMI 
# R2: 0.47
# plot(model)

fmi_gf <- fmi_gdp %>%
  dplyr::group_by(rgn_id) %>%
  dplyr::do({ 
    fmi_pred <- predict(model, newdata =.[c('gdppcppp')]) 
    data.frame(., fmi_pred) # do loop applies the model fitting and prediction to each country group
  }) %>% 
  dplyr::ungroup()

summary(gdppcppp_gf)
head(gdppcppp_gf)

```

# Alternative linear models 
## Correlation and model based on gapfilled, rescaled "score" data (calculated from gdppcppp, AO output product)
```{r}

ao_score <- read_csv(here("globalprep/ao/v2019/output/wb_gdppcppp_rescaled.csv"))

fmi_ao <- fmi_rgn %>% 
  left_join(ao_score, by=c("rgn_id","year")) %>% 
  select(-rgn_name) %>% 
  rename(ao_need_score = value)

ggplotly(ggplot(fmi_ao, aes(x = ao_need_score, y = fmi, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red"))

fmi_ao_plot <- fmi_ao %>% 
  select(fmi, ao_need_score)
  
pairs.panels(fmi_ao_plot, density = TRUE, cor=TRUE, lm=TRUE)

# Take logs and create scatter
# Log transformations adjust for skew/exponential relationships
log_fmi_ao <- log(fmi_ao_plot+1) %>% 
  rename(log_fmi=fmi) %>% 
  rename(log_ao_need_score=ao_need_score)

pairs.panels(log_fmi_ao, density = TRUE, cor=FALSE, lm=TRUE)

# Correlation test
cor.test(log_fmi_ao$log_fmi, log_fmi_ao$log_ao_need_score, method="pearson", alternative="two.sided")
# p <0.001, reject null that true correlation=0
# cor = ~0.597 

model_ao <- lm(fmi ~ ao_need_score, data=fmi_ao)
summary(model_ao)
anova(model_ao)
# p<0.001 for ao_need_score, reject null ; AO need is significant predictor of FMI 
# R2 = 0.3971
# plot(model)

fmi_gf_ao <- fmi_ao %>%
  dplyr::group_by(rgn_id) %>%
  dplyr::do({ 
    fmi_pred <- predict(model, newdata =.[c('ao_need_score')]) 
    data.frame(., fmi_pred) # do loop applies the model fitting and prediction to each country group
  }) %>% 
  dplyr::ungroup()


```

## Model based on Social Progress Index (SPI)
**Note: this data set only has 1100 data points compared to more for wgi and ao (data is only 2014-2018)**
```{r}
spi <- read_csv(here("globalprep/prs_res_spi/v2018/output/spi_res.csv"))

fmi_spi <- fmi_rgn %>% 
  left_join(spi, by=c("rgn_id","year")) %>% 
  select(-rgn_name)

ggplotly(ggplot(fmi_spi, aes(x = resilience_score, y = fmi, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red"))

fmi_spi_plot <- fmi_spi %>% 
  select(fmi, resilience_score)
  
pairs.panels(fmi_spi_plot, density = TRUE, cor=TRUE, lm=TRUE)

# Take logs and create scatter
# Log transformations adjust for skew/exponential relationships
log_fmi_spi <- log(fmi_spi_plot+1) %>% 
  rename(log_fmi=fmi) %>% 
  rename(log_resilience_score=resilience_score)

pairs.panels(log_fmi_spi, density = TRUE, cor=FALSE, lm=TRUE)

# Correlation test
cor.test(log_fmi_spi$log_fmi, log_fmi_spi$log_resilience_score, method="pearson", alternative="two.sided")
# p <0.001, reject null that true correlation=0
# cor = ~0.568 (decent strong positve)

model_spi <- lm(fmi ~ resilience_score, data=fmi_spi)
summary(model_spi)
anova(model_spi)
# p<0.001 for resilience_score, reject null ; SPI resilience score is significant predictor of FMI 
# R2 = 0.4204
# plot(model)

fmi_gf_spi <- fmi_spi %>%
  dplyr::group_by(rgn_id) %>%
  dplyr::do({ 
    fmi_pred <- predict(model, newdata =.[c('resilience_score')]) 
    data.frame(., fmi_pred) # do loop applies the model fitting and prediction to each country group
  }) %>% 
  dplyr::ungroup()

```

## Model based on world governance indicator score
**Note: v2019 data go through 2017**
```{r}
wgi <- read_csv(here("globalprep/prs_res_wgi/v2019/output/wgi_res.csv")) %>% 
dplyr::mutate(year = ifelse(stringr::str_detect(year,"2017"), "2018", year)) # making data from 2017 2018 instead so that it will merge with FMI data 

wgi$year <- as.numeric(wgi$year)

fmi_wgi <- fmi_rgn %>% 
  left_join(wgi, by=c("rgn_id","year")) %>% 
  select(-rgn_name)

ggplotly(ggplot(fmi_wgi, aes(x = resilience_score, y = fmi, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red"))

fmi_wgi_plot <- fmi_wgi %>% 
  select(fmi, resilience_score)

pairs.panels(fmi_wgi_plot, density = TRUE, cor=TRUE, lm=TRUE)

# Take logs and create scatter
# Log transformations adjust for skew/exponential relationships
log_fmi_wgi <- log(fmi_wgi_plot+1) %>% 
  rename(log_fmi=fmi) %>% 
  rename(log_resilience_score=resilience_score)

pairs.panels(log_fmi_wgi, density = TRUE, cor=FALSE, lm=TRUE)

# Correlation test
cor.test(log_fmi_wgi$log_fmi, log_fmi_wgi$log_resilience_score, method="pearson", alternative="two.sided")
# p <0.001, reject null that true correlation=0
# cor = ~0.585

model_wgi <- lm(fmi ~ resilience_score, data=fmi_wgi)
summary(model_wgi)
anova(model_wgi)
# p<0.001 for resilience_score, reject null ; SPI resilience score is significant predictor of FMI 
# R2 = 0.4298
# plot(model)

fmi_gf_wgi <- fmi_wgi %>%
  dplyr::group_by(rgn_id) %>%
  dplyr::do({ 
    fmi_pred <- predict(model, newdata =.[c('resilience_score')]) 
    data.frame(., fmi_pred) # do loop applies the model fitting and prediction to each country group
  }) %>% 
  dplyr::ungroup()
```


**How can we test to determine an acceptable level of difference between predicted an actual for the data we do have?**


## Model based on GDP per capita (constant 2010 US$; alternative to use current US$)
Source: https://data.worldbank.org/indicator/NY.GDP.PCAP.KD
**Can skip if you have already run this chunk and saved GDP data*
```{r}

# Extract data from WDI()

# check world bank website to see what years are available
yr_start = 1960
yr_end   = 2018


# get description of variables (NOTE: these descriptions appear out of date, they aren't in sync with the definitions of the World Bank):
indicators <-  data.frame(WDI_data[[1]])

# Get information about dataset 
indicators[grep("NY.GDP.PCAP.KD", indicators$indicator), ]  # constant dollars. grep helps identify rows to select based on a string. (used this data)


# download the data using the WDI package 
gdppc_raw <-  WDI(country = "all",
               indicator = "NY.GDP.PCAP.KD", 
               start = yr_start, end=yr_end)
summary(gdppc_raw)

# Can add data saving/re-importing steps in between here if we decide to use this data in the future

# Clean and wrangle GDP data
gdppc_clean <- gdppc_raw %>% 
  dplyr::select(country, value=NY.GDP.PCAP.KD, year) %>%
  dplyr::filter(year >= 2005) %>%
  tidyr::spread(year, value) %>%
    # spread to fill in potentially missing values with NA
  data.frame() %>% # this will add an X in front of the column names, allowing us to gather the values
  tidyr::gather(year, value, starts_with("X")) %>%
  dplyr::mutate(year = gsub("X", "", year)) %>% #substitute X for "" (nothing) in the column year
  dplyr::mutate(year = as.numeric(year)) #convert the year column into a numeric format

summary(gdppc_clean) # 273 NAs

# For the first gapfilling stage, if a region has only one value use this value for all years
gdppc_val <- gdppc_clean %>%
  dplyr::group_by(country) %>%
  dplyr::mutate(value_num = sum(!is.na(value))) %>% # counts the numbers of non-missing values for each country (logical TRUEs regarded as one)
  dplyr::filter(value_num > 0) %>%    # filter out the countries with no data between 2005 and 2018 
  dplyr::mutate(value_num_gf = ifelse(value_num==1, mean(value, na.rm=TRUE), NA)) %>%  # mean() function is used on regions with one year of data, applies that single value to all NAs for that region 
  dplyr::ungroup() %>%
  dplyr::mutate(value = ifelse(is.na(value), value_num_gf, value)) %>% # if no value is missing, leave it, otherwise gapfill
  dplyr::select(country, year, value, value_num)  # select just these columns; to eliminate extraneous value_num_gf column

summary(gdppc_val) # now only 39 NAs ! 


# Predict values using a linear regression with 'year' as an independent variable 
# Create new column with these predicted values
gdppc_gf <- gdppc_val %>%
  dplyr::group_by(country) %>%
  dplyr::do({ 
    mod <- lm(value ~ year, data =.)
    value_pred <- predict(mod, newdata =.[c('year')]) # value_pred = country-grouped mod$fitted.values?
    data.frame(., value_pred) # do loop applies the model fitting and prediction to each country group
  }) %>% 
  dplyr::ungroup()

summary(gdppc_gf)

# Fill in the remaining NA values using the predicted values
gdppc_gf2 <- gdppc_gf %>%
  dplyr::ungroup() %>% # why would we need to ungroup here?
  dplyr::mutate(gapfilled = ifelse(is.na(value), 1, 0)) %>% # Create column 'gapfilled', if value is currently NA, it will be gapfilled, indicated by a 1 in the gapfill column
  dplyr::mutate(gapfilled = ifelse(value_num == 1, 1, gapfilled)) %>% # if value_num is 1 it was gapfilled previously and gets a 1 in the gapfill column
  dplyr::mutate(value = ifelse(is.na(value), value_pred, value)) %>% # if NA in value column, input the value in value_pred column
  dplyr::mutate(method = ifelse(gapfilled==1, paste("lm based on N years data:", value_num, sep=" "), NA)) %>% # Create column 'method' that indicates method of gapfilling; this puts message "lm based..." even in some rows gapfilled with one year of data
  dplyr::mutate(method = ifelse(value_num == 1, "gapfilled using one year of data", method)) # this overwrites/corrects method "lm based..." for rows actually gapfilled with "one-year of data"" method
  
summary(gdppc_gf2) # no more NAs because everything has been gap-filled.

## Rescale values
# Values at the 95th Quantile or greater are given a rescaled score of '1' (the highest value)
gdppc_rescale <- gdppc_gf2 %>%
  dplyr::mutate(quantile_95 = quantile(value, probs=0.95)) %>% # gives a single value - the 95th quant (v2019=53582.14; without current year of data quant95=53208.67 - these should be different)
  dplyr::mutate(value_stand = value/quantile_95) %>% # where does value scale relative to 95th quantile
  dplyr::mutate(value_stand = ifelse(value_stand > 1, 1, value_stand)) %>% 
  dplyr::select(country, year, value, score=value_stand, gapfilled, method) # rename value_stand 'score'

summary(gdppc_rescale)

## Convert to OHI regions
d_stand_rgn <- name_2_rgn(df_in = gdppc_rescale, 
                       fld_name='country', 
                       flds_unique=c('year'))


# Combine the duplicate regions (we report these at lower resolution)
# In this case, we take the average score weighted by population
population_weights <- read.csv('https://raw.githubusercontent.com/OHI-Science/ohiprep_v2017/master/src/LookupTables/Pop_weight_ChinaSAR_USVIslPRico.csv')


# Weight the `score`, `value`, and `gapfilled` column by population
d_stand_rgn <- d_stand_rgn %>%
  dplyr::left_join(population_weights, by="country") %>% # does it make sense to backfill population data with static data?
  dplyr::mutate(population = ifelse(is.na(population), 1, population)) %>% # If no value available, input 1 (these values will not change)
  dplyr::group_by(rgn_id, year, method, gapfilled) %>% 
  dplyr::summarize(score = weighted.mean(score, population), # weight the single score value by pop.
            value = weighted.mean(value, population)) %>%
  ungroup() 

# check again:
data.frame(filter(d_stand_rgn, rgn_id == 209))


# Removed `Azerbaijan` (255) because the adjacent body of water is a sea not the ocean
d_stand_rgn <- d_stand_rgn %>%
  filter(rgn_id <= 250)

summary(d_stand_rgn) # no NAs


# save simplified cleaned gdppc data for existing regions
gdppc_data <- d_stand_rgn %>%
  select(rgn_id, year, value)

write_csv(gdppc_data, here("globalprep/res_fmi/v2019/intermediate/gdppc_ohi.csv"))



## Gapfilling regions with no GDP per capita data using UN georegions 

# Re-import cleaned/gapfilled gdppc data if you've already completed the above steps 
gdppc_data <- read_csv(here("globalprep/res_fmi/v2019/intermediate/gdppc_ohi.csv"))


# Create dataframe pairing each UN geopolitical region id with a year from 2005 to current
UNgeorgn()  

# Assign georegion labels to each region for each level (r0, r1, r2)
d_stand_gf <- data.frame(year=min(d_stand_rgn$year):max(d_stand_rgn$year)) %>% 
  merge(UNgeorgn, by=NULL) 

# Combine the two data frames by region id and year
# Calculate means across increasing geopolitical levels (e.g. r2, r1), using the highest resolution possible
d_stand_gf <- d_stand_gf %>%  
  left_join(d_stand_rgn, by = c("rgn_id", "year")) %>%
  group_by(r2_label, year) %>%
  mutate(r2_value = mean(score, na.rm=TRUE)) %>%
  ungroup() %>%
  group_by(r1_label, year) %>%
  mutate(r1_value = mean(score, na.rm=TRUE)) %>%
  ungroup() %>%
  group_by(r0_label, year) %>%
  mutate(r0_value = mean(score, na.rm=TRUE)) %>%
  ungroup()
summary(d_stand_gf) 

# Create vector to filter out low population regions so that they are NA 
low_pop()
low_pop <- low_pop %>%
  filter(est_population < 3000 | is.na(est_population)) #filter out regions that have populations > 3000 and keep NA values 
low_pop_vector <- c(low_pop$rgn_id) #make a vector of low population areas 

# For `score` cells that still have NA values (still several hundred):
# Check to see if r2 has a value, if so use that to gapfill `score`, otherwise use r1, otherwise use r0
d_stand_gf <- d_stand_gf %>%
  mutate(gapfilled = ifelse(is.na(score) & !is.na(r2_value), "1", gapfilled)) %>%
  mutate(method = ifelse(is.na(score) & !is.na(r2_value), "UN_geopolitical region avg, r2", method)) %>%
  mutate(score = ifelse(is.na(score), r2_value, score)) %>%
  mutate(gapfilled = ifelse(is.na(score) & !is.na(r1_value), "1", gapfilled)) %>%
  mutate(method = ifelse(is.na(score) & !is.na(r1_value), "UN_geopolitical region avg, r1", method)) %>%
  mutate(score = ifelse(is.na(score), r1_value, score)) %>%
  mutate(gapfilled = ifelse(is.na(score) & !is.na(r0_value), "1", gapfilled)) %>%
  mutate(method = ifelse(is.na(score) & !is.na(r0_value), "UN_geopolitical region avg, r0", method)) %>%
  mutate(score = ifelse(is.na(score), r0_value, score))
  

d_stand_gf$score[d_stand_gf$rgn_id %in% low_pop_vector] <- NA
#should now have NA values in score column for low popuation areas


# Save complete GDP per capita data 

final_gdppc <- d_stand_gf %>%
  select(rgn_id, year, value=score)

write_csv(final_gdppc, here("globalprep/res_fmi/v2019/output/wb_gdppc_rescaled.csv"))


# Save dataframe with gapfilled method and status information
final_gdppc_gf <- d_stand_gf %>%
  select(rgn_id, year, gapfilled, method)

write_csv(final_gdppc_gf, here("globalprep/res_fmi/v2019/output/wb_gdppc_rescaled_gf.csv"))
```


# Create linear model based on GDP per capita data 

```{r}
gdppc <- read_csv(here("globalprep/res_fmi/v2019/output/wb_gdppc_rescaled.csv"))

fmi_gdppc <- fmi_rgn %>% 
  left_join(gdppc, by=c("rgn_id","year")) %>% 
  select(-rgn_name) %>% 
  rename(gdppc = value)


ggplotly(ggplot(fmi_gdppc, aes(x = gdppc, y = fmi, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red"))

fmi_gdppc_plot <- fmi_gdppc %>% 
  select(fmi, gdppc)
  
pairs.panels(fmi_gdppc_plot, density = TRUE, cor=TRUE, lm=TRUE)

# Take logs and create scatter
# Log transformations adjust for skew/exponential relationships
log_fmi_gdppc <- log(fmi_gdppc_plot+1) %>% 
  rename(log_fmi=fmi) %>% 
  rename(log_gdppc=gdppc)

pairs.panels(log_fmi_gdppc, density = TRUE, cor=FALSE, lm=TRUE)

# Correlation test
cor.test(log_fmi_gdppc$log_fmi, log_fmi_gdppc$log_gdppc, method="pearson", alternative="two.sided")
# p <0.0001, reject null that true correlation=0
# cor = ~0.604

model_gdppc <- lm(fmi ~ gdppc, data=fmi_gdppc)
summary(model_gdppc)
# p<0.001 for gdppc, reject null ; GDP per capita is significant predictor of FMI 
# R2 = 0.3935 (but not a very strong predictor)
# plot(model)

# Create array of predicted FMI values 
fmi_gf_gdppc <- fmi_gdppc %>%
  dplyr::group_by(rgn_id) %>%
  dplyr::do({ 
    fmi_pred <- predict(model_gdppc, newdata =.[c('gdppc')]) 
    data.frame(., fmi_pred) # do loop applies the model fitting and prediction to each country group
  }) %>% 
  dplyr::ungroup()

# Plot predicted vs actual values 
ggplotly(ggplot(fmi_gf_gdppc, aes(x = fmi, y = fmi_pred, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red"))

```


# Create linear model based on UN georegions 

```{r}
source('https://raw.githubusercontent.com/OHI-Science/ohiprep_v2019/gh-pages/workflow/R/common.R')

# Create dataframe pairing each UN geopolitical region id with a year from 2005 to current
UNgeorgn()
georegions <- UNgeorgn %>% 
  select(rgn_id, rgn_name = rgn_label, r1_label, r2_label)


fmi_georegions <- fmi_rgn %>% 
  left_join(georegions, by=c("rgn_id","rgn_name"))


model_gr <- lm(fmi ~ r2_label, data=fmi_georegions)
summary(model_gr)
# p<0.001 for r2_label, reject null ; smaller georegions are significant predictor of FMI 
# R2 = 0.5951 (pretty good!)
# plot(model_gr)


# Create array of predicted FMI values 
fmi_gf_gr <- fmi_georegions %>%
  dplyr::group_by(rgn_id) %>%
  dplyr::do({ 
    fmi_pred <- predict(model_gr, newdata =.[c('r2_label')]) 
    data.frame(., fmi_pred) # do loop applies the model fitting and prediction to each country group
  }) %>% 
  dplyr::ungroup()

# Plot predicted vs actual values 
ggplotly(ggplot(fmi_gf_gr, aes(x = fmi, y = fmi_pred, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red"))

model_gl <- lm(fmi ~ r1_label, data=fmi_georegions)
summary(model_gl)
# p<0.001 for r1_label, reject null ; larger georegions are significant predictor of FMI 
# R2 = 0.467 
# plot(model)

# Create array of predicted FMI values:
fmi_gf_gl <- fmi_georegions %>%
  dplyr::group_by(rgn_id) %>%
  dplyr::do({ 
    fmi_pred <- predict(model_gl, newdata =.[c('r1_label')]) 
    data.frame(., fmi_pred) # do loop applies the model fitting and prediction to each country group
  }) %>% 
  dplyr::ungroup()

# Plot predicted vs actual values 
ggplotly(ggplot(fmi_gf_gl, aes(x = fmi, y = fmi_pred, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red"))
```


# Combined linear models and testing
Variables to combine to predict FMI: SPI, WGI, 
```{r}

all_layers <- fmi_ao %>% 
  left_join(fmi_wgi, by = c("rgn_id", "year", "fmi")) %>% 
  rename(wgi = resilience_score) %>% 
  left_join(fmi_spi, by = c("rgn_id", "year", "fmi")) %>% 
  rename(spi = resilience_score) %>% 
  left_join(fmi_georegions, by = c("rgn_id", "year", "fmi")) %>% 
  select(rgn_id, rgn_name, fmi, ao_need = ao_need_score, wgi, spi, r2_label)
  
mod1 <- lm(fmi ~ ao_need + wgi, data=all_layers)
summary(mod1)  # r^2 = 0.3962

mod2 <- lm(fmi ~ ao_need + spi, data=all_layers)
summary(mod2)  # r^2 = 0.3812

mod3 <- lm(fmi ~ ao_need + r2_label, data=all_layers)
summary(mod3)  # r^2 = 0.6246

mod4 <- lm(fmi ~ wgi + spi, data=all_layers)
summary(mod4)  # r^2 = 0.3517

mod5 <- lm(fmi ~ wgi + r2_label, data=all_layers)
summary(mod5)  # r^2 = 0.6152

mod6 <- lm(fmi ~ spi + r2_label, data=all_layers)
summary(mod6)  # r^2 = 0.6583

mod7 <- lm(fmi ~ ao_need + wgi + spi, data=all_layers)
summary(mod7)  # r^2 = 0.3832

mod8 <- lm(fmi ~ ao_need + wgi + r2_label, data=all_layers)
summary(mod8)  # r^2 = 0.6151

mod9 <- lm(fmi ~ ao_need + spi + r2_label, data=all_layers)
summary(mod9)  # r^2 = 0.6416

mod10 <- lm(fmi ~ wgi + spi + r2_label, data=all_layers)
summary(mod10)  # r^2 = 0.6442

mod11 <- lm(fmi ~ ao_need + wgi + spi + r2_label, data=all_layers)
summary(mod11)  # r^2 = 0.6257

AIC(mod1, mod2, mod3, mod4, mod5, mod6, mod7, mod8, mod9, mod10, mod11)

```



