---
title: 'OHI 2019 - Fisheries Management Index (Resilience)'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html' 
  pdf_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

# Summary
This document outlines the process for developing linear models to gapfill fisheries management index scores to use as a resilience layer for the 2019 global assessment.

# Data Sources

The following data are used:

## Fisheries Management Index Data 

[Fisheries Management Index](https://oursharedseas.com/2019-update/fisheries/#highchart-fisheries-fmindex) data were first introduced in a 2017 paper by Melnychuk et al., *[Fisheries Management Impacts on Target Species Status](https://doi.org/10.1073/pnas.1609915114)*. Scores range from 0-1 and rate management effectiveness of fisheries in distinct regions on a stock-by-stock basis. The scores are determined by expert surveys that characterize attributes of research, management, enforcement, and socioeconomic factors. The first survey was conducted in 2016 in 28 major fishing countries that collectively account for >80% of global catch. Another survey was performed in 2018 on 40 countries, the scores from which we are using for this OHI resilience layer. 

**Date retreived:** 12 July 2019

**Method:** Data are not accessible in csv format from website, so points were manually entered into excel and saved as a csv (found in v2019/raw). Because FMI scores only exist for 40 out of the 220 OHI regions, we trained linear models using scores from the AO need layer (rescaled GDP per capita per person purchasing power), GDP per capita, World Governance Index (WGI), Social Progress Indicator (SPI), and UN georegion labels to determine which would be most effective for gapfilling the missing data. After comparing the models, we determined SPI and UN georegions to be the best predictors of FMI, and used these models to gapfill FMI scores for the remaining 80 regions. 20 of these regions remain NA as they are uninhabited.


## SPI

*Note: These data were not updated for v2019 at the time of completing the FMI resilience layer.*

**Citation**: http://www.socialprogress.org/

Stern, S., A. Wares and T. Epner. 2018. Social Progress Index: 2018 Methodology Report.

**Source information**: 
http://www.socialprogress.org/ --> Download Data

**Date Downloaded**: 9/21/2018

**Time range**: 2014-2018

**Native data resolution**: country scores

**Format**:  Excel file

**Description**: Social Progress Index scores and components for countries.


## UN georegions

UNgeorgn() loads a dataframe from common.R with UN geopolitical designations, and is commonly used in OHI to gapfill missing data. The distinct regions are derived from the [United Nations Statistics Division](https://unstats.un.org/unsd/methodology/m49/). Each region is assigned four labels with increasing granularity/specificity: r0_label = World (1 level), r1_label = continental regions (7 levels: Oceania, Asia, Africa, Europe, Southern Islands, Latin America and the Caribbean, and Americas), r2_label = georegions (22 levels for subregions and intermediary regions).


# Updates from previous assessment
These data have not been updated since 2013, so this is an entirely new method for establishing resilience values. 


# Initial set-up code

```{r setup, message=FALSE, warning=FALSE, results="hide"}

#library(devtools)
#devtools::install_github("ohi-science/ohicore@dev")
library(ohicore)
library(tidyverse)
library(stringr)
library(WDI) # for accessing World Bank data 
library(here) 
library(plotly)
library(psych) # for correlation testing

source('https://raw.githubusercontent.com/OHI-Science/ohiprep_v2019/gh-pages/workflow/R/common.R')



```

# Load and wrangle FMI data 

```{r, message=FALSE, warning=FALSE, results="hide"}
fmi_raw <- read_csv(here("globalprep/res_fmi/v2019/raw/FMI_data_raw.csv")) %>%
  rename("2016" = fmi_2016) %>% 
  rename("2018" = fmi_2018) %>% 
  gather(key = "year", value = "fmi", -country)

# Add region ID
fmi_rgn <- name_2_rgn(df_in = fmi_raw, 
                       fld_name='country', 
                       flds_unique=c('fmi', 'year')) %>% 
  select(rgn_id, rgn_name, year, fmi) %>% 
  filter(year == 2018) # remove 2016 points so that they don't skew the model

fmi_rgn$year <- as.numeric(fmi_rgn$year)

```


## Load data from Social Progress Index (SPI) + UN Georegions, create linear models for gapfilling
```{r, message=FALSE, warning=FALSE, results="hide"}

## Load SPI data

spi <- read_csv(here("globalprep/prs_res_spi/v2018/output/spi_res.csv"))

fmi_spi <- fmi_rgn %>% 
  left_join(spi, by=c("rgn_id","year")) %>% 
  select(-rgn_name)

# Load UN georegion data 
UNgeorgn()

georegions <- UNgeorgn %>%
  select(rgn_id, rgn_name = rgn_label, r0_label, r1_label, r2_label)

fmi_georegions <- fmi_rgn %>% 
  left_join(georegions, by=c("rgn_id","rgn_name"))

all_layers <- fmi_spi %>% 
  rename(spi = resilience_score) %>% 
  left_join(fmi_georegions, by = c("rgn_id", "year", "fmi")) %>% 
  select(rgn_id, rgn_name, fmi, spi, r0_label, r2_label, r1_label)
  
mod6 <- lm(fmi ~ spi + r2_label, data=all_layers)
summary(mod6)  # r^2 = 0.6583
plot(predict(mod6), all_layers$fmi)
abline(0,1, col="red")

mod6.5 <- lm(fmi ~ spi + r1_label, data=all_layers)
summary(mod6.5)  # r^2 = 0.5776
plot(predict(mod6.5), all_layers$fmi)
abline(0,1, col="red")


```


# Gapfilling regions without FMI scores using best fit model
More information about the linear model process development can be found in fmi_model_compare.Rmd and the [Github issue from v2019](https://github.com/OHI-Science/globalfellows-issues/issues/91). 
```{r, message=FALSE, warning=FALSE, results="hide"}

fmi_gf <- UNgeorgn %>% 
  merge(spi) %>%
  filter(year==2018) %>% 
  left_join(fmi_rgn, by=c("rgn_id", "year")) %>%
  mutate(rgn_label = as.character(rgn_label)) %>% 
  mutate(rgn_label = ifelse(str_detect(rgn_label, "R_union"), "Reunion", rgn_label)) %>% 
  select(-rgn_name, -r0_label, spi=resilience_score)


# Create array of predicted FMI values using fmi ~ spi + r2_label

mod_r2 <- lm(fmi ~ r2_label + spi, data=fmi_gf)
## have to do the predict in a more complicated fashion because some r2 categories have no data, this returns an NA for these
fmi_gf$fmi_pred_r2 <- 
  sapply(1:nrow(fmi_gf), 
           function(i) 
               tryCatch(predict(mod_r2, fmi_gf[i,]), 
                           error=function(e) NA))

# get predictions for the regions not represented by r2 regions:
mod_r1 <- lm(fmi ~ r1_label + spi, data=fmi_gf)
## have to do the predict in a more complicated fashion because some r2 categories have no data, this returns an NA for these
fmi_gf$fmi_pred_r1 <- 
  sapply(1:nrow(fmi_gf), 
           function(i) 
               tryCatch(predict(mod_r1, fmi_gf[i,]), 
                           error=function(e) NA))

# final data and gapfilling recordkeeping

fmi_gf_all <- fmi_gf %>%
  dplyr::mutate(gapfilled = ifelse(is.na(fmi), "1", 0)) %>%
  dplyr::mutate(method = ifelse(is.na(fmi) & !is.na(fmi_pred_r2), "SPI + UN_geopolitical region r2", NA)) %>%
  dplyr::mutate(method = ifelse(is.na(fmi) & is.na(fmi_pred_r2), "SPI + UN_geopolitical region r1" , method)) %>%
  dplyr::mutate(fmi2 = ifelse(is.na(fmi), fmi_pred_r2, fmi)) %>% 
  dplyr::mutate(fmi2 = ifelse(is.na(fmi2), fmi_pred_r1, fmi2)) %>%
  dplyr::mutate(fmi = fmi2) %>%
  dplyr::select(-fmi2)


# make sure all low/no population regions are NA 
low_pop()
low_pop <- low_pop %>%
  filter(est_population < 3000 | is.na(est_population)) #filter out regions that have populations > 3000 and keep NA values 

fmi_gf_low_pop <- fmi_gf_all %>%
  dplyr::filter(rgn_id %in% low_pop$rgn_id)
summary(fmi_gf_low_pop)
  

## Make sure all other regions have an fmi value
fmi_gf_with_pop <- fmi_gf_all %>%
  dplyr::filter(!(rgn_id %in% low_pop$rgn_id))
summary(fmi_gf_with_pop)


# Correct gapfilling info to low pop regions
fmi_gf_all <- fmi_gf_all %>% 
  mutate(gapfilled = ifelse(rgn_id %in% low_pop$rgn_id, 0, gapfilled)) %>% 
  mutate(method = ifelse(rgn_id %in% low_pop$rgn_id, NA, method))

## format final data
fmi_final <- fmi_gf_all %>%
  select(rgn_id, year, value=fmi)
dim(fmi_final) # all regions represented
summary(fmi_final) # 20 NA values


# Save the data 
write_csv(fmi_final, here("globalprep/res_fmi/v2019/output/fmi_res.csv"))


# Save dataframe with gapfilled method and status information
final_fmi_gf <- fmi_gf_all %>%
  select(rgn_id, year, gapfilled, method)
write_csv(final_fmi_gf, here("globalprep/res_fmi/v2019/output/fmi_res_gf.csv"))


```

## Compare to Mora data from 2013

```{r, message=FALSE, warning=FALSE, echo=FALSE}

# Load old data (not sure which file to use for comparions?)

mora <- read_csv(here("globalprep/res_mora_ao/v2013/data/r_mora_2013a.csv"))

mora_v_fmi <- fmi_final %>% 
  rename(fmi_2019=value) %>% 
  left_join(mora, by="rgn_id") %>% 
  rename(mora=value)

# Compare gapfilled FMI scores with Mora data
ggplotly(ggplot(mora_v_fmi, aes(x = mora, y = fmi_2019, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red"))


# Not a great correlation here - compare non-gapfilled data only with Mora data:

mora_v_fmi2 <- fmi_rgn %>% 
  rename(fmi_2019_not_gf=fmi) %>% 
  left_join(mora, by="rgn_id") %>% 
  rename(mora=value)

ggplotly(ggplot(mora_v_fmi2, aes(x = mora, y = fmi_2019_not_gf, labels = rgn_id)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red"))

# Also not a great correlation; differences can be attributed to source data rather than gapfilling method. 
```

