---
title: "Coding Exercise for NCEAS Data Science Job Application"
output: html_document
---

## Instructions

You will have 24 hours to complete this activity. You will work directly in this document, using code chunks (gray parts) to wrangle, analyze, and visualize data according to the instructions. 

As necessary, document any responses to questions, as well as notes describing your process, in the code chunks and/or in the text of the Rmarkdown file.  

Once you have finished, please add your last name to the file name (e.g., frazier_coding_exercise.Rmd) and send to halpern@nceas.ucsb.edu.  

You will be evaluated on: coding ability, documentation of code, and communication of results.     


## Packages

Add the packages you will use to this code chunk.  We start by loading the `readr` package, which we use to read in the dataset below.
```{r setup}

## loading packages
library(readr)

```



## Data

You will be working with a dataset describing Ocean Health Index scores for:
 
  - 10 goals + 1 overall country score ("Index") 
  - 220 countries/regions (we use these terms interchangably)
  - 7 years (2012 - 2018)
  
We will read in the data from Github.com:

```{r data}

## read in OHI data:
ohi_data <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2019/gh-pages/Reference/Lessons_in_OHI/coding_exercise/scores.csv")

```


## Task 1

Help make these data more human readable by including the full goal name (vs. the 2 letter abbreviations) and the country name.  

Here is a table that includes the goal name and description: https://github.com/OHI-Science/ohi-global/blob/draft/global_supplement/tables/ten_goals.csv

Here is a table with the region names: https://github.com/OHI-Science/ohi-global/blob/draft/eez/spatial/regions_list.csv
(the rgn_id column in this file matches the region_id column in the ohi_data)

```{r human readable}

## Write code needed to update ohi_data to include the full goal name and the country name.




```


## Task 2

The above description of the data claims there are 220 countries with Ocean Health Index scores; however, a quick examination of these data reveals there are 221 countries in the data:

```{r countrycount}

length(unique(ohi_data$region_id))

```


One of the countries has no data.  Figure out which one.

```{r regions}

## Write code to identify the country with no data.



```


## Task 3

Create a plot showing the goal and Index scores for the United States of America from 2012 to 2017. 

```{r US}

## Code to create plot of USA of goal and Index scores across years



```


## Task 4

For each country and goal, calculate the average score across all years (i.e., each country should have N=11 scores when finished).

```{r across year avg}

## Code to summarize data



```


Create a plot that shows the distribution of Index scores of all countries.

```{r across year avg}

## Code to create a plot showing distribution of country Index scores



```


## Task 5

For each year and goal (including "Index"), determine the average score across all countries (The data frame should have 11*7=77 rows when finished).  

```{r global scores}

## Code to get global average scores for each goal and year


```

Create a plot showing average global scores for each goal (and the Index) from 2012 to 2018.   

```{r global scores plot}

## Code to create plot of average global goal scores across time for each goal/Index


```

Which goals tend to have high scores? Low scores?  Can you see any patterns over time?


## Task 6
Figure out which goals have the most NA values for scores in 2018.  Which ones have the least?  

```{r missing data}

## Explore missing data


```


