---
title: "Creating fishing pressure layers"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../ohiprep_v2018/src/templates/ohi_hdr.html'
pdf_document:
  toc: true
---

This script standardizes the annual catch rasters by net primary production.  NPP is first gapfilled using the average of neighboring cells. The standardized catch data are rescaled to range from 0-1 using the 99.99th quantile of all global raster cells across all years of data for each commercial catch category.  

```{r setup, message=F,warning=F}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'figs/',message = FALSE, eval=F, warning = FALSE)


library(seaaroundus)
library(raster)
library(readr)
library(stringr)
library(dplyr)
library(data.table)
library(parallel)
library(foreach)
library(doParallel)

source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/master/src/R/spatial_common.R")

years_of_data <- 2003:2014

# setwd("stressors/comm_fish")

# paralell processing
cl<-makeCluster(10)
registerDoParallel(cl)


```


# Primary production data

Load NPP data

```{r,eval=F}

npp_files <- list.files(file.path(dir_M,'git-annex/impact_acceleration/stressors/comm_fish/int/npp_annual'), full.names=TRUE)

npp_files <- npp_files[!(str_detect(npp_files, "1km"))]
npp_files <- npp_files[!(str_detect(npp_files, "_gf"))]
npp_files <- npp_files[grep(paste(years_of_data, collapse="|"), npp_files)]


```

Gapfill missing NPP raster data.  This is done by averaging the neighboring cells of raster cells with NA values and repeating until all ocean cells are gapfilled. 

```{r, eval=FALSE}

## Save a raster showing which areas were gapfilled
npp_rast <- raster(npp_files[1])
npp_rast[is.na(npp_rast)] <- 999
plot(npp_rast)

npp_rast_mol <- resample(npp_rast, ocean, method="ngb")
npp_rast_mol_mask <- mask(npp_rast_mol, ocean)  
writeRaster(npp_rast_mol_mask, file.path(dir_M,'/git-annex/impact_acceleration/stressors/comm_fish/int/npp_gf.tif'))


## Function to gapfill NA cells using mean of surrounding cells
gf_raster <- function(x){raster::focal(x, w = matrix(1,3,3), fun = mean, na.rm=TRUE, pad = TRUE, NAonly=TRUE)}


## Iterative gapfilling for each year of NPP data
foreach(npp = npp_files,.packages="dplyr") %dopar%{ # file_name = l[1]
#for (npp in npp_files){ # npp <- npp_files[1]

r <- raster::raster(npp)
yr <- stringr::str_sub(npp,-8,-5)

## Repeat until critical NA cells are filled
i <- 0

while (i <= 500){
r <- gf_raster(r)
i <- i + 1
print(i)
}
## take a look at gapfilling
# r[is.na(r)] <- 999
# plot(r)
## check to make sure that relevant cells are gapfilled at final resolution
# r_mol <- resample(r, ocean, method="ngb")
# r_mol_mask <- mask(r_mol, ocean)  
# plot(r_mol_mask)
# r_mol_mask

raster::writeRaster(r, file.path(dir_M, sprintf("/git-annex/impact_acceleration/stressors/comm_fish/int/npp_annual/annual_mean_npp_%s_gf.tif", yr)), overwrite=TRUE)

}

# get a list of the npp files

npp_files_gf <- list.files(file.path(dir_M,'git-annex/impact_acceleration/stressors/comm_fish/int/npp_annual'), full.names=TRUE, pattern = "gf")

npp_files_gf <- npp_files_gf[grep(paste(years_of_data, collapse="|"), npp_files_gf)]


```


# Standardize catch

`catch_npp_fun` is a function that reads in a catch raster, reprojects & resamples it then divides by npp for the same year and saves the output.

```{r catch, eval=FALSE}

catch_npp_fun <- function(file, layer){ # file = dem_d_files[1]  layer = 'dem_dest'
  
  catch <- raster(file)
  
  yr <- substr(file, nchar(file)-10, nchar(file)-7)
  
  npp <- npp_files_gf[str_detect(npp_files_gf, yr)] %>% 
    raster()
  
  c <- catch %>%
        resample(npp, method = 'ngb') %>%
        overlay(., npp, fun=function(x,y){x/y})
  
  writeRaster(c, filename = file.path(dir_M, sprintf('git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_npp_rasters/%s/%s_%s.tif', layer, layer, yr)), overwrite=TRUE) 
}

```

## Apply the function
```{r, eval=F}

years_filter <-  paste(years_of_data, collapse="|")

dem_d_files <- list.files(file.path(dir_M,'git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_rasters/dem_dest'), full.names=TRUE, pattern = "gf")
dem_d_files <- dem_d_files[grep(years_filter, dem_d_files)]
dem_d_files <- dem_d_files[-grep("all_catch", dem_d_files)]

dem_nd_hb    <-list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_rasters/dem_nondest_hb'), full.names=TRUE, pattern = "gf")
dem_nd_hb <- dem_nd_hb[grep(years_filter, dem_nd_hb)]

dem_nd_lb    <-list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_rasters/dem_nondest_lb'), full.names=TRUE, pattern = "gf")
dem_nd_lb <- dem_nd_lb[grep(years_filter, dem_nd_lb)]

pel_hb    <-list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_rasters/pel_hb'), full.names=TRUE, pattern = "gf")
pel_hb <- pel_hb[grep(years_filter, pel_hb)]

pel_lb    <-list.files(file.path(dir_M, 'git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_rasters/pel_lb'), full.names=TRUE, pattern = "gf")
pel_lb <- pel_lb[grep(years_filter, pel_lb)]

lapply(dem_d_files, catch_npp_fun, layer = 'dem_dest')
lapply(dem_nd_hb, catch_npp_fun, layer = 'dem_nondest_hb')
lapply(dem_nd_lb, catch_npp_fun, layer = 'dem_nondest_lb')
lapply(pel_hb, catch_npp_fun, layer = 'pel_hb')
lapply(pel_lb, catch_npp_fun, layer = 'pel_lb')

```


# Rescale

Create an ocean raster with same scale as the fishery/npp rasters to cut land values.
```{r, eval=FALSE}

catch_npp <- list.files(file.path(dir_M,'git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_npp_rasters'), full.names = TRUE, recursive = TRUE)
catch_npp <- raster(catch_npp[1])

ocean_low_res <- ocean %>%
  resample(catch_npp, method="ngb")

```


Determine the 99.99th quantile for each type of commercial fisheries.  This value is saved in the comm_fish_ref_points.csv.
```{r, eval=FALSE}

#list of lists

all <- list.files(file.path(dir_M,'git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_npp_rasters'), full.names = TRUE, recursive = TRUE)

layers <- c("dem_dest","dem_nondest_hb","dem_nondest_lb","pel_hb","pel_lb")

for(layer in layers){

  files <-  all[which(str_detect(all, as.character(layer)))]
  
  values  <-  files %>%
          raster::stack() %>%
          raster::mask(ocean_low_res) %>%
          raster::getValues() 
  values_vector <- as.vector(as.matrix(values)) 
  values_vector <- na.omit(values_vector)
  
  quant <-  quantile(values_vector, prob=0.9999, na.rm=TRUE)
  
  refs <- read.csv("int/comm_fish_ref_points.csv") %>%
    dplyr::mutate(ref_point = ifelse(pressure == layer, quant, ref_point))
  write.csv(refs, "int/comm_fish_ref_points.csv", row.names=FALSE)
  
print(layer)
}  

```

   
For each type of commercial fisheries, rescale each year's data according to reference point and save output.

```{r, eval=FALSE}

# relevant files
all <- list.files(file.path(dir_M,'git-annex/impact_acceleration/stressors/comm_fish/int/catch_annual_npp_rasters'), full.names = TRUE, recursive = TRUE)

ref_points <- read.csv("int/comm_fish_ref_points.csv")

# function to run through the 5 categories
scale_fun <- function(prs){ # prs <- "pel_hb"

    files <-  all[which(stringr::str_detect(all, prs))]
 cat(prs)
    quant <- ref_points$ref_point[ref_points$pressure == prs]
 cat(quant)

    for(ind_rast in files){ # ind_rast=files[1]
 
         yr = substr(ind_rast, nchar(ind_rast)-7, nchar(ind_rast)-4)
          
        raster::raster(ind_rast) %>%
          raster::calc(fun=function(x){ifelse(x<0, 0,
                                      ifelse(x>quant, 1, x/quant))}) %>%
          raster::resample(ocean, method = 'ngb') %>%
            raster::mask(ocean,
            filename = file.path(dir_M, sprintf('git-annex/impact_acceleration/stressors/comm_fish/final/%s/%s_%s_rescaled_mol.tif', prs, prs, yr)), overwrite=TRUE)  
        # check:
        # tmp <- raster(file.path(dir_M, "git-annex/impact_acceleration/stressors/comm_fish/final/pel_hb/pel_hb_2003_rescaled_mol.tif"))
        # plot(tmp)
    }
}

# run function for 5 commercial fishing categories

layers <- c("dem_dest","dem_nondest_hb","dem_nondest_lb","pel_hb","pel_lb")

foreach(layer = layers,.packages="dplyr") %dopar%{ # file_name = l[1]
#for(layer in layers){
scale_fun(prs = layer)
}

```
