---
title: "Exploring Watson Fisheries Data"
author: "Jamie Afflerbach"
output:
 html_document:
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

## Setup
```{r}
library(RColorBrewer)
library(raster)
library(tidyverse)

dir_M <- "/home/shares/ohi"

cols <-  rev(colorRampPalette(brewer.pal(9, 'Spectral'))(255)) # rainbow color scheme
```

## Read in Data

* Industrial Catch (1950 - 1954) - reported, iuu, and discard catch data for each cell location and unique identifier
* Master Index File (Index.csv) - information associated with the unique identifiers in the Industrial Catch data
* Spatial Cells Reference (Cell.csv) - contains geospatial information associated wtih the Industrial Catch data

```{r}
data <- readRDS(file.path(dir_M, "git-annex/globalprep/_raw_data/IMAS_GlobalFisheriesLandings/d2018/CatchInd_1950_1954.rds"))

DT::datatable(head(data))
```

Hmm, it's weird that the data provided is only catch per cell per year. No information on species, gear, country. Another dataset provided is `Index.csv`. Maybe this has more information.

```{r}
index <- read_csv(file.path(dir_M, "git-annex/globalprep/_raw_data/IMAS_GlobalFisheriesLandings/d2018/Index.csv"))

DT::datatable(head(index))
```

Add in geospatial information (lat/lon/area)

```{r}
cells <- read_csv(file.path(dir_M, "git-annex/globalprep/_raw_data/IMAS_GlobalFisheriesLandings/d2018/Cells.csv"))

DT::datatable(head(cells))
```


## Combine Datasets

It looks like the **ID** category is used to link the two datasets. I'm going to test mapping this by using the code I have from [Ocean Health Index](https://mazu.nceas.ucsb.edu/rstudio/files/github/ohiprep_v2018/globalprep/prs_fish/v2017/fishing_pressure_layers.html). I'll select just the year 2014.

```{r}
df <- data %>%
  left_join(index, by = "ID") %>%
  dplyr::filter(Year == 1954) %>%
  dplyr::select(-IndReported, -IndIUU, -IndDiscards, -NIndReported, -NIndIUU, -NIndDiscards) %>% 
  left_join(cells, by = "Cell")
  
DT::datatable(head(df))
```

Now the data is joined. The columns **IndReported**, **IndIUU**, and **IndDiscards** are redundant for each ID, representing the total reported, IUU and discarded catch for each country/year/taxon names.


## Test/Check Data

Test to see whether the values for **IndReported**, **IndIUU**, and **IndDiscards** are sum totals of Clams cockles arkshells in the Republic of Korea in year 1953.
```{r}
korea_clam <- data %>% 
  filter(ID == 491423) %>% 
  mutate(tot_Reported = sum(Reported),
         tot_IUU = sum(IUU),
         tot_Discards = sum(Discards))

DT::datatable(head(korea_clam))
```

## Map Data

To map total catch per cell per year we need to sum up total catch (including reported, iuu, and discards) per cell.

```{r}
map_df <- df %>%
  mutate(cell_catch = Reported + IUU + Discards) %>%
  group_by(Cell) %>%
  summarize(catch = sum(cell_catch))

DT::datatable(head(map_df))
```

Create template raster from `seaaroundus`
```{r}
library(seaaroundus)

## First get the template raster with a resolution of 0.5 degree cells.
## getcells() functon is from the seaaroundus package
saup_cells <- getcells("POLYGON ((-180 90, 180 90, 180 -90, -180 -90, -180 90))")

saup_rast <- raster(ncol=720, nrow=360)

saup_rast[] <- saup_cells
```

Plot raster
```{r}
r <- raster::subs(saup_rast, map_df, by = 1, which = 2, subsWithNA=TRUE) 
#plot(r, axes = F, main = "1954 Total Industrial Harvest (tons)", col = cols)
plot(r, main = "2015 Total Industrial Harvest (tons)", col = cols)
```

