---
title: 'OHI 2018 - Clean Water - Pathogen Pollution: Preparing Sanitation data'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../src/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: console
---


# Summary

This document describes the steps for preparing the pathogen pollution and pathogen pollution trend data layers for the 2018 global assessment.

The percentage of the population with access to improved sanitation facilities (World Health Organization and United Nations Children’s Fund, Joint Monitoring Programme, 2011) was used in combination with measurements of coastal population as a proxy for pathogens in coastal waters. Access to improved sanitation facilities is defined as the percentage of the population in a country with at least adequate access to disposal facilities that can effectively prevent human, animal, and insect contact with excreta. These data are a country-wide average (not specific to the coastal region). 

# Updates from previous assessment
Data source (WHO-UNICEF) updated the way they report the data. The 2018 assesment is calculated using the "At least Basic" variable because it is consistent with what have been used in previous assesmentes and the data is more complete. 

Other changes include:

1.  Improved gapfilling.  Previously, sanitation data was gapfilled using only UN georegion averages.  Now it is based on a: 1) uninhabited regions (perfect scores for these); 2) within country linear regression model to gapfill missing years for a region that otherwise has data; 3) UN georegion averages when no years of data are available.
2. instead of defining the reference point as the max value across all years with a 10% buffer (so the highest pressure was 0.9), we use the 99th quantile across year/region with no buffer (so the highest pressure is now 1).  We also constrain the reference point to the first 10 years of data (vs. all available years of data).
3. New population data
4. Updated sanitation data from WHO-UNICEF (although no new years)


**Consider for future assesments:** make the "Safely managed" data more complete, and reconsider whether these data would be better to use.

Definition of each variable according to the data source:
"At least basic": Use of improved facilities that are not shared with other households.

"Safely managed": Use of improved facilities that are not shared with other households and where excreta are safely disposed of in situ or transported and treated offsite



***

# Data Source
**Reference**: 
            https://washdata.org/data
            Updated July 2017

**Downloaded**: Downloaded 5/1/2018

**Description**:  Percentage of the National population that has access to improved facilities that are not shared with other households (National, At a basic level).

Access to improved sanitation facilities is defined as the percentage of the population within a country with at least adequate access to excreta disposal facilities that can effectively prevent human, animal, and insect contact with excreta.

**Native data resolution**: country-wide average (not specific to the coastal region)

**Time range**: 2000 - 2015

**Format**:  csv

***
  
# Methods
Percentage of people without sanitation was multiplied by the coastal population (within 25 miles of the coast) to estimate the number of people located along the coast without sanitation. This value was rescaled between 0 and 1 by dividing by the 99th quantile across all regions from years 2000 to 2009. 

***


# Setup

```{r setup, results="hide", message = FALSE, warning = FALSE, eval=FALSE}
## set working directory; comment out for knitting
#setwd("~/github/ohiprep_v2018/globalprep/prs_cw_pathogen/v2018")
#setwd("globalprep/prs_cw_pathogen/v2018")

##Packages
library(ohicore) 
devtools::install_github('ohi-science/ohicore@dev') # may require uninstall and reinstall
library(zoo)     # for na.locf: Last Observation Carried Forward
library(tidyverse)

```

# Import raw data

```{r raw_data, results="hide", message = FALSE, warning = FALSE, eval=FALSE}
##Directly from Mazu


##reads in the directory path to Mazu based on operating system 
source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/master/src/R/common.R")


##From Mazu
sani_raw <- read.csv(file.path(dir_M, "git-annex/globalprep/_raw_data/WHO_UNICEFF/d2018/JMP_2017_WLD.csv"),  header = FALSE, sep = ",", na.strings = c(NA, ''), skip = 2, stringsAsFactors = FALSE, strip.white = TRUE)


head(sani_raw)

```

# Methods

## Data wrangling
Selecting and naming the columns of interest. Scale the percentage of population with access to improved sanitation to proportion (from 0-1). Transform population and percentage into a numeric variable.

```{r data_wrangling, results="hide", message = FALSE, warning = FALSE, eval=FALSE}

sani <- sani_raw %>%
  dplyr::slice(-1) %>%                # cut first row with column names
  dplyr::select(country = V1,
         year = V3,
         pop = V4,
         basic_pct = V6) %>%
  dplyr::mutate(basic_pct = ifelse(basic_pct=="-", NA, basic_pct)) %>% 
  dplyr::mutate(pop = stringr::str_remove_all(pop, pattern = " ")) %>%
  dplyr::mutate_at(.vars = c("pop", "basic_pct", "year"), .funs = as.numeric)%>%
  dplyr::mutate(pop = pop * 1000,
          basic_prop = basic_pct/100)

```


Change names of regions to match names in ohicore and filter out regions that are not part of the OHI assesment or do not have data.

If after running 'name_2_rgn' (see next r chunk), there are some coastal regions that are not identified by `name_2rgn` function. They must be checked to determine how to best include them (or not include them).

```{r, results="hide", message = FALSE, warning = FALSE, eval=FALSE}

sani_improved <- sani %>%
   dplyr::mutate(country = ifelse(stringr::str_detect(country,"Cura\\Sao"), "Curacao", country))%>% 
   dplyr::mutate(country = ifelse(stringr::str_detect(country,"R\\Sunion"), "Reunion", country)) %>% 
   dplyr::mutate(country = ifelse(stringr::str_detect(country, "C\\Ste d\\SIvoire"), "Ivory Coast", country)) %>% 
   dplyr::mutate(country=ifelse(stringr::str_detect(country,"Hong Kong"), "Hong Kong", country)) %>% 
   dplyr::mutate(country=ifelse(stringr::str_detect(country,"Macao"), "Macao", country))

##For 2018 assesment
##Caribbean Netherlands are in a lower resolution than the OHI regions. They include: Bonaire, Sint Eustatius, Saba. However, there is no data Caribbean Netherlands so it is filter out.
sani_improved <- sani_improved %>% 
  dplyr::filter(country!= 'Caribbean Netherlands')

##If there were data for Caribbean Netherlands this is how to use the data of this region for Bonaire, Sint Eustatius, Saba, using the following code.

# CN <- filter(sani_improved, country=="Caribbean Netherlands") %>%
#   rename(country_old = country)
# CN_subregions <- data.frame(country_old = "Caribbean Netherlands",
#                             country = c("Bonaire", "Sint Eustatius", "Saba")) %>%
#   left_join(Carribean_neth) %>%
#   select(-country_old)
# sani_improved <- sani_improved %>%
#   filter(country != "Carribean Netherlands") %>%
#   rbind(CN_subregions)  

#Channel Islands correspond to OHI regions of Guernsey and Jersey. Here the data reported for Channel Islands is used for these two regions.

CI <- dplyr::filter(sani_improved, country=="Channel Islands") %>%
  dplyr::rename(country_old = country)
CI_subregions <- data.frame(country_old = "Channel Islands",
                            country = c("Guernsey", "Jersey")) %>%
  dplyr::left_join(CI) %>%
  dplyr::select(-country_old)

sani_improved <- sani_improved %>%
  dplyr::filter(country != "Channel Islands") %>%
  rbind(CI_subregions)  

```

Add rgn_id and merge duplicate regions using a mean weighted by population. 

```{r rgn_id, results="hide", message = FALSE, warning = FALSE, eval=FALSE}
rgn_sani <- name_2_rgn(sani_improved, 
                        fld_name     = 'country',
                        flds_unique  = c('country','year')) 

##Warning: These data were removed for not having any match in the lookup tables.
## Check list in warning to make sure all countries removed are not of interest for the OHI global assesment. In this case West Bank and Gaza Strip is not considered because it is a disputed area and Isle of Man is not consider an OHI region.

##Check for duplicate regions.
sort(table(rgn_sani$rgn_id)) #for this analysis, regions which are duplicated are: rgn_id = 209, 140, 116, 13

## Each region has 16 years of data in this case. If there are regions with more than 16 values is because in this database they are reported in a higher resolution than the OHI global assesment (eg: China, Hong Kong and Macao are just one region for the OHI global). For these regions a weighted average will be calculated as its final score.


## Calculating weighted means for the duplicated regions.

sani_final <- rgn_sani %>% 
  dplyr::group_by(rgn_id, rgn_name, year) %>% 
  dplyr::summarise(basic_sani_prop= weighted.mean(basic_prop, pop, na.rm = TRUE))

##Check for duplicate regions. At this point, all regions should have the same sample size.
sort(table(sani_final$rgn_id))

```

# Gapfilling

First step is to get an idea of what needs to be gapfilled.

```{r check, message = FALSE, warning = FALSE, eval=FALSE}

sani_gf <- sani_final %>% 
  dplyr::group_by(rgn_id, rgn_name) %>%
  dplyr::mutate(gf_count = sum(is.na(basic_sani_prop)))

summary(sani_gf)

## list of regions that will have gapfilling
dplyr::filter(sani_gf, gf_count>0) %>% 
  dplyr::select(rgn_id, rgn_name, gf_count) %>% 
  unique() %>%
  data.frame()

##Some regions have no data. Filter those out to be gapfilled later.

##Define number of years of data
years_data <- mean(table(sani_gf$rgn_id))

sani_gf <- sani_gf %>%
  dplyr::filter(gf_count != years_data)

```

## Gapfilling 1: Linear model
Use a linear model within country data to estimate missing years.

```{r gf_lm, results="hide", message = FALSE, warning = FALSE, eval=FALSE}

sani_gf_lm <- sani_gf %>%
  dplyr::group_by(rgn_id, rgn_name) %>%
  dplyr::do({
    mod <- lm(basic_sani_prop ~ year, data = .)
    gf_lm <- predict(mod, newdata = .[c('year')])
    data.frame(., gf_lm)
  }) %>%
  dplyr::ungroup()

summary(sani_gf_lm)

sani_gf_lm <- sani_gf_lm %>%
  mutate(gf_lm = ifelse(gf_lm > 1, 1, gf_lm)) %>% # constrain predictions to <=1 
  mutate(method = ifelse(is.na(basic_sani_prop), "lm prediction", NA)) %>%
  mutate(basic_sani_prop = ifelse(is.na(basic_sani_prop), gf_lm, basic_sani_prop))

```


## Gapfilling 2: Georegional averages

Georegional gapfilling for regions that do not have data.

```{r gf georgn, results="hide", message = FALSE, warning = FALSE, eval=FALSE}
##Bring from ohicore package georegions and georegions lables as variables
georegions       <- georegions
georegion_labels <- georegion_labels

year <- min(sani_gf_lm$year):max(sani_gf_lm$year) #defines the year range

sani_georgn_gf <- georegions %>%
  expand(year, georegions) %>%
  dplyr::left_join(sani_gf_lm, by = c('rgn_id', 'year')) %>%
  dplyr::left_join(georegion_labels, by = 'rgn_id') %>%
  dplyr::select(-r0) #r0 is the world 

##Identify uninhabited regions according to table in github
rgn_inhab <- read.csv('https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/master/src/LookupTables/rgn_uninhabited_islands.csv')

rgn_inhab <- rgn_inhab %>% 
  dplyr::filter(est_population < 100 | is.na(est_population)) %>%
  dplyr::mutate(inhab = 1) %>% 
  dplyr::select(rgn_id, inhab)

sani_georgn_gf <- sani_georgn_gf %>% 
  dplyr::left_join(rgn_inhab, by= 'rgn_id')


##Calculate two different gapfill columns using r2 and r1 UN geopolitical classification
sani_georgn_gf <- sani_georgn_gf %>%
  dplyr::group_by(year, r2) %>%
  dplyr::mutate(basic_sani_r2 = mean(basic_sani_prop, na.rm=TRUE)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(year, r1) %>%
  dplyr::mutate(basic_sani_r1 = mean(basic_sani_prop, na.rm=TRUE)) %>%
  dplyr::ungroup()%>%
  dplyr::arrange(rgn_id, year)


##First gapfill with r2, if no value available use r1; create column indicating whether value was gapfilled and if so, by what method. Give NA to inhabited regions
sani_georgn_gf <- sani_georgn_gf %>%
  dplyr::mutate(method = ifelse(is.na(basic_sani_prop) & !is.na(basic_sani_r2) & is.na(method) , "UN georegion avg. (r2)", method)) %>%
  dplyr::mutate(method = ifelse(is.na(basic_sani_prop) & is.na(basic_sani_r2) & !is.na(basic_sani_r1) & is.na(method), "UN georegion avg (r1)", method))%>%
  dplyr::mutate(method = ifelse(inhab %in% 1, 'uninhabited, perfect score', method)) %>% #label inhabited regions
  dplyr::mutate(basic_sani_prop = ifelse(is.na(basic_sani_prop) & !is.na(basic_sani_r2), basic_sani_r2, basic_sani_prop)) %>%
  dplyr::mutate(basic_sani_prop = ifelse(is.na(basic_sani_prop) & !is.na(basic_sani_r1), basic_sani_r1, basic_sani_prop)) %>%
  dplyr::select(rgn_id, rgn_label, year, basic_sani_prop, method, inhab)

##See regions that have not been gapfilled. 
dplyr::filter(sani_georgn_gf, is.na(basic_sani_prop)) %>% 
  dplyr::select(rgn_id, basic_sani_prop) %>% 
  unique() %>%
  data.frame() #NA values for inhabitated regions. 

```

## Gapfilling 3: Replace scores for uninhabited regions

Uninhabited regions get a perfect score.

```{r, results="hide", message = FALSE, warning = FALSE, eval=FALSE}
##Check for uninhabited rigions
dplyr::filter(sani_georgn_gf, !is.na(inhab)) %>%
  dplyr::select(rgn_id, rgn_label, basic_sani_prop, method) %>% 
  #unique() %>%
  data.frame() #NOTE: Some of the uninhabited regions have been gapfilled!


#Fill in all inhabited in rgn_inhab with perfect sanitation prop
sani_complete <- sani_georgn_gf %>% 
  dplyr::mutate(basic_sani_prop = ifelse(!is.na(inhab), 1, basic_sani_prop)) %>% 
  dplyr::mutate(gapfill = ifelse(is.na(method), 0, 1)) %>%
  dplyr::select(rgn_id, rgn_name = rgn_label, year, basic_sani_prop, gapfill, method)

summary(sani_complete)
```


# Save gapfilled data records

```{r, results="hide", message = FALSE, warning = FALSE, eval=FALSE}
#write.csv(sani_complete, "intermediate/sani_complete.csv", row.names=FALSE)

sani_complete <- read.csv("intermediate/sani_complete.csv")

# save gapfilling info
gf_data <- sani_complete %>%
  dplyr::select(rgn_id, year, gapfill, method)

#write.csv(gf_data, "output/po_pathogen_popdensity25mi_gf.csv", row.names=FALSE)

gf_data_trend <- sani_complete %>%
  dplyr::arrange(rgn_id, year) %>%
  dplyr::group_by(rgn_id) %>%
  dplyr::mutate(gapfill_5year = rollsum(gapfill, 5, align="right", fill=NA)) %>%
  dplyr::mutate(method = paste(na.exclude(unique(method)), collapse = ", ")) %>%
  dplyr::mutate(gapfill = gapfill_5year/5) %>%
  dplyr::select(rgn_id, year, gapfill, method)

#write.csv(gf_data, "output/po_pathogen_popdensity25mi_trend_gf.csv", row.names=FALSE)

```


# Standarizing sanitation data by population density

First calculate coastal population density (people/km^2^) is calculated by dividing the population within 25 miles of the coast by km^2^ within the 25 mile inland coastal area (yes! This is confusing because area is in km^2^, despite the boundary being 25 miles inland).

```{r, results="hide", message = FALSE, warning = FALSE, eval=FALSE}

## Population within 25 miles of coastline
population <- read.csv("../../mar_prs_population/v2018/output/mar_pop_25mi.csv") %>%
  dplyr::arrange(rgn_id, year)

#Note: there is a difference between the list of islands with no population in the mar_pop_25mi dataset with the uninhabited list in github. In the mar_pop_25mi, rgn_id 144 - Jan Mayen is not identified as a pop_0 region while in the github list it is. Population data also considers Antartica (rgn_id 213) (NOTE: Jan Mayen as a max population of 35-51 people, we will view this as uninhabited for these purposes, along with region 157)


#identify inhabited regions (popsum<100)
pop_0 <- population %>%
  dplyr::filter(popsum<100)
sort(table(pop_0$rgn_id))

#Read area 25mi inland to calculate population density
# (NOTE: this is confusing because it calculates the area in km2 for the 25 mile inland area)
area <- read.csv("../../mar_prs_population/v2018/int/area_km2_25mi.csv")

# People per km2 (for the 25 mile inland area)
pop_density <- population %>% 
  dplyr::left_join(area, by='rgn_id') %>% 
  dplyr::mutate(pop_per_km2= popsum/area_km2) %>% 
  dplyr::select(rgn_id, year, pop_per_km2)

##Save population density data
#write.csv(pop_density, "intermediate/pathogen_pop_density_25mi.csv", row.names=FALSE)

```

These data are transformed to a pressure, with a zero score indicating no pressure and 1 indicating the highest possible pressure. Given this we want to determine the number of people *without* access.

The number of people per km^2 without access to sanitation is calculated by:

1. converting proportion with access to sanitation to proportion without access to sanitation (i.e., 1 - proportion_with_access). 
2. The proportion without access is multiplied by the coastal population density.
3. Number of people without access are log transformed (ln(x+1))

```{r, message = FALSE, warning = FALSE, eval=FALSE}

unsani_pop <- sani_complete %>%  
    dplyr::select(rgn_id, rgn_name, year, basic_sani_prop) %>%
    dplyr::left_join(pop_density, 
              by=c('rgn_id', 'year')) %>%
    dplyr::mutate(propWO_x_pop     = (1 - basic_sani_prop) * pop_per_km2, # this calculates the population density of people without access (WO)
           propWO_x_pop_log = log(propWO_x_pop + 1)) # log is important because the skew was high otherwise

hist(unsani_pop$propWO_x_pop, main = "people without access")

hist(unsani_pop$propWO_x_pop_log, main = "log of people without access")

```

# Pressure Score

The reference point is the 99th quantile across all countries and years 2000-2009 as a reference point.

```{r pressure, eval=FALSE}

##Calculate reference point
ref_calc <- unsani_pop %>% 
  dplyr::filter(year %in% 2000:2009) %>% #years of reference
  ##summarise(ref= max(propWO_x_pop_log, na.rm = TRUE)*1.1) %>%  # old method
  dplyr::summarise(ref= quantile(propWO_x_pop_log, probs=c(0.99), na.rm = TRUE)) %>% 
  .$ref

##save to the master reference point list
##(edit file if necessary)
# master_refs <- read.csv("../../supplementary_information/v2018/reference_points_pressures.csv", stringsAsFactors = FALSE)
# master_refs$ref_point[master_refs$pressure == "Sanitation"] <- ref_calc
# write.csv(master_refs, "../../supplementary_information/v2018/reference_points_pressures.csv", row.names=FALSE)

master_refs <- read.csv("../../supplementary_information/v2018/reference_points_pressures.csv", stringsAsFactors = FALSE)
ref_value <- as.numeric(master_refs$ref_point[master_refs$pressure == "Sanitation"]) 
#ref_value <- 7.95 #v2018 data using old method
#ref_value <- 6.49  used in v2018

unsani_prs <- unsani_pop %>%
  dplyr::mutate(pressure_score = propWO_x_pop_log / ref_value) %>% 
  dplyr::mutate(pressure_score = ifelse(pressure_score>1, 1, pressure_score)) %>% #limits pressure scores not to be higher than 1
  dplyr::select(rgn_id, year, pressure_score) 

summary(unsani_prs)

#Save data pressure scores 
#write.csv(unsani_prs, "output/po_pathogen_popdensity25mi.csv", row.names=FALSE)


```


# Model Trend

Using `CalculateTrend` function form the `ohicore`, trend is calculated by applying a linear regression model to the pressuere scores using a window of 5 years of data. The solope of the linear regression (annual change in pressure) is then divided by the earliest year to get proportional change and then multiplied by 5 to get estimate trend on pressure in the next five years. 


```{r trend, eval=FALSE}
##Calculate trend using CalculateTrend()

##Define relevant years: Min and max year of data to calculate trend
assess_years <- 7  ##Years for which the trend is going to be calculated for. This will change every assessment year
maxyear <- max(unsani_prs$year) ##max year of data
minyear <- maxyear- assess_years +1 

trend_data <- data.frame() #create a data.frame to save trend socores


##For loop: calculates trend for all assess years within the corresponding 5 year window.
#focal_year is the year for which the trend is being calculated.
for(focal_year in minyear:maxyear){ #focal_year = 2009 

  trend_years <- (focal_year-4):focal_year #defines the 5 years window to calculate trend
  
  data_new <- unsani_prs %>% #format data to work in CalculateTrend()
    select(rgn_id, year, status=pressure_score)
  
trend_data_focal_year <- CalculateTrend(data_new, trend_years)

trend_data_focal_year <- trend_data_focal_year %>%
  mutate(year = focal_year) %>%
  select(rgn_id = region_id, year, trend=score) %>%
  data.frame()

trend_data <- rbind(trend_data, trend_data_focal_year) #add trend calculation to dataframe crearted outside the loop
}
summary(trend_data)

##Save trend data
#write.csv(trend_data, "output/po_pathogen_popdensity25mi_trend.csv", row.names=FALSE)

```



# Compare to previous years

### Trend data

There is not a great correspondence between this year and last year's trend estimates.  This is partially due to changes in the source data.  But mainly, it is due to the limited number of years used in last year's assessment to calculate trend.  

```{r, eval=FALSE}

new <- read.csv("output/po_pathogen_popdensity25mi_trend.csv")

old <- read.csv("../v2016/output/pathogens_popdensity25mi_trend_updated.csv")%>%
  select(rgn_id, year, old_trend = trend) %>%
  left_join(new, by=c('rgn_id', 'year'))

plot(old$old_trend, old$trend, xlim=c(-1,1), ylim=c(-1, 1))
abline(0,1, col="red")
#identify(old$old_trend, old$trend, labels = old$rgn_id)


data_old <- read.csv('../v2016/output/pathogens_popdensity25mi_updated.csv')
data_new <- read.csv('output/po_pathogen_popdensity25mi.csv')

filter(old, rgn_id==190)
filter(data_old, rgn_id==190)
filter(data_new, rgn_id==190)


```


### Pressure values
We checked the main discrepancies and these were due to changes in source data.  Fairly small changes in access can lead to fairly large changes in pressure scores, depending on the population. There are slightly higher pressure scores this year (indicated by points tending to be above the 1-1 red line) due to modifications of reference point calculations.

```{r, eval=FALSE}
new <- read.csv("output/po_pathogen_popdensity25mi.csv")

old <- read.csv("../v2016/output/pathogens_popdensity25mi_updated.csv") %>%
  select(rgn_id, year, old_pressure = pressure_score) %>%
  left_join(new, by=c('rgn_id', 'year'))

plot(old$old_pressure, old$pressure_score)
abline(0,1, col="red")
#identify(old$old_pressure, old$pressure_score, labels = old$rgn_id)

filter(old, rgn_id==125)

sani_complete <- read.csv("intermediate/sani_complete.csv")


old <- read.csv("../v2015/int/rgn_jmp_san_2015a_raw_prop_access.csv") %>%
  rename(old_access = access_pct)

# change in source data (Bermuda)
filter(old, rgn_id ==108)
filter(sani_complete, rgn_id == 108)

# Bahrain: relatively small change in source data, but given population, made a fairly large difference
filter(old, rgn_id ==52)
filter(sani_complete, rgn_id == 52)
filter(unsani_pop, rgn_id ==52)

# Malaysia: relatively small change in source data, but given population, made a fairly large difference
filter(old, rgn_id ==206)
filter(sani_complete, rgn_id == 206)
filter(unsani_pop, rgn_id ==206)

# Grenada: Change in source data
filter(old, rgn_id ==125)
filter(sani_complete, rgn_id == 125)
```

### Raw access data
(Note this also includes differences due to gapfilling from previous years).
```{r, eval=FALSE}

#compare raw data
old <- read.csv("../v2015/int/rgn_jmp_san_2015a_raw_prop_access.csv") %>%
  rename(old_access = access_pct)

new <- read.csv("intermediate/sani_complete.csv") %>%
  dplyr::select(rgn_id, year, new_access=basic_sani_prop) %>%
  left_join(old, by=c("rgn_id", "year"))

plot(new$new_access, new$old_access)
abline(0,1, col="red")
```


