---
title: 'OHI 2017: Commercial Fishing Pressure Layers '
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../src/templates/ohi_hdr.html'
pdf_document:
  toc: true
---

# Summary

The commercial fishing pressure layers are created from spatialized catch by gear data provided by Watson (2018), and net primary production data from the Vertically Generalized Production Model [(VGPM)](http://www.science.oregonstate.edu/ocean.productivity/) as described in [Behrenfeld and Falkowski (1997)](http://www.science.oregonstate.edu/ocean.productivity/references/L&O%201997a.pdf).

This script prepares and formats the IMAS Global Fisheries Catch raw data into intermediate data by combining CatchInd_XXXX_XXXX and CatchNInd_XXXX_XXXX rds files with Index.csv and Cells.csv as well as a single file with all years.

We create raster files of total annual global catch from the Watson data and correct for system productivity by dividing catch by NPP: [quick review](https://www.researchgate.net/publication/312614653_Reconciling_fisheries_catch_and_ocean_productivity).

# Updates from previous assessment

* No longer using gear type to classify catch data as high or low bycatch, since the IMAS data now provides values for **discards**, which we consider as bycatch. 
* Previously, the raw catch data was all in a single file. This year, we have to combine across three different data tables: catch data (CatchInd_XXXX_XXXX), master index file with country name, species, etc (Index.csv), and geospatial information (Cells.csv).
* Previously raw catch data was in tonnes/km^2^, but this year they are in tonnes. The area in km^2^ values are located in the Cells.csv file
* One other change from last year is that we will gapfill the missing NPP data prior to using

# Data Source

**Reference**: Watson, R. A. and Tidd, A. 2018. Mapping nearly a century and a half of global marine fishing: 1869–2015. Marine Policy, 93, pp. 171-177. [(Paper URL)](https://doi.org/10.1016/j.marpol.2018.04.023)

**Downloaded**: July 17, 2018 from [IMAS portal](http://data.imas.utas.edu.au/portal/search?uuid=ff1274e1-c0ab-411b-a8a2-5a12eb27f2c0) - clikc on download tab, step 3

**Description**:  Global fisheries landings data per cell separated by Industrial versus Non-Industrial catch, IUU, and discards.

**Native data resolution**:   

**Time range**: 1950 - 2015

**Format**:  CSV format

**Additional Information**: [Metadata](http://metadata.imas.utas.edu.au/geonetwork/srv/eng/metadata.show?uuid=ff1274e1-c0ab-411b-a8a2-5a12eb27f2c0), [Supplementary Material](https://ars.els-cdn.com/content/image/1-s2.0-S0308597X18300605-mmc1.docx)


***
  
# Setup

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}

library(tidyverse)
library(purrr)
library(here) # install.packages("here")
setwd(here::here("globalprep","prs_fish","v2018"))

source("https://rawgit.com/OHI-Science/ohiprep_v2018/master/src/R/common.R")

rawFolder <- file.path(dir_M, "git-annex/globalprep/_raw_data/IMAS_GlobalFisheriesLandings/d2018")
rastFolder <- paste0(dir_M,"/git-annex/globalprep/prs_fish/v2018/int/")

```

# New Global Fisheries Raw Data

Download the following from the **data_download.Rmd** script:

* Industrial Catch (1950 - 1954) - reported, iuu, and discard catch data for each cell location and unique identifier
* Master Index File (Index.csv) - information associated with the unique identifiers in the Industrial Catch data
* Spatial Cells Reference (Cell.csv) - contains geospatial information associated wtih the Industrial Catch data

# Explore Data

Read in Master Index file, Spatial cells, and a single 5-year Catch dataset. 

Filter master file to only include `ID`, `Year`, `CountryName`, `TaxonName`, `CommonName`, `FleetGearName` (I actually don't think we will need the `CountryName` or `FleetGearName`, but it might be interesting).

```{r}

## Master index file
master <- read.csv(file.path(rawFolder,"Index.csv")) %>% 
  select(ID, Year, CountryName, TaxonName, CommonName, FleetGearName)
DT::datatable(head(master))

## Spatial cells reference
spatialCells <- read.csv(file.path(rawFolder,"Cells.csv"))
DT::datatable(head(spatialCells))

```

```{r}

## Look at single catch file
data <- readRDS(file.path(rawFolder,"CatchInd_1950_1954.rds"))
DT::datatable(head(data))

```

Test to see whether the values for **IndReported**, **IndIUU**, and **IndDiscards** in the Master Index File are sum totals of Clams cockles arkshells in the Republic of Korea in year 1953.
```{r, eval=FALSE}

korea_clam <- data %>% 
  filter(ID == 491423) %>% 
  mutate(tot_Reported = sum(Reported),
         tot_IUU = sum(IUU),
         tot_Discards = sum(Discards))

master_check <- master %>%
  filter(ID == 491423)

## IndReported, IndIUU, IndDiscards should match tot_Reported, tot_IUU, and tot_Discards
DT::datatable(head(korea_clam))
head(master_check)

```

# Methods

## Wrangle 

Tidy Fisheries Files:
1. Combine the Master Index and Spatial Cells with the CatchInd and CatchNInd files.
2. Save each year of data as a separate file into: "globalprep/prs_fish/v2018/int/annual_catch"

**Function for Combining & Saving Files**: Create function to read in each file name, combine with Index and Cells data frame, and save each year of catch data into a separate file in mazu.

```{r tidy, eval=FALSE}

## Set function to read in each file, combine with Index.csv and Cells.csv, and save each year of data back into mazu
combine_fis <- function(x) {

  ## Read in the catch data
  ## Create a total Landings column
  ## Join with master and spatial cells file
  df <- readRDS(x) %>%
    mutate(Landings = Reported+IUU) %>% 
    left_join(master, by = "ID") %>% 
    left_join(spatialCells, by = "Cell")
  
  
  ## Save each individual year as a single file
  five_years <- sort(unique(df$Year)) 
  
  for(yr in five_years){
    print(yr) # will show you your progress
    
    single_yr_df <- df %>%
      filter(Year == yr)
    
    ## Save files with prefix CatchInd or CatchNInd
    ## Remove the suffix starting with '_' to get CatchInd or CatchNInd
    ind_Nind <- basename(x) %>% 
      tools::file_path_sans_ext() %>% 
      str_remove("_.*") # remove everything after the first underscore
    
    write_rds(single_yr_df, paste0(dir_M,"/git-annex/globalprep/prs_fish/v2018/int/annual_catch/", ind_Nind, "_", yr, ".rds"))
    
  }
  
  }

```

**Industrial Catch Data**

Create list of industrial catch file names and apply the `combine_fis` function to save each individual 5-year interval catch data.

```{r, eval=FALSE}

ind_files <- dir(file.path(dir_M,"git-annex/globalprep/prs_fish/v2018/int/annual_catch"), pattern ="CatchInd", full.names=TRUE)

## Wrangle and Save Each Year of Data Separately
indCatch <- map_df(ind_files, combine_fis)

```

**Non-industrial Catch Data**

Create list of non-industrial catch file names and apply the `combine_fis` function to save each individual 5-year interval catch data. Then, create a single file that has all years of data.

```{r, eval=FALSE}

nind_files <- dir(file.path(dir_M,"git-annex/globalprep/prs_fish/v2018/int/annual_catch"), pattern ="CatchNInd", full.names=TRUE)

## Wrangle and Save Each Year of Data Separately
nindCatch <- map_df(nind_files, combine_fis)

```


## Create Annual Catch Rasters

We want a landings raster and a discards raster for Industrial (commericial) and Non-Industrial (artisanal) data per year. 

Note: Annual catch per cell contains values in units of kg per km^2^. Since values in `Reported`, `Discards`, `Landings`, and `IUU` are in tonnes, they must be first converted to kg then divided by the `Area` column.

1. Setup 
```{r annRasters, eval=FALSE}

library(tidyverse)
library(parallel)
library(foreach)
library(doParallel)
library(raster)
library(rasterVis)
library(seaaroundus)
library(RColorBrewer)
library(cowplot)
library(stringr)
library(colorspace)
library(sp)

registerDoParallel(5) # registering cores for parallel processing

## color palette
cols = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255)) # rainbow color scheme
mytheme=rasterTheme(region=cols)

## Set template ocean raster and mollweide projection CRS
ocean <- raster::raster(file.path(dir_M, 'model/GL-NCEAS-Halpern2008/tmp/ocean.tif'))
mollCRS=crs('+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs')

options(scipen=999)

```

2. First get the template raster with a resolution of 0.5 degree cells. The getcells() function comes from the seaaroundus R package.

The values of these cells are the Cell ID numbers. In the fisheries catch dataset, these Cell ID numbers match up with the "Seq" numbers.

```{r, eval=FALSE}

saup_cells <- getcells("POLYGON ((-180 90, 180 90, 180 -90, -180 -90, -180 90))")

saup_rast <- raster(ncol=720, nrow=360)
saup_rast[] <- saup_cells
   
plot(saup_rast,col=cols,main = "SAUP Cell IDs")

```

3. Create commercial and artisanal rasters

* comm_landings_XXXX.tif
* comm_discards_XXXX.tif
* artisanal_landings_XXXX.tif
* artisanal_discards_XXXX.tif

Catch data is in tonnes, so need to convert to kg then divide by `Area` (km^2^) to get catch per area. Since `Cell` is the identifier for a specific 30-min spatial cell, we want to add up total landings and total discards grouped by `Cell`.

Note: Subset for data 2003 and onwards since NPP data starts at 2003. Since we will be taking 5-year averages, we will need years 1999-2003 to get the average value for year 2003.

```{r, eval=FALSE}

## Specify years of data, file locations, raster output location
years = c(1999:2015)
annual_files <- list.files(file.path(rawFolder, "int"), full.names=TRUE)


## Specify a list of arguments - commercial or artisanal - for reading in data and saving raster file name
raw_suffx <- list(catchind = "CatchInd", catchnind = "CatchNInd")
rast_prefx <- list(comm = "comm", artisanal = "artisanal")

## Function for totalling landings/discards and saves raster files 
catch2raster <- function(raw_suffx,rast_prefx){
  
foreach(yr = years) %dopar% { #yr = 2003
  
  ## find file path of the respective year of data
  yr <- as.character(yr)
  ## Select the respective year of industrial catch data
  dataname <- str_subset(annual_files, paste0(raw_suffx,"*.",yr))
  ## read in raw data
  raw_data <- readRDS(dataname)
 
  ## Total Landings per Cell
  landings <- raw_data %>%
    dplyr::mutate(Landings_CR = (Landings * 1000)/Area) %>% # convert to catch rate (kg/km^2)
    dplyr::group_by(Cell) %>%
    dplyr::summarise(cell_catch = sum(Landings_CR, na.rm=TRUE)) %>% # usually no NAs, but just in case
    dplyr::ungroup()

  ## Total Discards per Cell
  discards <- raw_data %>%
    dplyr::mutate(Discards_CR = (Discards * 1000)/Area) %>% 
    dplyr::group_by(Cell) %>%
    dplyr::summarise(cell_catch = sum(Discards_CR)) %>% 
    dplyr::ungroup()
  
  ## Rasterize Catch: swapping template cell values with those in dataframe
  raster::subs(saup_rast, landings, by = "Cell", which = "cell_catch", subsWithNA=TRUE, 
               filename = paste0(rastFolder, rast_prefx, '_landings/annual_catch_', yr ,'.tif'), overwrite=TRUE)
  raster::subs(saup_rast, discards, by = "Cell", which = "cell_catch", subsWithNA=TRUE, 
               filename = paste0(rastFolder, rast_prefx, '_discards/annual_catch_', yr ,'.tif'), overwrite=TRUE) 
  
  }
}

## Applies catch2raster function on commercial (industrial) then artisanal (non-industrial) files
create_rast <- map2(raw_suffx, rast_prefx, catch2raster)

```

## Check Rasters

Compare composite of newly created rasters with industrial landings figure created in [Watson (2018)](https://www.sciencedirect.com/science/article/pii/S0308597X18300605). See [Figure 2D](https://ars.els-cdn.com/content/image/1-s2.0-S0308597X18300605-gr2_lrg.jpg), which maps industrial landings in tonnes between 2000 and 2015.

```{r, eval=FALSE}

## Read in rasters for commercial landings between 2000 and 2015
commercial <- list.files(paste0(rastFolder,"comm_landings"),full.names = TRUE)

## Plot one year of data
catch <- raster(commercial[17])
plot(catch)
res(catch) # 0.5 degree cells

## Compare 2000-2015 industrial catch (log-transformed) to Watson figure
## create a raster stack from the input raster files (exclude 1999 which is commercial[1])
allRasters <- raster::stack(commercial[2:17])

## run the sum function on the raster stack - i.e. add (non-cumulatively) the rasters together
tot_Catch <- sum(allRasters)

## apply log to adjust for really large values
log_Catch <- raster::calc(tot_Catch, function(x){log(x+1)},
             filename = paste0(rastFolder,"datacheck/commLand_2000_2015_log.tif"), overwrite=TRUE)

## Plot & Save raster
png(filename="figs/log_IndCatch_2000_2015.png")
plot(log_Catch, col=cols, alpha=1)
dev.off()

## Check out some values or zoom in on an area
click(log_Catch)
zoom(log_Catch)


```

## Standardize Fisheries Catch

Use **net primary production** (NPP) data to correct global fisheries catch for spatial differences in ecosystem impact

* Prepare and gapfill NPP data in [**npp.Rmd**](https://mazu.nceas.ucsb.edu/rstudio/files/github/ohiprep_v2018/globalprep/prs_fish/v2018/npp.html)
* Resample NPP and fish catch rasters to ocean raster
* Correct fish catch with NPP (catch divided by NPP)

Read in gapfilled NPP rasters
```{r gapfill}

npp_files_gf <- list.files(file.path(dir_M,"git-annex/globalprep/prs_fish/v2018/VGPM_primary_productivity/int/annual_npp"), full.names=TRUE, pattern = "gf")

plot(raster(npp_files_gf[13]), col=cols, axes=FALSE, main = 'Net Primary Production (mg C/m2/day) \n 2015')

```

Create function standardizing catch by NPP

* read in a single year of catch data
* read in the same year of npp data
* transform npp raster to match the crs and res of the catch raster (coarser resolution will run faster)
* correct fisheries catch with npp raster (the land values should turn back to NAs since they are NA in catch rasters)
* save corrected fish catch rasters!!

Note: there are two methods I tested to reproject and resample the catch rasters to match the CRS and resolution of the gapfilled NPP rasters (#45). Both likely work the same as the output raster information and raster plot look the same. Neither prevent wrapping of the continents outside the globe, despite setting `over=TRUE`, probably because the gapfilled NPP raster is a bit strange. Will be masked later after resampling catch rasters down to 1-km resolution.
```{r, eval=FALSE}

catch_npp_fun <- function(file, layer){ # file = dem_d_files[1]  layer = 'dem_dest'
  
  catch <- raster(file)
  
  yr <- str_extract(basename(file),"(\\d)+") # extracts the year, double check yr
  
  npp <- npp_files_gf[str_detect(npp_files_gf, yr)] %>% 
    raster()
  
  ## Two methods of reprojecting and resampling to match the crs and resolution of the catch rasters to NPP raster (#45)
  # catch_resmp <- catch %>%
  #   projectRaster(npp, method = 'ngb', over=TRUE, filename = file.path(dir_M,"git-annex/globalprep/prs_fish/v2018/tmp/temp_resample1.tif"), overwrite=TRUE)
  catch_resmp <- catch %>%
    projectRaster(crs=crs(npp), method = 'ngb', over=TRUE, filename = file.path(dir_M,"git-annex/globalprep/prs_fish/v2018/tmp/temp_resample2-1.tif"), overwrite=TRUE) %>%
    resample(npp, method = 'ngb', filename = file.path(dir_M,"git-annex/globalprep/prs_fish/v2018/tmp/temp_resample2-2.tif"), overwrite = TRUE)
    
  catch_corr <- overlay(catch_resmp, npp, fun=function(x,y){x/y}, filename = file.path(dir_M, sprintf("git-annex/globalprep/prs_fish/v2018/int/%s/annual_catch_corr_%s.tif",layer,yr)), overwrite=TRUE)
  
}

```


Create corrected catch rasters - apply `catch_npp_fun` function. Resolution and projection will be matched with NPP rasters (8 x 10 km resolution and Mollweide coordinate system).

```{r corrCatch, eval=FALSE}

## Specify years of catch data that match with years available in NPP data
years_of_data <- 2003:2015
years_filter <-  paste(years_of_data, collapse="|")

## Get file names in each of the four catch categories
comm_land <- list.files(paste0(rastFolder,"comm_landings"), pattern = "annual_catch_20", full.names = TRUE)
comm_land <- comm_land[grep(years_filter, comm_land)]

comm_disc <- list.files(paste0(rastFolder,"comm_discards"), pattern = "annual_catch_20", full.names = TRUE)
comm_disc <- comm_disc[grep(years_filter, comm_disc)]
  
art_land <- list.files(paste0(rastFolder,"artisanal_landings"), pattern = "annual_catch_20", full.names = TRUE)
art_land <- art_land[grep(years_filter, art_land)]

art_disc <- list.files(paste0(rastFolder,"artisanal_discards"), pattern = "annual_catch_20", full.names = TRUE)
art_disc <- art_disc[grep(years_filter, art_disc)]

lapply(comm_land, catch_npp_fun, layer = 'comm_landings')
lapply(comm_disc, catch_npp_fun, layer = 'comm_discards')
lapply(art_land, catch_npp_fun, layer = 'artisanal_landings')
lapply(art_disc, catch_npp_fun, layer = 'artisanal_discards')


```

Check corrected raster

Check if map plots outside of the global ellipse:
- See example of issue in [stackoverflow](https://stackoverflow.com/questions/27535047/how-to-properly-project-and-plot-raster-in-r)
- If it is plotting weirdly, will eventually be fixed in the final output map, but may be confusing throughout the data processing steps

**MAP PLOTS OUTSIDE ELLIPSE (8/9/18)....**

```{r}

commL_corr <- list.files(paste0(rastFolder,'comm_landings'), pattern="corr", full.names=TRUE)
commD_corr <- list.files(paste0(rastFolder,'comm_discards'), pattern="corr", full.names=TRUE)
artL_corr <- list.files(paste0(rastFolder,'artisanal_landings'), pattern="corr", full.names=TRUE)
artD_corr <- list.files(paste0(rastFolder,'artisanal_discards'), pattern="corr", full.names=TRUE)

## Check out 2015 catch corrected raster
plot(raster(artD_corr[13]))
zoom(raster(commL_corr[13]))
click(raster(commL_corr[13]))

```

## Calculate 5-Year Means

Calculate 5-year means for commercial and artisanal landings & discards. Set up foreach loop to start with the oldest year of data (2003) and end on the 5th most recent year of data (2011).

Note: Created mean catch rasters on 8/6/18. Re-created mean catch rasters for commercial landings on 8/9/18.
```{r catchmeans, eval=FALSE}

registerDoParallel(5)

## check to see which years of data to run in the foreach() loop
# years_of_data[[1]]:years_of_data[[length(years_of_data)-4]]

foreach (i = 2003:2011) %dopar%{ # i = 2010
  
  yrs <- c(i:(i+4))
  
  out_CL <- commL_corr[which(str_detect(commL_corr, pattern = paste(yrs,collapse = '|')))] %>%
            stack() %>%
            calc(fun=function(x){mean(x, na.rm=TRUE)}) %>%
            calc(fun=function(x){log(x+1)}, filename = paste0(rastFolder, 'comm_landings/mean_catch_',yrs[1],'_',yrs[5],'.tif'), overwrite=TRUE)

  out_CD <- commD_corr[which(str_detect(commD_corr, pattern = paste(yrs,collapse = '|')))] %>%
            stack() %>%
            calc(fun=function(x){mean(x,na.rm=TRUE)}) %>%
            calc(fun=function(x){log(x+1)}, filename = paste0(rastFolder, 'comm_discards/mean_catch_',yrs[1],'_',yrs[5],'.tif'), overwrite=TRUE)

  out_AL <- artL_corr[which(str_detect(artL_corr, pattern = paste(yrs,collapse = '|')))] %>%
            stack()%>%
            calc(fun=function(x){mean(x,na.rm=T)})%>%
            calc(fun=function(x){log(x+1)}, filename = paste0(rastFolder, 'artisanal_landings/mean_catch_',yrs[1],'_',yrs[5],'.tif'), overwrite=T)

  out_AD <- artD_corr[which(str_detect(artD_corr, pattern = paste(yrs,collapse = '|')))] %>%
            stack()%>%
            calc(fun=function(x){mean(x,na.rm=T)})%>%
            calc(fun=function(x){log(x+1)}, filename = paste0(rastFolder, 'artisanal_discards/mean_catch_',yrs[1],'_',yrs[5],'.tif'), overwrite=T)

}

```

Check 5-year mean raster - make sure land values are NA not zero.

```{r}

mean_CL <- list.files(paste0(rastFolder,'comm_landings'), pattern="mean_catch_20", full.names=TRUE)
mean_CD <- list.files(paste0(rastFolder,'comm_discards'), pattern="mean_catch_20", full.names=TRUE)
mean_AL <- list.files(paste0(rastFolder,'artisanal_landings'), pattern="mean_catch_20", full.names=TRUE)

plot(raster(mean_CL[1]), col=cols)
plot(raster(mean_CD[1]), col=cols)
plot(raster(mean_AL[1]), col=cols)

```

## Reference Point

Find 99.99th quantile to use as reference point:

* `ref_clb` is reference point for Commercial Low Bycatch
* `ref_chb` is reference point for Commercial High Bycatch
* `ref_alb` is reference point for Artisanal Low Bycatch

### Commercial Low Bycatch

Get industrial landings catch values across all years 2003-2015.

```{r, eval=FALSE}

vals <- c()

for(i in 2007:2015){ # i = 2015
#print(i)
  m <- mean_CL[which(str_detect(mean_CL, pattern = paste0("_",i,".tif")))] %>%
    raster()%>%
    getValues()
  
 n <- m[!is.na(m)]

  vals <- c(vals, n)

}

ref_clb <- quantile(vals, prob = 0.9999, na.rm=T)  # 10.64
  
```

### Commercial High Bycatch

Get industrial discards values across all years 2003-2015.

```{r, eval=FALSE}

vals <- c()

for(i in 2007:2015){ # i = 2007
#print(i)
  m <- mean_CD[which(str_detect(mean_CD, pattern = paste0("_",i,".tif")))] %>%
    raster()%>%
    getValues()
  
 n <- m[!is.na(m)]

  vals <- c(vals, n)

}

ref_chb <- quantile(vals, prob = 0.9999, na.rm=T)  # 9.88

```

### Artisanal Low Bycatch

Get artisanal discards values across all years 2003-2015.

```{r, eval=FALSE}

vals <- c()

for(i in 2007:2015){ # i = 2007
#print(i)
  m <- mean_AL[which(str_detect(mean_AL, pattern = paste0("_",i,".tif")))] %>%
    raster()%>%
    getValues()
  
 n <- m[!is.na(m)]

  vals <- c(vals, n)

}

ref_alb <- quantile(vals, prob = 0.9999, na.rm=T)  # 10.72
```

Save your reference points to a data frame and save to file
```{r}

pressures <- c("fp_com_lb","fp_com_hb","fp_art_lb")
ref_point <- as.numeric(c(ref_clb,ref_chb,ref_alb))
ref_df <- as.data.frame(cbind(pressures,ref_point))
write.csv(ref_df, "int/fish_ref_points.csv", row.names=FALSE)

```

Note: all quantiles increased slightly from last year's assessment (see [v2017 prep](https://github.com/OHI-Science/ohiprep_v2018/blob/master/globalprep/prs_fish/v2017/fishing_pressure_layers.Rmd)).


## Rescale (0-1) & Resample (1-km res)

Use reference points `ref_clb` on commercial landings data, `ref_chb` on commercial discards data, and `ref_alb` on artisanal landings data.

Rasters are currently in resoution 8.35 km x 10.3 km (same as NPP rasters) and in Mollweide coordinate system. Convert to ocean raster 1-km x 1-km resolution.

```{r, eval=FALSE}

mean_CL <- list.files(paste0(rastFolder, 'comm_landings'), pattern = "mean_catch_20", full.names=TRUE)
mean_CD <- list.files(paste0(rastFolder, 'comm_discards'), pattern = "mean_catch_20", full.names=TRUE)
mean_AL <- list.files(paste0(rastFolder, 'artisanal_landings'), pattern = "mean_catch_20", full.names=TRUE)
#mean_AD <- list.files(paste0(rastFolder, 'artisanal_discards'), pattern = "mean_catch", full.names=TRUE)

foreach (i = 2003:2011) %dopar%{ # i = 2011
  
  yrs <- c(i:(i+4))

  out_CL <- mean_CL[which(str_detect(mean_CL, pattern = paste0("_",i,"_")))] %>%
            raster() %>%
            calc(fun=function(x){ifelse(x > ref_clb, 1, x/ref_clb)}) %>%
            calc(fun=function(x){ifelse(x < 0, 0, x)}) %>%
            resample(ocean, method = 'ngb', filename = paste0(rastFolder, 'comm_landings/mean_catch_1km_', yrs[1], '_', yrs[5], '.tif'), overwrite = TRUE)


  out_CD <- mean_CD[which(str_detect(mean_CD, pattern = paste0("_",i,"_")))] %>%
            raster() %>%
            calc(fun=function(x){ifelse(x > ref_chb, 1, x/ref_chb)}) %>%
            calc(fun=function(x){ifelse(x < 0, 0, x)}) %>%
            resample(ocean, method = 'ngb', filename = paste0(rastFolder, 'comm_discards/mean_catch_1km_', yrs[1], '_', yrs[5], '.tif'), overwrite = TRUE)



  out_AL <- mean_AL[which(str_detect(mean_AL, pattern = paste0("_",i,"_")))] %>%
            raster() %>%
            calc(fun = function(x){ifelse(x > ref_alb, 1, x/ref_alb)}) %>%
            calc(fun = function(x){ifelse(x < 0, 0, x)}) %>%
            resample(ocean, method = 'ngb', filename = paste0(rastFolder, 'artisanal_landings/mean_catch_1km_', yrs[1], '_', yrs[5], '.tif'), overwrite = TRUE) 

}

```

Check out intermediate results
```{r}

cd_2003_2007 <- raster(file.path(rastFolder, 'comm_discards/mean_catch_1km_2003_2007.tif'))
cd_2004_2008 <- raster(file.path(rastFolder, 'comm_discards/mean_catch_1km_2004_2008.tif'))
cd_2005_2009 <- raster(file.path(rastFolder, 'comm_discards/mean_catch_1km_2005_2009.tif'))
cd_2006_2010 <- raster(file.path(rastFolder, 'comm_discards/mean_catch_1km_2006_2010.tif'))
cd_2007_2011 <- raster(file.path(rastFolder, 'comm_discards/mean_catch_1km_2007_2011.tif'))
cd_2008_2012 <- raster(file.path(rastFolder, 'comm_discards/mean_catch_1km_2008_2012.tif'))
cd_2009_2013 <- raster(file.path(rastFolder, 'comm_discards/mean_catch_1km_2009_2013.tif'))
cd_2010_2014 <- raster(file.path(rastFolder, 'comm_discards/mean_catch_1km_2010_2014.tif'))
cd_2011_2015 <- raster(file.path(rastFolder, 'comm_discards/mean_catch_1km_2011_2015.tif'))
all_cd <- stack(cd_2003_2007, cd_2004_2008,cd_2005_2009, cd_2006_2010, cd_2007_2011, cd_2008_2012, cd_2009_2013, cd_2010_2014, cd_2011_2015)
plot(all_cd, col=cols, axes=FALSE ,box=FALSE)

```

## Mask with Ocean Raster

```{r}

foreach (i = 2003:2011) %dopar%{ # i = 2011
  
  yrs <- c(i:(i+4))
  
  ## Commercial Low Bycatch (Landings)
  cl_mask <- raster(paste0(rastFolder, 'comm_landings/mean_catch_1km_', yrs[1], '_', yrs[5], '.tif'))
  cl_output <- mask(cl_mask, ocean, filename = paste0(dir_M, '/git-annex/globalprep/prs_fish/v2018/output/comm_low_bycatch/lb_fish_pressure_', yrs[1], '_', yrs[5], '.tif'), overwrite = TRUE)
  
  ## Commercial High Bycatch (Discards)
  # cd_mask <- raster(paste0(rastFolder, 'comm_discards/mean_catch_1km_', yrs[1], '_', yrs[5], '.tif'))
  # cd_output <- mask(cd_mask, ocean, filename = paste0(dir_M, '/git-annex/globalprep/prs_fish/v2018/output/comm_high_bycatch/hb_fish_pressure_', yrs[1], '_', yrs[5], '.tif'), overwrite = TRUE)
  
  
  ## Artisanal Low Bycatch (Landings)
  art_mask <- raster(paste0(rastFolder, 'artisanal_landings/mean_catch_1km_', yrs[1], '_', yrs[5], '.tif'))
  art_output <- mask(art_mask, ocean, filename = paste0(dir_M, '/git-annex/globalprep/prs_fish/v2018/output/art_low_bycatch/art_fish_pressure_', yrs[1], '_', yrs[5], '.tif'), overwrite = TRUE)
  
}

```

Check masked rasters!
```{r}

outFolder <-  "/home/shares/ohi/git-annex/globalprep/prs_fish/v2018/output"
cd_2003_2007 <- raster(file.path(outFolder, 'art_low_bycatch/art_fish_pressure_2003_2007.tif'))
cd_2004_2008 <- raster(file.path(outFolder, 'art_low_bycatch/art_fish_pressure_2004_2008.tif'))
cd_2005_2009 <- raster(file.path(outFolder, 'art_low_bycatch/art_fish_pressure_2005_2009.tif'))
cd_2006_2010 <- raster(file.path(outFolder, 'art_low_bycatch/art_fish_pressure_2006_2010.tif'))
cd_2007_2011 <- raster(file.path(outFolder, 'art_low_bycatch/art_fish_pressure_2007_2011.tif'))
cd_2008_2012 <- raster(file.path(outFolder, 'art_low_bycatch/art_fish_pressure_2008_2012.tif'))
cd_2009_2013 <- raster(file.path(outFolder, 'art_low_bycatch/art_fish_pressure_2009_2013.tif'))
cd_2010_2014 <- raster(file.path(outFolder, 'art_low_bycatch/art_fish_pressure_2010_2014.tif'))
cd_2011_2015 <- raster(file.path(outFolder, 'art_low_bycatch/art_fish_pressure_2011_2015.tif'))
all_cd <- stack(cd_2003_2007,cd_2004_2008,cd_2005_2009, cd_2006_2010,cd_2007_2011, cd_2008_2012,cd_2009_2013,  cd_2010_2014,cd_2011_2015)
plot(all_cd, col=cols, axes=FALSE ,box=FALSE)

cd_2003_2007 <- raster(file.path(outFolder, 'comm_low_bycatch/lb_fish_pressure_2003_2007.tif'))
cd_2004_2008 <- raster(file.path(outFolder, 'comm_low_bycatch/lb_fish_pressure_2004_2008.tif'))
cd_2005_2009 <- raster(file.path(outFolder, 'comm_low_bycatch/lb_fish_pressure_2005_2009.tif'))
cd_2006_2010 <- raster(file.path(outFolder, 'comm_low_bycatch/lb_fish_pressure_2006_2010.tif'))
cd_2007_2011 <- raster(file.path(outFolder, 'comm_low_bycatch/lb_fish_pressure_2007_2011.tif'))
cd_2008_2012 <- raster(file.path(outFolder, 'comm_low_bycatch/lb_fish_pressure_2008_2012.tif'))
cd_2009_2013 <- raster(file.path(outFolder, 'comm_low_bycatch/lb_fish_pressure_2009_2013.tif'))
cd_2010_2014 <- raster(file.path(outFolder, 'comm_low_bycatch/lb_fish_pressure_2010_2014.tif'))
cd_2011_2015 <- raster(file.path(outFolder, 'comm_low_bycatch/lb_fish_pressure_2011_2015.tif'))
all_cd <- stack(cd_2003_2007,cd_2004_2008,cd_2005_2009, cd_2006_2010,cd_2007_2011, cd_2008_2012,cd_2009_2013,  cd_2010_2014,cd_2011_2015)
plot(all_cd, col=cols, axes=FALSE ,box=FALSE)
```

# Extract Data from Rasters

Summary: Combine all years of data for catch types **low bycatch commercial**, **high bycatch commercial**, and **low bycatch artisanal** and extract the average pressure data for each region. Do some tidying and save the three outputs.

## Get OHI Raster/Zonal Data:

* **regions_eez_with_fao_ant.tif** - This includes all the ocean regions (eez/fao/antarctica), but the raster cell values correspond to the rgn_ant_id in regions_2017_update.  This file is most often used to extract pressure values for each region.
* **regionData.csv** - has data for spatial id's used in raster

```{r, eval=FALSE}

## Grab OHI base data for pressures
zones <- raster(file.path(dir_M, "git-annex/globalprep/spatial/v2017/regions_eez_with_fao_ant.tif")) 
rgn_data <- read.csv(file.path(dir_M, "git-annex/globalprep/spatial/d2014/data/rgn_mol_raster_1km/regionData.csv"))

```

## Commercial Low Bycatch Data:

```{r, eval=FALSE}

## combine all years
rasts_lb <- list.files(file.path(outFolder,'comm_low_bycatch'), full =TRUE)
press_stack_lb <- stack(rasts_lb)

## extract data for each region:
regions_stats_lb <- zonal(press_stack_lb,  zones, fun="mean", na.rm=TRUE, progress="text")
regions_stats2_lb <- data.frame(regions_stats_lb) 

write.csv(regions_stats2_lb, "int/comm_low_bycatch.csv", row.names=FALSE)

```

Compare zone IDs between intermediate data with **regionData.csv**. Should be no difference in `ant_id` zone ids and the `zone` column in the new data frame.

```{r, eval=FALSE}
setdiff(regions_stats2_lb$zone, rgn_data$ant_id) 
setdiff(rgn_data$ant_id, regions_stats2_lb$zone)
```

Tidy and create output data. 

* Preserve all rows in `regions_stats2_lb` even if it doesn't have a match in `rgn_data`.
* Gather all years of data together
* Fix up year column, selecting for last year of data in the 5-year interval
```{r}

data <- merge(rgn_data, regions_stats2_lb, all.y=TRUE, by.x="rgn_id", by.y="zone") %>%
  tidyr::gather("year", "pressure_score", starts_with("lb")) 

DT::datatable(data) # take a look

lb_data <- data %>%
  dplyr::mutate(year = stringr::str_sub(year, -4, -1)) %>%
  dplyr::mutate(year = as.numeric(year)) %>%
  dplyr::filter(rgn_id <= 250) %>%
  dplyr::select(rgn_id, year, pressure_score)

write.csv(lb_data, "output/comm_lb.csv", row.names=FALSE)

```

Take a look at output of low bycatch data.

Note: Bouvet Island has NAs for all years.
```{r}

summary(lb_data)
filter(lb_data, is.na(pressure_score))

```

## Commercial High Bycatch Data:

```{r}

## combine all years
rasts_hb <- list.files(file.path(outFolder,'comm_high_bycatch'), full = TRUE)
press_stack_hb <- stack(rasts_hb)

# extract data for each region:
regions_stats_hb <- zonal(press_stack_hb,  zones, fun="mean", na.rm=TRUE, progress="text")
regions_stats2_hb <- data.frame(regions_stats_hb)

write.csv(regions_stats2_hb, "int/comm_high_bycatch.csv", row.names = FALSE)

```

Compare zone IDs between intermediate data with **regionData.csv**. Should be no difference in `ant_id` zone ids and the `zone` column in the new data frame.

```{r}
setdiff(regions_stats2_hb$zone, rgn_data$ant_id) 
setdiff(rgn_data$ant_id, regions_stats2_hb$zone) 
```

Tidy and create output data
```{r}
data <- merge(rgn_data, regions_stats2_hb, all.y=TRUE, by.x="rgn_id", by.y="zone") %>%
  tidyr::gather("year", "pressure_score", starts_with("hb")) 

DT::datatable(data) # take a look

hb_data <- data %>%
  dplyr::mutate(year = stringr::str_sub(year, -4, -1)) %>%
  dplyr::mutate(year = as.numeric(year)) %>%
  dplyr::filter(rgn_id <= 250) %>%
  dplyr::select(rgn_id, year, pressure_score) # na values 71, 72, 74, 75, 188, 215

write.csv(hb_data, "output/comm_hb.csv", row.names=FALSE)

```

## Artisanal Low Bycatch:

```{r}

## combine all years
rasts_art <- list.files(file.path(outFolder,'art_low_bycatch'), full = TRUE)
pressure_stack_art <- stack(rasts_art)

# extract data for each region:
regions_stats_art <- zonal(pressure_stack_art,  zones, fun="mean", na.rm=TRUE, progress="text")
regions_stats2_art <- data.frame(regions_stats_art)
write.csv(regions_stats2_art, "int/art_low_bycatch.csv", row.names = FALSE)
```

Compare zone IDs between intermediate data with **regionData.csv**. Should be no difference in `ant_id` zone ids and the `zone` column in the new data frame.

```{r}
setdiff(regions_stats2_art$zone, rgn_data$ant_id) # antarctica regions are in there, makes sense....no land
setdiff(rgn_data$ant_id, regions_stats2_art$zone) # 213 is in there, that makes sense (Antarctica)
```

Tidy and create output data
```{r}

data <- merge(rgn_data, regions_stats2_art, all.y=TRUE, by.x="rgn_id", by.y="zone") %>%
  tidyr::gather("year", "pressure_score", starts_with("art")) 

art_data <- data %>%
  dplyr::mutate(year = stringr::str_sub(year, -4, -1)) %>%
  dplyr::mutate(year = as.numeric(year)) %>%
  dplyr::filter(rgn_id <= 250) %>%
  dplyr::select(rgn_id, year, pressure_score)

write.csv(art_data, "output/art_lb.csv", row.names=FALSE)

```

# Data check
```{r}

## compare with previous year's data

new <- read.csv("output/comm_lb.csv") %>%
  filter(year == 2014) %>%
  dplyr::select(rgn_id, new_pressure_score = pressure_score) 

old_new <- read.csv("../v2017/output/lb.csv") %>%
  filter(year == 2014) %>% 
  left_join(new, by = 'rgn_id')

plot(old$new_pressure_score, old$pressure_score)
abline(0, 1, col="red")


new <- read.csv("output/hb.csv") %>%
  filter(year == 2010) %>%
  dplyr::select(rgn_id, new_pressure_score = pressure_score) 

old <- read.csv("../v2016/output/comm_fish_hb_2016.csv") %>%
  left_join(new, by = 'rgn_id')

plot(old$new_pressure_score, old$pressure_score)
abline(0, 1, col="red")


new <- read.csv("output/art.csv") %>%
  filter(year == 2010) %>%
  dplyr::select(rgn_id, new_pressure_score = pressure_score) 

old <- read.csv("../v2016/output/artisanal_fish_lb_2016.csv") %>%
  left_join(new, by = 'rgn_id')

plot(old$new_pressure_score, old$pressure_score)
abline(0, 1, col="red")

```



