---
title: 'Prepping Primary Productivity data for Commerical Fishing Pressure Layers '
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../ohiprep_v2018/src/templates/ohi_hdr.html'
pdf_document:
  toc: true
---

#Summary

This script calculates mean annual net primary production values globally from 2003 to 2015. This data is used to standardize commercial fishing catch to produce [fishing pressure layers for OHI](https://cdn.rawgit.com/OHI-Science/ohiprep/master/globalprep/prs_fish/v2016/fishing_pressure_layers.html) and Cumulative Human Impacts.


***

#Data Source

**Reference**: Behrenfeld, M.J. and Falkowski, P.G., 1997. Photosynthetic rates derived from satellite‐based chlorophyll concentration. Limnology and oceanography, 42(1), pp.1-20. 

**Downloaded**: July 30, 2016

**Description**:  Monthly Net Primary Production  (mg C / m2 / day)

**Native data resolution**: 0.083 x 0.083 degree grid   

**Time range**: 2002 - 2015, monthly. Only partial data provided for 2002. OHI uses 2003 - 2015.

**Format**:  XYZ format

***
  
# Setup

```{r setup, message = F,warning = F}

library(fields)
library(raster)
library(doParallel)
library(foreach)
library(parallel)
library(RColorBrewer)
library(dplyr)

source("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2018/master/src/R/common.R")
source('vgpm_func.R')

cols = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255)) # rainbow color scheme

#paths

dir_git <- file.path(dir_M,'git-annex/impact_acceleration/stressors/comm_fish/int')
```

```{r unzip, eval=F}
#unzip all files (held in git-annex not marine_threats on Mazu)
tars <- list.files(file.path(dir_M,'git-annex/globalprep/_raw_data/VGPM_primary_productivity/d2016'),full.names = TRUE, pattern = '.tar')

lapply(tars, untar, exdir = file.path(dir_M, 'git-annex/globalprep/_raw_data/VGPM_primary_productivity/d2016'))

```

# .xyz to .tiff

This code sources a function, `vgpm.raster` that transforms the raw .xyz files into GeoTIFFs.

```{r runFunction, eval = F}

files = list.files(file.path(dir_M,'git-annex/globalprep/_raw_data/VGPM_primary_productivity/d2016'), full.names=TRUE, pattern = '.gz')

registerDoParallel(12)

foreach (file = files) %dopar%{
  
  print(file)
  vgpm.raster(file, w.lon, e.lon, n.lat, s.lat, log = TRUE, 
              color = tim.colors(30))
}


```

# Calculate mean annual NPP

Calculate the mean annual Net Primary Production per year and save as rasters.

```{r, eval = F}

#global ocean raster at 1km

ocean = raster(file.path(dir_M,'model/GL-NCEAS-Halpern2008/tmp/ocean.tif'))


# set mollweide projection
moll_crs = CRS("+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs")

registerDoParallel(7)

foreach (i = 2004:2015) %dopar%{ #i=2012
  
  print(i)
  
  l    = list.files(file.path(dir_git,'npp_rasterized_rawdata'), pattern=as.character(i), full.names=TRUE) %>%
          stack() %>%
          calc(fun=function(x){mean(x, na.rm = TRUE)}) %>%
            projectRaster(., crs=moll_crs, over = TRUE, filename = file.path(dir_git,sprintf('npp_annual/annual_mean_npp_%s.tif', i)), overwrite=TRUE)
  
  
  out  = raster::resample(l, ocean, method='ngb',
                 filename = file.path(dir_git,sprintf('npp_annual/annual_mean_npp_moll_1km_%s.tif', i)),overwrite=TRUE)
  
}

```

# Output

```{r output}

raster(file.path(dir_M, 'git-annex/impact_acceleration/stressors/comm_fish/int/npp_annual/annual_mean_npp_moll_1km_2015.tif')) %>% 
  plot(col=cols, box=FALSE, axes=FALSE, main = 'Mean Net Primary Productivity (mg C/m2/day)')

```

***

##Citation information  
Behrenfeld, M.J. and Falkowski, P.G., 1997. Photosynthetic rates derived from satellite‐based chlorophyll concentration. Limnology and oceanography, 42(1), pp.1-20.

Downloaded from http://www.science.oregonstate.edu/ocean.productivity/standard.product.php