---
title: "Global Fisheries Catch"
output: html_document
---

# Summary

This script prepares and formats the IMAS Global Fisheries Catch raw data into intermediate data by combining CatchInd_XXXX_XXXX and CatchNInd_XXXX_XXXX rds files wtih Index.csv and Cells.csv as well as a single file with all years.

# Updates from previous assessment

No longer using gear type to classify catch data as high or low bycatch, since the IMAS data now provides values for **discards**, which we consider as bycatch. Previously, the raw catch data was all in a single file. This year, we have to combine across three different data tables: catch data (CatchInd_XXXX_XXXX), master index file with country name, species, etc (Index.csv), and geospatial information (Cells.csv).

# Data Source

**Reference**: Watson, R. A. and Tidd, A. 2018. Mapping nearly a century and a half of global marine fishing: 1869–2015. Marine Policy, 93, pp. 171-177. [(Paper URL)](https://doi.org/10.1016/j.marpol.2018.04.023)

**Downloaded**: July 17, 2018 from [IMAS portal](http://data.imas.utas.edu.au/portal/search?uuid=ff1274e1-c0ab-411b-a8a2-5a12eb27f2c0) - clikc on download tab, step 3

**Description**:  Global fisheries landings data per cell separated by Industrial versus Non-Industrial catch, IUU, and discards.

**Native data resolution**:   

**Time range**: 1950 - 2015

**Format**:  CSV format

**Additional Information**: [Metadata](http://metadata.imas.utas.edu.au/geonetwork/srv/eng/metadata.show?uuid=ff1274e1-c0ab-411b-a8a2-5a12eb27f2c0), [Supplementary Material](https://ars.els-cdn.com/content/image/1-s2.0-S0308597X18300605-mmc1.docx)


***
  
# Setup

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}

library(tidyverse)
library(purrr)

source("https://rawgit.com/OHI-Science/ohiprep_v2018/master/src/R/common.R")

rawFolder <- file.path(dir_M, "git-annex/globalprep/_raw_data/IMAS_GlobalFisheriesLandings/d2018")

```

# Import Raw Data

_See data_download.R_ if new data is available. 

Filter master file to only include `ID`, `Year`, `CountryName`, `TaxonName`, `CommonName`, `FleetGearName` (I actually don't think we will need the `CountryName` or `FleetGearName`, but it might be interesting).

```{r}
master <- read.csv(file.path(rawFolder,"Index.csv")) %>% 
  select(ID, Year, CountryName, TaxonName, CommonName, FleetGearName)

spatialCells <- read.csv(file.path(rawFolder,"Cells.csv"))
```

# Preliminary Exploration

_See explore_watson_data.Rmd_

# Methods

## Wrangle 

Combine the Master Index and Spatial Cells with the CatchInd and CatchNInd files.

### Function for Combining Files

Create function to read in each file name, combine with Index and Cells data frame, and save everything into one single data file.

```{r}

## Create empty data frame to combine all catch information
bigDataFrame <- data.frame()

## Set function to read in each file, combine with Index.csv and Cells.csv, and save back into mazu
combine_fis <- function(x) {
  
  ## This will be the name of the data frame (w/o extension)
  name <- basename(x) %>% 
    tools::file_path_sans_ext()
  ## Save the data frame into R's memory
  name <- readRDS(x)
  
  ## Create a totals column
  df <- name %>%
    mutate(Landings = Reported+IUU) %>% 
    left_join(master, by = "ID") %>% 
    left_join(spatialCells, by = "Cell")
  
  ## Save intermediate files into mazu with same name as raw
  saveRDS(df, file.path(rawFolder,"int",basename(x)))
  
  ## Add catch data to single data frame
  rbind(df,bigDataFrame)
   
}

```

### Industrial Catch Data

Create list of industrial catch file names and apply the `combine_fis` function to save each individual 5-year interval catch data. Then, create a single file that has all years of data.

The `combine_fis` process should take about 7 minutes. Saving all years into a single file with `saveRDS` will take about 4 minutes.
```{r}

ind_files <- dir(rawFolder, pattern ="CatchInd", full.names=TRUE)

## Wrangle and save each 5-year interval of data
indCatch <- map_df(ind_files, combine_fis)
## Save all years into single file
saveRDS(indCatch, file.path(rawFolder,"int/CatchInd_all.rds"))

```

### Non-industrial Catch Data

```{r}

nind_files <- dir(rawFolder, pattern ="CatchNInd", full.names=TRUE)

## Apply the combine_fis function, which saves each 5-year catch data
nindCatch <- map_df(nind_files, combine_fis)
## Save the 14 separate catch files into a single one
saveRDS(nindCatch, file.path(rawFolder,"int/CatchNInd_all.rds"))

```

