---
title: 'OHI 2018: LSP, Rasterize WDPA polygons'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohiprep/src/templates/ohi_hdr.html'
pdf_document:
  toc: true
---

# Summary

* Setup WDPA shapefile for lasting special places by filtering out non-"Designated" protected areas, and filtering out "non-MPA programmatic management plans" (US protected areas that distort scores).
* Rasterize the result using `fasterize::fasterize()` (not `gdal_rast2()` as in previous assessments) and save to disk using `writeRasterBlocks()` defined in this script.

View lsp_data_prep.Rmd for full description of Lasting Special Places data preparation.

# Updates from previous assessment

Using `fasterize::fasterize()` to rasterize the WDPA polygons, rather than ArcGIS.  This allows the entire process to be completed in R. 

Using `fasterize()` argument `fun=min` to assign oldest year to an area where there are overlapping polygons. Prior assessment re-ordered the status year with the oldest year at the end, so that `gdalUtils::gdal_rasterize()` in `gdal_rast2()` would work correctly. Both achieve the same results.

***

# Setup

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = FALSE, message = FALSE, warning = FALSE, eval=FALSE)
library(rgdal)
library(gdalUtils)
library(tidyverse)
source('~/github/ohiprep_v2018/src/R/common.R')

goal     <- 'globalprep/lsp'
scenario <- 'v2018'
dir_goal_anx   <- file.path(dir_M, 'git-annex', goal, scenario) 
dir_goal  <- file.path('~/github/ohiprep_v2018', goal, scenario)
dir_data <- file.path(dir_M, 'git-annex/globalprep', '_raw_data',
                         'wdpa_mpa/d2018')

## set up provenance tracking for this script:
# library(provRmd); prov_setup() #devtools::install_github('oharac/provRmd', build_vignettes = TRUE)

```


# Methods

## Filter WDPA Shapefile 

Read in the polygons from the WDPA dataset; filter as needed.

``` {r filter_and_reorder_poly}

shp_raw     <- file.path(dir_data, 'WDPA_June2018-shapefile', 'WDPA_June2018-shapefile-polygons')
shp_reorder <- file.path(dir_data, 'shps', 'WDPA_June2018_shp_ordered')

if(!file.exists(paste0(shp_reorder, '.shp'))) {
  message('No shp found for filtered/reordered WDPA database')
  ## Read in the raw shapefile (3.6 GB); 472 sec on Mazu
  message('Reading in raw shapefile: \n  ', shp_raw)
  ptm <- proc.time()
  wdpa_poly <- rgdal::readOGR(dsn = dirname(shp_raw), 
                       layer = basename(shp_raw),
                       stringsAsFactors = FALSE)
  message('elapsed: ', (proc.time() - ptm)[3])
  
  ## filter polygons
  # glimpse(wdpa_poly@data) ## check class of STATUS_YR
  wdpa_poly@data <- wdpa_poly@data %>%
    setNames(tolower(names(.))) %>%
    dplyr::select(wdpaid, name, orig_name, 
           desig, desig_eng, desig_type,
           iucn_cat, 
           marine, no_take, no_tk_area, 
           status, status_yr, 
           mang_auth, mang_plan, verif,
           sub_loc, parent_iso, iso3) %>%
    dplyr::mutate(status_yr = as.integer(status_yr))
  
  # wdpa_poly@data$status %>% unique()
  ## [1] "Designated"   "Proposed"     "Inscribed"    "Not Reported" "Adopted"     

  # x <- wdpa_poly@data %>% filter(str_detect(tolower(mang_plan), 'non-mpa program'))
  ## 84 observations
  wdpa_poly <- wdpa_poly[wdpa_poly@data$status == 'Designated', ]
  wdpa_poly <- wdpa_poly[!str_detect(tolower(wdpa_poly@data$mang_plan), 'non-mpa program'), ]
  
  message('Writing filtered WDPA polygons to: \n  ', shp_reorder)
  ## probably gonna take half an hour to write?
  ptm <- proc.time()
  writeOGR(wdpa_poly, 
           dsn = dirname(shp_reorder), layer = basename(shp_reorder),
           driver = 'ESRI Shapefile')
  message('elapsed: ', (proc.time() - ptm)[3])
  ## lots of warnings similar to:
  ##   "Warning 1: Value 555593627 of field WDPAID of feature 507 not successfully written."
  ## warning ignored: WDPAID field not used in analysis
  
  rm('wdpa_poly') ### clean up memory
  
} else {
  ## filtered shapefile exists; confirm
  message('Filtered shapefile exists: \n  ', shp_reorder)
}

```

## Transform to Mollweide Projection

Transform ordered polygons to Mollweide and save as new polygons.

``` {r transform_poly}

shp_xformed <- file.path(dir_data, 'shps', 'WDPA_June2018_shp_xformed')

if(!file.exists(paste0(shp_xformed, '.shp'))) {
  message('No shp found for filtered/reordered/transformed WDPA database')
  
  if(!exists('wdpa_poly1')) {
    
    message('loading wdpa polygons (filtered and ordered)')
    shp_reorder <- file.path(dir_data, 'shps', 'WDPA_June2018_shp_ordered')
    wdpa_poly1 <- readOGR(dsn = dirname(shp_reorder), 
                          layer = basename(shp_reorder),
                          stringsAsFactors = FALSE)
  }

  message('Spatial transforming WDPA polygons to Mollweide')
  crs_mol <- CRS('+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs')
  ptm <- proc.time()
  wdpa_poly2 <- spTransform(wdpa_poly1, crs_mol)
  message('elapsed: ', (proc.time() - ptm)[3])  ### elapsed: 251 s on Mazu
  
  message('Writing filtered/reorderedtransformed WDPA polygons to: \n  ', shp_xformed)
  ptm <- proc.time()
  writeOGR(wdpa_poly2, dsn = dirname(shp_xformed), layer = basename(shp_xformed),
           driver = 'ESRI Shapefile')
  message('elapsed: ', (proc.time() - ptm)[3])
  ## elapsed: 734.635 s on Mazu
  
  rm('wdpa_poly1', 'wdpa_poly2') ### clean up the memory
  
} else {
  ## transformed shapefile exists
  message('Transformed WDPA shapefile exists at \n  ', shp_xformed)
  message('Go inspect it to make sure it looks good!')
}

```

### Define some raster functions.

Here we define `gdal_rast2()` function as a wrapper for `gdalUtils::gdal_rasterize()`, which has some clunky arguments.  This was originally written for OHIBC.

Here we also define `writeRasterBlocks()` to write out very large rasters chunk-by-chunk to avoid an issue in which `writeRaster()` fails due to inability to work with long vectors (over 2 billion elements?).

``` {r define_gdal_rast2}
## from OHIBC...


writeRasterBlocks <- function(rast, filename) {
  bs <- blockSize(rast)
  rast <- writeStart(rast, filename, overwrite = TRUE)
  ptm <- proc.time()
  for (i in 1:bs$n) { # i <- 1
    message('... writing block ', i, ' of ', bs$n, ' to ', basename(filename), 
        ' (', (proc.time() - ptm)[3], ' sec)')
    v <- getValues(rast, row = bs$row[i], nrows = bs$nrows[i] )
    rast <- writeValues(rast, v, bs$row[i])
  }
  rast <- writeStop(rast)
  # return(rast)
}



```


## Shapefile to Raster: `fasterize()`

`fasterize()` from the `fasterize` package takes advantage of Simple Features objects from the `sf` package, rather than objects from the `sp` package.  It is considerably faster; it returned a completed raster in ten minutes.  However, saving the very large (18GB) resulting raster proved problematic.  The `writeRasterBlocks()` function defined above helped get around that problem though still took over an hour to write the raster to disk.

``` {r fasterize_wdpa, eval = TRUE}

library(sf)
library(fasterize) ### devtools::install_github('ecohealthalliance/fasterize')
rast_wdpa_file   <- file.path(dir_goal_anx, 'rast', 'wdpa_2018_moll_500m.tif')

if(!file.exists(rast_wdpa_file)) {
shp_xformed_file <- file.path(dir_data, 'shps', 'WDPA_June2018_shp_xformed')
  
  ptm <- proc.time()
  wdpa_poly <- sf::st_read(dsn = dirname(shp_xformed_file),
                           layer = basename(shp_xformed_file))
  cat('elapsed: ', (proc.time() - ptm)[3])
  ### 47-53 seconds to read in the polygons (about 472 sec for readOGR)

  rast_base <- raster::raster(file.path(dir_M, 'git-annex/globalprep/spatial/d2014',
                                'data/rgn_mol_raster_500m',
                                'rgn_inland1km_mol_500mcell.tif'))
  
  # test_poly <- sf::st_read(path.expand('~/github/ohibc/prep/spatial'),
  #                             'ohibc_offshore_3nm',
  #                             stringsAsFactors = FALSE) %>%
  #   st_transform(rast_base@crs@projargs)
  # poly_ext <- raster::extent(as.vector(st_bbox(test_poly)))
  # 
  # rast_base <- raster::crop(rast_base, poly_ext)
  # test_poly <- wdpa_poly[1:100, ]
  
  ## For any overlapping polygons, assign the oldest (minimum) status year to that area
  ptm <- proc.time()
  rast_wdpa <- fasterize(wdpa_poly, rast_base, field = 'status_yr', fun = 'min')
  cat('fasterize elapsed: ', (proc.time() - ptm)[3]) 
  ### 358 - 403 seconds?! holy crap, if it rasterized properly...
  
  # ptm <- proc.time()
  # raster::writeRaster(rast_wdpa, rast_wdpa_file)
  # cat('writeRaster elapsed: ', (proc.time() - ptm)[3])
  ### writeRaster failed on an issue with large vectors...
  ### see below about writing block by block approach.
  
  library(raster)
  # rast <- rast_wdpa
  # filename <- rast_wdpa_file
  ptm <- proc.time()
  x <- writeRasterBlocks(rast_wdpa, rast_wdpa_file)
  message('writeRaster elapsed: ', (proc.time() - ptm)[3])
  
}


```

## Data Checking

Compare shapfile 2017 v 2018.

```{r}

library(sf)
library(raster)

## Read in reordered shapefile
# shp_reorder <- file.path(dir_data, 'shps', 'WDPA_June2018_shp_ordered')
# shp_reorder17 <- file.path(dir_data, '../d2017', 'shps', 'WDPA_May2017_shp_ordered')
  
## Read in as shapefile
# wdpa_order <- rgdal::readOGR(dsn = dirname(shp_reorder), 
#                        layer = basename(shp_reorder),
#                        stringsAsFactors = FALSE)
# 
# wdpa_order17 <- rgdal::readOGR(dsn = dirname(shp_reorder17), 
#                        layer = basename(shp_reorder17),
#                        stringsAsFactors = FALSE)

## Subtract wdpa_order17 from wdpa_order to view differences



## Read in as sf
wdpa_sf <- st_read(dsn = file.path(dir_data, 'shps'), layer = "WDPA_June2018_shp_ordered")
wdpa_sf_17 <- st_read(dsn = file.path(dir_data, '../d2017', 'shps'), layer = 'WDPA_May2017_shp_ordered')

plot(wdpa_sf[1])
wdpa_diff <- st_difference(wdpa_sf[1], wdpa_sf_17[1])
plot(wdpa_diff)
ggplot(wdpa_diff) +
  geom_sf()


## Convert back to 'Spatial' object
# wdpa_order <- as(variable, "Spatial")


```

