---
title: 'OHI 2019: LSP Updates Check '
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
---

```{r setup, warning=FALSE, message=FALSE, echo=FALSE}

knitr::opts_chunk$set(eval=FALSE)

library(raster)
library(readr)
library(tidyverse)

dir_goal <- "~/github/ohiprep_v2019/globalprep/lsp/v2019"
```

# Summary

***
Check outliers in output files from Lasting Special Places data prep and compare to Protected Planet's [monthly updates](https://www.protectedplanet.net/c/monthly-updates/) webpage. 

Check outliers (diff > 0.05) in Status plot:


rgn ID    |   rgn name  | status diff  
--------- | -------------- |------- 
  81      | Cyprus            |  0.217 -> 0.404
  107     | Clipperton Island |  0 -> 0.5    
  132     | Colombia          |  0.633 -> 0.9    
  197     | Cameroon          |  0.743 -> 1 
  29      | Mayotte           |  0.933 -> 1
  218     | Canada            | 0.247 -> 0.332

# Check Output Files

Regions with differences after adding non-resilience LSP layers

**If no difference, may need to check whether there are additional protected areas within the EEZ not in the coastal or offshore regions.**
``` {r outlier_checks, messages=F, warnings=F}

mpa2019 <- read_csv(file.path(dir_goal, 'output/lsp_prot_area_offshore3nm.csv'))
tpa2019 <- read_csv(file.path(dir_goal, 'output/lsp_prot_area_inland1km.csv'))

marea2019 <- read_csv(file.path(dir_goal, 'output/rgn_area_offshore3nm.csv'))
tarea2019 <- read_csv(file.path(dir_goal, 'output/rgn_area_inland1km.csv'))

mpa2018 <- read_csv(file.path(dir_goal, '../v2018/output/lsp_prot_area_offshore3nm.csv')) #%>%
#  group_by(rgn_id) %>%
#  mutate(a_prot_3nm = cumsum(a_prot_3nm)) 
tpa2018 <- read_csv(file.path(dir_goal, '../v2018/output/lsp_prot_area_inland1km.csv')) # %>%
#  group_by(rgn_id) %>%
#  mutate(a_prot_1km = cumsum(a_prot_1km)) 
marea2018 <- read_csv(file.path(dir_goal, '../v2018/output/rgn_area_offshore3nm.csv'))
tarea2018 <- read_csv(file.path(dir_goal, '../v2018/output/rgn_area_inland1km.csv'))

#lsp_v2017 <- read_csv('~/github/ohi-global/eez2013/scores.csv') 

## Check a few regions from non-resilience LSP layers
check_rgn <- c(29, 81, 107, 132, 197, 218)

checks <- mpa2019 %>% filter(rgn_id %in% check_rgn) %>% filter(year > 2000) %>% rename(prot_3nm_2019 = a_prot_3nm) %>%
  left_join(tpa2019 %>% filter(rgn_id %in% check_rgn) %>% filter(year > 2000) %>% rename(prot_1km_2019 = a_prot_1km),
            by = c('rgn_id', 'year')) %>%
  left_join(mpa2018 %>% filter(rgn_id %in% check_rgn) %>% filter(year > 2000) %>% rename(prot_3nm_2018 = a_prot_3nm),
            by = c('rgn_id', 'year')) %>%
  left_join(tpa2018 %>% filter(rgn_id %in% check_rgn) %>% filter(year > 2000) %>% rename(prot_1km_2018 = a_prot_1km),
            by = c('rgn_id', 'year')) %>%
  left_join(marea2019 %>% filter(rgn_id %in% check_rgn) %>% rename(marea19 = area),
            by = c('rgn_id')) %>%
  left_join(tarea2019 %>% filter(rgn_id %in% check_rgn) %>% rename(tarea19 = area),
            by = c('rgn_id')) %>%
  left_join(marea2018 %>% filter(rgn_id %in% check_rgn) %>% rename(marea18 = area),
            by = c('rgn_id')) %>%
  left_join(tarea2018 %>% filter(rgn_id %in% check_rgn) %>% rename(tarea18 = area),
            by = c('rgn_id'))

DT::datatable(checks)
```

# Check Intermediate Files

Regions with differences after adding resilience LSP layers. These intermediate files contain both **total area** and **total protected area** within the offshore 3-nm, coastal 1-km, and EEZ region (after processing with zonal stats files and `calc_areas()`).

**If no difference, may need to check whether there are additional protected areas within the EEZ not in the coastal or offshore regions.**

Combine assessment year 2018 and 2019 data
```{r}

## Regions to check: Mayotte, Cyprus, Clipperton Island, Colombia, Cameroon, Canada 
check_rgn <- c(29, 81, 107, 132, 197, 218)

## 2018 offshore 3-nm and coastal 1-km data
off_2019 <- read_csv(file.path(dir_goal, 'int/area_protected_3nm.csv')) %>% 
  filter(year < 2019)
coastal_2019 <- read_csv(file.path(dir_goal, 'int/area_protected_1km.csv')) %>% 
  filter(year < 2019)

## 2017 offshore 3-nm and coastal 1-km data
off_2018 <- read_csv(file.path(dir_goal, '../v2018/int/area_protected_3nm.csv'))
coastal_2018 <- read_csv(file.path(dir_goal, '../v2018/int/area_protected_1km.csv')) 

## Join the above 4 datasets, rename columns to reflect type and assessment year, then summarize the change in values (or area of MPA)
checks <- off_2019 %>% filter(rgn_id %in% check_rgn) %>% rename(prot_3nm_2019 = a_prot_km2, tot_3nm_2019 = a_tot_km2) %>%
  left_join(coastal_2019 %>% filter(rgn_id %in% check_rgn) %>% rename(prot_1km_2019 = a_prot_km2, tot_1km_2019 = a_tot_km2), by = c('rgn_id', 'rgn_name', 'year')) %>%
  left_join(off_2018 %>% filter(rgn_id %in% check_rgn) %>% rename(prot_3nm_2018 = a_prot_km2, tot_3nm_2018 = a_tot_km2), by = c('rgn_id', 'rgn_name', 'year')) %>%
  left_join(coastal_2018 %>% filter(rgn_id %in% check_rgn) %>% rename(prot_1km_2018 = a_prot_km2, tot_1km_2018 = a_tot_km2), by = c('rgn_id', 'rgn_name', 'year')) %>% 
  select(rgn_id, rgn_name, year, tot_3nm_2019, prot_3nm_2019, tot_3nm_2018, prot_3nm_2018, tot_1km_2019, prot_1km_2019, tot_1km_2018, prot_1km_2018)

```

***



