---
title: "25mile_inland_raster"
output: html_document
---

Molly noticed some potential weirdness in the 25 mile inland buffer we have been using for the mariculture goal and an intertidal habitat destruction pressure layer.  I am going to improve the spatial file.


The raster layer has dimensions of 934.4789 m.  This means that about 43 cells equal about 40 km.  Which corresponds to about 25 miles.  I am going to buffer our ocean raster by 43 cells to obtain 25 miles inland. 
```{r}

library(raster)
library(fasterize)
library(dplyr)
library(here)

source('https://raw.githubusercontent.com/OHI-Science/ohiprep_v2019/gh-pages/workflow/R/common.R')

ohi_rasters()
regions_shape()
region_data()

tmp <- boundaries(ocean, type="outer", progress="text")
plot(tmp)
extent_tmp <- zoom(tmp)
plot(regions[1], add=TRUE, col=NA)

for(i in 1:43){
  if(i==1){
    tmp <- boundaries(ocean, type="outer", progress="text")
  } else{ 
tmp <- boundaries(tmp, type="outer", progress="text")}
# plot(crop(tmp, extent_tmp))
# plot(regions[1], add=TRUE, col=NA)
i=i+1
print(i)
}

writeRaster(tmp, file.path(dir_M, "git-annex/globalprep/spatial/v2019/tmp/ocean_plus25milebuffer.tif"))

```


Next step is to crop out the ocean stuff, leaving only the 25 miles inland area.

```{r}

# outer layer of buffered raster values equal 1 while the other layers are 0, given the
# boundaries function methods.  Convert the 1 values to zero.

tmp <- raster::raster(file.path(dir_M, "git-annex/globalprep/spatial/v2019/tmp/ocean_plus25milebuffer.tif"))


tmp[tmp > 0] <- 0

plot(regions[1], add=TRUE, col=NA)

## create raster of regions so we can ID 

eez_land_raster <- fasterize::fasterize(regions, tmp, field="rgn_ant_id")

# To get ID of inland buffer as well as ocean regions: sum the buffered raster with the eez_land_raster
# this should eliminate the NA regions

rast_stack <- stack(tmp, eez_land_raster)

calc(rast_stack, fun=sum, 
     filename = file.path(dir_M, "git-annex/globalprep/spatial/v2019/ocean_plus25mile_inland.tif"),
     progress="text")

```

Check results. (which look good!)
```{r}

buffer <- raster(file.path(dir_M, "git-annex/globalprep/spatial/v2019/ocean_plus25mile_inland.tif"))

plot(buffer)
click(buffer)
plot(crop(tmp, extent_tmp))
plot(regions[1], add=TRUE, col=NA)

## One more check (just want to see what the inland raster looks like without the ocean area included)
regions_land <- regions[regions$rgn_type %in% c("land", "land-disputed"),]
regions_land_raster <- fasterize::fasterize(regions_land, tmp, field = "rgn_id") 

inland_only <- raster::mask(tmp, regions_land_raster, progress="text")

rast_stack <- stack(inland_only, regions_land_raster)

calc(rast_stack, fun=sum, 
     filename = file.path(dir_M, "git-annex/globalprep/spatial/v2019/25mile_inland.tif"),
     progress="text")

inland_only <- raster::raster(file.path(dir_M, "git-annex/globalprep/spatial/v2019/25mile_inland.tif"))
plot(inland_only)

area_25mi_inland <- freq(inland_only, progress="text")
area_25mi_inland_format <- data.frame(area_25mi_inland) %>%
  select(rgn_id = value, raster_cells = count) %>%
  mutate(area_km2 = raster_cells*0.8732508) %>%
  select(rgn_id, area_km2)
  
write.csv(area_25mi_inland_format, here("globalprep/spatial/v2019/output/area_km2_25mi_inland.csv"), row.names=FALSE)

old_area <- read.csv("globalprep/mar_prs_population/v2018/int/area_km2_25mi.csv") %>%
  select(rgn_id, old_area_km2 = area_km2)
area_compare <- left_join(area_25mi_inland_format, old_area, by="rgn_id") %>%
  filter(rgn_id != 213, 
         rgn_id <= 250)

library(ggplot2)
library(plotly)
compare_plot <- ggplot(area_compare, aes(x=area_km2, y=old_area_km2, text=rgn_id))+
  geom_point() + 
  geom_abline(slope=1, intercept=0, color="red")

ggplotly(compare_plot) # 73 (Russia) and 218 (Canada) much higher, this makes sense looking at the previous version of the 25 mile inland data.
```