---
title: 'OHI 2019: Seagrass Condition'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

# Summary

This script generates the condition of seagrass for each OHI region. 


## Updates from previous assessment
Creating an actual script to calculate this.

***
## Data Source 

**Downloaded**: 07/24/2019

**Description**:  
Accelerating loss of seagrasses across the globe threatens coastal ecosystems
https://www.pnas.org/content/pnas/suppl/2009/07/08/0905620106.DCSupplemental/ST1_PDF.pdf


This dataset shows the global distribution of seagrass area. It is a per site, per country, per year dataset. 

**Time range**: 1879-2006


***
# Methods
Reference condition was calculated as the mean of the three oldest years between 1975-1985, or the two earliest years if needed. If data meeting these conditions was not available, we fitted a linear model to all data points, and then used the mean of the predicted values for 1979-1981 as the reference condition. For the current condition we used the mean of the three most recent years after 2000 or the two most recent years. If condition data satisfying these constraints were still not available, we fitted a linear model to all data points, provided that there were at least three data points and then used the mean of the predicted values for 2008-2010 as the current condition and the mean of the predicted values for 1979-1981 as the reference condition. Otherwise, we used neighboring (adjacent) regional averages, weighted by habitat area, or averages weighted by habitat area using seagrass geographical regions as defined by Hemminga and Duarte (2000). We did not project beyond a 15-year timeframe. 


## Setup
``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(readr)      # for read_csv()
library(here)
library(tidyverse)

source(file.path('~/github/ohiprep_v2019/workflow/R/common.R'))

goal     <- 'globalprep/hab_seagrass/v2019'
dir_git  <- file.path('~/github/ohiprep_v2019', goal)
dir_wcmc <- file.path(file.path(dir_M, 'git-annex/globalprep/_raw_data/wcmc_seagrass'))
dir_waycott <- file.path(dir_git, '/waycott_data')
```

**Read in Waycott et al. 2009 data**
```{r}
waycott <- read_csv(file.path(dir_waycott, 'waycott_edited.csv'))
```

